### 2013-02-07 LaTeX/RhodiumConcl.tex
pushing up latex file structure

\section{Conclusion}
\label{conclusion}

If NATO had not intervened, the expected duration of the conflict would have been much shorter. But the rebels would have lost, so there is a trade-off.