### 2014-03-12 LaTeX/RhodiumConcl.tex
Model updates, slight outline updates

\section{Conclusion}
\label{conclusion}

\begin{itemize}
	\item Note that sub-state conflict location may have implications for contagion.  Are regional economic consequences of civil war not a function of civil war existence but proximity of the conflict to cities or resources in neighboring states?
\end{itemize}