### 2014-03-13 LaTeX/RhodiumConcl.tex
mods to outline

\section{Conclusion}
\label{conclusion}

Contributed to our understanding of why some civil conflicts impact economic performance more severely than others using a novel spatial approach.

Note that sub-state conflict location may have implications for contagion.  Are regional economic consequences of civil war not a function of civil war existence but proximity of the conflict to cities or resources in neighboring states?