### 2014-04-02 LaTeX/PreezeText.tex
preeze draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\titlepage
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Motivating Question}

\begin{block}{Puzzle}
High variation in economic performance between countries in the midst of internal armed conflict
\begin{itemize}
\item Mexico
\item India
\item Nicaragua
\end{itemize}
\end{block}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Literature Review}

\begin{itemize}
	\item Civil War $\rightarrow$ Economic Performance
	\item Economic Performance $\rightarrow$ Civil War
	\item Disaggregating Civil Wars
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Disaggregating Conflict and Growth}

\begin{block}{Our argument}
State economic performance in countries experiencing civil conflict is conditional on the proximity of conflict from major population, economic, and resource centers
\end{block}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{FARC guerillas \& the Colombian government}

\begin{quote}FARC's strategy and [beliefs have] always been to make economic pressure on both, multinational companies and the Colombian government. This has been done by attacking oil and natural gas infrastructure affecting companies such as Pacific Rubiales Energy, Oxy and Ecopetrol. For non-fuel related international companies with subsidiaries in Colombia, such as Goodyear, Nestle, Microsoft, Toyota, among others, FARC’s modus operandi was mainly racketeering, kidnappings and extortion. (Flannery 2012)\end{quote}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Conflict \& City Data}

\vspace{-0.25cm}
\begin{figure}[ht]
  \centering
  \includegraphics[width=1\textwidth]{CityConfMap-crop}
\end{figure}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Aggregating to Country-Year}

\begin{figure}[ht]
	\centering
	\includegraphics[width=.6\textwidth]{colombiaMap-crop}
\end{figure}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Estimation Approach}

\begin{align*}
	\% \Delta GDP_{i,t} &= \beta_{1}(Upper \; Income_{i,t}) \\
	& \;+\; \beta_{2}(Conflict \; Intensity_{i,t}) \;+\; \beta_{3}(Ln(Conflict \; Area)_{i,t}) \\
	& \;+\; \beta_{4}(Ln(Min. \; Conflict \; Dist.)_{i,t}) \;+\; \beta_{5}(Conflict \; Type_{i,t}) \\
	& \;+\; \beta_{6}(Conflict \; Duration_{i,t}) \;+\; \beta_{7}(Inflation_{i,t-1}) \\
	& \;+\; \beta_{8}(Ln(Land \; Area)_{i,t}) \;+\; \alpha_{i} + \gamma_{t} + \mu_{i,t}
\end{align*}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Findings}

\begin{figure}[ht]
	\centering
	\resizebox{.8\textwidth}{!}{\input{mod3CoefPlot.tex}}
\end{figure}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Substantive Effects}

\begin{figure}[ht]
	\centering
	\resizebox{.9\textwidth}{!}{\input{mod3SimPlot.tex}}
\end{figure}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Conclusions \& Next Steps}

\begin{block}{Conclusions}
\begin{itemize}
	\item Novel approach in the study of civil conflict to distinguish between spatially dissimilar events
	\item Proximity of conflict to major cities explains disparate economic outcomes
\end{itemize}
\end{block}

\begin{block}{Next Steps}
\begin{itemize}
\item Account for other economically important locations (e.g., oil fields)
\item Alternative methods to address aggregation problem
\item More refined analysis within a country using subnational economic data
\end{itemize}
\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%