### 2014-09-23 LaTeX/mCapSimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-09-23 11:40:21
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 39.69, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 39.69, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 57.05,107.99) --
	( 61.80,109.63) --
	( 66.56,111.28) --
	( 71.31,112.93) --
	( 76.07,114.58) --
	( 80.82,116.22) --
	( 85.58,117.87) --
	( 90.34,119.52) --
	( 95.09,121.16) --
	( 99.85,122.81) --
	(104.60,124.46) --
	(109.36,126.11) --
	(114.11,127.75) --
	(118.87,129.40) --
	(123.63,131.05) --
	(128.38,132.70) --
	(133.14,134.34) --
	(137.89,135.99) --
	(142.65,137.64) --
	(147.40,139.28) --
	(152.16,140.93) --
	(156.92,142.58) --
	(161.67,144.23) --
	(166.43,145.87) --
	(171.18,147.52) --
	(175.94,149.17) --
	(180.70,150.81) --
	(185.45,152.46) --
	(190.21,154.11) --
	(194.96,155.76) --
	(199.72,157.40) --
	(204.47,159.05) --
	(209.23,160.70) --
	(213.99,162.34) --
	(218.74,163.99) --
	(223.50,165.64) --
	(228.25,167.29) --
	(233.01,168.93) --
	(237.76,170.58) --
	(242.52,172.23) --
	(247.28,173.88) --
	(252.03,175.52) --
	(256.79,177.17) --
	(261.54,178.82) --
	(266.30,180.46) --
	(271.05,182.11) --
	(275.81,183.76) --
	(280.57,185.41) --
	(285.32,187.05) --
	(290.08,188.70) --
	(294.83,190.35) --
	(299.59,191.99) --
	(304.35,193.64) --
	(309.10,195.29) --
	(313.86,196.94) --
	(318.61,198.58) --
	(323.37,200.23) --
	(328.12,201.88) --
	(332.88,203.52) --
	(337.64,205.17) --
	(342.39,206.82) --
	(347.15,208.47) --
	(351.90,210.11) --
	(356.66,211.76) --
	(361.41,213.41) --
	(366.17,215.05) --
	(370.93,216.70) --
	(375.68,218.35) --
	(380.44,220.00) --
	(385.19,221.64) --
	(389.95,223.29) --
	(394.70,224.94) --
	(399.46,226.59) --
	(404.22,228.23);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 57.05,170.30) --
	( 61.80,170.75) --
	( 66.56,171.43) --
	( 71.31,172.00) --
	( 76.07,172.73) --
	( 80.82,173.39) --
	( 85.58,174.13) --
	( 90.34,174.65) --
	( 95.09,175.20) --
	( 99.85,175.83) --
	(104.60,176.50) --
	(109.36,177.16) --
	(114.11,177.80) --
	(118.87,178.59) --
	(123.63,179.29) --
	(128.38,179.81) --
	(133.14,180.41) --
	(137.89,181.20) --
	(142.65,181.89) --
	(147.40,182.78) --
	(152.16,183.41) --
	(156.92,183.93) --
	(161.67,184.69) --
	(166.43,185.42) --
	(171.18,186.20) --
	(175.94,187.05) --
	(180.70,188.02) --
	(185.45,188.60) --
	(190.21,189.50) --
	(194.96,190.46) --
	(199.72,191.24) --
	(204.47,192.15) --
	(209.23,193.07) --
	(213.99,194.11) --
	(218.74,195.08) --
	(223.50,196.00) --
	(228.25,196.95) --
	(233.01,198.08) --
	(237.76,199.37) --
	(242.52,200.49) --
	(247.28,201.79) --
	(252.03,203.07) --
	(256.79,204.30) --
	(261.54,205.54) --
	(266.30,207.04) --
	(271.05,208.33) --
	(275.81,209.80) --
	(280.57,211.29) --
	(285.32,212.83) --
	(290.08,214.53) --
	(294.83,216.23) --
	(299.59,217.85) --
	(304.35,219.60) --
	(309.10,221.58) --
	(313.86,223.41) --
	(318.61,225.32) --
	(323.37,227.31) --
	(328.12,229.50) --
	(332.88,231.70) --
	(337.64,233.85) --
	(342.39,235.89) --
	(347.15,238.03) --
	(351.90,240.17) --
	(356.66,242.38) --
	(361.41,244.58) --
	(366.17,246.95) --
	(370.93,249.06) --
	(375.68,251.34) --
	(380.44,253.72) --
	(385.19,255.97) --
	(389.95,258.60) --
	(394.70,261.10) --
	(399.46,263.65) --
	(404.22,265.99) --
	(404.22,189.23) --
	(399.46,188.51) --
	(394.70,187.90) --
	(389.95,187.02) --
	(385.19,186.13) --
	(380.44,185.09) --
	(375.68,184.39) --
	(370.93,183.61) --
	(366.17,182.69) --
	(361.41,181.83) --
	(356.66,180.81) --
	(351.90,179.77) --
	(347.15,178.57) --
	(342.39,177.44) --
	(337.64,176.32) --
	(332.88,175.29) --
	(328.12,174.08) --
	(323.37,172.90) --
	(318.61,171.57) --
	(313.86,170.24) --
	(309.10,168.76) --
	(304.35,167.31) --
	(299.59,165.82) --
	(294.83,164.52) --
	(290.08,162.91) --
	(285.32,161.16) --
	(280.57,159.46) --
	(275.81,157.79) --
	(271.05,155.95) --
	(266.30,154.07) --
	(261.54,152.24) --
	(256.79,150.23) --
	(252.03,148.25) --
	(247.28,146.22) --
	(242.52,143.99) --
	(237.76,141.83) --
	(233.01,139.84) --
	(228.25,137.63) --
	(223.50,135.49) --
	(218.74,133.10) --
	(213.99,130.62) --
	(209.23,128.31) --
	(204.47,125.87) --
	(199.72,123.59) --
	(194.96,121.22) --
	(190.21,118.75) --
	(185.45,116.43) --
	(180.70,114.03) --
	(175.94,111.47) --
	(171.18,108.99) --
	(166.43,106.38) --
	(161.67,103.74) --
	(156.92,101.22) --
	(152.16, 98.74) --
	(147.40, 96.24) --
	(142.65, 93.73) --
	(137.89, 91.21) --
	(133.14, 88.50) --
	(128.38, 85.77) --
	(123.63, 83.16) --
	(118.87, 80.48) --
	(114.11, 77.79) --
	(109.36, 75.08) --
	(104.60, 72.30) --
	( 99.85, 69.52) --
	( 95.09, 66.86) --
	( 90.34, 64.14) --
	( 85.58, 61.43) --
	( 80.82, 58.71) --
	( 76.07, 55.94) --
	( 71.31, 53.21) --
	( 66.56, 50.50) --
	( 61.80, 47.80) --
	( 57.05, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 57.05,159.56) --
	( 61.80,160.25) --
	( 66.56,160.98) --
	( 71.31,161.86) --
	( 76.07,162.52) --
	( 80.82,163.37) --
	( 85.58,164.21) --
	( 90.34,164.90) --
	( 95.09,165.74) --
	( 99.85,166.75) --
	(104.60,167.59) --
	(109.36,168.46) --
	(114.11,169.19) --
	(118.87,170.16) --
	(123.63,170.99) --
	(128.38,171.84) --
	(133.14,172.69) --
	(137.89,173.59) --
	(142.65,174.41) --
	(147.40,175.17) --
	(152.16,176.19) --
	(156.92,177.17) --
	(161.67,178.09) --
	(166.43,178.94) --
	(171.18,179.97) --
	(175.94,180.98) --
	(180.70,181.99) --
	(185.45,182.96) --
	(190.21,183.82) --
	(194.96,184.88) --
	(199.72,185.88) --
	(204.47,186.83) --
	(209.23,187.76) --
	(213.99,188.99) --
	(218.74,190.13) --
	(223.50,191.26) --
	(228.25,192.31) --
	(233.01,193.53) --
	(237.76,194.62) --
	(242.52,195.84) --
	(247.28,197.18) --
	(252.03,198.48) --
	(256.79,199.68) --
	(261.54,201.07) --
	(266.30,202.56) --
	(271.05,204.06) --
	(275.81,205.62) --
	(280.57,207.24) --
	(285.32,209.00) --
	(290.08,210.51) --
	(294.83,212.14) --
	(299.59,213.87) --
	(304.35,215.72) --
	(309.10,217.52) --
	(313.86,219.34) --
	(318.61,221.26) --
	(323.37,223.16) --
	(328.12,225.10) --
	(332.88,227.06) --
	(337.64,228.96) --
	(342.39,231.01) --
	(347.15,233.07) --
	(351.90,235.12) --
	(356.66,237.31) --
	(361.41,239.43) --
	(366.17,241.62) --
	(370.93,243.83) --
	(375.68,246.04) --
	(380.44,248.34) --
	(385.19,250.61) --
	(389.95,252.92) --
	(394.70,255.34) --
	(399.46,257.65) --
	(404.22,260.08) --
	(404.22,196.18) --
	(399.46,195.16) --
	(394.70,194.11) --
	(389.95,193.22) --
	(385.19,192.29) --
	(380.44,191.22) --
	(375.68,190.29) --
	(370.93,189.27) --
	(366.17,188.31) --
	(361.41,187.21) --
	(356.66,185.94) --
	(351.90,184.86) --
	(347.15,183.77) --
	(342.39,182.69) --
	(337.64,181.59) --
	(332.88,180.34) --
	(328.12,179.09) --
	(323.37,177.75) --
	(318.61,176.29) --
	(313.86,174.79) --
	(309.10,173.33) --
	(304.35,171.86) --
	(299.59,170.35) --
	(294.83,168.77) --
	(290.08,167.23) --
	(285.32,165.59) --
	(280.57,163.88) --
	(275.81,162.14) --
	(271.05,160.29) --
	(266.30,158.47) --
	(261.54,156.48) --
	(256.79,154.63) --
	(252.03,152.76) --
	(247.28,150.64) --
	(242.52,148.69) --
	(237.76,146.57) --
	(233.01,144.54) --
	(228.25,142.40) --
	(223.50,140.17) --
	(218.74,137.98) --
	(213.99,135.99) --
	(209.23,133.82) --
	(204.47,131.54) --
	(199.72,129.37) --
	(194.96,127.03) --
	(190.21,124.77) --
	(185.45,122.49) --
	(180.70,120.16) --
	(175.94,117.80) --
	(171.18,115.40) --
	(166.43,113.13) --
	(161.67,110.68) --
	(156.92,108.42) --
	(152.16,105.89) --
	(147.40,103.39) --
	(142.65,100.83) --
	(137.89, 98.32) --
	(133.14, 95.77) --
	(128.38, 93.33) --
	(123.63, 90.81) --
	(118.87, 88.34) --
	(114.11, 85.90) --
	(109.36, 83.33) --
	(104.60, 80.84) --
	( 99.85, 78.40) --
	( 95.09, 75.83) --
	( 90.34, 73.27) --
	( 85.58, 70.75) --
	( 80.82, 68.24) --
	( 76.07, 65.78) --
	( 71.31, 63.28) --
	( 66.56, 60.67) --
	( 61.80, 58.16) --
	( 57.05, 55.65) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.69, 34.03) --
	( 39.69,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57,101.49) {3.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57,183.81) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57,266.13) {4.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.69, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 41.46, 20.31) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 89.02, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (136.58, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (184.13, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (231.69, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (279.25, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.81, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (374.36, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (230.63,  4.82) {Ln(Min. Cap Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}