### 2014-09-23 LaTeX/mCitySimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-09-23 11:39:59
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 39.69, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 39.69, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 57.05,110.46) --
	( 62.08,112.14) --
	( 67.11,113.83) --
	( 72.14,115.51) --
	( 77.17,117.20) --
	( 82.20,118.88) --
	( 87.23,120.57) --
	( 92.27,122.25) --
	( 97.30,123.94) --
	(102.33,125.62) --
	(107.36,127.30) --
	(112.39,128.99) --
	(117.42,130.67) --
	(122.45,132.36) --
	(127.49,134.04) --
	(132.52,135.73) --
	(137.55,137.41) --
	(142.58,139.10) --
	(147.61,140.78) --
	(152.64,142.47) --
	(157.67,144.15) --
	(162.71,145.84) --
	(167.74,147.52) --
	(172.77,149.20) --
	(177.80,150.89) --
	(182.83,152.57) --
	(187.86,154.26) --
	(192.89,155.94) --
	(197.93,157.63) --
	(202.96,159.31) --
	(207.99,161.00) --
	(213.02,162.68) --
	(218.05,164.37) --
	(223.08,166.05) --
	(228.12,167.74) --
	(233.15,169.42) --
	(238.18,171.11) --
	(243.21,172.79) --
	(248.24,174.47) --
	(253.27,176.16) --
	(258.30,177.84) --
	(263.34,179.53) --
	(268.37,181.21) --
	(273.40,182.90) --
	(278.43,184.58) --
	(283.46,186.27) --
	(288.49,187.95) --
	(293.52,189.64) --
	(298.56,191.32) --
	(303.59,193.01) --
	(308.62,194.69) --
	(313.65,196.38) --
	(318.68,198.06) --
	(323.71,199.74) --
	(328.74,201.43) --
	(333.78,203.11) --
	(338.81,204.80) --
	(343.84,206.48) --
	(348.87,208.17) --
	(353.90,209.85) --
	(358.93,211.54) --
	(363.96,213.22) --
	(369.00,214.91) --
	(374.03,216.59) --
	(379.06,218.28) --
	(384.09,219.96) --
	(389.12,221.65) --
	(394.15,223.33) --
	(399.18,225.01) --
	(404.22,226.70);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 57.05,174.43) --
	( 62.08,174.80) --
	( 67.11,175.43) --
	( 72.14,175.95) --
	( 77.17,176.45) --
	( 82.20,177.01) --
	( 87.23,177.60) --
	( 92.27,178.07) --
	( 97.30,178.83) --
	(102.33,179.63) --
	(107.36,180.19) --
	(112.39,180.68) --
	(117.42,181.22) --
	(122.45,181.91) --
	(127.49,182.61) --
	(132.52,183.22) --
	(137.55,183.85) --
	(142.58,184.72) --
	(147.61,185.27) --
	(152.64,185.85) --
	(157.67,186.68) --
	(162.71,187.40) --
	(167.74,188.19) --
	(172.77,188.79) --
	(177.80,189.72) --
	(182.83,190.53) --
	(187.86,191.58) --
	(192.89,192.32) --
	(197.93,193.15) --
	(202.96,194.08) --
	(207.99,195.20) --
	(213.02,196.17) --
	(218.05,197.04) --
	(223.08,198.12) --
	(228.12,199.15) --
	(233.15,200.19) --
	(238.18,201.29) --
	(243.21,202.47) --
	(248.24,203.82) --
	(253.27,205.19) --
	(258.30,206.48) --
	(263.34,207.77) --
	(268.37,209.20) --
	(273.40,210.73) --
	(278.43,212.20) --
	(283.46,213.76) --
	(288.49,215.30) --
	(293.52,216.81) --
	(298.56,218.76) --
	(303.59,220.71) --
	(308.62,222.73) --
	(313.65,224.51) --
	(318.68,226.51) --
	(323.71,228.57) --
	(328.74,230.64) --
	(333.78,232.65) --
	(338.81,234.73) --
	(343.84,236.98) --
	(348.87,239.31) --
	(353.90,241.60) --
	(358.93,243.92) --
	(363.96,246.40) --
	(369.00,248.74) --
	(374.03,251.08) --
	(379.06,253.40) --
	(384.09,255.89) --
	(389.12,258.42) --
	(394.15,260.84) --
	(399.18,263.30) --
	(404.22,265.99) --
	(404.22,186.70) --
	(399.18,185.91) --
	(394.15,185.08) --
	(389.12,184.30) --
	(384.09,183.58) --
	(379.06,182.70) --
	(374.03,181.87) --
	(369.00,180.89) --
	(363.96,179.93) --
	(358.93,178.91) --
	(353.90,178.04) --
	(348.87,176.86) --
	(343.84,175.89) --
	(338.81,174.74) --
	(333.78,173.45) --
	(328.74,172.28) --
	(323.71,171.02) --
	(318.68,169.43) --
	(313.65,167.97) --
	(308.62,166.54) --
	(303.59,165.18) --
	(298.56,163.70) --
	(293.52,162.08) --
	(288.49,160.29) --
	(283.46,158.60) --
	(278.43,156.91) --
	(273.40,155.18) --
	(268.37,153.23) --
	(263.34,151.42) --
	(258.30,149.61) --
	(253.27,147.65) --
	(248.24,145.45) --
	(243.21,143.23) --
	(238.18,141.02) --
	(233.15,138.72) --
	(228.12,136.35) --
	(223.08,133.80) --
	(218.05,131.34) --
	(213.02,128.83) --
	(207.99,126.45) --
	(202.96,124.00) --
	(197.93,121.57) --
	(192.89,119.03) --
	(187.86,116.41) --
	(182.83,113.90) --
	(177.80,111.23) --
	(172.77,108.70) --
	(167.74,105.98) --
	(162.71,103.34) --
	(157.67,100.79) --
	(152.64, 98.10) --
	(147.61, 95.49) --
	(142.58, 92.64) --
	(137.55, 89.85) --
	(132.52, 87.02) --
	(127.49, 84.25) --
	(122.45, 81.44) --
	(117.42, 78.59) --
	(112.39, 75.96) --
	(107.36, 73.32) --
	(102.33, 70.44) --
	( 97.30, 67.62) --
	( 92.27, 64.83) --
	( 87.23, 61.80) --
	( 82.20, 58.96) --
	( 77.17, 56.21) --
	( 72.14, 53.52) --
	( 67.11, 50.65) --
	( 62.08, 47.79) --
	( 57.05, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 57.05,165.69) --
	( 62.08,166.26) --
	( 67.11,166.92) --
	( 72.14,167.65) --
	( 77.17,168.34) --
	( 82.20,169.03) --
	( 87.23,169.66) --
	( 92.27,170.38) --
	( 97.30,171.07) --
	(102.33,171.75) --
	(107.36,172.44) --
	(112.39,173.12) --
	(117.42,173.84) --
	(122.45,174.45) --
	(127.49,175.12) --
	(132.52,175.81) --
	(137.55,176.59) --
	(142.58,177.35) --
	(147.61,178.25) --
	(152.64,179.27) --
	(157.67,180.18) --
	(162.71,181.11) --
	(167.74,181.96) --
	(172.77,182.76) --
	(177.80,183.71) --
	(182.83,184.54) --
	(187.86,185.47) --
	(192.89,186.43) --
	(197.93,187.51) --
	(202.96,188.46) --
	(207.99,189.58) --
	(213.02,190.58) --
	(218.05,191.75) --
	(223.08,192.84) --
	(228.12,194.00) --
	(233.15,195.12) --
	(238.18,196.45) --
	(243.21,197.65) --
	(248.24,199.02) --
	(253.27,200.34) --
	(258.30,201.75) --
	(263.34,203.21) --
	(268.37,204.61) --
	(273.40,206.24) --
	(278.43,207.82) --
	(283.46,209.30) --
	(288.49,210.93) --
	(293.52,212.69) --
	(298.56,214.53) --
	(303.59,216.36) --
	(308.62,218.21) --
	(313.65,220.05) --
	(318.68,222.00) --
	(323.71,224.00) --
	(328.74,225.95) --
	(333.78,227.98) --
	(338.81,230.07) --
	(343.84,232.25) --
	(348.87,234.41) --
	(353.90,236.60) --
	(358.93,238.76) --
	(363.96,240.91) --
	(369.00,243.13) --
	(374.03,245.40) --
	(379.06,247.74) --
	(384.09,250.09) --
	(389.12,252.61) --
	(394.15,255.21) --
	(399.18,257.56) --
	(404.22,259.93) --
	(404.22,192.89) --
	(399.18,192.12) --
	(394.15,191.18) --
	(389.12,190.22) --
	(384.09,189.28) --
	(379.06,188.36) --
	(374.03,187.32) --
	(369.00,186.30) --
	(363.96,185.20) --
	(358.93,184.18) --
	(353.90,182.99) --
	(348.87,181.90) --
	(343.84,180.64) --
	(338.81,179.59) --
	(333.78,178.30) --
	(328.74,176.91) --
	(323.71,175.49) --
	(318.68,174.05) --
	(313.65,172.59) --
	(308.62,171.06) --
	(303.59,169.50) --
	(298.56,167.92) --
	(293.52,166.28) --
	(288.49,164.66) --
	(283.46,162.97) --
	(278.43,161.38) --
	(273.40,159.49) --
	(268.37,157.55) --
	(263.34,155.74) --
	(258.30,153.81) --
	(253.27,151.84) --
	(248.24,149.82) --
	(243.21,147.69) --
	(238.18,145.73) --
	(233.15,143.60) --
	(228.12,141.48) --
	(223.08,139.15) --
	(218.05,136.75) --
	(213.02,134.58) --
	(207.99,132.15) --
	(202.96,129.81) --
	(197.93,127.34) --
	(192.89,124.90) --
	(187.86,122.54) --
	(182.83,119.97) --
	(177.80,117.68) --
	(172.77,115.15) --
	(167.74,112.73) --
	(162.71,110.22) --
	(157.67,107.87) --
	(152.64,105.34) --
	(147.61,102.80) --
	(142.58,100.16) --
	(137.55, 97.55) --
	(132.52, 94.94) --
	(127.49, 92.32) --
	(122.45, 89.72) --
	(117.42, 87.08) --
	(112.39, 84.45) --
	(107.36, 81.80) --
	(102.33, 79.16) --
	( 97.30, 76.57) --
	( 92.27, 74.01) --
	( 87.23, 71.42) --
	( 82.20, 68.63) --
	( 77.17, 65.97) --
	( 72.14, 63.37) --
	( 67.11, 60.74) --
	( 62.08, 58.05) --
	( 57.05, 55.38) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.69, 34.03) --
	( 39.69,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57, 52.86) {3.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57, 96.58) {3.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57,140.30) {3.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57,184.02) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57,227.74) {4.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 32.57,271.45) {4.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.69, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 40.56, 20.31) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 90.87, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (141.19, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (191.50, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (241.82, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (292.13, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (342.45, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (392.76, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (230.63,  4.82) {Ln(Min. City Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}