### 2014-09-25 LaTeX/mCapCoefPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-09-25 14:01:56
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] (137.46, 34.03) rectangle (277.03,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] (137.46, 34.03) rectangle (277.03,277.03);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (207.65, 49.88) -- (215.05, 49.88);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (198.55, 76.30) -- (233.74, 76.30);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (176.10,102.71) -- (190.46,102.71);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (154.85,129.12) -- (247.54,129.12);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (172.69,155.53) -- (201.44,155.53);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (207.34,181.95) -- (208.56,181.95);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (197.01,208.36) -- (234.79,208.36);

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (183.40,234.77) -- (211.73,234.77);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (209.89,261.19) -- (222.37,261.19);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (208.25, 49.88) -- (214.45, 49.88);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (201.38, 76.30) -- (230.91, 76.30);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (177.26,102.71) -- (189.31,102.71);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (162.30,129.12) -- (240.08,129.12);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (175.00,155.53) -- (199.13,155.53);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (207.43,181.95) -- (208.46,181.95);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (200.05,208.36) -- (231.75,208.36);

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (185.67,234.77) -- (209.45,234.77);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (210.89,261.19) -- (221.37,261.19);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}
\definecolor[named]{fillColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round,fill=fillColor] (207.25, 34.03) -- (207.25,277.03);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (216.13,261.19) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (197.56,234.77) circle (  2.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (215.90,208.36) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (207.95,181.95) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.07,155.53) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (201.19,129.12) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.28,102.71) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (216.15, 76.30) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (211.35, 49.88) circle (  2.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.05, 48.56) --
	(215.05, 51.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.05, 49.88) --
	(207.65, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.65, 48.56) --
	(207.65, 51.20);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (233.74, 74.97) --
	(233.74, 77.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (233.74, 76.30) --
	(198.55, 76.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.55, 74.97) --
	(198.55, 77.62);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.46,101.39) --
	(190.46,104.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.46,102.71) --
	(176.10,102.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (176.10,101.39) --
	(176.10,104.03);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.54,127.80) --
	(247.54,130.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.54,129.12) --
	(154.85,129.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.85,127.80) --
	(154.85,130.44);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.44,154.21) --
	(201.44,156.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.44,155.53) --
	(172.69,155.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.69,154.21) --
	(172.69,156.86);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.6pt,line join=round] (208.56,180.63) --
	(208.56,183.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (208.56,181.95) --
	(207.34,181.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.34,180.63) --
	(207.34,183.27);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (234.79,207.04) --
	(234.79,209.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (234.79,208.36) --
	(197.01,208.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.01,207.04) --
	(197.01,209.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.73,233.45) --
	(211.73,236.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.73,234.77) --
	(183.40,234.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.40,233.45) --
	(183.40,236.09);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.6pt,line join=round] (222.37,259.87) --
	(222.37,262.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (222.37,261.19) --
	(209.89,261.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.89,259.87) --
	(209.89,262.51);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.46, 34.03) --
	(137.46,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35, 46.58) {World GDP Growth$_{t}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35, 72.99) {Democracy$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35, 99.40) {Ln(Inflation)$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35,125.82) {Upper Income};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35,152.23) {Area$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35,178.64) {Duration$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35,205.06) {Type$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35,231.47) {Intensity$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (130.35,257.88) {Ln(Min. Capital Dist.)$_{t-1}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.46, 34.03) --
	(277.03, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (143.80, 20.31) {-0.004};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (175.53, 20.31) {-0.002};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (207.25, 20.31) {0.000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (238.97, 20.31) {0.002};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (270.69, 20.31) {0.004};
\end{scope}
\end{tikzpicture}