### 2014-09-25 LaTeX/mCapSimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-09-25 14:22:37
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 52.48, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 52.48, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 69.26,107.02) --
	( 76.56,109.63) --
	( 80.61,111.07) --
	( 80.74,111.12) --
	( 93.47,115.66) --
	(101.51,118.53) --
	(106.88,120.45) --
	(109.99,121.56) --
	(116.04,123.72) --
	(124.61,126.77) --
	(153.74,137.17) --
	(166.44,141.70) --
	(169.66,142.85) --
	(182.48,147.43) --
	(189.23,149.83) --
	(198.25,153.05) --
	(202.82,154.68) --
	(205.60,155.68) --
	(205.69,155.71) --
	(206.30,155.92) --
	(207.92,156.50) --
	(209.67,157.13) --
	(214.56,158.87) --
	(214.75,158.94) --
	(219.02,160.46) --
	(220.78,161.09) --
	(224.96,162.59) --
	(226.35,163.08) --
	(226.45,163.12) --
	(226.69,163.20) --
	(227.16,163.37) --
	(229.07,164.05) --
	(229.22,164.10) --
	(229.80,164.31) --
	(230.46,164.55) --
	(231.37,164.87) --
	(232.05,165.12) --
	(232.29,165.20) --
	(232.34,165.22) --
	(232.73,165.36) --
	(237.64,167.11) --
	(239.57,167.80) --
	(239.75,167.86) --
	(240.75,168.22) --
	(241.66,168.54) --
	(242.52,168.85) --
	(243.48,169.19) --
	(244.41,169.52) --
	(245.54,169.93) --
	(245.55,169.93) --
	(247.34,170.57) --
	(247.36,170.58) --
	(248.05,170.82) --
	(248.78,171.09) --
	(249.23,171.25) --
	(249.60,171.38) --
	(252.59,172.45) --
	(255.95,173.64) --
	(257.34,174.14) --
	(257.79,174.30) --
	(257.94,174.35) --
	(258.24,174.46) --
	(259.46,174.90) --
	(259.93,175.06) --
	(261.11,175.48) --
	(261.30,175.55) --
	(261.73,175.70) --
	(263.06,176.18) --
	(264.22,176.59) --
	(266.18,177.29) --
	(267.22,177.66) --
	(267.29,177.69) --
	(268.47,178.11) --
	(269.59,178.51) --
	(270.46,178.82) --
	(271.09,179.05) --
	(271.37,179.15) --
	(271.86,179.32) --
	(271.89,179.33) --
	(272.02,179.38) --
	(272.38,179.51) --
	(272.44,179.53) --
	(273.54,179.92) --
	(273.54,179.92) --
	(274.54,180.28) --
	(274.62,180.31) --
	(274.74,180.35) --
	(274.93,180.42) --
	(275.09,180.47) --
	(275.27,180.54) --
	(275.51,180.62) --
	(275.85,180.75) --
	(276.03,180.81) --
	(276.19,180.86) --
	(276.46,180.96) --
	(276.61,181.01) --
	(278.14,181.56) --
	(278.91,181.83) --
	(278.99,181.86) --
	(279.48,182.04) --
	(279.57,182.07) --
	(279.84,182.17) --
	(280.03,182.23) --
	(280.82,182.52) --
	(281.34,182.70) --
	(281.36,182.71) --
	(281.96,182.92) --
	(282.46,183.10) --
	(282.56,183.14) --
	(283.13,183.34) --
	(283.17,183.36) --
	(283.65,183.53) --
	(283.74,183.56) --
	(285.14,184.06) --
	(285.23,184.09) --
	(285.29,184.11) --
	(285.32,184.13) --
	(285.65,184.24) --
	(286.10,184.40) --
	(286.16,184.42) --
	(286.46,184.53) --
	(286.60,184.58) --
	(286.65,184.60) --
	(286.71,184.62) --
	(286.73,184.63) --
	(286.83,184.66) --
	(286.86,184.67) --
	(287.36,184.85) --
	(287.70,184.97) --
	(287.78,185.00) --
	(289.22,185.51) --
	(289.39,185.58) --
	(289.45,185.60) --
	(289.83,185.73) --
	(289.89,185.75) --
	(290.32,185.91) --
	(290.49,185.97) --
	(290.52,185.98) --
	(290.79,186.08) --
	(290.84,186.09) --
	(290.85,186.10) --
	(290.85,186.10) --
	(290.88,186.11) --
	(290.92,186.12) --
	(291.46,186.31) --
	(291.68,186.39) --
	(292.03,186.52) --
	(292.66,186.74) --
	(292.90,186.83) --
	(293.68,187.11) --
	(294.05,187.24) --
	(294.33,187.34) --
	(294.72,187.48) --
	(294.89,187.54) --
	(294.98,187.57) --
	(295.14,187.63) --
	(295.44,187.73) --
	(295.54,187.77) --
	(295.80,187.86) --
	(296.23,188.02) --
	(297.04,188.31) --
	(297.20,188.36) --
	(297.32,188.41) --
	(297.39,188.43) --
	(297.92,188.62) --
	(298.07,188.67) --
	(298.46,188.81) --
	(298.47,188.82) --
	(298.54,188.84) --
	(298.62,188.87) --
	(299.15,189.06) --
	(299.26,189.10) --
	(299.54,189.20) --
	(299.58,189.21) --
	(300.71,189.62) --
	(301.41,189.86) --
	(302.13,190.12) --
	(302.61,190.29) --
	(303.45,190.59) --
	(303.55,190.63) --
	(303.59,190.64) --
	(303.66,190.67) --
	(303.89,190.75) --
	(303.92,190.76) --
	(304.13,190.84) --
	(304.31,190.90) --
	(304.32,190.90) --
	(304.63,191.01) --
	(305.44,191.30) --
	(306.03,191.51) --
	(306.65,191.74) --
	(307.10,191.90) --
	(307.15,191.91) --
	(307.34,191.98) --
	(307.36,191.99) --
	(307.48,192.03) --
	(307.57,192.06) --
	(307.71,192.11) --
	(307.78,192.14) --
	(307.86,192.17) --
	(308.14,192.27) --
	(308.17,192.28) --
	(308.40,192.36) --
	(309.29,192.68) --
	(309.65,192.80) --
	(309.66,192.81) --
	(309.77,192.85) --
	(309.85,192.88) --
	(310.26,193.02) --
	(310.42,193.08) --
	(310.88,193.25) --
	(310.97,193.28) --
	(311.25,193.38) --
	(312.09,193.68) --
	(312.22,193.72) --
	(312.41,193.79) --
	(313.42,194.15) --
	(313.62,194.22) --
	(313.78,194.28) --
	(314.05,194.38) --
	(314.50,194.54) --
	(314.58,194.57) --
	(314.67,194.60) --
	(315.27,194.81) --
	(315.36,194.84) --
	(315.70,194.96) --
	(315.87,195.03) --
	(315.98,195.06) --
	(316.35,195.20) --
	(316.56,195.27) --
	(316.72,195.33) --
	(317.21,195.50) --
	(317.49,195.60) --
	(317.56,195.63) --
	(318.84,196.09) --
	(318.87,196.09) --
	(319.28,196.24) --
	(319.56,196.34) --
	(319.75,196.41) --
	(319.97,196.49) --
	(320.23,196.58) --
	(320.29,196.60) --
	(320.33,196.62) --
	(320.94,196.84) --
	(321.01,196.86) --
	(321.22,196.93) --
	(321.25,196.94) --
	(321.36,196.98) --
	(321.75,197.12) --
	(323.15,197.62) --
	(323.66,197.81) --
	(325.17,198.34) --
	(325.77,198.56) --
	(325.89,198.60) --
	(326.94,198.97) --
	(327.01,199.00) --
	(327.29,199.10) --
	(328.29,199.46) --
	(328.36,199.48) --
	(328.82,199.65) --
	(329.10,199.75) --
	(329.71,199.96) --
	(330.08,200.10) --
	(330.10,200.10) --
	(330.14,200.12) --
	(330.37,200.20) --
	(330.66,200.30) --
	(331.19,200.49) --
	(331.66,200.66) --
	(331.92,200.75) --
	(332.49,200.96) --
	(333.01,201.14) --
	(334.21,201.57) --
	(334.41,201.64) --
	(334.89,201.81) --
	(337.42,202.72) --
	(337.45,202.72) --
	(337.58,202.77) --
	(337.70,202.81) --
	(337.71,202.82) --
	(337.80,202.85) --
	(339.18,203.34) --
	(339.28,203.38) --
	(341.35,204.12) --
	(341.75,204.26) --
	(342.17,204.41) --
	(342.48,204.52) --
	(342.53,204.54) --
	(342.69,204.60) --
	(342.85,204.65) --
	(343.03,204.72) --
	(343.17,204.77) --
	(343.37,204.84) --
	(343.45,204.87) --
	(343.75,204.97) --
	(343.88,205.02) --
	(344.23,205.14) --
	(345.06,205.44) --
	(346.03,205.79) --
	(346.49,205.95) --
	(347.23,206.21) --
	(347.38,206.27) --
	(348.49,206.66) --
	(348.58,206.70) --
	(349.13,206.89) --
	(349.81,207.14) --
	(350.10,207.24) --
	(350.24,207.29) --
	(350.58,207.41) --
	(351.85,207.86) --
	(352.05,207.94) --
	(352.77,208.19) --
	(353.17,208.34) --
	(353.42,208.43) --
	(353.71,208.53) --
	(354.71,208.88) --
	(355.56,209.19) --
	(357.56,209.90) --
	(357.72,209.96) --
	(357.97,210.05) --
	(358.00,210.06) --
	(358.30,210.17) --
	(358.37,210.19) --
	(358.57,210.26) --
	(358.60,210.27) --
	(358.96,210.40) --
	(358.97,210.41) --
	(359.44,210.57) --
	(359.71,210.67) --
	(359.84,210.72) --
	(359.89,210.73) --
	(359.95,210.75) --
	(360.12,210.82) --
	(360.50,210.95) --
	(360.53,210.96) --
	(361.15,211.18) --
	(361.24,211.21) --
	(361.57,211.33) --
	(362.32,211.60) --
	(362.86,211.79) --
	(362.88,211.80) --
	(362.99,211.84) --
	(363.37,211.98) --
	(365.15,212.61) --
	(365.51,212.74) --
	(365.58,212.76) --
	(365.58,212.76) --
	(366.24,213.00) --
	(366.27,213.01) --
	(366.79,213.19) --
	(367.09,213.30) --
	(367.28,213.37) --
	(367.50,213.45) --
	(367.74,213.54) --
	(368.30,213.73) --
	(368.30,213.73) --
	(368.38,213.76) --
	(368.49,213.80) --
	(368.90,213.95) --
	(369.46,214.15) --
	(370.56,214.54) --
	(370.74,214.60) --
	(370.85,214.64) --
	(372.47,215.22) --
	(373.39,215.55) --
	(373.79,215.69) --
	(374.21,215.84) --
	(375.70,216.37) --
	(378.93,217.53) --
	(380.42,218.06) --
	(385.52,219.88) --
	(386.08,220.08) --
	(386.54,220.24) --
	(386.90,220.37) --
	(387.31,220.52) --
	(387.53,220.60) --
	(388.24,220.85) --
	(388.35,220.89) --
	(388.69,221.01) --
	(388.71,221.02) --
	(388.75,221.03) --
	(388.90,221.08) --
	(389.17,221.18) --
	(389.73,221.38) --
	(390.59,221.69) --
	(391.22,221.91) --
	(391.77,222.11) --
	(391.82,222.13) --
	(392.15,222.25) --
	(392.41,222.34) --
	(396.33,223.74) --
	(397.13,224.02) --
	(397.28,224.07) --
	(404.32,226.59) --
	(404.80,226.76);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 69.26,167.58) --
	( 76.56,168.49) --
	( 80.61,169.02) --
	( 80.74,169.02) --
	( 93.47,170.57) --
	(101.51,171.61) --
	(106.88,172.38) --
	(109.99,172.78) --
	(116.04,173.71) --
	(124.61,175.00) --
	(153.74,179.40) --
	(166.44,181.37) --
	(169.66,181.90) --
	(182.48,184.32) --
	(189.23,185.73) --
	(198.25,187.43) --
	(202.82,188.36) --
	(205.60,189.03) --
	(205.69,189.04) --
	(206.30,189.17) --
	(207.92,189.53) --
	(209.67,189.86) --
	(214.56,191.01) --
	(214.75,191.06) --
	(219.02,191.93) --
	(220.78,192.25) --
	(224.96,193.40) --
	(226.35,193.80) --
	(226.45,193.83) --
	(226.69,193.88) --
	(227.16,193.97) --
	(229.07,194.38) --
	(229.22,194.47) --
	(229.80,194.62) --
	(230.46,194.79) --
	(231.37,195.05) --
	(232.05,195.24) --
	(232.29,195.30) --
	(232.34,195.31) --
	(232.73,195.42) --
	(237.64,196.73) --
	(239.57,197.26) --
	(239.75,197.33) --
	(240.75,197.60) --
	(241.66,197.84) --
	(242.52,198.05) --
	(243.48,198.31) --
	(244.41,198.51) --
	(245.54,198.84) --
	(245.55,198.85) --
	(247.34,199.26) --
	(247.36,199.27) --
	(248.05,199.47) --
	(248.78,199.61) --
	(249.23,199.75) --
	(249.60,199.85) --
	(252.59,200.68) --
	(255.95,201.50) --
	(257.34,201.84) --
	(257.79,201.94) --
	(257.94,202.01) --
	(258.24,202.10) --
	(259.46,202.47) --
	(259.93,202.59) --
	(261.11,202.97) --
	(261.30,203.05) --
	(261.73,203.20) --
	(263.06,203.72) --
	(264.22,204.06) --
	(266.18,204.56) --
	(267.22,204.89) --
	(267.29,204.91) --
	(268.47,205.25) --
	(269.59,205.73) --
	(270.46,206.09) --
	(271.09,206.28) --
	(271.37,206.36) --
	(271.86,206.51) --
	(271.89,206.52) --
	(272.02,206.57) --
	(272.38,206.68) --
	(272.44,206.71) --
	(273.54,207.03) --
	(273.54,207.03) --
	(274.54,207.35) --
	(274.62,207.37) --
	(274.74,207.40) --
	(274.93,207.48) --
	(275.09,207.54) --
	(275.27,207.59) --
	(275.51,207.69) --
	(275.85,207.76) --
	(276.03,207.79) --
	(276.19,207.84) --
	(276.46,207.94) --
	(276.61,207.99) --
	(278.14,208.40) --
	(278.91,208.64) --
	(278.99,208.66) --
	(279.48,208.87) --
	(279.57,208.92) --
	(279.84,208.96) --
	(280.03,208.99) --
	(280.82,209.17) --
	(281.34,209.31) --
	(281.36,209.32) --
	(281.96,209.48) --
	(282.46,209.66) --
	(282.56,209.69) --
	(283.13,209.95) --
	(283.17,209.97) --
	(283.65,210.14) --
	(283.74,210.15) --
	(285.14,210.63) --
	(285.23,210.67) --
	(285.29,210.69) --
	(285.32,210.71) --
	(285.65,210.81) --
	(286.10,210.98) --
	(286.16,211.01) --
	(286.46,211.13) --
	(286.60,211.18) --
	(286.65,211.20) --
	(286.71,211.21) --
	(286.73,211.21) --
	(286.83,211.23) --
	(286.86,211.24) --
	(287.36,211.42) --
	(287.70,211.53) --
	(287.78,211.57) --
	(289.22,212.06) --
	(289.39,212.10) --
	(289.45,212.13) --
	(289.83,212.25) --
	(289.89,212.27) --
	(290.32,212.44) --
	(290.49,212.49) --
	(290.52,212.50) --
	(290.79,212.58) --
	(290.84,212.60) --
	(290.85,212.60) --
	(290.85,212.60) --
	(290.88,212.61) --
	(290.92,212.62) --
	(291.46,212.83) --
	(291.68,212.92) --
	(292.03,213.07) --
	(292.66,213.31) --
	(292.90,213.41) --
	(293.68,213.72) --
	(294.05,213.86) --
	(294.33,213.95) --
	(294.72,214.08) --
	(294.89,214.17) --
	(294.98,214.23) --
	(295.14,214.31) --
	(295.44,214.42) --
	(295.54,214.45) --
	(295.80,214.55) --
	(296.23,214.74) --
	(297.04,215.07) --
	(297.20,215.13) --
	(297.32,215.18) --
	(297.39,215.20) --
	(297.92,215.38) --
	(298.07,215.42) --
	(298.46,215.59) --
	(298.47,215.59) --
	(298.54,215.62) --
	(298.62,215.65) --
	(299.15,215.80) --
	(299.26,215.83) --
	(299.54,215.92) --
	(299.58,215.93) --
	(300.71,216.45) --
	(301.41,216.73) --
	(302.13,217.04) --
	(302.61,217.20) --
	(303.45,217.53) --
	(303.55,217.56) --
	(303.59,217.58) --
	(303.66,217.61) --
	(303.89,217.71) --
	(303.92,217.72) --
	(304.13,217.80) --
	(304.31,217.89) --
	(304.32,217.89) --
	(304.63,218.07) --
	(305.44,218.39) --
	(306.03,218.64) --
	(306.65,218.85) --
	(307.10,219.05) --
	(307.15,219.08) --
	(307.34,219.17) --
	(307.36,219.18) --
	(307.48,219.22) --
	(307.57,219.26) --
	(307.71,219.29) --
	(307.78,219.32) --
	(307.86,219.33) --
	(308.14,219.41) --
	(308.17,219.43) --
	(308.40,219.53) --
	(309.29,219.89) --
	(309.65,220.03) --
	(309.66,220.04) --
	(309.77,220.06) --
	(309.85,220.08) --
	(310.26,220.25) --
	(310.42,220.31) --
	(310.88,220.51) --
	(310.97,220.54) --
	(311.25,220.65) --
	(312.09,220.89) --
	(312.22,220.94) --
	(312.41,221.03) --
	(313.42,221.40) --
	(313.62,221.47) --
	(313.78,221.52) --
	(314.05,221.67) --
	(314.50,221.90) --
	(314.58,221.94) --
	(314.67,221.97) --
	(315.27,222.21) --
	(315.36,222.23) --
	(315.70,222.38) --
	(315.87,222.44) --
	(315.98,222.48) --
	(316.35,222.60) --
	(316.56,222.70) --
	(316.72,222.78) --
	(317.21,222.97) --
	(317.49,223.13) --
	(317.56,223.16) --
	(318.84,223.77) --
	(318.87,223.77) --
	(319.28,223.88) --
	(319.56,224.01) --
	(319.75,224.09) --
	(319.97,224.18) --
	(320.23,224.27) --
	(320.29,224.29) --
	(320.33,224.31) --
	(320.94,224.63) --
	(321.01,224.66) --
	(321.22,224.74) --
	(321.25,224.75) --
	(321.36,224.78) --
	(321.75,224.91) --
	(323.15,225.56) --
	(323.66,225.78) --
	(325.17,226.53) --
	(325.77,226.82) --
	(325.89,226.87) --
	(326.94,227.29) --
	(327.01,227.33) --
	(327.29,227.43) --
	(328.29,227.87) --
	(328.36,227.90) --
	(328.82,228.09) --
	(329.10,228.24) --
	(329.71,228.50) --
	(330.08,228.68) --
	(330.10,228.68) --
	(330.14,228.70) --
	(330.37,228.80) --
	(330.66,228.94) --
	(331.19,229.21) --
	(331.66,229.41) --
	(331.92,229.55) --
	(332.49,229.78) --
	(333.01,230.05) --
	(334.21,230.64) --
	(334.41,230.77) --
	(334.89,230.98) --
	(337.42,232.04) --
	(337.45,232.05) --
	(337.58,232.13) --
	(337.70,232.18) --
	(337.71,232.18) --
	(337.80,232.22) --
	(339.18,232.90) --
	(339.28,232.95) --
	(341.35,233.86) --
	(341.75,234.03) --
	(342.17,234.21) --
	(342.48,234.37) --
	(342.53,234.40) --
	(342.69,234.46) --
	(342.85,234.53) --
	(343.03,234.60) --
	(343.17,234.65) --
	(343.37,234.74) --
	(343.45,234.77) --
	(343.75,234.90) --
	(343.88,234.96) --
	(344.23,235.11) --
	(345.06,235.47) --
	(346.03,235.91) --
	(346.49,236.10) --
	(347.23,236.53) --
	(347.38,236.62) --
	(348.49,237.12) --
	(348.58,237.18) --
	(349.13,237.46) --
	(349.81,237.76) --
	(350.10,237.92) --
	(350.24,237.98) --
	(350.58,238.16) --
	(351.85,238.80) --
	(352.05,238.90) --
	(352.77,239.24) --
	(353.17,239.44) --
	(353.42,239.54) --
	(353.71,239.66) --
	(354.71,240.11) --
	(355.56,240.57) --
	(357.56,241.53) --
	(357.72,241.60) --
	(357.97,241.72) --
	(358.00,241.74) --
	(358.30,241.88) --
	(358.37,241.91) --
	(358.57,242.01) --
	(358.60,242.02) --
	(358.96,242.21) --
	(358.97,242.22) --
	(359.44,242.45) --
	(359.71,242.56) --
	(359.84,242.65) --
	(359.89,242.68) --
	(359.95,242.70) --
	(360.12,242.82) --
	(360.50,242.98) --
	(360.53,242.99) --
	(361.15,243.34) --
	(361.24,243.39) --
	(361.57,243.56) --
	(362.32,243.88) --
	(362.86,244.15) --
	(362.88,244.16) --
	(362.99,244.21) --
	(363.37,244.41) --
	(365.15,245.33) --
	(365.51,245.52) --
	(365.58,245.55) --
	(365.58,245.55) --
	(366.24,245.88) --
	(366.27,245.90) --
	(366.79,246.19) --
	(367.09,246.30) --
	(367.28,246.39) --
	(367.50,246.52) --
	(367.74,246.66) --
	(368.30,246.92) --
	(368.30,246.92) --
	(368.38,246.95) --
	(368.49,247.00) --
	(368.90,247.19) --
	(369.46,247.45) --
	(370.56,248.00) --
	(370.74,248.06) --
	(370.85,248.13) --
	(372.47,248.84) --
	(373.39,249.28) --
	(373.79,249.49) --
	(374.21,249.72) --
	(375.70,250.52) --
	(378.93,252.19) --
	(380.42,253.06) --
	(385.52,255.81) --
	(386.08,256.12) --
	(386.54,256.36) --
	(386.90,256.56) --
	(387.31,256.78) --
	(387.53,256.89) --
	(388.24,257.30) --
	(388.35,257.37) --
	(388.69,257.57) --
	(388.71,257.58) --
	(388.75,257.60) --
	(388.90,257.69) --
	(389.17,257.84) --
	(389.73,258.09) --
	(390.59,258.54) --
	(391.22,258.88) --
	(391.77,259.16) --
	(391.82,259.17) --
	(392.15,259.38) --
	(392.41,259.49) --
	(396.33,261.54) --
	(397.13,261.95) --
	(397.28,262.02) --
	(404.32,265.72) --
	(404.80,265.99) --
	(404.80,187.19) --
	(404.32,187.06) --
	(397.28,186.09) --
	(397.13,186.06) --
	(396.33,185.88) --
	(392.41,185.16) --
	(392.15,185.08) --
	(391.82,184.98) --
	(391.77,184.96) --
	(391.22,184.83) --
	(390.59,184.77) --
	(389.73,184.61) --
	(389.17,184.48) --
	(388.90,184.43) --
	(388.75,184.40) --
	(388.71,184.40) --
	(388.69,184.39) --
	(388.35,184.33) --
	(388.24,184.32) --
	(387.53,184.20) --
	(387.31,184.16) --
	(386.90,184.09) --
	(386.54,183.99) --
	(386.08,183.85) --
	(385.52,183.74) --
	(380.42,182.74) --
	(378.93,182.37) --
	(375.70,181.71) --
	(374.21,181.28) --
	(373.79,181.13) --
	(373.39,181.02) --
	(372.47,180.83) --
	(370.85,180.50) --
	(370.74,180.49) --
	(370.56,180.44) --
	(369.46,180.14) --
	(368.90,179.97) --
	(368.49,179.90) --
	(368.38,179.87) --
	(368.30,179.85) --
	(368.30,179.85) --
	(367.74,179.77) --
	(367.50,179.74) --
	(367.28,179.72) --
	(367.09,179.69) --
	(366.79,179.59) --
	(366.27,179.50) --
	(366.24,179.50) --
	(365.58,179.36) --
	(365.58,179.36) --
	(365.51,179.34) --
	(365.15,179.23) --
	(363.37,178.83) --
	(362.99,178.73) --
	(362.88,178.70) --
	(362.86,178.69) --
	(362.32,178.55) --
	(361.57,178.28) --
	(361.24,178.19) --
	(361.15,178.17) --
	(360.53,178.07) --
	(360.50,178.06) --
	(360.12,177.98) --
	(359.95,177.95) --
	(359.89,177.93) --
	(359.84,177.92) --
	(359.71,177.88) --
	(359.44,177.80) --
	(358.97,177.71) --
	(358.96,177.71) --
	(358.60,177.64) --
	(358.57,177.63) --
	(358.37,177.60) --
	(358.30,177.60) --
	(358.00,177.51) --
	(357.97,177.49) --
	(357.72,177.41) --
	(357.56,177.37) --
	(355.56,176.98) --
	(354.71,176.78) --
	(353.71,176.53) --
	(353.42,176.47) --
	(353.17,176.42) --
	(352.77,176.32) --
	(352.05,176.22) --
	(351.85,176.14) --
	(350.58,175.77) --
	(350.24,175.68) --
	(350.10,175.64) --
	(349.81,175.57) --
	(349.13,175.50) --
	(348.58,175.34) --
	(348.49,175.31) --
	(347.38,175.04) --
	(347.23,175.00) --
	(346.49,174.84) --
	(346.03,174.73) --
	(345.06,174.54) --
	(344.23,174.31) --
	(343.88,174.22) --
	(343.75,174.18) --
	(343.45,174.11) --
	(343.37,174.09) --
	(343.17,174.02) --
	(343.03,173.97) --
	(342.85,173.91) --
	(342.69,173.86) --
	(342.53,173.81) --
	(342.48,173.80) --
	(342.17,173.68) --
	(341.75,173.58) --
	(341.35,173.51) --
	(339.28,173.04) --
	(339.18,173.03) --
	(337.80,172.73) --
	(337.71,172.72) --
	(337.70,172.72) --
	(337.58,172.70) --
	(337.45,172.64) --
	(337.42,172.63) --
	(334.89,171.98) --
	(334.41,171.89) --
	(334.21,171.81) --
	(333.01,171.45) --
	(332.49,171.27) --
	(331.92,171.05) --
	(331.66,170.98) --
	(331.19,170.85) --
	(330.66,170.74) --
	(330.37,170.62) --
	(330.14,170.56) --
	(330.10,170.55) --
	(330.08,170.54) --
	(329.71,170.44) --
	(329.10,170.30) --
	(328.82,170.26) --
	(328.36,170.16) --
	(328.29,170.15) --
	(327.29,169.86) --
	(327.01,169.78) --
	(326.94,169.75) --
	(325.89,169.42) --
	(325.77,169.39) --
	(325.17,169.23) --
	(323.66,168.80) --
	(323.15,168.62) --
	(321.75,168.21) --
	(321.36,168.09) --
	(321.25,168.05) --
	(321.22,168.04) --
	(321.01,167.98) --
	(320.94,167.93) --
	(320.33,167.77) --
	(320.29,167.76) --
	(320.23,167.74) --
	(319.97,167.67) --
	(319.75,167.57) --
	(319.56,167.52) --
	(319.28,167.46) --
	(318.87,167.33) --
	(318.84,167.33) --
	(317.56,166.94) --
	(317.49,166.92) --
	(317.21,166.84) --
	(316.72,166.71) --
	(316.56,166.68) --
	(316.35,166.61) --
	(315.98,166.48) --
	(315.87,166.43) --
	(315.70,166.38) --
	(315.36,166.28) --
	(315.27,166.25) --
	(314.67,166.07) --
	(314.58,166.04) --
	(314.50,166.02) --
	(314.05,165.88) --
	(313.78,165.83) --
	(313.62,165.78) --
	(313.42,165.72) --
	(312.41,165.41) --
	(312.22,165.34) --
	(312.09,165.30) --
	(311.25,165.06) --
	(310.97,165.00) --
	(310.88,164.98) --
	(310.42,164.86) --
	(310.26,164.82) --
	(309.85,164.69) --
	(309.77,164.66) --
	(309.66,164.62) --
	(309.65,164.61) --
	(309.29,164.48) --
	(308.40,164.29) --
	(308.17,164.21) --
	(308.14,164.19) --
	(307.86,164.11) --
	(307.78,164.09) --
	(307.71,164.08) --
	(307.57,164.04) --
	(307.48,164.01) --
	(307.36,163.96) --
	(307.34,163.95) --
	(307.15,163.88) --
	(307.10,163.87) --
	(306.65,163.74) --
	(306.03,163.55) --
	(305.44,163.36) --
	(304.63,163.10) --
	(304.32,162.99) --
	(304.31,162.99) --
	(304.13,162.94) --
	(303.92,162.88) --
	(303.89,162.87) --
	(303.66,162.79) --
	(303.59,162.77) --
	(303.55,162.76) --
	(303.45,162.74) --
	(302.61,162.56) --
	(302.13,162.46) --
	(301.41,162.29) --
	(300.71,162.10) --
	(299.58,161.75) --
	(299.54,161.74) --
	(299.26,161.63) --
	(299.15,161.58) --
	(298.62,161.31) --
	(298.54,161.31) --
	(298.47,161.28) --
	(298.46,161.28) --
	(298.07,161.14) --
	(297.92,161.07) --
	(297.39,160.94) --
	(297.32,160.90) --
	(297.20,160.85) --
	(297.04,160.83) --
	(296.23,160.63) --
	(295.80,160.53) --
	(295.54,160.46) --
	(295.44,160.41) --
	(295.14,160.30) --
	(294.98,160.22) --
	(294.89,160.18) --
	(294.72,160.12) --
	(294.33,159.97) --
	(294.05,159.87) --
	(293.68,159.78) --
	(292.90,159.50) --
	(292.66,159.42) --
	(292.03,159.17) --
	(291.68,159.07) --
	(291.46,158.95) --
	(290.92,158.76) --
	(290.88,158.74) --
	(290.85,158.72) --
	(290.85,158.72) --
	(290.84,158.72) --
	(290.79,158.69) --
	(290.52,158.56) --
	(290.49,158.55) --
	(290.32,158.46) --
	(289.89,158.32) --
	(289.83,158.30) --
	(289.45,158.18) --
	(289.39,158.16) --
	(289.22,158.11) --
	(287.78,157.56) --
	(287.70,157.54) --
	(287.36,157.49) --
	(286.86,157.37) --
	(286.83,157.36) --
	(286.73,157.33) --
	(286.71,157.31) --
	(286.65,157.28) --
	(286.60,157.25) --
	(286.46,157.17) --
	(286.16,157.04) --
	(286.10,157.03) --
	(285.65,156.93) --
	(285.32,156.82) --
	(285.29,156.80) --
	(285.23,156.77) --
	(285.14,156.75) --
	(283.74,156.26) --
	(283.65,156.23) --
	(283.17,156.09) --
	(283.13,156.08) --
	(282.56,155.90) --
	(282.46,155.87) --
	(281.96,155.72) --
	(281.36,155.53) --
	(281.34,155.52) --
	(280.82,155.34) --
	(280.03,155.03) --
	(279.84,154.97) --
	(279.57,154.88) --
	(279.48,154.85) --
	(278.99,154.69) --
	(278.91,154.66) --
	(278.14,154.34) --
	(276.61,153.82) --
	(276.46,153.75) --
	(276.19,153.66) --
	(276.03,153.58) --
	(275.85,153.49) --
	(275.51,153.32) --
	(275.27,153.24) --
	(275.09,153.18) --
	(274.93,153.13) --
	(274.74,153.07) --
	(274.62,153.01) --
	(274.54,152.96) --
	(273.54,152.47) --
	(273.54,152.47) --
	(272.44,152.06) --
	(272.38,152.04) --
	(272.02,151.89) --
	(271.89,151.85) --
	(271.86,151.85) --
	(271.37,151.64) --
	(271.09,151.55) --
	(270.46,151.24) --
	(269.59,150.90) --
	(268.47,150.47) --
	(267.29,149.94) --
	(267.22,149.89) --
	(266.18,149.49) --
	(264.22,148.70) --
	(263.06,148.21) --
	(261.73,147.75) --
	(261.30,147.63) --
	(261.11,147.56) --
	(259.93,147.06) --
	(259.46,146.86) --
	(258.24,146.37) --
	(257.94,146.22) --
	(257.79,146.15) --
	(257.34,145.93) --
	(255.95,145.29) --
	(252.59,143.92) --
	(249.60,142.66) --
	(249.23,142.52) --
	(248.78,142.28) --
	(248.05,141.96) --
	(247.36,141.73) --
	(247.34,141.72) --
	(245.55,140.68) --
	(245.54,140.67) --
	(244.41,140.23) --
	(243.48,139.81) --
	(242.52,139.40) --
	(241.66,139.00) --
	(240.75,138.64) --
	(239.75,138.24) --
	(239.57,138.16) --
	(237.64,137.39) --
	(232.73,135.16) --
	(232.34,134.96) --
	(232.29,134.94) --
	(232.05,134.85) --
	(231.37,134.57) --
	(230.46,134.14) --
	(229.80,133.84) --
	(229.22,133.54) --
	(229.07,133.47) --
	(227.16,132.57) --
	(226.69,132.33) --
	(226.45,132.19) --
	(226.35,132.15) --
	(224.96,131.49) --
	(220.78,129.59) --
	(219.02,128.61) --
	(214.75,126.43) --
	(214.56,126.34) --
	(209.67,123.91) --
	(207.92,122.94) --
	(206.30,122.24) --
	(205.69,121.90) --
	(205.60,121.85) --
	(202.82,120.36) --
	(198.25,118.05) --
	(189.23,113.19) --
	(182.48,109.70) --
	(169.66,102.54) --
	(166.44,100.78) --
	(153.74, 94.08) --
	(124.61, 77.55) --
	(116.04, 72.56) --
	(109.99, 69.23) --
	(106.88, 67.49) --
	(101.51, 64.29) --
	( 93.47, 59.68) --
	( 80.74, 52.08) --
	( 80.61, 52.00) --
	( 76.56, 49.55) --
	( 69.26, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 69.26,158.06) --
	( 76.56,159.37) --
	( 80.61,160.10) --
	( 80.74,160.12) --
	( 93.47,161.99) --
	(101.51,163.36) --
	(106.88,164.36) --
	(109.99,164.97) --
	(116.04,165.98) --
	(124.61,167.56) --
	(153.74,173.30) --
	(166.44,175.73) --
	(169.66,176.32) --
	(182.48,179.03) --
	(189.23,180.36) --
	(198.25,182.33) --
	(202.82,183.39) --
	(205.60,184.05) --
	(205.69,184.08) --
	(206.30,184.21) --
	(207.92,184.53) --
	(209.67,184.97) --
	(214.56,186.05) --
	(214.75,186.10) --
	(219.02,187.06) --
	(220.78,187.55) --
	(224.96,188.64) --
	(226.35,188.99) --
	(226.45,189.01) --
	(226.69,189.04) --
	(227.16,189.17) --
	(229.07,189.76) --
	(229.22,189.83) --
	(229.80,189.97) --
	(230.46,190.13) --
	(231.37,190.40) --
	(232.05,190.62) --
	(232.29,190.67) --
	(232.34,190.69) --
	(232.73,190.81) --
	(237.64,192.19) --
	(239.57,192.68) --
	(239.75,192.72) --
	(240.75,192.97) --
	(241.66,193.18) --
	(242.52,193.45) --
	(243.48,193.69) --
	(244.41,194.00) --
	(245.54,194.36) --
	(245.55,194.36) --
	(247.34,194.92) --
	(247.36,194.92) --
	(248.05,195.09) --
	(248.78,195.28) --
	(249.23,195.39) --
	(249.60,195.46) --
	(252.59,196.28) --
	(255.95,197.25) --
	(257.34,197.64) --
	(257.79,197.74) --
	(257.94,197.77) --
	(258.24,197.86) --
	(259.46,198.20) --
	(259.93,198.33) --
	(261.11,198.71) --
	(261.30,198.77) --
	(261.73,198.92) --
	(263.06,199.31) --
	(264.22,199.66) --
	(266.18,200.26) --
	(267.22,200.59) --
	(267.29,200.62) --
	(268.47,201.00) --
	(269.59,201.32) --
	(270.46,201.65) --
	(271.09,201.82) --
	(271.37,201.90) --
	(271.86,202.05) --
	(271.89,202.06) --
	(272.02,202.07) --
	(272.38,202.18) --
	(272.44,202.20) --
	(273.54,202.60) --
	(273.54,202.61) --
	(274.54,202.96) --
	(274.62,202.98) --
	(274.74,203.01) --
	(274.93,203.06) --
	(275.09,203.12) --
	(275.27,203.18) --
	(275.51,203.24) --
	(275.85,203.34) --
	(276.03,203.40) --
	(276.19,203.45) --
	(276.46,203.57) --
	(276.61,203.62) --
	(278.14,204.15) --
	(278.91,204.42) --
	(278.99,204.45) --
	(279.48,204.60) --
	(279.57,204.61) --
	(279.84,204.69) --
	(280.03,204.74) --
	(280.82,204.98) --
	(281.34,205.16) --
	(281.36,205.17) --
	(281.96,205.44) --
	(282.46,205.62) --
	(282.56,205.65) --
	(283.13,205.83) --
	(283.17,205.85) --
	(283.65,206.01) --
	(283.74,206.04) --
	(285.14,206.56) --
	(285.23,206.60) --
	(285.29,206.62) --
	(285.32,206.63) --
	(285.65,206.72) --
	(286.10,206.84) --
	(286.16,206.85) --
	(286.46,206.97) --
	(286.60,207.03) --
	(286.65,207.05) --
	(286.71,207.07) --
	(286.73,207.08) --
	(286.83,207.11) --
	(286.86,207.13) --
	(287.36,207.33) --
	(287.70,207.49) --
	(287.78,207.52) --
	(289.22,208.10) --
	(289.39,208.18) --
	(289.45,208.21) --
	(289.83,208.33) --
	(289.89,208.36) --
	(290.32,208.55) --
	(290.49,208.59) --
	(290.52,208.60) --
	(290.79,208.71) --
	(290.84,208.71) --
	(290.85,208.71) --
	(290.85,208.72) --
	(290.88,208.73) --
	(290.92,208.74) --
	(291.46,208.93) --
	(291.68,209.00) --
	(292.03,209.14) --
	(292.66,209.32) --
	(292.90,209.42) --
	(293.68,209.75) --
	(294.05,209.81) --
	(294.33,209.92) --
	(294.72,210.07) --
	(294.89,210.14) --
	(294.98,210.18) --
	(295.14,210.26) --
	(295.44,210.36) --
	(295.54,210.39) --
	(295.80,210.51) --
	(296.23,210.69) --
	(297.04,210.99) --
	(297.20,211.06) --
	(297.32,211.11) --
	(297.39,211.13) --
	(297.92,211.35) --
	(298.07,211.41) --
	(298.46,211.58) --
	(298.47,211.58) --
	(298.54,211.61) --
	(298.62,211.64) --
	(299.15,211.84) --
	(299.26,211.88) --
	(299.54,211.99) --
	(299.58,212.00) --
	(300.71,212.49) --
	(301.41,212.70) --
	(302.13,213.03) --
	(302.61,213.21) --
	(303.45,213.49) --
	(303.55,213.52) --
	(303.59,213.53) --
	(303.66,213.55) --
	(303.89,213.61) --
	(303.92,213.61) --
	(304.13,213.70) --
	(304.31,213.78) --
	(304.32,213.78) --
	(304.63,213.91) --
	(305.44,214.22) --
	(306.03,214.40) --
	(306.65,214.65) --
	(307.10,214.83) --
	(307.15,214.85) --
	(307.34,214.92) --
	(307.36,214.92) --
	(307.48,214.98) --
	(307.57,215.01) --
	(307.71,215.07) --
	(307.78,215.10) --
	(307.86,215.13) --
	(308.14,215.24) --
	(308.17,215.26) --
	(308.40,215.36) --
	(309.29,215.75) --
	(309.65,215.90) --
	(309.66,215.90) --
	(309.77,215.94) --
	(309.85,215.96) --
	(310.26,216.13) --
	(310.42,216.19) --
	(310.88,216.36) --
	(310.97,216.39) --
	(311.25,216.48) --
	(312.09,216.80) --
	(312.22,216.84) --
	(312.41,216.92) --
	(313.42,217.35) --
	(313.62,217.46) --
	(313.78,217.52) --
	(314.05,217.62) --
	(314.50,217.82) --
	(314.58,217.86) --
	(314.67,217.90) --
	(315.27,218.11) --
	(315.36,218.13) --
	(315.70,218.31) --
	(315.87,218.39) --
	(315.98,218.43) --
	(316.35,218.58) --
	(316.56,218.68) --
	(316.72,218.76) --
	(317.21,218.94) --
	(317.49,219.07) --
	(317.56,219.10) --
	(318.84,219.67) --
	(318.87,219.68) --
	(319.28,219.82) --
	(319.56,219.91) --
	(319.75,219.99) --
	(319.97,220.10) --
	(320.23,220.22) --
	(320.29,220.24) --
	(320.33,220.26) --
	(320.94,220.45) --
	(321.01,220.47) --
	(321.22,220.57) --
	(321.25,220.59) --
	(321.36,220.64) --
	(321.75,220.84) --
	(323.15,221.38) --
	(323.66,221.56) --
	(325.17,222.18) --
	(325.77,222.40) --
	(325.89,222.48) --
	(326.94,222.97) --
	(327.01,223.01) --
	(327.29,223.13) --
	(328.29,223.54) --
	(328.36,223.57) --
	(328.82,223.76) --
	(329.10,223.87) --
	(329.71,224.11) --
	(330.08,224.27) --
	(330.10,224.28) --
	(330.14,224.30) --
	(330.37,224.41) --
	(330.66,224.51) --
	(331.19,224.72) --
	(331.66,224.93) --
	(331.92,225.05) --
	(332.49,225.26) --
	(333.01,225.49) --
	(334.21,226.03) --
	(334.41,226.11) --
	(334.89,226.30) --
	(337.42,227.37) --
	(337.45,227.38) --
	(337.58,227.44) --
	(337.70,227.50) --
	(337.71,227.50) --
	(337.80,227.54) --
	(339.18,228.14) --
	(339.28,228.20) --
	(341.35,229.04) --
	(341.75,229.22) --
	(342.17,229.38) --
	(342.48,229.55) --
	(342.53,229.57) --
	(342.69,229.66) --
	(342.85,229.73) --
	(343.03,229.82) --
	(343.17,229.89) --
	(343.37,229.97) --
	(343.45,230.02) --
	(343.75,230.20) --
	(343.88,230.27) --
	(344.23,230.45) --
	(345.06,230.84) --
	(346.03,231.26) --
	(346.49,231.43) --
	(347.23,231.80) --
	(347.38,231.88) --
	(348.49,232.39) --
	(348.58,232.43) --
	(349.13,232.71) --
	(349.81,233.08) --
	(350.10,233.20) --
	(350.24,233.26) --
	(350.58,233.40) --
	(351.85,233.92) --
	(352.05,233.99) --
	(352.77,234.32) --
	(353.17,234.51) --
	(353.42,234.63) --
	(353.71,234.76) --
	(354.71,235.15) --
	(355.56,235.61) --
	(357.56,236.47) --
	(357.72,236.54) --
	(357.97,236.65) --
	(358.00,236.66) --
	(358.30,236.82) --
	(358.37,236.86) --
	(358.57,236.95) --
	(358.60,236.97) --
	(358.96,237.17) --
	(358.97,237.18) --
	(359.44,237.44) --
	(359.71,237.56) --
	(359.84,237.62) --
	(359.89,237.63) --
	(359.95,237.66) --
	(360.12,237.73) --
	(360.50,237.88) --
	(360.53,237.89) --
	(361.15,238.15) --
	(361.24,238.18) --
	(361.57,238.32) --
	(362.32,238.66) --
	(362.86,238.92) --
	(362.88,238.93) --
	(362.99,238.99) --
	(363.37,239.16) --
	(365.15,240.02) --
	(365.51,240.19) --
	(365.58,240.23) --
	(365.58,240.23) --
	(366.24,240.61) --
	(366.27,240.63) --
	(366.79,240.88) --
	(367.09,241.02) --
	(367.28,241.13) --
	(367.50,241.25) --
	(367.74,241.35) --
	(368.30,241.57) --
	(368.30,241.57) --
	(368.38,241.60) --
	(368.49,241.65) --
	(368.90,241.82) --
	(369.46,242.09) --
	(370.56,242.57) --
	(370.74,242.67) --
	(370.85,242.73) --
	(372.47,243.56) --
	(373.39,244.01) --
	(373.79,244.21) --
	(374.21,244.42) --
	(375.70,245.15) --
	(378.93,246.76) --
	(380.42,247.53) --
	(385.52,250.10) --
	(386.08,250.38) --
	(386.54,250.63) --
	(386.90,250.79) --
	(387.31,251.04) --
	(387.53,251.17) --
	(388.24,251.54) --
	(388.35,251.60) --
	(388.69,251.73) --
	(388.71,251.74) --
	(388.75,251.76) --
	(388.90,251.84) --
	(389.17,252.00) --
	(389.73,252.28) --
	(390.59,252.69) --
	(391.22,253.03) --
	(391.77,253.31) --
	(391.82,253.33) --
	(392.15,253.48) --
	(392.41,253.61) --
	(396.33,255.53) --
	(397.13,255.92) --
	(397.28,256.01) --
	(404.32,259.58) --
	(404.80,259.80) --
	(404.80,193.63) --
	(404.32,193.52) --
	(397.28,192.09) --
	(397.13,192.06) --
	(396.33,191.92) --
	(392.41,191.02) --
	(392.15,190.97) --
	(391.82,190.88) --
	(391.77,190.87) --
	(391.22,190.75) --
	(390.59,190.60) --
	(389.73,190.40) --
	(389.17,190.27) --
	(388.90,190.21) --
	(388.75,190.18) --
	(388.71,190.18) --
	(388.69,190.17) --
	(388.35,190.08) --
	(388.24,190.06) --
	(387.53,189.90) --
	(387.31,189.87) --
	(386.90,189.79) --
	(386.54,189.71) --
	(386.08,189.64) --
	(385.52,189.44) --
	(380.42,188.27) --
	(378.93,187.91) --
	(375.70,187.16) --
	(374.21,186.81) --
	(373.79,186.72) --
	(373.39,186.63) --
	(372.47,186.48) --
	(370.85,186.09) --
	(370.74,186.05) --
	(370.56,185.99) --
	(369.46,185.78) --
	(368.90,185.66) --
	(368.49,185.60) --
	(368.38,185.57) --
	(368.30,185.54) --
	(368.30,185.54) --
	(367.74,185.37) --
	(367.50,185.30) --
	(367.28,185.23) --
	(367.09,185.20) --
	(366.79,185.18) --
	(366.27,185.04) --
	(366.24,185.04) --
	(365.58,184.87) --
	(365.58,184.87) --
	(365.51,184.85) --
	(365.15,184.74) --
	(363.37,184.26) --
	(362.99,184.16) --
	(362.88,184.14) --
	(362.86,184.13) --
	(362.32,183.99) --
	(361.57,183.78) --
	(361.24,183.69) --
	(361.15,183.67) --
	(360.53,183.58) --
	(360.50,183.57) --
	(360.12,183.47) --
	(359.95,183.42) --
	(359.89,183.40) --
	(359.84,183.40) --
	(359.71,183.37) --
	(359.44,183.28) --
	(358.97,183.13) --
	(358.96,183.12) --
	(358.60,182.99) --
	(358.57,182.99) --
	(358.37,182.96) --
	(358.30,182.95) --
	(358.00,182.90) --
	(357.97,182.90) --
	(357.72,182.84) --
	(357.56,182.81) --
	(355.56,182.34) --
	(354.71,182.07) --
	(353.71,181.78) --
	(353.42,181.68) --
	(353.17,181.60) --
	(352.77,181.50) --
	(352.05,181.35) --
	(351.85,181.33) --
	(350.58,181.06) --
	(350.24,181.00) --
	(350.10,180.98) --
	(349.81,180.92) --
	(349.13,180.77) --
	(348.58,180.62) --
	(348.49,180.60) --
	(347.38,180.24) --
	(347.23,180.21) --
	(346.49,179.99) --
	(346.03,179.86) --
	(345.06,179.60) --
	(344.23,179.40) --
	(343.88,179.29) --
	(343.75,179.26) --
	(343.45,179.17) --
	(343.37,179.16) --
	(343.17,179.09) --
	(343.03,179.05) --
	(342.85,179.02) --
	(342.69,178.99) --
	(342.53,178.94) --
	(342.48,178.93) --
	(342.17,178.85) --
	(341.75,178.74) --
	(341.35,178.63) --
	(339.28,178.08) --
	(339.18,178.06) --
	(337.80,177.70) --
	(337.71,177.67) --
	(337.70,177.67) --
	(337.58,177.64) --
	(337.45,177.61) --
	(337.42,177.61) --
	(334.89,176.97) --
	(334.41,176.82) --
	(334.21,176.78) --
	(333.01,176.45) --
	(332.49,176.29) --
	(331.92,176.16) --
	(331.66,176.11) --
	(331.19,175.98) --
	(330.66,175.82) --
	(330.37,175.74) --
	(330.14,175.66) --
	(330.10,175.65) --
	(330.08,175.64) --
	(329.71,175.56) --
	(329.10,175.37) --
	(328.82,175.26) --
	(328.36,175.14) --
	(328.29,175.12) --
	(327.29,174.83) --
	(327.01,174.77) --
	(326.94,174.76) --
	(325.89,174.48) --
	(325.77,174.45) --
	(325.17,174.31) --
	(323.66,173.93) --
	(323.15,173.78) --
	(321.75,173.36) --
	(321.36,173.20) --
	(321.25,173.15) --
	(321.22,173.15) --
	(321.01,173.11) --
	(320.94,173.10) --
	(320.33,172.92) --
	(320.29,172.91) --
	(320.23,172.88) --
	(319.97,172.81) --
	(319.75,172.74) --
	(319.56,172.67) --
	(319.28,172.58) --
	(318.87,172.45) --
	(318.84,172.44) --
	(317.56,172.09) --
	(317.49,172.08) --
	(317.21,172.00) --
	(316.72,171.83) --
	(316.56,171.77) --
	(316.35,171.69) --
	(315.98,171.61) --
	(315.87,171.59) --
	(315.70,171.54) --
	(315.36,171.47) --
	(315.27,171.46) --
	(314.67,171.26) --
	(314.58,171.24) --
	(314.50,171.22) --
	(314.05,171.12) --
	(313.78,171.03) --
	(313.62,170.98) --
	(313.42,170.90) --
	(312.41,170.53) --
	(312.22,170.46) --
	(312.09,170.42) --
	(311.25,170.19) --
	(310.97,170.09) --
	(310.88,170.07) --
	(310.42,169.91) --
	(310.26,169.87) --
	(309.85,169.77) --
	(309.77,169.74) --
	(309.66,169.71) --
	(309.65,169.71) --
	(309.29,169.57) --
	(308.40,169.24) --
	(308.17,169.17) --
	(308.14,169.15) --
	(307.86,169.07) --
	(307.78,169.05) --
	(307.71,169.02) --
	(307.57,168.96) --
	(307.48,168.94) --
	(307.36,168.90) --
	(307.34,168.89) --
	(307.15,168.83) --
	(307.10,168.82) --
	(306.65,168.65) --
	(306.03,168.41) --
	(305.44,168.24) --
	(304.63,167.94) --
	(304.32,167.86) --
	(304.31,167.86) --
	(304.13,167.83) --
	(303.92,167.75) --
	(303.89,167.74) --
	(303.66,167.64) --
	(303.59,167.61) --
	(303.55,167.59) --
	(303.45,167.57) --
	(302.61,167.26) --
	(302.13,167.11) --
	(301.41,166.77) --
	(300.71,166.59) --
	(299.58,166.27) --
	(299.54,166.26) --
	(299.26,166.14) --
	(299.15,166.11) --
	(298.62,165.91) --
	(298.54,165.90) --
	(298.47,165.90) --
	(298.46,165.89) --
	(298.07,165.78) --
	(297.92,165.74) --
	(297.39,165.53) --
	(297.32,165.52) --
	(297.20,165.47) --
	(297.04,165.42) --
	(296.23,165.17) --
	(295.80,165.04) --
	(295.54,164.95) --
	(295.44,164.92) --
	(295.14,164.82) --
	(294.98,164.77) --
	(294.89,164.75) --
	(294.72,164.69) --
	(294.33,164.51) --
	(294.05,164.40) --
	(293.68,164.30) --
	(292.90,164.04) --
	(292.66,163.95) --
	(292.03,163.71) --
	(291.68,163.61) --
	(291.46,163.54) --
	(290.92,163.33) --
	(290.88,163.31) --
	(290.85,163.29) --
	(290.85,163.29) --
	(290.84,163.29) --
	(290.79,163.29) --
	(290.52,163.19) --
	(290.49,163.17) --
	(290.32,163.10) --
	(289.89,162.93) --
	(289.83,162.92) --
	(289.45,162.82) --
	(289.39,162.80) --
	(289.22,162.75) --
	(287.78,162.30) --
	(287.70,162.28) --
	(287.36,162.17) --
	(286.86,161.98) --
	(286.83,161.97) --
	(286.73,161.93) --
	(286.71,161.92) --
	(286.65,161.89) --
	(286.60,161.87) --
	(286.46,161.80) --
	(286.16,161.68) --
	(286.10,161.65) --
	(285.65,161.51) --
	(285.32,161.41) --
	(285.29,161.40) --
	(285.23,161.39) --
	(285.14,161.35) --
	(283.74,160.86) --
	(283.65,160.83) --
	(283.17,160.65) --
	(283.13,160.63) --
	(282.56,160.45) --
	(282.46,160.44) --
	(281.96,160.22) --
	(281.36,159.99) --
	(281.34,159.98) --
	(280.82,159.75) --
	(280.03,159.42) --
	(279.84,159.35) --
	(279.57,159.24) --
	(279.48,159.20) --
	(278.99,158.99) --
	(278.91,158.96) --
	(278.14,158.66) --
	(276.61,158.07) --
	(276.46,158.02) --
	(276.19,157.93) --
	(276.03,157.90) --
	(275.85,157.83) --
	(275.51,157.70) --
	(275.27,157.60) --
	(275.09,157.55) --
	(274.93,157.51) --
	(274.74,157.44) --
	(274.62,157.39) --
	(274.54,157.35) --
	(273.54,156.94) --
	(273.54,156.94) --
	(272.44,156.53) --
	(272.38,156.51) --
	(272.02,156.41) --
	(271.89,156.36) --
	(271.86,156.35) --
	(271.37,156.14) --
	(271.09,156.07) --
	(270.46,155.83) --
	(269.59,155.51) --
	(268.47,155.04) --
	(267.29,154.56) --
	(267.22,154.54) --
	(266.18,154.10) --
	(264.22,153.32) --
	(263.06,152.86) --
	(261.73,152.31) --
	(261.30,152.12) --
	(261.11,152.03) --
	(259.93,151.57) --
	(259.46,151.36) --
	(258.24,150.94) --
	(257.94,150.85) --
	(257.79,150.79) --
	(257.34,150.60) --
	(255.95,150.01) --
	(252.59,148.62) --
	(249.60,147.41) --
	(249.23,147.23) --
	(248.78,147.03) --
	(248.05,146.76) --
	(247.36,146.49) --
	(247.34,146.48) --
	(245.55,145.70) --
	(245.54,145.69) --
	(244.41,145.19) --
	(243.48,144.81) --
	(242.52,144.42) --
	(241.66,144.05) --
	(240.75,143.64) --
	(239.75,143.12) --
	(239.57,143.07) --
	(237.64,142.25) --
	(232.73,140.08) --
	(232.34,139.90) --
	(232.29,139.87) --
	(232.05,139.75) --
	(231.37,139.44) --
	(230.46,139.01) --
	(229.80,138.72) --
	(229.22,138.43) --
	(229.07,138.33) --
	(227.16,137.42) --
	(226.69,137.17) --
	(226.45,137.06) --
	(226.35,137.03) --
	(224.96,136.50) --
	(220.78,134.55) --
	(219.02,133.70) --
	(214.75,131.69) --
	(214.56,131.61) --
	(209.67,129.15) --
	(207.92,128.41) --
	(206.30,127.62) --
	(205.69,127.37) --
	(205.60,127.33) --
	(202.82,125.94) --
	(198.25,123.69) --
	(189.23,119.35) --
	(182.48,116.07) --
	(169.66,109.49) --
	(166.44,107.79) --
	(153.74,101.09) --
	(124.61, 85.71) --
	(116.04, 81.15) --
	(109.99, 77.69) --
	(106.88, 75.97) --
	(101.51, 73.00) --
	( 93.47, 68.72) --
	( 80.74, 61.97) --
	( 80.61, 61.89) --
	( 76.56, 59.73) --
	( 69.26, 55.71) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 69.26, 34.03) -- ( 69.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 76.56, 34.03) -- ( 76.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 80.61, 34.03) -- ( 80.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 80.74, 34.03) -- ( 80.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 93.47, 34.03) -- ( 93.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (101.51, 34.03) -- (101.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.88, 34.03) -- (106.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (109.99, 34.03) -- (109.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (116.04, 34.03) -- (116.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (124.61, 34.03) -- (124.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (153.74, 34.03) -- (153.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (166.44, 34.03) -- (166.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (169.66, 34.03) -- (169.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (182.48, 34.03) -- (182.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (189.23, 34.03) -- (189.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (198.25, 34.03) -- (198.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (202.82, 34.03) -- (202.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.60, 34.03) -- (205.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.69, 34.03) -- (205.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (206.30, 34.03) -- (206.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (207.92, 34.03) -- (207.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (209.67, 34.03) -- (209.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.56, 34.03) -- (214.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.75, 34.03) -- (214.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.02, 34.03) -- (219.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.78, 34.03) -- (220.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (224.96, 34.03) -- (224.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.35, 34.03) -- (226.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.45, 34.03) -- (226.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.69, 34.03) -- (226.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (227.16, 34.03) -- (227.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.07, 34.03) -- (229.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.22, 34.03) -- (229.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.80, 34.03) -- (229.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (230.46, 34.03) -- (230.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.37, 34.03) -- (231.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.05, 34.03) -- (232.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.29, 34.03) -- (232.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.34, 34.03) -- (232.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.73, 34.03) -- (232.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (237.64, 34.03) -- (237.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.57, 34.03) -- (239.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.75, 34.03) -- (239.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.75, 34.03) -- (240.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (241.66, 34.03) -- (241.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.52, 34.03) -- (242.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (243.48, 34.03) -- (243.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (244.41, 34.03) -- (244.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (245.54, 34.03) -- (245.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (245.55, 34.03) -- (245.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.34, 34.03) -- (247.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.36, 34.03) -- (247.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.05, 34.03) -- (248.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.78, 34.03) -- (248.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.23, 34.03) -- (249.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.60, 34.03) -- (249.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.59, 34.03) -- (252.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (255.95, 34.03) -- (255.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (257.34, 34.03) -- (257.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (257.79, 34.03) -- (257.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (257.94, 34.03) -- (257.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.24, 34.03) -- (258.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (259.46, 34.03) -- (259.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (259.93, 34.03) -- (259.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.11, 34.03) -- (261.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.30, 34.03) -- (261.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.73, 34.03) -- (261.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (263.06, 34.03) -- (263.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.22, 34.03) -- (264.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.18, 34.03) -- (266.18, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.22, 34.03) -- (267.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.29, 34.03) -- (267.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.47, 34.03) -- (268.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.59, 34.03) -- (269.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (270.46, 34.03) -- (270.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.09, 34.03) -- (271.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.37, 34.03) -- (271.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.86, 34.03) -- (271.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.89, 34.03) -- (271.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.02, 34.03) -- (272.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.38, 34.03) -- (272.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.44, 34.03) -- (272.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.54, 34.03) -- (273.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.54, 34.03) -- (273.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.54, 34.03) -- (274.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.62, 34.03) -- (274.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.74, 34.03) -- (274.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.93, 34.03) -- (274.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.09, 34.03) -- (275.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.27, 34.03) -- (275.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.51, 34.03) -- (275.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.85, 34.03) -- (275.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.03, 34.03) -- (276.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.19, 34.03) -- (276.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.46, 34.03) -- (276.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.61, 34.03) -- (276.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.14, 34.03) -- (278.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.91, 34.03) -- (278.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.99, 34.03) -- (278.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.48, 34.03) -- (279.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.57, 34.03) -- (279.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.84, 34.03) -- (279.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.03, 34.03) -- (280.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.82, 34.03) -- (280.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.34, 34.03) -- (281.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.36, 34.03) -- (281.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.96, 34.03) -- (281.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.46, 34.03) -- (282.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.56, 34.03) -- (282.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.13, 34.03) -- (283.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.17, 34.03) -- (283.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.65, 34.03) -- (283.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.74, 34.03) -- (283.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.14, 34.03) -- (285.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.23, 34.03) -- (285.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.29, 34.03) -- (285.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.32, 34.03) -- (285.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.65, 34.03) -- (285.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.10, 34.03) -- (286.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.16, 34.03) -- (286.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.46, 34.03) -- (286.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.60, 34.03) -- (286.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.65, 34.03) -- (286.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.71, 34.03) -- (286.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.73, 34.03) -- (286.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.83, 34.03) -- (286.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.86, 34.03) -- (286.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.36, 34.03) -- (287.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.70, 34.03) -- (287.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.78, 34.03) -- (287.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.22, 34.03) -- (289.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.39, 34.03) -- (289.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.45, 34.03) -- (289.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.83, 34.03) -- (289.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.89, 34.03) -- (289.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.32, 34.03) -- (290.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.49, 34.03) -- (290.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.52, 34.03) -- (290.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.79, 34.03) -- (290.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.84, 34.03) -- (290.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.85, 34.03) -- (290.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.85, 34.03) -- (290.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.88, 34.03) -- (290.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.92, 34.03) -- (290.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.46, 34.03) -- (291.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.68, 34.03) -- (291.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.03, 34.03) -- (292.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.66, 34.03) -- (292.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.90, 34.03) -- (292.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.68, 34.03) -- (293.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.05, 34.03) -- (294.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.33, 34.03) -- (294.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.72, 34.03) -- (294.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.89, 34.03) -- (294.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.98, 34.03) -- (294.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.14, 34.03) -- (295.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.44, 34.03) -- (295.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.54, 34.03) -- (295.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.80, 34.03) -- (295.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.23, 34.03) -- (296.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.04, 34.03) -- (297.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.20, 34.03) -- (297.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.32, 34.03) -- (297.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.39, 34.03) -- (297.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.92, 34.03) -- (297.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.07, 34.03) -- (298.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.46, 34.03) -- (298.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.47, 34.03) -- (298.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.54, 34.03) -- (298.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.62, 34.03) -- (298.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.15, 34.03) -- (299.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.26, 34.03) -- (299.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.54, 34.03) -- (299.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.58, 34.03) -- (299.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.71, 34.03) -- (300.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.41, 34.03) -- (301.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.13, 34.03) -- (302.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.61, 34.03) -- (302.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.45, 34.03) -- (303.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.55, 34.03) -- (303.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.59, 34.03) -- (303.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.66, 34.03) -- (303.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.89, 34.03) -- (303.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.92, 34.03) -- (303.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.13, 34.03) -- (304.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.31, 34.03) -- (304.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.32, 34.03) -- (304.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.63, 34.03) -- (304.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.44, 34.03) -- (305.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.03, 34.03) -- (306.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.65, 34.03) -- (306.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.10, 34.03) -- (307.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.15, 34.03) -- (307.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.34, 34.03) -- (307.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.36, 34.03) -- (307.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.48, 34.03) -- (307.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.57, 34.03) -- (307.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.71, 34.03) -- (307.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.78, 34.03) -- (307.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.86, 34.03) -- (307.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.14, 34.03) -- (308.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.17, 34.03) -- (308.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.40, 34.03) -- (308.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.29, 34.03) -- (309.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.65, 34.03) -- (309.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.66, 34.03) -- (309.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.77, 34.03) -- (309.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.85, 34.03) -- (309.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.26, 34.03) -- (310.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.42, 34.03) -- (310.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.88, 34.03) -- (310.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.97, 34.03) -- (310.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.25, 34.03) -- (311.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.09, 34.03) -- (312.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.22, 34.03) -- (312.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.41, 34.03) -- (312.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.42, 34.03) -- (313.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.62, 34.03) -- (313.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.78, 34.03) -- (313.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.05, 34.03) -- (314.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.50, 34.03) -- (314.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.58, 34.03) -- (314.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.67, 34.03) -- (314.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.27, 34.03) -- (315.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.36, 34.03) -- (315.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.70, 34.03) -- (315.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.87, 34.03) -- (315.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.98, 34.03) -- (315.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.35, 34.03) -- (316.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.56, 34.03) -- (316.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.72, 34.03) -- (316.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.21, 34.03) -- (317.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.49, 34.03) -- (317.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.56, 34.03) -- (317.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.84, 34.03) -- (318.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.87, 34.03) -- (318.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.28, 34.03) -- (319.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.56, 34.03) -- (319.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.75, 34.03) -- (319.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.97, 34.03) -- (319.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.23, 34.03) -- (320.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.29, 34.03) -- (320.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.33, 34.03) -- (320.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.94, 34.03) -- (320.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.01, 34.03) -- (321.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.22, 34.03) -- (321.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.25, 34.03) -- (321.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.36, 34.03) -- (321.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.75, 34.03) -- (321.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.15, 34.03) -- (323.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.66, 34.03) -- (323.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.17, 34.03) -- (325.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.77, 34.03) -- (325.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.89, 34.03) -- (325.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.94, 34.03) -- (326.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.01, 34.03) -- (327.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.29, 34.03) -- (327.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.29, 34.03) -- (328.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.36, 34.03) -- (328.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.82, 34.03) -- (328.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.10, 34.03) -- (329.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.71, 34.03) -- (329.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.08, 34.03) -- (330.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.10, 34.03) -- (330.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.14, 34.03) -- (330.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.37, 34.03) -- (330.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.66, 34.03) -- (330.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.19, 34.03) -- (331.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.66, 34.03) -- (331.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.92, 34.03) -- (331.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (332.49, 34.03) -- (332.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.01, 34.03) -- (333.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (334.21, 34.03) -- (334.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (334.41, 34.03) -- (334.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (334.89, 34.03) -- (334.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.42, 34.03) -- (337.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.45, 34.03) -- (337.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.58, 34.03) -- (337.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.70, 34.03) -- (337.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.71, 34.03) -- (337.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.80, 34.03) -- (337.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.18, 34.03) -- (339.18, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.28, 34.03) -- (339.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.35, 34.03) -- (341.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.75, 34.03) -- (341.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.17, 34.03) -- (342.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.48, 34.03) -- (342.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.53, 34.03) -- (342.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.69, 34.03) -- (342.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.85, 34.03) -- (342.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.03, 34.03) -- (343.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.17, 34.03) -- (343.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.37, 34.03) -- (343.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.45, 34.03) -- (343.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.75, 34.03) -- (343.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.88, 34.03) -- (343.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.23, 34.03) -- (344.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.06, 34.03) -- (345.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.03, 34.03) -- (346.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.49, 34.03) -- (346.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (347.23, 34.03) -- (347.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (347.38, 34.03) -- (347.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.49, 34.03) -- (348.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.58, 34.03) -- (348.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.13, 34.03) -- (349.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.81, 34.03) -- (349.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.10, 34.03) -- (350.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.24, 34.03) -- (350.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.58, 34.03) -- (350.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.85, 34.03) -- (351.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (352.05, 34.03) -- (352.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (352.77, 34.03) -- (352.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (353.17, 34.03) -- (353.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (353.42, 34.03) -- (353.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (353.71, 34.03) -- (353.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.71, 34.03) -- (354.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.56, 34.03) -- (355.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.56, 34.03) -- (357.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.72, 34.03) -- (357.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.97, 34.03) -- (357.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.00, 34.03) -- (358.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.30, 34.03) -- (358.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.37, 34.03) -- (358.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.57, 34.03) -- (358.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.60, 34.03) -- (358.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.96, 34.03) -- (358.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.97, 34.03) -- (358.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.44, 34.03) -- (359.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.71, 34.03) -- (359.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.84, 34.03) -- (359.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.89, 34.03) -- (359.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.95, 34.03) -- (359.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.12, 34.03) -- (360.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.50, 34.03) -- (360.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.53, 34.03) -- (360.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (361.15, 34.03) -- (361.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (361.24, 34.03) -- (361.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (361.57, 34.03) -- (361.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.32, 34.03) -- (362.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.86, 34.03) -- (362.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.88, 34.03) -- (362.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.99, 34.03) -- (362.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.37, 34.03) -- (363.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.15, 34.03) -- (365.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.51, 34.03) -- (365.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.58, 34.03) -- (365.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.58, 34.03) -- (365.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.24, 34.03) -- (366.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.27, 34.03) -- (366.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.79, 34.03) -- (366.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (367.09, 34.03) -- (367.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (367.28, 34.03) -- (367.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (367.50, 34.03) -- (367.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (367.74, 34.03) -- (367.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.30, 34.03) -- (368.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.30, 34.03) -- (368.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.38, 34.03) -- (368.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.49, 34.03) -- (368.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.90, 34.03) -- (368.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (369.46, 34.03) -- (369.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (370.56, 34.03) -- (370.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (370.74, 34.03) -- (370.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (370.85, 34.03) -- (370.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.47, 34.03) -- (372.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.39, 34.03) -- (373.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.79, 34.03) -- (373.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.21, 34.03) -- (374.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (375.70, 34.03) -- (375.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.93, 34.03) -- (378.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (380.42, 34.03) -- (380.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.52, 34.03) -- (385.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.08, 34.03) -- (386.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.54, 34.03) -- (386.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.90, 34.03) -- (386.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.31, 34.03) -- (387.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.53, 34.03) -- (387.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.24, 34.03) -- (388.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.35, 34.03) -- (388.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.69, 34.03) -- (388.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.71, 34.03) -- (388.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.75, 34.03) -- (388.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.90, 34.03) -- (388.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.17, 34.03) -- (389.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.73, 34.03) -- (389.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.59, 34.03) -- (390.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (391.22, 34.03) -- (391.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (391.77, 34.03) -- (391.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (391.82, 34.03) -- (391.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (392.15, 34.03) -- (392.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (392.41, 34.03) -- (392.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (396.33, 34.03) -- (396.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (397.13, 34.03) -- (397.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (397.28, 34.03) -- (397.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.32, 34.03) -- (404.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.80, 34.03) -- (404.80, 41.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.48, 34.03) --
	( 52.48,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37, 87.37) {-0.002};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37,145.86) {0.000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37,204.34) {0.002};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37,262.83) {0.004};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.48, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 53.04, 20.31) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 98.87, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (144.70, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (190.53, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (236.36, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (282.19, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (328.02, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (373.85, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (237.03,  4.82) {Ln(Min. Cap Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}