### 2014-09-25 LaTeX/mCityCoefPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-09-25 14:01:51
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] (124.66, 34.03) rectangle (277.03,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] (124.66, 34.03) rectangle (277.04,277.03);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (201.54, 49.88) -- (209.61, 49.88);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (193.96, 76.30) -- (232.45, 76.30);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (167.42,102.71) -- (183.20,102.71);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (143.88,129.12) -- (245.67,129.12);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (162.34,155.53) -- (193.68,155.53);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (200.93,181.95) -- (202.27,181.95);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (189.78,208.36) -- (231.19,208.36);

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (174.23,234.77) -- (205.28,234.77);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (203.17,261.19) -- (217.19,261.19);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (202.19, 49.88) -- (208.96, 49.88);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (197.05, 76.30) -- (229.35, 76.30);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (168.69,102.71) -- (181.93,102.71);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (152.06,129.12) -- (237.49,129.12);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (164.86,155.53) -- (191.16,155.53);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (201.04,181.95) -- (202.16,181.95);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (193.11,208.36) -- (227.86,208.36);

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (176.73,234.77) -- (202.79,234.77);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (204.30,261.19) -- (216.06,261.19);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}
\definecolor[named]{fillColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round,fill=fillColor] (200.85, 34.03) -- (200.85,277.03);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (210.18,261.19) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (189.76,234.77) circle (  2.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (210.49,208.36) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (201.60,181.95) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (178.01,155.53) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (194.77,129.12) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.31,102.71) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (213.20, 76.30) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (205.57, 49.88) circle (  2.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.61, 48.56) --
	(209.61, 51.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.61, 49.88) --
	(201.54, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.54, 48.56) --
	(201.54, 51.20);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.45, 74.97) --
	(232.45, 77.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.45, 76.30) --
	(193.96, 76.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.96, 74.97) --
	(193.96, 77.62);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.20,101.39) --
	(183.20,104.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.20,102.71) --
	(167.42,102.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.42,101.39) --
	(167.42,104.03);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.67,127.80) --
	(245.67,130.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.67,129.12) --
	(143.88,129.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.88,127.80) --
	(143.88,130.44);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.68,154.21) --
	(193.68,156.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (193.68,155.53) --
	(162.34,155.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.34,154.21) --
	(162.34,156.86);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.27,180.63) --
	(202.27,183.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.27,181.95) --
	(200.93,181.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.93,180.63) --
	(200.93,183.27);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.19,207.04) --
	(231.19,209.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.19,208.36) --
	(189.78,208.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.78,207.04) --
	(189.78,209.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.28,233.45) --
	(205.28,236.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.28,234.77) --
	(174.23,234.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (174.23,233.45) --
	(174.23,236.09);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.19,259.87) --
	(217.19,262.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.19,261.19) --
	(203.17,261.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.17,259.87) --
	(203.17,262.51);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] (124.66, 34.03) --
	(124.66,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55, 46.58) {World GDP Growth$_{t}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55, 72.99) {Democracy$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55, 99.40) {Ln(Inflation)$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,125.82) {Upper Income};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,152.23) {Area$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,178.64) {Duration$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,205.06) {Type$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,231.47) {Intensity$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,257.88) {Ln(Min. City Dist.)$_{t-1}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] (124.66, 34.03) --
	(277.03, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (131.59, 20.31) {-0.004};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (166.22, 20.31) {-0.002};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (200.85, 20.31) {0.000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (235.48, 20.31) {0.002};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (270.11, 20.31) {0.004};
\end{scope}
\end{tikzpicture}