### 2014-09-25 LaTeX/mCitySimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-09-25 14:22:36
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 52.48, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 52.48, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 69.26,108.08) --
	( 76.94,110.75) --
	( 81.21,112.23) --
	( 81.35,112.28) --
	( 94.75,116.93) --
	(103.21,119.87) --
	(108.87,121.84) --
	(112.14,122.97) --
	(118.51,125.19) --
	(123.77,127.01) --
	(127.53,128.32) --
	(130.04,129.19) --
	(138.05,131.97) --
	(158.20,138.97) --
	(171.57,143.61) --
	(171.70,143.66) --
	(174.96,144.79) --
	(175.86,145.11) --
	(185.97,148.62) --
	(188.46,149.48) --
	(195.56,151.95) --
	(205.05,155.24) --
	(205.06,155.25) --
	(209.59,156.82) --
	(209.87,156.92) --
	(211.41,157.45) --
	(212.10,157.69) --
	(212.80,157.94) --
	(212.89,157.97) --
	(213.53,158.19) --
	(213.69,158.24) --
	(215.24,158.78) --
	(220.60,160.64) --
	(222.23,161.21) --
	(222.43,161.28) --
	(224.02,161.83) --
	(228.32,163.33) --
	(228.78,163.49) --
	(233.22,165.03) --
	(235.00,165.65) --
	(235.50,165.82) --
	(236.76,166.26) --
	(238.27,166.78) --
	(238.97,167.03) --
	(239.94,167.36) --
	(240.65,167.61) --
	(240.90,167.70) --
	(240.95,167.71) --
	(241.36,167.86) --
	(244.68,169.01) --
	(246.54,169.65) --
	(247.54,170.00) --
	(248.75,170.42) --
	(249.31,170.62) --
	(250.77,171.12) --
	(251.67,171.44) --
	(252.68,171.79) --
	(253.66,172.13) --
	(254.85,172.54) --
	(254.86,172.54) --
	(255.67,172.83) --
	(256.23,173.02) --
	(256.77,173.21) --
	(257.49,173.46) --
	(257.71,173.54) --
	(258.26,173.73) --
	(258.74,173.89) --
	(259.12,174.02) --
	(259.38,174.11) --
	(259.99,174.33) --
	(260.56,174.52) --
	(260.79,174.60) --
	(262.27,175.12) --
	(262.38,175.16) --
	(263.62,175.59) --
	(263.78,175.64) --
	(265.81,176.35) --
	(267.16,176.82) --
	(267.27,176.85) --
	(267.90,177.08) --
	(269.02,177.46) --
	(269.24,177.54) --
	(269.50,177.63) --
	(270.00,177.80) --
	(271.24,178.23) --
	(271.44,178.30) --
	(271.47,178.31) --
	(271.89,178.46) --
	(272.97,178.83) --
	(273.00,178.84) --
	(273.30,178.95) --
	(274.10,179.23) --
	(274.51,179.37) --
	(275.15,179.59) --
	(275.85,179.83) --
	(276.45,180.04) --
	(276.57,180.09) --
	(277.67,180.47) --
	(277.74,180.49) --
	(278.30,180.69) --
	(278.99,180.93) --
	(279.37,181.06) --
	(280.17,181.34) --
	(281.01,181.63) --
	(281.75,181.88) --
	(282.04,181.99) --
	(282.56,182.17) --
	(282.59,182.17) --
	(282.73,182.22) --
	(283.17,182.38) --
	(283.67,182.55) --
	(284.33,182.78) --
	(284.38,182.80) --
	(284.78,182.94) --
	(285.47,183.18) --
	(285.59,183.22) --
	(285.79,183.29) --
	(285.93,183.34) --
	(285.96,183.35) --
	(286.40,183.50) --
	(286.76,183.63) --
	(286.94,183.69) --
	(287.11,183.75) --
	(287.40,183.85) --
	(287.56,183.90) --
	(287.64,183.93) --
	(288.93,184.38) --
	(289.17,184.46) --
	(289.98,184.74) --
	(290.06,184.77) --
	(290.15,184.80) --
	(290.68,184.98) --
	(290.90,185.06) --
	(291.07,185.12) --
	(291.16,185.15) --
	(291.99,185.44) --
	(292.53,185.63) --
	(292.56,185.64) --
	(293.19,185.86) --
	(293.72,186.04) --
	(294.42,186.28) --
	(294.46,186.30) --
	(295.07,186.51) --
	(295.58,186.69) --
	(296.54,187.02) --
	(296.64,187.06) --
	(296.96,187.17) --
	(297.07,187.21) --
	(297.55,187.37) --
	(297.61,187.39) --
	(297.93,187.50) --
	(298.07,187.55) --
	(298.13,187.57) --
	(298.19,187.59) --
	(298.21,187.60) --
	(298.32,187.64) --
	(298.88,187.83) --
	(299.55,188.07) --
	(300.83,188.51) --
	(301.02,188.58) --
	(301.08,188.60) --
	(301.48,188.74) --
	(301.54,188.76) --
	(301.64,188.79) --
	(301.73,188.83) --
	(302.00,188.92) --
	(302.17,188.98) --
	(302.49,189.09) --
	(302.51,189.10) --
	(302.54,189.11) --
	(302.55,189.11) --
	(302.59,189.12) --
	(302.62,189.13) --
	(302.72,189.17) --
	(302.80,189.20) --
	(302.80,189.20) --
	(303.19,189.33) --
	(303.42,189.41) --
	(303.79,189.54) --
	(304.00,189.61) --
	(304.46,189.77) --
	(305.92,190.28) --
	(306.03,190.32) --
	(306.63,190.53) --
	(307.38,190.79) --
	(307.50,190.83) --
	(307.76,190.92) --
	(308.06,191.02) --
	(308.22,191.08) --
	(308.71,191.25) --
	(309.07,191.37) --
	(309.23,191.43) --
	(309.24,191.43) --
	(309.36,191.48) --
	(309.39,191.49) --
	(309.44,191.50) --
	(310.04,191.71) --
	(310.57,191.90) --
	(310.65,191.92) --
	(310.74,191.95) --
	(310.97,192.03) --
	(311.29,192.15) --
	(311.41,192.19) --
	(312.93,192.71) --
	(312.98,192.73) --
	(313.47,192.90) --
	(313.56,192.93) --
	(314.18,193.15) --
	(314.39,193.22) --
	(314.46,193.25) --
	(314.61,193.30) --
	(314.93,193.41) --
	(315.17,193.49) --
	(315.66,193.66) --
	(315.81,193.72) --
	(315.92,193.75) --
	(316.04,193.79) --
	(316.27,193.88) --
	(316.28,193.88) --
	(316.31,193.89) --
	(316.40,193.92) --
	(316.53,193.97) --
	(316.54,193.97) --
	(316.73,194.03) --
	(316.73,194.03) --
	(317.05,194.15) --
	(317.72,194.38) --
	(317.91,194.44) --
	(318.05,194.49) --
	(318.31,194.58) --
	(318.59,194.68) --
	(318.98,194.82) --
	(319.05,194.84) --
	(319.19,194.89) --
	(319.36,194.95) --
	(319.71,195.07) --
	(319.91,195.14) --
	(319.93,195.15) --
	(320.06,195.19) --
	(320.17,195.23) --
	(320.20,195.24) --
	(320.30,195.27) --
	(320.38,195.30) --
	(320.46,195.33) --
	(320.79,195.44) --
	(320.87,195.47) --
	(321.03,195.53) --
	(321.25,195.60) --
	(321.56,195.71) --
	(321.69,195.76) --
	(321.81,195.80) --
	(321.97,195.85) --
	(322.34,195.98) --
	(322.35,195.99) --
	(322.38,196.00) --
	(322.44,196.02) --
	(322.47,196.03) --
	(322.55,196.06) --
	(322.98,196.21) --
	(323.15,196.26) --
	(323.39,196.35) --
	(323.51,196.39) --
	(323.64,196.43) --
	(323.74,196.47) --
	(324.03,196.57) --
	(324.31,196.67) --
	(324.91,196.88) --
	(326.25,197.34) --
	(326.31,197.36) --
	(326.44,197.41) --
	(326.69,197.49) --
	(326.98,197.59) --
	(327.12,197.64) --
	(327.38,197.73) --
	(327.45,197.76) --
	(327.62,197.82) --
	(328.36,198.07) --
	(328.71,198.20) --
	(328.89,198.26) --
	(329.00,198.30) --
	(329.34,198.41) --
	(330.21,198.72) --
	(330.77,198.91) --
	(330.92,198.96) --
	(331.77,199.26) --
	(332.05,199.35) --
	(332.77,199.61) --
	(333.03,199.70) --
	(333.16,199.74) --
	(333.59,199.89) --
	(335.17,200.44) --
	(335.96,200.71) --
	(336.21,200.80) --
	(336.98,201.07) --
	(337.10,201.11) --
	(337.64,201.30) --
	(339.32,201.88) --
	(339.85,202.06) --
	(340.50,202.29) --
	(340.63,202.33) --
	(341.25,202.55) --
	(341.62,202.68) --
	(341.97,202.80) --
	(342.00,202.81) --
	(342.05,202.83) --
	(342.45,202.97) --
	(342.53,202.99) --
	(343.04,203.17) --
	(343.17,203.22) --
	(343.85,203.45) --
	(343.87,203.46) --
	(343.95,203.49) --
	(344.16,203.56) --
	(344.46,203.67) --
	(344.63,203.72) --
	(344.93,203.83) --
	(345.02,203.86) --
	(345.07,203.88) --
	(345.10,203.89) --
	(345.51,204.03) --
	(345.57,204.05) --
	(346.09,204.23) --
	(346.39,204.34) --
	(347.80,204.83) --
	(348.20,204.96) --
	(348.41,205.04) --
	(348.87,205.20) --
	(348.92,205.21) --
	(349.60,205.45) --
	(351.33,206.05) --
	(351.43,206.09) --
	(351.56,206.13) --
	(351.74,206.20) --
	(351.87,206.24) --
	(353.20,206.70) --
	(353.44,206.78) --
	(353.54,206.82) --
	(354.59,207.18) --
	(357.17,208.08) --
	(357.38,208.15) --
	(357.90,208.33) --
	(357.93,208.34) --
	(358.21,208.44) --
	(358.75,208.63) --
	(359.57,208.92) --
	(359.62,208.93) --
	(360.33,209.18) --
	(360.36,209.19) --
	(360.85,209.36) --
	(361.40,209.55) --
	(362.06,209.78) --
	(362.35,209.88) --
	(362.48,209.93) --
	(362.58,209.96) --
	(363.30,210.21) --
	(363.33,210.22) --
	(363.40,210.24) --
	(364.54,210.64) --
	(365.21,210.87) --
	(366.77,211.42) --
	(366.99,211.49) --
	(368.17,211.90) --
	(368.73,212.10) --
	(372.78,213.50) --
	(373.25,213.67) --
	(374.25,214.01) --
	(374.25,214.01) --
	(374.55,214.12) --
	(374.76,214.19) --
	(375.05,214.29) --
	(375.30,214.38) --
	(375.66,214.50) --
	(375.91,214.59) --
	(376.54,214.81) --
	(376.88,214.93) --
	(377.80,215.25) --
	(378.36,215.44) --
	(378.50,215.49) --
	(378.91,215.63) --
	(379.95,215.99) --
	(381.22,216.44) --
	(381.23,216.44) --
	(381.92,216.68) --
	(381.95,216.69) --
	(382.82,216.99) --
	(383.24,217.14) --
	(383.51,217.23) --
	(384.09,217.43) --
	(384.17,217.46) --
	(384.29,217.50) --
	(384.73,217.65) --
	(386.66,218.32) --
	(386.77,218.36) --
	(390.32,219.59) --
	(394.80,221.15) --
	(396.85,221.86) --
	(404.07,224.37) --
	(404.80,224.62);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 69.26,174.00) --
	( 76.94,174.91) --
	( 81.21,175.28) --
	( 81.35,175.28) --
	( 94.75,176.82) --
	(103.21,177.55) --
	(108.87,178.21) --
	(112.14,178.55) --
	(118.51,179.36) --
	(123.77,179.97) --
	(127.53,180.45) --
	(130.04,180.72) --
	(138.05,181.83) --
	(158.20,184.48) --
	(171.57,186.12) --
	(171.70,186.12) --
	(174.96,186.43) --
	(175.86,186.59) --
	(185.97,188.24) --
	(188.46,188.71) --
	(195.56,189.94) --
	(205.05,191.47) --
	(205.06,191.47) --
	(209.59,192.51) --
	(209.87,192.54) --
	(211.41,192.84) --
	(212.10,193.02) --
	(212.80,193.13) --
	(212.89,193.15) --
	(213.53,193.30) --
	(213.69,193.33) --
	(215.24,193.64) --
	(220.60,194.57) --
	(222.23,194.95) --
	(222.43,194.98) --
	(224.02,195.30) --
	(228.32,196.22) --
	(228.78,196.37) --
	(233.22,197.33) --
	(235.00,197.73) --
	(235.50,197.81) --
	(236.76,198.12) --
	(238.27,198.42) --
	(238.97,198.51) --
	(239.94,198.70) --
	(240.65,198.90) --
	(240.90,198.95) --
	(240.95,198.96) --
	(241.36,199.05) --
	(244.68,199.85) --
	(246.54,200.21) --
	(247.54,200.41) --
	(248.75,200.77) --
	(249.31,200.94) --
	(250.77,201.35) --
	(251.67,201.62) --
	(252.68,201.84) --
	(253.66,202.15) --
	(254.85,202.44) --
	(254.86,202.44) --
	(255.67,202.67) --
	(256.23,202.80) --
	(256.77,202.98) --
	(257.49,203.17) --
	(257.71,203.20) --
	(258.26,203.36) --
	(258.74,203.50) --
	(259.12,203.66) --
	(259.38,203.76) --
	(259.99,203.84) --
	(260.56,203.97) --
	(260.79,204.00) --
	(262.27,204.53) --
	(262.38,204.57) --
	(263.62,204.89) --
	(263.78,204.92) --
	(265.81,205.57) --
	(267.16,205.98) --
	(267.27,206.00) --
	(267.90,206.19) --
	(269.02,206.47) --
	(269.24,206.53) --
	(269.50,206.63) --
	(270.00,206.83) --
	(271.24,207.27) --
	(271.44,207.33) --
	(271.47,207.33) --
	(271.89,207.39) --
	(272.97,207.77) --
	(273.00,207.78) --
	(273.30,207.93) --
	(274.10,208.20) --
	(274.51,208.30) --
	(275.15,208.47) --
	(275.85,208.69) --
	(276.45,208.85) --
	(276.57,208.88) --
	(277.67,209.23) --
	(277.74,209.26) --
	(278.30,209.41) --
	(278.99,209.61) --
	(279.37,209.72) --
	(280.17,209.90) --
	(281.01,210.17) --
	(281.75,210.39) --
	(282.04,210.51) --
	(282.56,210.73) --
	(282.59,210.74) --
	(282.73,210.80) --
	(283.17,210.98) --
	(283.67,211.13) --
	(284.33,211.34) --
	(284.38,211.36) --
	(284.78,211.50) --
	(285.47,211.75) --
	(285.59,211.80) --
	(285.79,211.87) --
	(285.93,211.89) --
	(285.96,211.90) --
	(286.40,212.01) --
	(286.76,212.10) --
	(286.94,212.15) --
	(287.11,212.19) --
	(287.40,212.28) --
	(287.56,212.36) --
	(287.64,212.40) --
	(288.93,212.86) --
	(289.17,212.95) --
	(289.98,213.19) --
	(290.06,213.22) --
	(290.15,213.26) --
	(290.68,213.45) --
	(290.90,213.53) --
	(291.07,213.59) --
	(291.16,213.62) --
	(291.99,213.93) --
	(292.53,214.13) --
	(292.56,214.14) --
	(293.19,214.38) --
	(293.72,214.62) --
	(294.42,214.91) --
	(294.46,214.93) --
	(295.07,215.21) --
	(295.58,215.42) --
	(296.54,215.77) --
	(296.64,215.82) --
	(296.96,215.91) --
	(297.07,215.95) --
	(297.55,216.15) --
	(297.61,216.18) --
	(297.93,216.29) --
	(298.07,216.34) --
	(298.13,216.36) --
	(298.19,216.38) --
	(298.21,216.39) --
	(298.32,216.43) --
	(298.88,216.63) --
	(299.55,216.84) --
	(300.83,217.35) --
	(301.02,217.42) --
	(301.08,217.44) --
	(301.48,217.63) --
	(301.54,217.65) --
	(301.64,217.68) --
	(301.73,217.71) --
	(302.00,217.78) --
	(302.17,217.84) --
	(302.49,217.96) --
	(302.51,217.97) --
	(302.54,217.98) --
	(302.55,217.98) --
	(302.59,218.00) --
	(302.62,218.02) --
	(302.72,218.07) --
	(302.80,218.10) --
	(302.80,218.10) --
	(303.19,218.24) --
	(303.42,218.32) --
	(303.79,218.45) --
	(304.00,218.54) --
	(304.46,218.72) --
	(305.92,219.26) --
	(306.03,219.30) --
	(306.63,219.55) --
	(307.38,219.82) --
	(307.50,219.86) --
	(307.76,219.99) --
	(308.06,220.05) --
	(308.22,220.15) --
	(308.71,220.34) --
	(309.07,220.51) --
	(309.23,220.58) --
	(309.24,220.58) --
	(309.36,220.63) --
	(309.39,220.64) --
	(309.44,220.65) --
	(310.04,220.93) --
	(310.57,221.18) --
	(310.65,221.20) --
	(310.74,221.23) --
	(310.97,221.37) --
	(311.29,221.50) --
	(311.41,221.54) --
	(312.93,222.17) --
	(312.98,222.19) --
	(313.47,222.35) --
	(313.56,222.39) --
	(314.18,222.61) --
	(314.39,222.66) --
	(314.46,222.67) --
	(314.61,222.71) --
	(314.93,222.88) --
	(315.17,222.95) --
	(315.66,223.09) --
	(315.81,223.13) --
	(315.92,223.16) --
	(316.04,223.21) --
	(316.27,223.29) --
	(316.28,223.29) --
	(316.31,223.29) --
	(316.40,223.32) --
	(316.53,223.37) --
	(316.54,223.38) --
	(316.73,223.47) --
	(316.73,223.47) --
	(317.05,223.60) --
	(317.72,223.94) --
	(317.91,224.00) --
	(318.05,224.03) --
	(318.31,224.13) --
	(318.59,224.22) --
	(318.98,224.34) --
	(319.05,224.36) --
	(319.19,224.40) --
	(319.36,224.48) --
	(319.71,224.64) --
	(319.91,224.74) --
	(319.93,224.75) --
	(320.06,224.83) --
	(320.17,224.87) --
	(320.20,224.88) --
	(320.30,224.92) --
	(320.38,224.96) --
	(320.46,224.99) --
	(320.79,225.13) --
	(320.87,225.16) --
	(321.03,225.25) --
	(321.25,225.37) --
	(321.56,225.45) --
	(321.69,225.50) --
	(321.81,225.57) --
	(321.97,225.64) --
	(322.34,225.80) --
	(322.35,225.81) --
	(322.38,225.82) --
	(322.44,225.86) --
	(322.47,225.88) --
	(322.55,225.93) --
	(322.98,226.04) --
	(323.15,226.10) --
	(323.39,226.21) --
	(323.51,226.27) --
	(323.64,226.34) --
	(323.74,226.38) --
	(324.03,226.53) --
	(324.31,226.67) --
	(324.91,226.96) --
	(326.25,227.49) --
	(326.31,227.51) --
	(326.44,227.55) --
	(326.69,227.67) --
	(326.98,227.78) --
	(327.12,227.86) --
	(327.38,228.02) --
	(327.45,228.05) --
	(327.62,228.10) --
	(328.36,228.30) --
	(328.71,228.45) --
	(328.89,228.55) --
	(329.00,228.61) --
	(329.34,228.76) --
	(330.21,229.08) --
	(330.77,229.33) --
	(330.92,229.40) --
	(331.77,229.82) --
	(332.05,229.95) --
	(332.77,230.21) --
	(333.03,230.32) --
	(333.16,230.38) --
	(333.59,230.58) --
	(335.17,231.30) --
	(335.96,231.66) --
	(336.21,231.80) --
	(336.98,232.18) --
	(337.10,232.25) --
	(337.64,232.45) --
	(339.32,233.28) --
	(339.85,233.56) --
	(340.50,233.89) --
	(340.63,233.96) --
	(341.25,234.28) --
	(341.62,234.47) --
	(341.97,234.60) --
	(342.00,234.62) --
	(342.05,234.66) --
	(342.45,234.88) --
	(342.53,234.91) --
	(343.04,235.11) --
	(343.17,235.17) --
	(343.85,235.44) --
	(343.87,235.45) --
	(343.95,235.48) --
	(344.16,235.58) --
	(344.46,235.71) --
	(344.63,235.81) --
	(344.93,235.95) --
	(345.02,235.98) --
	(345.07,236.01) --
	(345.10,236.02) --
	(345.51,236.18) --
	(345.57,236.22) --
	(346.09,236.43) --
	(346.39,236.59) --
	(347.80,237.20) --
	(348.20,237.34) --
	(348.41,237.44) --
	(348.87,237.68) --
	(348.92,237.70) --
	(349.60,238.04) --
	(351.33,238.84) --
	(351.43,238.88) --
	(351.56,238.95) --
	(351.74,239.05) --
	(351.87,239.12) --
	(353.20,239.73) --
	(353.44,239.87) --
	(353.54,239.91) --
	(354.59,240.36) --
	(357.17,241.77) --
	(357.38,241.84) --
	(357.90,242.12) --
	(357.93,242.13) --
	(358.21,242.26) --
	(358.75,242.53) --
	(359.57,242.91) --
	(359.62,242.93) --
	(360.33,243.29) --
	(360.36,243.30) --
	(360.85,243.52) --
	(361.40,243.78) --
	(362.06,244.10) --
	(362.35,244.25) --
	(362.48,244.30) --
	(362.58,244.35) --
	(363.30,244.68) --
	(363.33,244.70) --
	(363.40,244.73) --
	(364.54,245.27) --
	(365.21,245.59) --
	(366.77,246.33) --
	(366.99,246.43) --
	(368.17,246.98) --
	(368.73,247.27) --
	(372.78,249.35) --
	(373.25,249.62) --
	(374.25,250.09) --
	(374.25,250.09) --
	(374.55,250.26) --
	(374.76,250.38) --
	(375.05,250.52) --
	(375.30,250.65) --
	(375.66,250.84) --
	(375.91,251.00) --
	(376.54,251.30) --
	(376.88,251.43) --
	(377.80,251.88) --
	(378.36,252.13) --
	(378.50,252.21) --
	(378.91,252.40) --
	(379.95,252.87) --
	(381.22,253.47) --
	(381.23,253.47) --
	(381.92,253.81) --
	(381.95,253.82) --
	(382.82,254.27) --
	(383.24,254.49) --
	(383.51,254.65) --
	(384.09,254.97) --
	(384.17,255.02) --
	(384.29,255.09) --
	(384.73,255.34) --
	(386.66,256.33) --
	(386.77,256.39) --
	(390.32,258.27) --
	(394.80,260.52) --
	(396.85,261.63) --
	(404.07,265.49) --
	(404.80,265.99) --
	(404.80,183.67) --
	(404.07,183.52) --
	(396.85,182.44) --
	(394.80,181.97) --
	(390.32,181.18) --
	(386.77,180.54) --
	(386.66,180.52) --
	(384.73,180.06) --
	(384.29,179.98) --
	(384.17,179.97) --
	(384.09,179.97) --
	(383.51,179.91) --
	(383.24,179.86) --
	(382.82,179.77) --
	(381.95,179.60) --
	(381.92,179.60) --
	(381.23,179.45) --
	(381.22,179.45) --
	(379.95,179.17) --
	(378.91,178.92) --
	(378.50,178.83) --
	(378.36,178.78) --
	(377.80,178.72) --
	(376.88,178.52) --
	(376.54,178.43) --
	(375.91,178.35) --
	(375.66,178.34) --
	(375.30,178.26) --
	(375.05,178.23) --
	(374.76,178.17) --
	(374.55,178.11) --
	(374.25,178.04) --
	(374.25,178.03) --
	(373.25,177.72) --
	(372.78,177.61) --
	(368.73,176.81) --
	(368.17,176.63) --
	(366.99,176.37) --
	(366.77,176.32) --
	(365.21,176.05) --
	(364.54,175.92) --
	(363.40,175.69) --
	(363.33,175.66) --
	(363.30,175.66) --
	(362.58,175.51) --
	(362.48,175.51) --
	(362.35,175.47) --
	(362.06,175.40) --
	(361.40,175.29) --
	(360.85,175.14) --
	(360.36,174.95) --
	(360.33,174.95) --
	(359.62,174.91) --
	(359.57,174.90) --
	(358.75,174.64) --
	(358.21,174.57) --
	(357.93,174.52) --
	(357.90,174.51) --
	(357.38,174.41) --
	(357.17,174.33) --
	(354.59,173.96) --
	(353.54,173.77) --
	(353.44,173.75) --
	(353.20,173.71) --
	(351.87,173.38) --
	(351.74,173.35) --
	(351.56,173.33) --
	(351.43,173.30) --
	(351.33,173.27) --
	(349.60,172.85) --
	(348.92,172.68) --
	(348.87,172.67) --
	(348.41,172.62) --
	(348.20,172.56) --
	(347.80,172.44) --
	(346.39,172.18) --
	(346.09,172.06) --
	(345.57,171.89) --
	(345.51,171.87) --
	(345.10,171.77) --
	(345.07,171.76) --
	(345.02,171.75) --
	(344.93,171.72) --
	(344.63,171.64) --
	(344.46,171.61) --
	(344.16,171.57) --
	(343.95,171.54) --
	(343.87,171.53) --
	(343.85,171.53) --
	(343.17,171.40) --
	(343.04,171.38) --
	(342.53,171.23) --
	(342.45,171.23) --
	(342.05,171.18) --
	(342.00,171.16) --
	(341.97,171.15) --
	(341.62,171.04) --
	(341.25,170.97) --
	(340.63,170.85) --
	(340.50,170.83) --
	(339.85,170.68) --
	(339.32,170.51) --
	(337.64,169.94) --
	(337.10,169.80) --
	(336.98,169.80) --
	(336.21,169.64) --
	(335.96,169.52) --
	(335.17,169.32) --
	(333.59,168.83) --
	(333.16,168.72) --
	(333.03,168.69) --
	(332.77,168.66) --
	(332.05,168.46) --
	(331.77,168.38) --
	(330.92,168.18) --
	(330.77,168.15) --
	(330.21,168.06) --
	(329.34,167.81) --
	(329.00,167.74) --
	(328.89,167.72) --
	(328.71,167.68) --
	(328.36,167.60) --
	(327.62,167.36) --
	(327.45,167.31) --
	(327.38,167.30) --
	(327.12,167.23) --
	(326.98,167.20) --
	(326.69,167.12) --
	(326.44,167.04) --
	(326.31,167.00) --
	(326.25,166.99) --
	(324.91,166.70) --
	(324.31,166.61) --
	(324.03,166.54) --
	(323.74,166.46) --
	(323.64,166.42) --
	(323.51,166.37) --
	(323.39,166.35) --
	(323.15,166.30) --
	(322.98,166.27) --
	(322.55,166.11) --
	(322.47,166.08) --
	(322.44,166.07) --
	(322.38,166.04) --
	(322.35,166.04) --
	(322.34,166.04) --
	(321.97,165.94) --
	(321.81,165.90) --
	(321.69,165.87) --
	(321.56,165.83) --
	(321.25,165.73) --
	(321.03,165.68) --
	(320.87,165.63) --
	(320.79,165.58) --
	(320.46,165.51) --
	(320.38,165.50) --
	(320.30,165.48) --
	(320.20,165.46) --
	(320.17,165.45) --
	(320.06,165.42) --
	(319.93,165.38) --
	(319.91,165.37) --
	(319.71,165.28) --
	(319.36,165.11) --
	(319.19,165.08) --
	(319.05,165.05) --
	(318.98,165.03) --
	(318.59,164.93) --
	(318.31,164.84) --
	(318.05,164.78) --
	(317.91,164.76) --
	(317.72,164.72) --
	(317.05,164.59) --
	(316.73,164.53) --
	(316.73,164.53) --
	(316.54,164.50) --
	(316.53,164.50) --
	(316.40,164.47) --
	(316.31,164.45) --
	(316.28,164.44) --
	(316.27,164.44) --
	(316.04,164.36) --
	(315.92,164.33) --
	(315.81,164.29) --
	(315.66,164.24) --
	(315.17,164.10) --
	(314.93,164.02) --
	(314.61,163.92) --
	(314.46,163.86) --
	(314.39,163.83) --
	(314.18,163.75) --
	(313.56,163.59) --
	(313.47,163.56) --
	(312.98,163.44) --
	(312.93,163.43) --
	(311.41,163.01) --
	(311.29,162.97) --
	(310.97,162.86) --
	(310.74,162.80) --
	(310.65,162.79) --
	(310.57,162.77) --
	(310.04,162.61) --
	(309.44,162.38) --
	(309.39,162.37) --
	(309.36,162.36) --
	(309.24,162.34) --
	(309.23,162.34) --
	(309.07,162.31) --
	(308.71,162.22) --
	(308.22,162.06) --
	(308.06,162.00) --
	(307.76,161.88) --
	(307.50,161.76) --
	(307.38,161.70) --
	(306.63,161.40) --
	(306.03,161.19) --
	(305.92,161.16) --
	(304.46,160.70) --
	(304.00,160.53) --
	(303.79,160.51) --
	(303.42,160.35) --
	(303.19,160.23) --
	(302.80,160.09) --
	(302.80,160.09) --
	(302.72,160.06) --
	(302.62,160.03) --
	(302.59,160.01) --
	(302.55,160.00) --
	(302.54,160.00) --
	(302.51,159.98) --
	(302.49,159.98) --
	(302.17,159.86) --
	(302.00,159.79) --
	(301.73,159.71) --
	(301.64,159.67) --
	(301.54,159.63) --
	(301.48,159.61) --
	(301.08,159.53) --
	(301.02,159.52) --
	(300.83,159.48) --
	(299.55,159.10) --
	(298.88,158.93) --
	(298.32,158.72) --
	(298.21,158.66) --
	(298.19,158.66) --
	(298.13,158.62) --
	(298.07,158.60) --
	(297.93,158.55) --
	(297.61,158.41) --
	(297.55,158.38) --
	(297.07,158.23) --
	(296.96,158.20) --
	(296.64,158.11) --
	(296.54,158.10) --
	(295.58,157.82) --
	(295.07,157.63) --
	(294.46,157.42) --
	(294.42,157.41) --
	(293.72,157.19) --
	(293.19,157.02) --
	(292.56,156.85) --
	(292.53,156.84) --
	(291.99,156.75) --
	(291.16,156.46) --
	(291.07,156.43) --
	(290.90,156.37) --
	(290.68,156.28) --
	(290.15,156.09) --
	(290.06,156.07) --
	(289.98,156.04) --
	(289.17,155.82) --
	(288.93,155.75) --
	(287.64,155.36) --
	(287.56,155.32) --
	(287.40,155.25) --
	(287.11,155.16) --
	(286.94,155.09) --
	(286.76,155.04) --
	(286.40,154.93) --
	(285.96,154.73) --
	(285.93,154.72) --
	(285.79,154.65) --
	(285.59,154.56) --
	(285.47,154.54) --
	(284.78,154.30) --
	(284.38,154.16) --
	(284.33,154.14) --
	(283.67,153.90) --
	(283.17,153.69) --
	(282.73,153.53) --
	(282.59,153.50) --
	(282.56,153.49) --
	(282.04,153.22) --
	(281.75,153.16) --
	(281.01,152.85) --
	(280.17,152.58) --
	(279.37,152.33) --
	(278.99,152.24) --
	(278.30,151.99) --
	(277.74,151.83) --
	(277.67,151.80) --
	(276.57,151.29) --
	(276.45,151.23) --
	(275.85,150.95) --
	(275.15,150.75) --
	(274.51,150.62) --
	(274.10,150.45) --
	(273.30,150.09) --
	(273.00,149.95) --
	(272.97,149.94) --
	(271.89,149.56) --
	(271.47,149.38) --
	(271.44,149.38) --
	(271.24,149.28) --
	(270.00,148.75) --
	(269.50,148.57) --
	(269.24,148.50) --
	(269.02,148.44) --
	(267.90,147.97) --
	(267.27,147.77) --
	(267.16,147.74) --
	(265.81,147.22) --
	(263.78,146.42) --
	(263.62,146.35) --
	(262.38,145.77) --
	(262.27,145.73) --
	(260.79,145.17) --
	(260.56,145.05) --
	(259.99,144.87) --
	(259.38,144.57) --
	(259.12,144.44) --
	(258.74,144.28) --
	(258.26,144.06) --
	(257.71,143.89) --
	(257.49,143.80) --
	(256.77,143.49) --
	(256.23,143.20) --
	(255.67,143.00) --
	(254.86,142.66) --
	(254.85,142.65) --
	(253.66,142.12) --
	(252.68,141.77) --
	(251.67,141.30) --
	(250.77,140.91) --
	(249.31,140.38) --
	(248.75,140.16) --
	(247.54,139.62) --
	(246.54,139.15) --
	(244.68,138.39) --
	(241.36,136.98) --
	(240.95,136.85) --
	(240.90,136.83) --
	(240.65,136.72) --
	(239.94,136.44) --
	(238.97,136.03) --
	(238.27,135.71) --
	(236.76,135.03) --
	(235.50,134.47) --
	(235.00,134.30) --
	(233.22,133.55) --
	(228.78,131.62) --
	(228.32,131.44) --
	(224.02,129.53) --
	(222.43,128.74) --
	(222.23,128.64) --
	(220.60,127.76) --
	(215.24,125.24) --
	(213.69,124.40) --
	(213.53,124.34) --
	(212.89,124.01) --
	(212.80,123.97) --
	(212.10,123.55) --
	(211.41,123.27) --
	(209.87,122.48) --
	(209.59,122.31) --
	(205.06,119.98) --
	(205.05,119.98) --
	(195.56,115.19) --
	(188.46,111.38) --
	(185.97,110.03) --
	(175.86,104.77) --
	(174.96,104.36) --
	(171.70,102.75) --
	(171.57,102.68) --
	(158.20, 95.45) --
	(138.05, 84.52) --
	(130.04, 80.01) --
	(127.53, 78.54) --
	(123.77, 76.37) --
	(118.51, 73.39) --
	(112.14, 69.76) --
	(108.87, 67.97) --
	(103.21, 64.72) --
	( 94.75, 59.85) --
	( 81.35, 52.16) --
	( 81.21, 52.08) --
	( 76.94, 49.62) --
	( 69.26, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 69.26,162.69) --
	( 76.94,163.86) --
	( 81.21,164.49) --
	( 81.35,164.51) --
	( 94.75,166.70) --
	(103.21,167.92) --
	(108.87,168.87) --
	(112.14,169.36) --
	(118.51,170.37) --
	(123.77,171.22) --
	(127.53,171.70) --
	(130.04,172.14) --
	(138.05,173.52) --
	(158.20,177.08) --
	(171.57,179.46) --
	(171.70,179.49) --
	(174.96,180.16) --
	(175.86,180.26) --
	(185.97,182.09) --
	(188.46,182.56) --
	(195.56,184.04) --
	(205.05,185.94) --
	(205.06,185.94) --
	(209.59,186.79) --
	(209.87,186.84) --
	(211.41,187.13) --
	(212.10,187.25) --
	(212.80,187.40) --
	(212.89,187.41) --
	(213.53,187.51) --
	(213.69,187.55) --
	(215.24,187.91) --
	(220.60,189.21) --
	(222.23,189.53) --
	(222.43,189.58) --
	(224.02,189.88) --
	(228.32,190.86) --
	(228.78,190.94) --
	(233.22,192.19) --
	(235.00,192.66) --
	(235.50,192.79) --
	(236.76,193.15) --
	(238.27,193.50) --
	(238.97,193.69) --
	(239.94,193.90) --
	(240.65,194.05) --
	(240.90,194.11) --
	(240.95,194.13) --
	(241.36,194.25) --
	(244.68,195.12) --
	(246.54,195.56) --
	(247.54,195.84) --
	(248.75,196.16) --
	(249.31,196.34) --
	(250.77,196.74) --
	(251.67,197.02) --
	(252.68,197.28) --
	(253.66,197.57) --
	(254.85,197.95) --
	(254.86,197.95) --
	(255.67,198.20) --
	(256.23,198.38) --
	(256.77,198.52) --
	(257.49,198.70) --
	(257.71,198.75) --
	(258.26,198.88) --
	(258.74,199.00) --
	(259.12,199.11) --
	(259.38,199.21) --
	(259.99,199.39) --
	(260.56,199.55) --
	(260.79,199.64) --
	(262.27,200.15) --
	(262.38,200.20) --
	(263.62,200.57) --
	(263.78,200.65) --
	(265.81,201.31) --
	(267.16,201.67) --
	(267.27,201.70) --
	(267.90,201.87) --
	(269.02,202.14) --
	(269.24,202.18) --
	(269.50,202.25) --
	(270.00,202.43) --
	(271.24,202.77) --
	(271.44,202.85) --
	(271.47,202.86) --
	(271.89,202.98) --
	(272.97,203.23) --
	(273.00,203.24) --
	(273.30,203.29) --
	(274.10,203.53) --
	(274.51,203.67) --
	(275.15,203.94) --
	(275.85,204.13) --
	(276.45,204.30) --
	(276.57,204.32) --
	(277.67,204.64) --
	(277.74,204.66) --
	(278.30,204.87) --
	(278.99,205.11) --
	(279.37,205.22) --
	(280.17,205.50) --
	(281.01,205.81) --
	(281.75,206.07) --
	(282.04,206.16) --
	(282.56,206.35) --
	(282.59,206.36) --
	(282.73,206.41) --
	(283.17,206.58) --
	(283.67,206.75) --
	(284.33,206.98) --
	(284.38,207.00) --
	(284.78,207.17) --
	(285.47,207.40) --
	(285.59,207.43) --
	(285.79,207.48) --
	(285.93,207.52) --
	(285.96,207.53) --
	(286.40,207.71) --
	(286.76,207.81) --
	(286.94,207.90) --
	(287.11,207.95) --
	(287.40,208.07) --
	(287.56,208.10) --
	(287.64,208.13) --
	(288.93,208.48) --
	(289.17,208.54) --
	(289.98,208.82) --
	(290.06,208.86) --
	(290.15,208.88) --
	(290.68,209.05) --
	(290.90,209.11) --
	(291.07,209.16) --
	(291.16,209.18) --
	(291.99,209.47) --
	(292.53,209.68) --
	(292.56,209.68) --
	(293.19,209.95) --
	(293.72,210.12) --
	(294.42,210.36) --
	(294.46,210.38) --
	(295.07,210.61) --
	(295.58,210.83) --
	(296.54,211.25) --
	(296.64,211.30) --
	(296.96,211.44) --
	(297.07,211.49) --
	(297.55,211.64) --
	(297.61,211.66) --
	(297.93,211.79) --
	(298.07,211.84) --
	(298.13,211.86) --
	(298.19,211.88) --
	(298.21,211.88) --
	(298.32,211.91) --
	(298.88,212.06) --
	(299.55,212.34) --
	(300.83,212.79) --
	(301.02,212.84) --
	(301.08,212.86) --
	(301.48,212.99) --
	(301.54,213.01) --
	(301.64,213.04) --
	(301.73,213.07) --
	(302.00,213.13) --
	(302.17,213.19) --
	(302.49,213.31) --
	(302.51,213.32) --
	(302.54,213.33) --
	(302.55,213.33) --
	(302.59,213.34) --
	(302.62,213.35) --
	(302.72,213.37) --
	(302.80,213.40) --
	(302.80,213.40) --
	(303.19,213.59) --
	(303.42,213.69) --
	(303.79,213.78) --
	(304.00,213.85) --
	(304.46,214.02) --
	(305.92,214.57) --
	(306.03,214.62) --
	(306.63,214.86) --
	(307.38,215.16) --
	(307.50,215.19) --
	(307.76,215.29) --
	(308.06,215.37) --
	(308.22,215.41) --
	(308.71,215.58) --
	(309.07,215.76) --
	(309.23,215.83) --
	(309.24,215.83) --
	(309.36,215.86) --
	(309.39,215.87) --
	(309.44,215.89) --
	(310.04,216.08) --
	(310.57,216.34) --
	(310.65,216.38) --
	(310.74,216.42) --
	(310.97,216.52) --
	(311.29,216.64) --
	(311.41,216.70) --
	(312.93,217.33) --
	(312.98,217.35) --
	(313.47,217.50) --
	(313.56,217.54) --
	(314.18,217.74) --
	(314.39,217.80) --
	(314.46,217.82) --
	(314.61,217.89) --
	(314.93,218.08) --
	(315.17,218.17) --
	(315.66,218.43) --
	(315.81,218.48) --
	(315.92,218.51) --
	(316.04,218.55) --
	(316.27,218.63) --
	(316.28,218.63) --
	(316.31,218.64) --
	(316.40,218.67) --
	(316.53,218.73) --
	(316.54,218.74) --
	(316.73,218.82) --
	(316.73,218.83) --
	(317.05,218.99) --
	(317.72,219.22) --
	(317.91,219.30) --
	(318.05,219.35) --
	(318.31,219.45) --
	(318.59,219.60) --
	(318.98,219.81) --
	(319.05,219.85) --
	(319.19,219.89) --
	(319.36,219.93) --
	(319.71,220.04) --
	(319.91,220.14) --
	(319.93,220.15) --
	(320.06,220.21) --
	(320.17,220.24) --
	(320.20,220.24) --
	(320.30,220.28) --
	(320.38,220.31) --
	(320.46,220.35) --
	(320.79,220.45) --
	(320.87,220.47) --
	(321.03,220.52) --
	(321.25,220.63) --
	(321.56,220.77) --
	(321.69,220.83) --
	(321.81,220.87) --
	(321.97,220.93) --
	(322.34,221.05) --
	(322.35,221.06) --
	(322.38,221.06) --
	(322.44,221.08) --
	(322.47,221.09) --
	(322.55,221.12) --
	(322.98,221.32) --
	(323.15,221.38) --
	(323.39,221.47) --
	(323.51,221.54) --
	(323.64,221.60) --
	(323.74,221.65) --
	(324.03,221.78) --
	(324.31,221.90) --
	(324.91,222.19) --
	(326.25,222.75) --
	(326.31,222.77) --
	(326.44,222.85) --
	(326.69,222.97) --
	(326.98,223.07) --
	(327.12,223.12) --
	(327.38,223.22) --
	(327.45,223.25) --
	(327.62,223.32) --
	(328.36,223.62) --
	(328.71,223.74) --
	(328.89,223.84) --
	(329.00,223.88) --
	(329.34,224.03) --
	(330.21,224.37) --
	(330.77,224.65) --
	(330.92,224.71) --
	(331.77,224.99) --
	(332.05,225.10) --
	(332.77,225.35) --
	(333.03,225.45) --
	(333.16,225.52) --
	(333.59,225.70) --
	(335.17,226.36) --
	(335.96,226.68) --
	(336.21,226.79) --
	(336.98,227.14) --
	(337.10,227.18) --
	(337.64,227.38) --
	(339.32,228.03) --
	(339.85,228.20) --
	(340.50,228.45) --
	(340.63,228.52) --
	(341.25,228.77) --
	(341.62,228.90) --
	(341.97,229.05) --
	(342.00,229.06) --
	(342.05,229.08) --
	(342.45,229.24) --
	(342.53,229.28) --
	(343.04,229.50) --
	(343.17,229.54) --
	(343.85,229.85) --
	(343.87,229.86) --
	(343.95,229.87) --
	(344.16,229.97) --
	(344.46,230.12) --
	(344.63,230.20) --
	(344.93,230.36) --
	(345.02,230.39) --
	(345.07,230.42) --
	(345.10,230.43) --
	(345.51,230.61) --
	(345.57,230.63) --
	(346.09,230.84) --
	(346.39,230.97) --
	(347.80,231.60) --
	(348.20,231.77) --
	(348.41,231.82) --
	(348.87,232.01) --
	(348.92,232.02) --
	(349.60,232.28) --
	(351.33,233.12) --
	(351.43,233.18) --
	(351.56,233.22) --
	(351.74,233.28) --
	(351.87,233.36) --
	(353.20,233.97) --
	(353.44,234.08) --
	(353.54,234.11) --
	(354.59,234.56) --
	(357.17,235.70) --
	(357.38,235.79) --
	(357.90,235.98) --
	(357.93,236.00) --
	(358.21,236.08) --
	(358.75,236.31) --
	(359.57,236.60) --
	(359.62,236.62) --
	(360.33,237.04) --
	(360.36,237.05) --
	(360.85,237.22) --
	(361.40,237.45) --
	(362.06,237.74) --
	(362.35,237.87) --
	(362.48,237.93) --
	(362.58,237.98) --
	(363.30,238.32) --
	(363.33,238.34) --
	(363.40,238.37) --
	(364.54,238.92) --
	(365.21,239.23) --
	(366.77,239.98) --
	(366.99,240.10) --
	(368.17,240.64) --
	(368.73,240.88) --
	(372.78,242.85) --
	(373.25,243.04) --
	(374.25,243.52) --
	(374.25,243.53) --
	(374.55,243.71) --
	(374.76,243.79) --
	(375.05,243.91) --
	(375.30,244.01) --
	(375.66,244.18) --
	(375.91,244.31) --
	(376.54,244.60) --
	(376.88,244.76) --
	(377.80,245.25) --
	(378.36,245.52) --
	(378.50,245.59) --
	(378.91,245.79) --
	(379.95,246.29) --
	(381.22,246.90) --
	(381.23,246.90) --
	(381.92,247.24) --
	(381.95,247.25) --
	(382.82,247.67) --
	(383.24,247.86) --
	(383.51,248.01) --
	(384.09,248.23) --
	(384.17,248.29) --
	(384.29,248.34) --
	(384.73,248.57) --
	(386.66,249.57) --
	(386.77,249.61) --
	(390.32,251.41) --
	(394.80,253.54) --
	(396.85,254.58) --
	(404.07,257.98) --
	(404.80,258.32) --
	(404.80,189.64) --
	(404.07,189.50) --
	(396.85,188.00) --
	(394.80,187.59) --
	(390.32,186.83) --
	(386.77,186.36) --
	(386.66,186.36) --
	(384.73,185.98) --
	(384.29,185.90) --
	(384.17,185.88) --
	(384.09,185.87) --
	(383.51,185.77) --
	(383.24,185.74) --
	(382.82,185.65) --
	(381.95,185.48) --
	(381.92,185.47) --
	(381.23,185.34) --
	(381.22,185.34) --
	(379.95,185.10) --
	(378.91,184.79) --
	(378.50,184.70) --
	(378.36,184.66) --
	(377.80,184.56) --
	(376.88,184.37) --
	(376.54,184.29) --
	(375.91,184.19) --
	(375.66,184.15) --
	(375.30,184.06) --
	(375.05,184.03) --
	(374.76,183.96) --
	(374.55,183.91) --
	(374.25,183.85) --
	(374.25,183.85) --
	(373.25,183.61) --
	(372.78,183.51) --
	(368.73,182.59) --
	(368.17,182.50) --
	(366.99,182.24) --
	(366.77,182.20) --
	(365.21,181.90) --
	(364.54,181.78) --
	(363.40,181.57) --
	(363.33,181.55) --
	(363.30,181.54) --
	(362.58,181.32) --
	(362.48,181.30) --
	(362.35,181.27) --
	(362.06,181.18) --
	(361.40,181.08) --
	(360.85,180.92) --
	(360.36,180.79) --
	(360.33,180.78) --
	(359.62,180.60) --
	(359.57,180.58) --
	(358.75,180.41) --
	(358.21,180.32) --
	(357.93,180.28) --
	(357.90,180.27) --
	(357.38,180.13) --
	(357.17,180.06) --
	(354.59,179.48) --
	(353.54,179.19) --
	(353.44,179.17) --
	(353.20,179.12) --
	(351.87,178.72) --
	(351.74,178.69) --
	(351.56,178.65) --
	(351.43,178.61) --
	(351.33,178.58) --
	(349.60,178.18) --
	(348.92,177.96) --
	(348.87,177.94) --
	(348.41,177.84) --
	(348.20,177.79) --
	(347.80,177.68) --
	(346.39,177.30) --
	(346.09,177.23) --
	(345.57,177.09) --
	(345.51,177.07) --
	(345.10,176.98) --
	(345.07,176.97) --
	(345.02,176.96) --
	(344.93,176.96) --
	(344.63,176.89) --
	(344.46,176.83) --
	(344.16,176.75) --
	(343.95,176.68) --
	(343.87,176.67) --
	(343.85,176.67) --
	(343.17,176.56) --
	(343.04,176.53) --
	(342.53,176.35) --
	(342.45,176.33) --
	(342.05,176.22) --
	(342.00,176.20) --
	(341.97,176.19) --
	(341.62,176.12) --
	(341.25,176.04) --
	(340.63,175.87) --
	(340.50,175.83) --
	(339.85,175.63) --
	(339.32,175.51) --
	(337.64,175.18) --
	(337.10,175.07) --
	(336.98,175.03) --
	(336.21,174.86) --
	(335.96,174.80) --
	(335.17,174.60) --
	(333.59,174.17) --
	(333.16,174.04) --
	(333.03,174.01) --
	(332.77,173.95) --
	(332.05,173.81) --
	(331.77,173.77) --
	(330.92,173.51) --
	(330.77,173.45) --
	(330.21,173.33) --
	(329.34,172.99) --
	(329.00,172.89) --
	(328.89,172.87) --
	(328.71,172.79) --
	(328.36,172.72) --
	(327.62,172.54) --
	(327.45,172.47) --
	(327.38,172.46) --
	(327.12,172.39) --
	(326.98,172.34) --
	(326.69,172.26) --
	(326.44,172.25) --
	(326.31,172.21) --
	(326.25,172.19) --
	(324.91,171.82) --
	(324.31,171.68) --
	(324.03,171.60) --
	(323.74,171.54) --
	(323.64,171.51) --
	(323.51,171.46) --
	(323.39,171.42) --
	(323.15,171.32) --
	(322.98,171.25) --
	(322.55,171.10) --
	(322.47,171.07) --
	(322.44,171.06) --
	(322.38,171.05) --
	(322.35,171.04) --
	(322.34,171.04) --
	(321.97,170.92) --
	(321.81,170.88) --
	(321.69,170.84) --
	(321.56,170.80) --
	(321.25,170.71) --
	(321.03,170.65) --
	(320.87,170.61) --
	(320.79,170.59) --
	(320.46,170.48) --
	(320.38,170.46) --
	(320.30,170.44) --
	(320.20,170.41) --
	(320.17,170.41) --
	(320.06,170.37) --
	(319.93,170.32) --
	(319.91,170.31) --
	(319.71,170.26) --
	(319.36,170.17) --
	(319.19,170.11) --
	(319.05,170.05) --
	(318.98,170.02) --
	(318.59,169.91) --
	(318.31,169.82) --
	(318.05,169.74) --
	(317.91,169.70) --
	(317.72,169.65) --
	(317.05,169.42) --
	(316.73,169.34) --
	(316.73,169.34) --
	(316.54,169.30) --
	(316.53,169.30) --
	(316.40,169.27) --
	(316.31,169.24) --
	(316.28,169.23) --
	(316.27,169.22) --
	(316.04,169.15) --
	(315.92,169.11) --
	(315.81,169.07) --
	(315.66,169.02) --
	(315.17,168.87) --
	(314.93,168.80) --
	(314.61,168.70) --
	(314.46,168.64) --
	(314.39,168.61) --
	(314.18,168.54) --
	(313.56,168.33) --
	(313.47,168.29) --
	(312.98,168.12) --
	(312.93,168.10) --
	(311.41,167.66) --
	(311.29,167.62) --
	(310.97,167.57) --
	(310.74,167.49) --
	(310.65,167.48) --
	(310.57,167.46) --
	(310.04,167.25) --
	(309.44,167.10) --
	(309.39,167.09) --
	(309.36,167.07) --
	(309.24,167.03) --
	(309.23,167.03) --
	(309.07,166.97) --
	(308.71,166.82) --
	(308.22,166.66) --
	(308.06,166.62) --
	(307.76,166.54) --
	(307.50,166.45) --
	(307.38,166.41) --
	(306.63,166.19) --
	(306.03,165.97) --
	(305.92,165.94) --
	(304.46,165.44) --
	(304.00,165.34) --
	(303.79,165.28) --
	(303.42,165.16) --
	(303.19,165.08) --
	(302.80,164.97) --
	(302.80,164.97) --
	(302.72,164.95) --
	(302.62,164.92) --
	(302.59,164.91) --
	(302.55,164.90) --
	(302.54,164.90) --
	(302.51,164.89) --
	(302.49,164.88) --
	(302.17,164.79) --
	(302.00,164.74) --
	(301.73,164.67) --
	(301.64,164.65) --
	(301.54,164.62) --
	(301.48,164.61) --
	(301.08,164.50) --
	(301.02,164.48) --
	(300.83,164.41) --
	(299.55,163.99) --
	(298.88,163.70) --
	(298.32,163.50) --
	(298.21,163.45) --
	(298.19,163.45) --
	(298.13,163.43) --
	(298.07,163.41) --
	(297.93,163.34) --
	(297.61,163.23) --
	(297.55,163.21) --
	(297.07,163.03) --
	(296.96,163.00) --
	(296.64,162.91) --
	(296.54,162.88) --
	(295.58,162.53) --
	(295.07,162.36) --
	(294.46,162.18) --
	(294.42,162.18) --
	(293.72,161.95) --
	(293.19,161.77) --
	(292.56,161.53) --
	(292.53,161.52) --
	(291.99,161.37) --
	(291.16,161.08) --
	(291.07,161.06) --
	(290.90,160.99) --
	(290.68,160.94) --
	(290.15,160.77) --
	(290.06,160.74) --
	(289.98,160.71) --
	(289.17,160.47) --
	(288.93,160.43) --
	(287.64,159.98) --
	(287.56,159.96) --
	(287.40,159.91) --
	(287.11,159.83) --
	(286.94,159.78) --
	(286.76,159.73) --
	(286.40,159.61) --
	(285.96,159.45) --
	(285.93,159.44) --
	(285.79,159.39) --
	(285.59,159.31) --
	(285.47,159.26) --
	(284.78,159.01) --
	(284.38,158.87) --
	(284.33,158.84) --
	(283.67,158.61) --
	(283.17,158.41) --
	(282.73,158.24) --
	(282.59,158.19) --
	(282.56,158.18) --
	(282.04,158.00) --
	(281.75,157.91) --
	(281.01,157.60) --
	(280.17,157.31) --
	(279.37,157.02) --
	(278.99,156.86) --
	(278.30,156.58) --
	(277.74,156.39) --
	(277.67,156.37) --
	(276.57,156.01) --
	(276.45,155.98) --
	(275.85,155.79) --
	(275.15,155.49) --
	(274.51,155.32) --
	(274.10,155.21) --
	(273.30,154.88) --
	(273.00,154.78) --
	(272.97,154.77) --
	(271.89,154.37) --
	(271.47,154.17) --
	(271.44,154.15) --
	(271.24,154.06) --
	(270.00,153.51) --
	(269.50,153.33) --
	(269.24,153.19) --
	(269.02,153.08) --
	(267.90,152.64) --
	(267.27,152.43) --
	(267.16,152.39) --
	(265.81,151.89) --
	(263.78,151.02) --
	(263.62,150.94) --
	(262.38,150.41) --
	(262.27,150.37) --
	(260.79,149.85) --
	(260.56,149.75) --
	(259.99,149.51) --
	(259.38,149.23) --
	(259.12,149.14) --
	(258.74,148.98) --
	(258.26,148.80) --
	(257.71,148.57) --
	(257.49,148.48) --
	(256.77,148.18) --
	(256.23,147.96) --
	(255.67,147.76) --
	(254.86,147.39) --
	(254.85,147.39) --
	(253.66,146.86) --
	(252.68,146.41) --
	(251.67,146.06) --
	(250.77,145.70) --
	(249.31,145.09) --
	(248.75,144.87) --
	(247.54,144.32) --
	(246.54,143.90) --
	(244.68,143.00) --
	(241.36,141.60) --
	(240.95,141.37) --
	(240.90,141.35) --
	(240.65,141.25) --
	(239.94,140.97) --
	(238.97,140.59) --
	(238.27,140.29) --
	(236.76,139.59) --
	(235.50,139.08) --
	(235.00,138.89) --
	(233.22,138.10) --
	(228.78,136.06) --
	(228.32,135.82) --
	(224.02,133.94) --
	(222.43,133.19) --
	(222.23,133.10) --
	(220.60,132.33) --
	(215.24,129.80) --
	(213.69,129.08) --
	(213.53,129.01) --
	(212.89,128.71) --
	(212.80,128.67) --
	(212.10,128.34) --
	(211.41,128.11) --
	(209.87,127.43) --
	(209.59,127.31) --
	(205.06,125.25) --
	(205.05,125.25) --
	(195.56,120.80) --
	(188.46,117.41) --
	(185.97,116.14) --
	(175.86,111.07) --
	(174.96,110.63) --
	(171.70,108.95) --
	(171.57,108.88) --
	(158.20,102.11) --
	(138.05, 91.81) --
	(130.04, 87.70) --
	(127.53, 86.38) --
	(123.77, 84.36) --
	(118.51, 81.72) --
	(112.14, 78.35) --
	(108.87, 76.65) --
	(103.21, 73.60) --
	( 94.75, 69.16) --
	( 81.35, 62.32) --
	( 81.21, 62.25) --
	( 76.94, 59.97) --
	( 69.26, 55.82) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 69.26, 34.03) -- ( 69.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 76.94, 34.03) -- ( 76.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 81.21, 34.03) -- ( 81.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 81.35, 34.03) -- ( 81.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 94.75, 34.03) -- ( 94.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (103.21, 34.03) -- (103.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (108.87, 34.03) -- (108.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (112.14, 34.03) -- (112.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (118.51, 34.03) -- (118.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (123.77, 34.03) -- (123.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.53, 34.03) -- (127.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (130.04, 34.03) -- (130.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (138.05, 34.03) -- (138.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (158.20, 34.03) -- (158.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (171.57, 34.03) -- (171.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (171.70, 34.03) -- (171.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (174.96, 34.03) -- (174.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (175.86, 34.03) -- (175.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (185.97, 34.03) -- (185.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (188.46, 34.03) -- (188.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.56, 34.03) -- (195.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.05, 34.03) -- (205.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.06, 34.03) -- (205.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (209.59, 34.03) -- (209.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (209.87, 34.03) -- (209.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (211.41, 34.03) -- (211.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (212.10, 34.03) -- (212.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (212.80, 34.03) -- (212.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (212.89, 34.03) -- (212.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.53, 34.03) -- (213.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.69, 34.03) -- (213.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (215.24, 34.03) -- (215.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.60, 34.03) -- (220.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (222.23, 34.03) -- (222.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (222.43, 34.03) -- (222.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (224.02, 34.03) -- (224.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (228.32, 34.03) -- (228.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (228.78, 34.03) -- (228.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (233.22, 34.03) -- (233.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (235.00, 34.03) -- (235.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (235.50, 34.03) -- (235.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.76, 34.03) -- (236.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (238.27, 34.03) -- (238.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (238.97, 34.03) -- (238.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.94, 34.03) -- (239.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.65, 34.03) -- (240.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.90, 34.03) -- (240.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.95, 34.03) -- (240.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (241.36, 34.03) -- (241.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (244.68, 34.03) -- (244.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (246.54, 34.03) -- (246.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.54, 34.03) -- (247.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.75, 34.03) -- (248.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.31, 34.03) -- (249.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.77, 34.03) -- (250.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.67, 34.03) -- (251.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.68, 34.03) -- (252.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.66, 34.03) -- (253.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (254.85, 34.03) -- (254.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (254.86, 34.03) -- (254.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (255.67, 34.03) -- (255.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.23, 34.03) -- (256.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.77, 34.03) -- (256.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (257.49, 34.03) -- (257.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (257.71, 34.03) -- (257.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.26, 34.03) -- (258.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.74, 34.03) -- (258.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (259.12, 34.03) -- (259.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (259.38, 34.03) -- (259.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (259.99, 34.03) -- (259.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.56, 34.03) -- (260.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.79, 34.03) -- (260.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.27, 34.03) -- (262.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.38, 34.03) -- (262.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (263.62, 34.03) -- (263.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (263.78, 34.03) -- (263.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (265.81, 34.03) -- (265.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.16, 34.03) -- (267.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.27, 34.03) -- (267.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.90, 34.03) -- (267.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.02, 34.03) -- (269.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.24, 34.03) -- (269.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.50, 34.03) -- (269.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (270.00, 34.03) -- (270.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.24, 34.03) -- (271.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.44, 34.03) -- (271.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.47, 34.03) -- (271.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.89, 34.03) -- (271.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.97, 34.03) -- (272.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.00, 34.03) -- (273.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.30, 34.03) -- (273.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.10, 34.03) -- (274.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.51, 34.03) -- (274.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.15, 34.03) -- (275.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.85, 34.03) -- (275.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.45, 34.03) -- (276.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.57, 34.03) -- (276.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.67, 34.03) -- (277.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.74, 34.03) -- (277.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.30, 34.03) -- (278.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.99, 34.03) -- (278.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.37, 34.03) -- (279.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.17, 34.03) -- (280.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.01, 34.03) -- (281.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.75, 34.03) -- (281.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.04, 34.03) -- (282.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.56, 34.03) -- (282.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.59, 34.03) -- (282.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.73, 34.03) -- (282.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.17, 34.03) -- (283.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.67, 34.03) -- (283.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.33, 34.03) -- (284.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.38, 34.03) -- (284.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.78, 34.03) -- (284.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.47, 34.03) -- (285.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.59, 34.03) -- (285.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.79, 34.03) -- (285.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.93, 34.03) -- (285.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.96, 34.03) -- (285.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.40, 34.03) -- (286.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.76, 34.03) -- (286.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.94, 34.03) -- (286.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.11, 34.03) -- (287.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.40, 34.03) -- (287.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.56, 34.03) -- (287.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.64, 34.03) -- (287.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.93, 34.03) -- (288.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.17, 34.03) -- (289.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.98, 34.03) -- (289.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.06, 34.03) -- (290.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.15, 34.03) -- (290.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.68, 34.03) -- (290.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.90, 34.03) -- (290.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.07, 34.03) -- (291.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.16, 34.03) -- (291.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.99, 34.03) -- (291.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.53, 34.03) -- (292.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.56, 34.03) -- (292.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.19, 34.03) -- (293.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.72, 34.03) -- (293.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.42, 34.03) -- (294.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.46, 34.03) -- (294.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.07, 34.03) -- (295.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.58, 34.03) -- (295.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.54, 34.03) -- (296.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.64, 34.03) -- (296.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.96, 34.03) -- (296.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.07, 34.03) -- (297.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.55, 34.03) -- (297.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.61, 34.03) -- (297.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.93, 34.03) -- (297.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.07, 34.03) -- (298.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.13, 34.03) -- (298.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.19, 34.03) -- (298.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.21, 34.03) -- (298.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.32, 34.03) -- (298.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.88, 34.03) -- (298.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.55, 34.03) -- (299.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.83, 34.03) -- (300.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.02, 34.03) -- (301.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.08, 34.03) -- (301.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.48, 34.03) -- (301.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.54, 34.03) -- (301.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.64, 34.03) -- (301.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.73, 34.03) -- (301.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.00, 34.03) -- (302.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.17, 34.03) -- (302.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.49, 34.03) -- (302.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.51, 34.03) -- (302.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.54, 34.03) -- (302.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.55, 34.03) -- (302.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.59, 34.03) -- (302.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.62, 34.03) -- (302.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.72, 34.03) -- (302.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.80, 34.03) -- (302.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.80, 34.03) -- (302.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.19, 34.03) -- (303.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.42, 34.03) -- (303.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.79, 34.03) -- (303.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.00, 34.03) -- (304.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.46, 34.03) -- (304.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.92, 34.03) -- (305.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.03, 34.03) -- (306.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.63, 34.03) -- (306.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.38, 34.03) -- (307.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.50, 34.03) -- (307.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.76, 34.03) -- (307.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.06, 34.03) -- (308.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.22, 34.03) -- (308.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.71, 34.03) -- (308.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.07, 34.03) -- (309.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.23, 34.03) -- (309.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.24, 34.03) -- (309.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.36, 34.03) -- (309.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.39, 34.03) -- (309.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.44, 34.03) -- (309.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.04, 34.03) -- (310.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.57, 34.03) -- (310.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.65, 34.03) -- (310.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.74, 34.03) -- (310.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.97, 34.03) -- (310.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.29, 34.03) -- (311.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.41, 34.03) -- (311.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.93, 34.03) -- (312.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.98, 34.03) -- (312.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.47, 34.03) -- (313.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.56, 34.03) -- (313.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.18, 34.03) -- (314.18, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.39, 34.03) -- (314.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.46, 34.03) -- (314.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.61, 34.03) -- (314.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.93, 34.03) -- (314.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.17, 34.03) -- (315.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.66, 34.03) -- (315.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.81, 34.03) -- (315.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.92, 34.03) -- (315.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.04, 34.03) -- (316.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.27, 34.03) -- (316.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.28, 34.03) -- (316.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.31, 34.03) -- (316.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.40, 34.03) -- (316.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.53, 34.03) -- (316.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.54, 34.03) -- (316.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.73, 34.03) -- (316.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.73, 34.03) -- (316.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.05, 34.03) -- (317.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.72, 34.03) -- (317.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.91, 34.03) -- (317.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.05, 34.03) -- (318.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.31, 34.03) -- (318.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.59, 34.03) -- (318.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.98, 34.03) -- (318.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.05, 34.03) -- (319.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.19, 34.03) -- (319.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.36, 34.03) -- (319.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.71, 34.03) -- (319.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.91, 34.03) -- (319.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.93, 34.03) -- (319.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.06, 34.03) -- (320.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.17, 34.03) -- (320.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.20, 34.03) -- (320.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.30, 34.03) -- (320.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.38, 34.03) -- (320.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.46, 34.03) -- (320.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.79, 34.03) -- (320.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.87, 34.03) -- (320.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.03, 34.03) -- (321.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.25, 34.03) -- (321.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.56, 34.03) -- (321.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.69, 34.03) -- (321.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.81, 34.03) -- (321.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.97, 34.03) -- (321.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.34, 34.03) -- (322.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.35, 34.03) -- (322.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.38, 34.03) -- (322.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.44, 34.03) -- (322.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.47, 34.03) -- (322.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.55, 34.03) -- (322.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.98, 34.03) -- (322.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.15, 34.03) -- (323.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.39, 34.03) -- (323.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.51, 34.03) -- (323.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.64, 34.03) -- (323.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.74, 34.03) -- (323.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.03, 34.03) -- (324.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.31, 34.03) -- (324.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.91, 34.03) -- (324.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.25, 34.03) -- (326.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.31, 34.03) -- (326.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.44, 34.03) -- (326.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.69, 34.03) -- (326.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.98, 34.03) -- (326.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.12, 34.03) -- (327.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.38, 34.03) -- (327.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.45, 34.03) -- (327.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.62, 34.03) -- (327.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.36, 34.03) -- (328.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.71, 34.03) -- (328.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.89, 34.03) -- (328.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.00, 34.03) -- (329.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.34, 34.03) -- (329.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.21, 34.03) -- (330.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.77, 34.03) -- (330.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.92, 34.03) -- (330.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.77, 34.03) -- (331.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (332.05, 34.03) -- (332.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (332.77, 34.03) -- (332.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.03, 34.03) -- (333.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.16, 34.03) -- (333.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.59, 34.03) -- (333.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.17, 34.03) -- (335.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.96, 34.03) -- (335.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.21, 34.03) -- (336.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.98, 34.03) -- (336.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.10, 34.03) -- (337.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.64, 34.03) -- (337.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.32, 34.03) -- (339.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.85, 34.03) -- (339.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.50, 34.03) -- (340.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.63, 34.03) -- (340.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.25, 34.03) -- (341.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.62, 34.03) -- (341.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.97, 34.03) -- (341.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.00, 34.03) -- (342.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.05, 34.03) -- (342.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.45, 34.03) -- (342.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.53, 34.03) -- (342.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.04, 34.03) -- (343.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.17, 34.03) -- (343.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.85, 34.03) -- (343.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.87, 34.03) -- (343.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.95, 34.03) -- (343.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.16, 34.03) -- (344.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.46, 34.03) -- (344.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.63, 34.03) -- (344.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.93, 34.03) -- (344.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.02, 34.03) -- (345.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.07, 34.03) -- (345.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.10, 34.03) -- (345.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.51, 34.03) -- (345.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.57, 34.03) -- (345.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.09, 34.03) -- (346.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.39, 34.03) -- (346.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (347.80, 34.03) -- (347.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.20, 34.03) -- (348.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.41, 34.03) -- (348.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.87, 34.03) -- (348.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.92, 34.03) -- (348.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.60, 34.03) -- (349.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.33, 34.03) -- (351.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.43, 34.03) -- (351.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.56, 34.03) -- (351.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.74, 34.03) -- (351.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.87, 34.03) -- (351.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (353.20, 34.03) -- (353.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (353.44, 34.03) -- (353.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (353.54, 34.03) -- (353.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.59, 34.03) -- (354.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.17, 34.03) -- (357.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.38, 34.03) -- (357.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.90, 34.03) -- (357.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.93, 34.03) -- (357.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.21, 34.03) -- (358.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.75, 34.03) -- (358.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.57, 34.03) -- (359.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.62, 34.03) -- (359.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.33, 34.03) -- (360.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.36, 34.03) -- (360.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.85, 34.03) -- (360.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (361.40, 34.03) -- (361.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.06, 34.03) -- (362.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.35, 34.03) -- (362.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.48, 34.03) -- (362.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.58, 34.03) -- (362.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.30, 34.03) -- (363.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.33, 34.03) -- (363.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.40, 34.03) -- (363.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.54, 34.03) -- (364.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.21, 34.03) -- (365.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.77, 34.03) -- (366.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.99, 34.03) -- (366.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.17, 34.03) -- (368.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.73, 34.03) -- (368.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.78, 34.03) -- (372.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.25, 34.03) -- (373.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.25, 34.03) -- (374.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.25, 34.03) -- (374.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.55, 34.03) -- (374.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.76, 34.03) -- (374.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (375.05, 34.03) -- (375.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (375.30, 34.03) -- (375.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (375.66, 34.03) -- (375.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (375.91, 34.03) -- (375.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.54, 34.03) -- (376.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.88, 34.03) -- (376.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (377.80, 34.03) -- (377.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.36, 34.03) -- (378.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.50, 34.03) -- (378.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.91, 34.03) -- (378.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (379.95, 34.03) -- (379.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.22, 34.03) -- (381.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.23, 34.03) -- (381.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.92, 34.03) -- (381.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.95, 34.03) -- (381.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.82, 34.03) -- (382.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.24, 34.03) -- (383.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.51, 34.03) -- (383.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.09, 34.03) -- (384.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.17, 34.03) -- (384.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.29, 34.03) -- (384.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.73, 34.03) -- (384.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.66, 34.03) -- (386.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.77, 34.03) -- (386.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.32, 34.03) -- (390.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (394.80, 34.03) -- (394.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (396.85, 34.03) -- (396.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.07, 34.03) -- (404.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.80, 34.03) -- (404.80, 41.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.48, 34.03) --
	( 52.48,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37, 80.71) {-0.002};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37,142.88) {0.000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37,205.04) {0.002};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 45.37,267.21) {0.004};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.48, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (100.44, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (148.69, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (196.93, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (245.18, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (293.43, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (341.68, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (389.93, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (237.03,  4.82) {Ln(Min. City Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}