### 2014-10-06 LaTeX/RhodiumTheory.tex
theory section done?

\section{Spatial Distribution of Cities and Violence}
\label{theory}

The extant literature has made important strides in explaining the relationship between economic growth and civil conflict. Yet, none to our knowledge have explored the effects of subnational overlaps in the spatial distribution of cities and conflicts. We argue that this overlap is key to explaining variation in macroeconomic performance during civil conflict.

Our argument on the proximity of conflict to cities should not be confused with arguments focusing on simply the area covered by a conflict. Our argument explicitly differs in the hypothesized mechanism through which conflict affects economic performance. We do not disagree that the spread of a conflict could impact state economic prospects, but we argue that a large conflict area is not a necessary condition for poor economic performance. Conflict area is only one possible proxy for overall destructiveness. However, conflicts with smaller spatial areas can be similarly disruptive if they are centered near urban centers. In fact, we anticipate measures of conflict proximity relative to cities rather than spread to be more appropriate to test the macroeconomic effects of conflict.

% \cite[p. 137]{quigley:1998}, we should use page citations everywhere we reference long quotes or not at all

That cities are central to the development of national economic performance is well supported in the long literature on the importance of urban centers to national economies. \citet{myrdal:sitohang:1957} were among the first to argue that once cities reach a certain size through the process of urbanization, they tend to become self-reinforcing growth centers through a process Myrdal termed as ``cumulative causation''. \citet{chenery:syrquin:1975} provided support for this idea by showing that sharp declines in fertility and substantial increases in growth per capita typically follow cases of urbanization. Jane Jacobs, a progenitor of the move towards focusing on urban centers, went farther than most in her time by arguing that cities are the primary motivators of state economies and should be given primacy over the nation state in economic analysis \citep{jacobs:1969,jacobs:1984}.

In the past three decades, as the pace of urbanization has accelerated, economists have devoted even more energy to investigating the role of city-level economic drivers \citep{lucas:1988, ciccone:hall:1996, begg:1999, henderson:wang:2007}. Most notably, \citeauthor{krugman:1991a} in his seminal 1991 article, ``Increasing Returns and Economic Geography'', convincingly showed the importance of studying economic outcomes within a spatial context. A key finding of Krugman's work into economic geography has been the importance of agglomeration effects, the idea that ``activities tend to cluster where markets are large and markets become larger there where activities cluster'' \citep{krugman:1997}. \citet{henderson:2000} provides empirical evidence for why these agglomeration effects would manifest by showing that spatial clustering promotes economic efficiency in a variety of ways from simply enabling savings in transport costs to more efficiently functioning labor markets.

In addition to agglomeration effects, there is an even simpler story, which can be traced back to \citet{marshall:1920}, that states geographic clustering of firms promotes valuable learning and exchange between actors. \citet{lucas:1988} provides a formal treatment of this argument and shows that the accumulation of human capital generates positive spillovers, where if even one worker acquires a new skill then spatially proximate workers would all become more productive. Empirical estimates of agglomeration effects indicate that a doubling of employment density correspond to a 5\% increase in labor productivity \citep{ciccone:hall:1996,ciccone:2002}. \citet{glaeser:etal:1992} examine this problem empirically and show that cities are important for economic growth precisely because they lead to the knowledge spillovers that are important for innovation. Similar empirical results have been found by many others, which has led to a growing consensus on how cities can serve as important knowledge hubs for a national economy \citep{jaffe:etal:1993, glaeser:1994, firestone:2010}.

% Further \citet{quigley:1998} concludes that ``large cities have been and will continue to be an important source of economic growth.''

The econometric evidence on the importance of cities has crucial implications for our understanding of the spatial structure of subnational economies. Namely, at the subnational level the economy can be thought of as consisting of ``lumps'' of productivity and swaths of areas that contribute little to macroeconomic growth \citep{venables:2005}. This implies that subnational conflicts will have heterogeneous effects on economic growth given their spatial proximity to economically relevant centers such as cities.

\cite{glaeser:shapiro:2002} discuss the historic role of cities in warfare and their current role with respect to terrorism. Cities, they argue, once provided safe haven in the midst of conflict; defense of a concentrated population is easier than defense of a dispersed population. However, as the tactics of warfare have shifted over the centuries, the relationship between violent conflict and cities has become more complex. Their dense populations also tempt belligerents and maximize the impact to cost ratio of a given violent action. Additionally, warfare destroys transportation infrastructure, which can interfere both with commerce and with rebuilding during a conflict. In particular, conflict affects citizens ability and willingness to participate in commerce. When major population centers are threatened by violence, residents will be less likely to engage in economically productive activities. Violence near major population centers not only threatens residents directly, but impedes business by threatening trade between the population center and other cities or rural areas.

We extend this line of research by conceptualizing civil wars not as homogeneous national phenomenon, but as a diverse class of violent conflict with properties that distinguish the effects of one conflict from another. Since urban economies are responsible for a disproportionate share of national economic performance, civil conflicts in or near these engines of commerce should likewise exert a disproportionate influence on state performance.

\subsection{Descriptive Cases}

This hypothesis does not seem unknown to armed actors. The guerilla group Fuerzas Armadas Revolucionarios de Colombia (FARC) appears to have internalized these mechanisms. In 1998 and 1999, the organization moved its violent operations from mostly rural areas of Colombia into major cities and near to the capital (Petras and Brescia 2000). This coincided with economic strain caused by the implementation of an IMF/World Bank structural readjustment program. However, the timing was likely not coincidental. FARC advocates a number of political and economic reforms and chooses targets strategically related to these objectives.

figure \ref{fig:columbiaMap}

\begin{figure}[ht]
	\centering
	\includegraphics[width=.5\textwidth]{colombiaMap-crop}
	\caption{This map illustrates the geographic distribution of all conflicts, according to the PRIO Conflict Site Dataset, and major cities listed in The World Almanac from 1989 to 2008. Countries for which no armed conflicts are recorded are shaded in grey.}
	\label{fig:columbiaMap}
\end{figure}

Forbes magazine, reporting on peace talks between FARC guerillas and the Colombian government in 2012, wrote:

\begin{quote}
	FARC's strategy and [beliefs have] always been to make economic pressure on both, multinational companies and the Colombian government. This has been done by attacking oil and natural gas infrastructure affecting companies such as Pacific Rubiales Energy, Oxy and Ecopetrol. For non-fuel related international companies with subsidiaries in Colombia, such as Goodyear, Nestle, Microsoft, Toyota, among others, FARC’s modus operandi was mainly racketeering, kidnappings and extortion. (Flannery 2012)
\end{quote}

By targeting economic centers and resource infrastructure, FARC can strain Colombia's economy, frighten investors, and bolster support from poor and rural workers sensitive to wealth disparity in the country. \cite{rabasa:chalk:2001} identify a three-pronged strategy pursued by FARC in the 1990s: to consolidate power in coca-growing regions, to conduct military operations in economically valuable areas, and to isolate major cities from the rest of the country by limiting communication and travel between them.

While FARC seems to exploit its ability to target areas of economic importance including cities, other insurgencies tend to be more peripheral. India, for instance, has faced challenges by armed groups in its north-east for half a century. However, this area of India is remote, primarily agrarian, and relatively less populous than other parts of India. Indeed, for much of this period, India has experienced relatively robust economic growth. Mexico, now nearly a decade into a violent and complicated conflict between several organized criminal enterprises and the federal government, has maintained healthy economic performance. For much of this time, the cartel violence generally occurred in rural areas along drug trafficking routes and not within major cities. \cite{beittel:2011} writes in 2011 that ``drug trafficking-related killings remain concentrated in a relatively few cities,'' but notes that the violence was beginning to expand in geographic range by 2010.

\begin{figure}[ht]
	\centering
	\includegraphics[width=.5\textwidth]{indiaMap-crop}
	\caption{This map illustrates the geographic distribution of all conflicts, according to the PRIO Conflict Site Dataset, and major cities listed in The World Almanac from 1989 to 2008. Countries for which no armed conflicts are recorded are shaded in grey.}
	\label{fig:indiaMap}
\end{figure}

We turn now to an empirical test of our proposition. Using a dataset that includes 510 country-years of civil war from 71 different countries, we seek to determine how the geography of conflict impacts state economic outcomes.