### 2014-10-06 LaTeX/mCapSimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-10-06 19:44:45
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.97,105.62) --
	( 60.61,108.24) --
	( 64.85,109.70) --
	( 64.98,109.74) --
	( 78.30,114.31) --
	( 86.71,117.20) --
	( 92.34,119.13) --
	( 95.59,120.24) --
	(101.92,122.42) --
	(110.88,125.49) --
	(154.64,140.51) --
	(158.01,141.66) --
	(171.43,146.26) --
	(178.49,148.69) --
	(187.92,151.92) --
	(192.71,153.57) --
	(195.62,154.56) --
	(195.71,154.59) --
	(196.35,154.81) --
	(198.04,155.39) --
	(199.87,156.02) --
	(204.99,157.78) --
	(205.19,157.85) --
	(209.65,159.38) --
	(211.50,160.01) --
	(215.87,161.51) --
	(217.33,162.01) --
	(217.43,162.05) --
	(217.68,162.13) --
	(218.17,162.30) --
	(220.17,162.99) --
	(220.33,163.04) --
	(220.93,163.25) --
	(221.63,163.49) --
	(222.58,163.81) --
	(223.29,164.06) --
	(223.54,164.14) --
	(223.59,164.16) --
	(224.00,164.30) --
	(229.14,166.07) --
	(231.15,166.76) --
	(231.34,166.82) --
	(232.39,167.18) --
	(233.34,167.51) --
	(234.25,167.82) --
	(235.24,168.16) --
	(236.22,168.49) --
	(237.40,168.90) --
	(237.41,168.90) --
	(239.29,169.55) --
	(239.31,169.55) --
	(240.03,169.80) --
	(240.80,170.06) --
	(241.27,170.23) --
	(241.65,170.36) --
	(242.95,170.80) --
	(244.78,171.43) --
	(248.30,172.64) --
	(249.74,173.13) --
	(250.22,173.30) --
	(250.38,173.35) --
	(250.69,173.46) --
	(251.97,173.90) --
	(252.46,174.07) --
	(253.69,174.49) --
	(253.89,174.55) --
	(254.34,174.71) --
	(255.74,175.19) --
	(256.94,175.60) --
	(258.99,176.31) --
	(260.08,176.68) --
	(260.16,176.71) --
	(261.39,177.13) --
	(262.57,177.53) --
	(263.47,177.84) --
	(264.14,178.07) --
	(264.43,178.17) --
	(264.94,178.35) --
	(265.11,178.40) --
	(265.48,178.53) --
	(265.54,178.55) --
	(266.69,178.95) --
	(266.70,178.95) --
	(267.74,179.31) --
	(267.83,179.34) --
	(267.95,179.38) --
	(268.15,179.45) --
	(268.32,179.51) --
	(268.51,179.57) --
	(268.75,179.66) --
	(269.12,179.78) --
	(269.30,179.84) --
	(269.75,180.00) --
	(269.91,180.05) --
	(271.51,180.60) --
	(272.31,180.88) --
	(272.40,180.91) --
	(272.91,181.08) --
	(273.01,181.11) --
	(273.29,181.21) --
	(273.48,181.28) --
	(274.31,181.56) --
	(274.85,181.75) --
	(274.88,181.76) --
	(275.50,181.97) --
	(276.03,182.15) --
	(276.14,182.19) --
	(276.73,182.39) --
	(276.77,182.41) --
	(277.28,182.58) --
	(277.37,182.61) --
	(278.83,183.11) --
	(278.93,183.15) --
	(278.99,183.17) --
	(279.03,183.18) --
	(279.36,183.30) --
	(279.83,183.46) --
	(279.90,183.48) --
	(280.21,183.59) --
	(280.36,183.64) --
	(280.41,183.66) --
	(280.47,183.68) --
	(280.50,183.68) --
	(280.60,183.72) --
	(280.63,183.73) --
	(281.16,183.91) --
	(281.52,184.03) --
	(281.60,184.06) --
	(283.10,184.58) --
	(283.28,184.64) --
	(283.34,184.66) --
	(283.74,184.80) --
	(283.80,184.82) --
	(284.26,184.97) --
	(284.43,185.03) --
	(284.75,185.14) --
	(284.80,185.16) --
	(284.81,185.16) --
	(284.81,185.16) --
	(284.84,185.18) --
	(284.88,185.19) --
	(285.44,185.38) --
	(285.67,185.46) --
	(286.04,185.59) --
	(286.70,185.81) --
	(286.95,185.90) --
	(287.76,186.18) --
	(288.16,186.31) --
	(288.45,186.41) --
	(288.86,186.55) --
	(289.03,186.61) --
	(289.13,186.65) --
	(289.30,186.70) --
	(289.61,186.81) --
	(289.71,186.85) --
	(289.99,186.94) --
	(290.44,187.10) --
	(291.29,187.39) --
	(291.45,187.44) --
	(291.58,187.49) --
	(292.20,187.70) --
	(292.36,187.75) --
	(292.77,187.89) --
	(292.78,187.90) --
	(292.85,187.92) --
	(292.94,187.95) --
	(293.49,188.14) --
	(293.61,188.18) --
	(293.90,188.28) --
	(293.94,188.30) --
	(295.12,188.70) --
	(295.85,188.95) --
	(296.61,189.21) --
	(297.11,189.38) --
	(297.99,189.68) --
	(298.09,189.72) --
	(298.13,189.74) --
	(298.21,189.76) --
	(298.45,189.84) --
	(298.48,189.85) --
	(298.71,189.93) --
	(298.89,190.00) --
	(299.22,190.11) --
	(300.07,190.40) --
	(300.69,190.61) --
	(301.34,190.83) --
	(301.81,191.00) --
	(301.86,191.01) --
	(302.06,191.08) --
	(302.08,191.09) --
	(302.20,191.13) --
	(302.30,191.16) --
	(302.45,191.22) --
	(302.52,191.24) --
	(302.61,191.27) --
	(302.89,191.37) --
	(302.93,191.38) --
	(303.17,191.46) --
	(304.10,191.78) --
	(304.31,191.85) --
	(304.47,191.91) --
	(304.48,191.91) --
	(304.60,191.95) --
	(304.68,191.98) --
	(305.11,192.13) --
	(305.28,192.19) --
	(305.76,192.35) --
	(305.86,192.39) --
	(306.15,192.49) --
	(307.03,192.79) --
	(307.16,192.83) --
	(307.37,192.90) --
	(308.42,193.27) --
	(308.63,193.34) --
	(308.79,193.39) --
	(309.08,193.49) --
	(309.55,193.65) --
	(309.72,193.71) --
	(310.36,193.93) --
	(310.45,193.96) --
	(310.80,194.08) --
	(310.98,194.14) --
	(311.09,194.18) --
	(311.49,194.32) --
	(311.71,194.39) --
	(311.87,194.45) --
	(312.39,194.63) --
	(312.68,194.73) --
	(312.75,194.75) --
	(314.09,195.21) --
	(314.12,195.22) --
	(314.55,195.37) --
	(314.84,195.47) --
	(315.05,195.54) --
	(315.28,195.62) --
	(315.54,195.71) --
	(315.61,195.73) --
	(315.65,195.75) --
	(316.29,195.97) --
	(316.37,195.99) --
	(316.58,196.06) --
	(316.61,196.07) --
	(316.72,196.11) --
	(317.13,196.25) --
	(318.60,196.76) --
	(319.14,196.94) --
	(320.71,197.48) --
	(321.34,197.70) --
	(321.47,197.74) --
	(322.56,198.12) --
	(322.64,198.14) --
	(322.93,198.24) --
	(323.98,198.60) --
	(324.06,198.63) --
	(324.53,198.79) --
	(324.82,198.89) --
	(325.46,199.11) --
	(325.85,199.24) --
	(325.87,199.25) --
	(325.91,199.27) --
	(326.16,199.35) --
	(326.46,199.45) --
	(327.01,199.64) --
	(327.50,199.81) --
	(327.78,199.91) --
	(328.37,200.11) --
	(328.92,200.30) --
	(330.17,200.73) --
	(330.38,200.80) --
	(330.88,200.97) --
	(333.53,201.88) --
	(333.56,201.89) --
	(333.69,201.94) --
	(333.82,201.98) --
	(333.83,201.98) --
	(335.38,202.51) --
	(335.47,202.55) --
	(337.65,203.29) --
	(338.50,203.58) --
	(338.83,203.70) --
	(338.87,203.71) --
	(339.04,203.77) --
	(339.21,203.83) --
	(339.40,203.89) --
	(339.55,203.94) --
	(339.76,204.02) --
	(339.84,204.04) --
	(340.15,204.15) --
	(340.28,204.20) --
	(340.65,204.32) --
	(341.52,204.62) --
	(342.53,204.97) --
	(343.79,205.40) --
	(343.95,205.45) --
	(345.11,205.85) --
	(345.20,205.89) --
	(345.78,206.08) --
	(346.49,206.33) --
	(346.80,206.43) --
	(346.94,206.48) --
	(347.30,206.60) --
	(348.63,207.06) --
	(348.84,207.13) --
	(349.59,207.39) --
	(350.01,207.53) --
	(350.27,207.62) --
	(350.57,207.73) --
	(351.61,208.08) --
	(352.50,208.39) --
	(354.60,209.11) --
	(354.77,209.17) --
	(355.03,209.26) --
	(355.06,209.27) --
	(355.38,209.38) --
	(355.44,209.40) --
	(355.66,209.47) --
	(355.69,209.48) --
	(356.06,209.61) --
	(356.08,209.62) --
	(356.57,209.78) --
	(356.86,209.88) --
	(356.99,209.93) --
	(357.04,209.94) --
	(357.10,209.97) --
	(357.28,210.03) --
	(357.68,210.17) --
	(357.71,210.18) --
	(358.36,210.40) --
	(358.45,210.43) --
	(358.80,210.55) --
	(359.58,210.82) --
	(360.14,211.01) --
	(360.17,211.02) --
	(360.28,211.06) --
	(360.68,211.20) --
	(362.54,211.83) --
	(362.92,211.96) --
	(362.99,211.99) --
	(362.99,211.99) --
	(363.68,212.22) --
	(363.71,212.23) --
	(364.25,212.42) --
	(364.57,212.53) --
	(364.77,212.60) --
	(365.00,212.68) --
	(365.26,212.76) --
	(365.84,212.96) --
	(365.84,212.96) --
	(365.92,212.99) --
	(366.04,213.03) --
	(366.47,213.18) --
	(367.05,213.38) --
	(368.20,213.78) --
	(368.39,213.84) --
	(368.50,213.88) --
	(370.21,214.46) --
	(371.17,214.79) --
	(371.58,214.94) --
	(372.02,215.09) --
	(373.58,215.62) --
	(376.96,216.78) --
	(378.52,217.32) --
	(383.85,219.15) --
	(384.44,219.35) --
	(384.92,219.51) --
	(385.30,219.64) --
	(385.73,219.79) --
	(385.96,219.87) --
	(386.70,220.12) --
	(386.81,220.16) --
	(387.17,220.28) --
	(387.19,220.29) --
	(387.24,220.31) --
	(387.39,220.36) --
	(387.67,220.46) --
	(388.26,220.66) --
	(389.16,220.97) --
	(389.82,221.19) --
	(390.40,221.39) --
	(390.44,221.41) --
	(390.79,221.53) --
	(391.07,221.62) --
	(395.16,223.02) --
	(396.00,223.31) --
	(396.15,223.37) --
	(403.52,225.89) --
	(404.02,226.07);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 52.97,165.20) --
	( 60.61,166.06) --
	( 64.85,166.50) --
	( 64.98,166.52) --
	( 78.30,168.34) --
	( 86.71,169.62) --
	( 92.34,170.49) --
	( 95.59,171.09) --
	(101.92,171.91) --
	(110.88,173.03) --
	(154.64,179.87) --
	(158.01,180.61) --
	(171.43,182.79) --
	(178.49,184.14) --
	(187.92,186.16) --
	(192.71,186.96) --
	(195.62,187.49) --
	(195.71,187.52) --
	(196.35,187.67) --
	(198.04,187.95) --
	(199.87,188.32) --
	(204.99,189.26) --
	(205.19,189.29) --
	(209.65,190.35) --
	(211.50,190.74) --
	(215.87,191.88) --
	(217.33,192.18) --
	(217.43,192.18) --
	(217.68,192.24) --
	(218.17,192.39) --
	(220.17,192.85) --
	(220.33,192.86) --
	(220.93,193.03) --
	(221.63,193.24) --
	(222.58,193.40) --
	(223.29,193.54) --
	(223.54,193.60) --
	(223.59,193.62) --
	(224.00,193.67) --
	(229.14,194.98) --
	(231.15,195.48) --
	(231.34,195.52) --
	(232.39,195.71) --
	(233.34,195.91) --
	(234.25,196.12) --
	(235.24,196.41) --
	(236.22,196.59) --
	(237.40,196.88) --
	(237.41,196.88) --
	(239.29,197.35) --
	(239.31,197.35) --
	(240.03,197.54) --
	(240.80,197.71) --
	(241.27,197.83) --
	(241.65,197.94) --
	(242.95,198.33) --
	(244.78,198.83) --
	(248.30,199.74) --
	(249.74,200.06) --
	(250.22,200.25) --
	(250.38,200.30) --
	(250.69,200.38) --
	(251.97,200.79) --
	(252.46,200.89) --
	(253.69,201.25) --
	(253.89,201.30) --
	(254.34,201.47) --
	(255.74,201.84) --
	(256.94,202.18) --
	(258.99,202.81) --
	(260.08,203.21) --
	(260.16,203.23) --
	(261.39,203.62) --
	(262.57,204.01) --
	(263.47,204.17) --
	(264.14,204.36) --
	(264.43,204.44) --
	(264.94,204.64) --
	(265.11,204.71) --
	(265.48,204.84) --
	(265.54,204.86) --
	(266.69,205.27) --
	(266.70,205.27) --
	(267.74,205.67) --
	(267.83,205.69) --
	(267.95,205.73) --
	(268.15,205.80) --
	(268.32,205.86) --
	(268.51,205.91) --
	(268.75,206.03) --
	(269.12,206.13) --
	(269.30,206.19) --
	(269.75,206.29) --
	(269.91,206.34) --
	(271.51,206.93) --
	(272.31,207.26) --
	(272.40,207.29) --
	(272.91,207.46) --
	(273.01,207.49) --
	(273.29,207.60) --
	(273.48,207.63) --
	(274.31,207.88) --
	(274.85,208.05) --
	(274.88,208.05) --
	(275.50,208.31) --
	(276.03,208.51) --
	(276.14,208.54) --
	(276.73,208.71) --
	(276.77,208.72) --
	(277.28,208.92) --
	(277.37,208.95) --
	(278.83,209.44) --
	(278.93,209.48) --
	(278.99,209.50) --
	(279.03,209.51) --
	(279.36,209.57) --
	(279.83,209.71) --
	(279.90,209.72) --
	(280.21,209.80) --
	(280.36,209.87) --
	(280.41,209.89) --
	(280.47,209.91) --
	(280.50,209.92) --
	(280.60,209.95) --
	(280.63,209.96) --
	(281.16,210.14) --
	(281.52,210.28) --
	(281.60,210.32) --
	(283.10,210.87) --
	(283.28,210.96) --
	(283.34,210.98) --
	(283.74,211.11) --
	(283.80,211.13) --
	(284.26,211.28) --
	(284.43,211.32) --
	(284.75,211.44) --
	(284.80,211.46) --
	(284.81,211.46) --
	(284.81,211.46) --
	(284.84,211.48) --
	(284.88,211.49) --
	(285.44,211.69) --
	(285.67,211.76) --
	(286.04,211.88) --
	(286.70,212.11) --
	(286.95,212.19) --
	(287.76,212.60) --
	(288.16,212.75) --
	(288.45,212.88) --
	(288.86,212.98) --
	(289.03,213.01) --
	(289.13,213.03) --
	(289.30,213.12) --
	(289.61,213.30) --
	(289.71,213.32) --
	(289.99,213.40) --
	(290.44,213.53) --
	(291.29,213.87) --
	(291.45,213.92) --
	(291.58,213.96) --
	(292.20,214.19) --
	(292.36,214.26) --
	(292.77,214.38) --
	(292.78,214.39) --
	(292.85,214.43) --
	(292.94,214.45) --
	(293.49,214.66) --
	(293.61,214.72) --
	(293.90,214.85) --
	(293.94,214.86) --
	(295.12,215.33) --
	(295.85,215.63) --
	(296.61,215.90) --
	(297.11,216.10) --
	(297.99,216.47) --
	(298.09,216.52) --
	(298.13,216.53) --
	(298.21,216.57) --
	(298.45,216.67) --
	(298.48,216.68) --
	(298.71,216.78) --
	(298.89,216.86) --
	(299.22,217.00) --
	(300.07,217.37) --
	(300.69,217.67) --
	(301.34,217.95) --
	(301.81,218.15) --
	(301.86,218.17) --
	(302.06,218.25) --
	(302.08,218.25) --
	(302.20,218.30) --
	(302.30,218.33) --
	(302.45,218.39) --
	(302.52,218.41) --
	(302.61,218.44) --
	(302.89,218.54) --
	(302.93,218.56) --
	(303.17,218.62) --
	(304.10,218.98) --
	(304.31,219.07) --
	(304.47,219.12) --
	(304.48,219.12) --
	(304.60,219.15) --
	(304.68,219.18) --
	(305.11,219.34) --
	(305.28,219.40) --
	(305.76,219.62) --
	(305.86,219.66) --
	(306.15,219.79) --
	(307.03,220.15) --
	(307.16,220.21) --
	(307.37,220.28) --
	(308.42,220.71) --
	(308.63,220.77) --
	(308.79,220.81) --
	(309.08,220.91) --
	(309.55,221.07) --
	(309.72,221.11) --
	(310.36,221.34) --
	(310.45,221.38) --
	(310.80,221.53) --
	(310.98,221.58) --
	(311.09,221.61) --
	(311.49,221.78) --
	(311.71,221.89) --
	(311.87,221.97) --
	(312.39,222.22) --
	(312.68,222.34) --
	(312.75,222.36) --
	(314.09,222.97) --
	(314.12,222.98) --
	(314.55,223.10) --
	(314.84,223.21) --
	(315.05,223.29) --
	(315.28,223.38) --
	(315.54,223.48) --
	(315.61,223.51) --
	(315.65,223.53) --
	(316.29,223.77) --
	(316.37,223.80) --
	(316.58,223.88) --
	(316.61,223.90) --
	(316.72,223.94) --
	(317.13,224.10) --
	(318.60,224.77) --
	(319.14,224.99) --
	(320.71,225.74) --
	(321.34,226.00) --
	(321.47,226.05) --
	(322.56,226.60) --
	(322.64,226.65) --
	(322.93,226.81) --
	(323.98,227.18) --
	(324.06,227.23) --
	(324.53,227.46) --
	(324.82,227.58) --
	(325.46,227.82) --
	(325.85,227.92) --
	(325.87,227.93) --
	(325.91,227.95) --
	(326.16,228.06) --
	(326.46,228.20) --
	(327.01,228.46) --
	(327.50,228.65) --
	(327.78,228.76) --
	(328.37,229.01) --
	(328.92,229.25) --
	(330.17,229.78) --
	(330.38,229.89) --
	(330.88,230.12) --
	(333.53,231.30) --
	(333.56,231.31) --
	(333.69,231.37) --
	(333.82,231.43) --
	(333.83,231.43) --
	(335.38,232.11) --
	(335.47,232.16) --
	(337.65,233.11) --
	(338.50,233.49) --
	(338.83,233.64) --
	(338.87,233.66) --
	(339.04,233.73) --
	(339.21,233.80) --
	(339.40,233.89) --
	(339.55,233.95) --
	(339.76,234.05) --
	(339.84,234.09) --
	(340.15,234.23) --
	(340.28,234.29) --
	(340.65,234.46) --
	(341.52,234.87) --
	(342.53,235.38) --
	(343.79,235.96) --
	(343.95,236.03) --
	(345.11,236.65) --
	(345.20,236.70) --
	(345.78,236.95) --
	(346.49,237.30) --
	(346.80,237.45) --
	(346.94,237.52) --
	(347.30,237.70) --
	(348.63,238.33) --
	(348.84,238.43) --
	(349.59,238.82) --
	(350.01,239.02) --
	(350.27,239.13) --
	(350.57,239.26) --
	(351.61,239.72) --
	(352.50,240.14) --
	(354.60,241.18) --
	(354.77,241.25) --
	(355.03,241.37) --
	(355.06,241.39) --
	(355.38,241.55) --
	(355.44,241.59) --
	(355.66,241.70) --
	(355.69,241.71) --
	(356.06,241.93) --
	(356.08,241.93) --
	(356.57,242.19) --
	(356.86,242.34) --
	(356.99,242.41) --
	(357.04,242.44) --
	(357.10,242.47) --
	(357.28,242.56) --
	(357.68,242.77) --
	(357.71,242.78) --
	(358.36,243.06) --
	(358.45,243.11) --
	(358.80,243.27) --
	(359.58,243.64) --
	(360.14,243.89) --
	(360.17,243.90) --
	(360.28,243.94) --
	(360.68,244.14) --
	(362.54,245.11) --
	(362.92,245.32) --
	(362.99,245.36) --
	(362.99,245.36) --
	(363.68,245.73) --
	(363.71,245.75) --
	(364.25,246.04) --
	(364.57,246.19) --
	(364.77,246.29) --
	(365.00,246.40) --
	(365.26,246.54) --
	(365.84,246.88) --
	(365.84,246.88) --
	(365.92,246.92) --
	(366.04,246.98) --
	(366.47,247.19) --
	(367.05,247.46) --
	(368.20,247.96) --
	(368.39,248.07) --
	(368.50,248.14) --
	(370.21,248.93) --
	(371.17,249.41) --
	(371.58,249.67) --
	(372.02,249.86) --
	(373.58,250.61) --
	(376.96,252.31) --
	(378.52,252.97) --
	(383.85,255.72) --
	(384.44,256.00) --
	(384.92,256.23) --
	(385.30,256.42) --
	(385.73,256.63) --
	(385.96,256.74) --
	(386.70,257.14) --
	(386.81,257.19) --
	(387.17,257.37) --
	(387.19,257.37) --
	(387.24,257.39) --
	(387.39,257.47) --
	(387.67,257.63) --
	(388.26,257.95) --
	(389.16,258.36) --
	(389.82,258.71) --
	(390.40,258.98) --
	(390.44,259.01) --
	(390.79,259.17) --
	(391.07,259.30) --
	(395.16,261.28) --
	(396.00,261.69) --
	(396.15,261.76) --
	(403.52,265.67) --
	(404.02,265.99) --
	(404.02,186.16) --
	(403.52,186.07) --
	(396.15,184.62) --
	(396.00,184.56) --
	(395.16,184.43) --
	(391.07,183.78) --
	(390.79,183.76) --
	(390.44,183.72) --
	(390.40,183.71) --
	(389.82,183.64) --
	(389.16,183.56) --
	(388.26,183.43) --
	(387.67,183.32) --
	(387.39,183.25) --
	(387.24,183.21) --
	(387.19,183.20) --
	(387.17,183.20) --
	(386.81,183.11) --
	(386.70,183.09) --
	(385.96,182.98) --
	(385.73,182.95) --
	(385.30,182.87) --
	(384.92,182.82) --
	(384.44,182.79) --
	(383.85,182.69) --
	(378.52,181.63) --
	(376.96,181.35) --
	(373.58,180.87) --
	(372.02,180.66) --
	(371.58,180.64) --
	(371.17,180.60) --
	(370.21,180.44) --
	(368.50,180.16) --
	(368.39,180.14) --
	(368.20,180.08) --
	(367.05,179.83) --
	(366.47,179.68) --
	(366.04,179.56) --
	(365.92,179.53) --
	(365.84,179.51) --
	(365.84,179.51) --
	(365.26,179.35) --
	(365.00,179.28) --
	(364.77,179.25) --
	(364.57,179.24) --
	(364.25,179.17) --
	(363.71,179.04) --
	(363.68,179.03) --
	(362.99,178.85) --
	(362.99,178.85) --
	(362.92,178.84) --
	(362.54,178.74) --
	(360.68,178.40) --
	(360.28,178.30) --
	(360.17,178.31) --
	(360.14,178.31) --
	(359.58,178.27) --
	(358.80,178.08) --
	(358.45,178.01) --
	(358.36,177.98) --
	(357.71,177.80) --
	(357.68,177.80) --
	(357.28,177.74) --
	(357.10,177.72) --
	(357.04,177.72) --
	(356.99,177.71) --
	(356.86,177.70) --
	(356.57,177.64) --
	(356.08,177.49) --
	(356.06,177.49) --
	(355.69,177.37) --
	(355.66,177.36) --
	(355.44,177.32) --
	(355.38,177.31) --
	(355.06,177.26) --
	(355.03,177.25) --
	(354.77,177.22) --
	(354.60,177.19) --
	(352.50,176.92) --
	(351.61,176.70) --
	(350.57,176.49) --
	(350.27,176.47) --
	(350.01,176.37) --
	(349.59,176.23) --
	(348.84,176.03) --
	(348.63,175.98) --
	(347.30,175.74) --
	(346.94,175.67) --
	(346.80,175.64) --
	(346.49,175.53) --
	(345.78,175.33) --
	(345.20,175.23) --
	(345.11,175.20) --
	(343.95,174.89) --
	(343.79,174.84) --
	(342.53,174.62) --
	(341.52,174.45) --
	(340.65,174.27) --
	(340.28,174.16) --
	(340.15,174.10) --
	(339.84,174.03) --
	(339.76,174.02) --
	(339.55,173.99) --
	(339.40,173.93) --
	(339.21,173.87) --
	(339.04,173.82) --
	(338.87,173.77) --
	(338.83,173.76) --
	(338.50,173.69) --
	(337.65,173.44) --
	(335.47,172.96) --
	(335.38,172.95) --
	(333.83,172.52) --
	(333.82,172.52) --
	(333.69,172.48) --
	(333.56,172.45) --
	(333.53,172.44) --
	(330.88,171.81) --
	(330.38,171.72) --
	(330.17,171.68) --
	(328.92,171.37) --
	(328.37,171.15) --
	(327.78,170.98) --
	(327.50,170.92) --
	(327.01,170.77) --
	(326.46,170.62) --
	(326.16,170.57) --
	(325.91,170.51) --
	(325.87,170.51) --
	(325.85,170.51) --
	(325.46,170.40) --
	(324.82,170.21) --
	(324.53,170.21) --
	(324.06,170.15) --
	(323.98,170.13) --
	(322.93,169.85) --
	(322.64,169.82) --
	(322.56,169.82) --
	(321.47,169.46) --
	(321.34,169.40) --
	(320.71,169.24) --
	(319.14,168.74) --
	(318.60,168.64) --
	(317.13,168.18) --
	(316.72,168.05) --
	(316.61,168.01) --
	(316.58,168.00) --
	(316.37,167.93) --
	(316.29,167.90) --
	(315.65,167.68) --
	(315.61,167.67) --
	(315.54,167.65) --
	(315.28,167.58) --
	(315.05,167.51) --
	(314.84,167.42) --
	(314.55,167.33) --
	(314.12,167.22) --
	(314.09,167.21) --
	(312.75,166.87) --
	(312.68,166.86) --
	(312.39,166.79) --
	(311.87,166.57) --
	(311.71,166.51) --
	(311.49,166.45) --
	(311.09,166.35) --
	(310.98,166.33) --
	(310.80,166.29) --
	(310.45,166.23) --
	(310.36,166.21) --
	(309.72,166.00) --
	(309.55,165.96) --
	(309.08,165.82) --
	(308.79,165.72) --
	(308.63,165.67) --
	(308.42,165.62) --
	(307.37,165.33) --
	(307.16,165.25) --
	(307.03,165.22) --
	(306.15,165.04) --
	(305.86,164.98) --
	(305.76,164.96) --
	(305.28,164.84) --
	(305.11,164.79) --
	(304.68,164.71) --
	(304.60,164.69) --
	(304.48,164.65) --
	(304.47,164.64) --
	(304.31,164.58) --
	(304.10,164.53) --
	(303.17,164.22) --
	(302.93,164.15) --
	(302.89,164.14) --
	(302.61,164.07) --
	(302.52,164.04) --
	(302.45,164.03) --
	(302.30,164.00) --
	(302.20,163.98) --
	(302.08,163.95) --
	(302.06,163.95) --
	(301.86,163.90) --
	(301.81,163.89) --
	(301.34,163.78) --
	(300.69,163.54) --
	(300.07,163.35) --
	(299.22,163.00) --
	(298.89,162.91) --
	(298.71,162.85) --
	(298.48,162.77) --
	(298.45,162.76) --
	(298.21,162.69) --
	(298.13,162.66) --
	(298.09,162.65) --
	(297.99,162.62) --
	(297.11,162.35) --
	(296.61,162.22) --
	(295.85,161.97) --
	(295.12,161.73) --
	(293.94,161.38) --
	(293.90,161.37) --
	(293.61,161.27) --
	(293.49,161.22) --
	(292.94,161.03) --
	(292.85,161.01) --
	(292.78,160.99) --
	(292.77,160.98) --
	(292.36,160.87) --
	(292.20,160.82) --
	(291.58,160.61) --
	(291.45,160.57) --
	(291.29,160.52) --
	(290.44,160.26) --
	(289.99,160.13) --
	(289.71,160.04) --
	(289.61,160.00) --
	(289.30,159.87) --
	(289.13,159.80) --
	(289.03,159.75) --
	(288.86,159.69) --
	(288.45,159.56) --
	(288.16,159.43) --
	(287.76,159.26) --
	(286.95,158.94) --
	(286.70,158.84) --
	(286.04,158.62) --
	(285.67,158.49) --
	(285.44,158.44) --
	(284.88,158.24) --
	(284.84,158.23) --
	(284.81,158.22) --
	(284.81,158.22) --
	(284.80,158.21) --
	(284.75,158.20) --
	(284.43,158.12) --
	(284.26,158.06) --
	(283.80,157.92) --
	(283.74,157.91) --
	(283.34,157.81) --
	(283.28,157.79) --
	(283.10,157.72) --
	(281.60,157.19) --
	(281.52,157.15) --
	(281.16,157.04) --
	(280.63,156.85) --
	(280.60,156.83) --
	(280.50,156.79) --
	(280.47,156.78) --
	(280.41,156.75) --
	(280.36,156.73) --
	(280.21,156.68) --
	(279.90,156.59) --
	(279.83,156.58) --
	(279.36,156.45) --
	(279.03,156.32) --
	(278.99,156.30) --
	(278.93,156.27) --
	(278.83,156.20) --
	(277.37,155.69) --
	(277.28,155.66) --
	(276.77,155.47) --
	(276.73,155.45) --
	(276.14,155.25) --
	(276.03,155.23) --
	(275.50,155.09) --
	(274.88,154.85) --
	(274.85,154.84) --
	(274.31,154.63) --
	(273.48,154.33) --
	(273.29,154.26) --
	(273.01,154.14) --
	(272.91,154.08) --
	(272.40,153.94) --
	(272.31,153.91) --
	(271.51,153.68) --
	(269.91,153.18) --
	(269.75,153.12) --
	(269.30,153.01) --
	(269.12,152.95) --
	(268.75,152.83) --
	(268.51,152.72) --
	(268.32,152.61) --
	(268.15,152.54) --
	(267.95,152.45) --
	(267.83,152.40) --
	(267.74,152.36) --
	(266.70,151.94) --
	(266.69,151.93) --
	(265.54,151.50) --
	(265.48,151.48) --
	(265.11,151.39) --
	(264.94,151.31) --
	(264.43,151.11) --
	(264.14,151.03) --
	(263.47,150.81) --
	(262.57,150.47) --
	(261.39,150.00) --
	(260.16,149.54) --
	(260.08,149.51) --
	(258.99,149.05) --
	(256.94,148.31) --
	(255.74,147.96) --
	(254.34,147.35) --
	(253.89,147.15) --
	(253.69,147.06) --
	(252.46,146.62) --
	(251.97,146.47) --
	(250.69,145.91) --
	(250.38,145.79) --
	(250.22,145.74) --
	(249.74,145.63) --
	(248.30,145.02) --
	(244.78,143.52) --
	(242.95,142.71) --
	(241.65,142.20) --
	(241.27,142.05) --
	(240.80,141.82) --
	(240.03,141.56) --
	(239.31,141.28) --
	(239.29,141.27) --
	(237.41,140.47) --
	(237.40,140.46) --
	(236.22,140.02) --
	(235.24,139.60) --
	(234.25,139.14) --
	(233.34,138.79) --
	(232.39,138.25) --
	(231.34,137.82) --
	(231.15,137.77) --
	(229.14,136.99) --
	(224.00,134.73) --
	(223.59,134.56) --
	(223.54,134.54) --
	(223.29,134.42) --
	(222.58,134.07) --
	(221.63,133.64) --
	(220.93,133.31) --
	(220.33,132.98) --
	(220.17,132.91) --
	(218.17,132.08) --
	(217.68,131.82) --
	(217.43,131.70) --
	(217.33,131.66) --
	(215.87,130.99) --
	(211.50,128.79) --
	(209.65,127.84) --
	(205.19,125.79) --
	(204.99,125.71) --
	(199.87,123.16) --
	(198.04,122.29) --
	(196.35,121.40) --
	(195.71,121.10) --
	(195.62,121.07) --
	(192.71,119.57) --
	(187.92,117.14) --
	(178.49,112.49) --
	(171.43,108.90) --
	(158.01,101.95) --
	(154.64,100.20) --
	(110.88, 76.90) --
	(101.92, 72.02) --
	( 95.59, 68.42) --
	( 92.34, 66.68) --
	( 86.71, 63.70) --
	( 78.30, 59.08) --
	( 64.98, 51.79) --
	( 64.85, 51.71) --
	( 60.61, 49.49) --
	( 52.97, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 52.97,155.78) --
	( 60.61,157.02) --
	( 64.85,157.72) --
	( 64.98,157.76) --
	( 78.30,159.89) --
	( 86.71,161.16) --
	( 92.34,162.02) --
	( 95.59,162.58) --
	(101.92,163.75) --
	(110.88,165.37) --
	(154.64,173.25) --
	(158.01,173.83) --
	(171.43,176.44) --
	(178.49,177.89) --
	(187.92,179.93) --
	(192.71,180.94) --
	(195.62,181.57) --
	(195.71,181.60) --
	(196.35,181.78) --
	(198.04,182.20) --
	(199.87,182.59) --
	(204.99,183.69) --
	(205.19,183.74) --
	(209.65,184.88) --
	(211.50,185.32) --
	(215.87,186.31) --
	(217.33,186.70) --
	(217.43,186.73) --
	(217.68,186.79) --
	(218.17,186.95) --
	(220.17,187.43) --
	(220.33,187.47) --
	(220.93,187.59) --
	(221.63,187.78) --
	(222.58,188.03) --
	(223.29,188.16) --
	(223.54,188.25) --
	(223.59,188.27) --
	(224.00,188.40) --
	(229.14,189.71) --
	(231.15,190.26) --
	(231.34,190.31) --
	(232.39,190.62) --
	(233.34,190.89) --
	(234.25,191.19) --
	(235.24,191.47) --
	(236.22,191.82) --
	(237.40,192.18) --
	(237.41,192.18) --
	(239.29,192.72) --
	(239.31,192.73) --
	(240.03,192.94) --
	(240.80,193.15) --
	(241.27,193.33) --
	(241.65,193.46) --
	(242.95,193.75) --
	(244.78,194.25) --
	(248.30,195.33) --
	(249.74,195.76) --
	(250.22,195.94) --
	(250.38,196.01) --
	(250.69,196.16) --
	(251.97,196.55) --
	(252.46,196.68) --
	(253.69,197.01) --
	(253.89,197.07) --
	(254.34,197.22) --
	(255.74,197.69) --
	(256.94,198.07) --
	(258.99,198.62) --
	(260.08,198.96) --
	(260.16,198.98) --
	(261.39,199.44) --
	(262.57,199.78) --
	(263.47,200.03) --
	(264.14,200.24) --
	(264.43,200.36) --
	(264.94,200.54) --
	(265.11,200.57) --
	(265.48,200.70) --
	(265.54,200.72) --
	(266.69,201.09) --
	(266.70,201.09) --
	(267.74,201.45) --
	(267.83,201.46) --
	(267.95,201.52) --
	(268.15,201.60) --
	(268.32,201.64) --
	(268.51,201.73) --
	(268.75,201.81) --
	(269.12,201.95) --
	(269.30,202.00) --
	(269.75,202.18) --
	(269.91,202.21) --
	(271.51,202.79) --
	(272.31,203.06) --
	(272.40,203.09) --
	(272.91,203.23) --
	(273.01,203.25) --
	(273.29,203.34) --
	(273.48,203.41) --
	(274.31,203.68) --
	(274.85,203.89) --
	(274.88,203.90) --
	(275.50,204.08) --
	(276.03,204.23) --
	(276.14,204.27) --
	(276.73,204.46) --
	(276.77,204.47) --
	(277.28,204.64) --
	(277.37,204.66) --
	(278.83,205.17) --
	(278.93,205.20) --
	(278.99,205.23) --
	(279.03,205.25) --
	(279.36,205.31) --
	(279.83,205.47) --
	(279.90,205.50) --
	(280.21,205.61) --
	(280.36,205.65) --
	(280.41,205.68) --
	(280.47,205.71) --
	(280.50,205.71) --
	(280.60,205.74) --
	(280.63,205.75) --
	(281.16,205.91) --
	(281.52,206.01) --
	(281.60,206.03) --
	(283.10,206.58) --
	(283.28,206.66) --
	(283.34,206.68) --
	(283.74,206.84) --
	(283.80,206.86) --
	(284.26,207.01) --
	(284.43,207.07) --
	(284.75,207.16) --
	(284.80,207.18) --
	(284.81,207.18) --
	(284.81,207.18) --
	(284.84,207.19) --
	(284.88,207.20) --
	(285.44,207.45) --
	(285.67,207.54) --
	(286.04,207.68) --
	(286.70,207.89) --
	(286.95,207.97) --
	(287.76,208.27) --
	(288.16,208.43) --
	(288.45,208.57) --
	(288.86,208.73) --
	(289.03,208.78) --
	(289.13,208.82) --
	(289.30,208.87) --
	(289.61,208.98) --
	(289.71,209.02) --
	(289.99,209.12) --
	(290.44,209.29) --
	(291.29,209.59) --
	(291.45,209.64) --
	(291.58,209.68) --
	(292.20,209.90) --
	(292.36,209.95) --
	(292.77,210.11) --
	(292.78,210.12) --
	(292.85,210.14) --
	(292.94,210.16) --
	(293.49,210.35) --
	(293.61,210.39) --
	(293.90,210.51) --
	(293.94,210.53) --
	(295.12,210.90) --
	(295.85,211.14) --
	(296.61,211.47) --
	(297.11,211.69) --
	(297.99,212.07) --
	(298.09,212.10) --
	(298.13,212.11) --
	(298.21,212.12) --
	(298.45,212.21) --
	(298.48,212.23) --
	(298.71,212.30) --
	(298.89,212.38) --
	(299.22,212.51) --
	(300.07,212.87) --
	(300.69,213.09) --
	(301.34,213.35) --
	(301.81,213.52) --
	(301.86,213.54) --
	(302.06,213.61) --
	(302.08,213.62) --
	(302.20,213.66) --
	(302.30,213.71) --
	(302.45,213.79) --
	(302.52,213.81) --
	(302.61,213.85) --
	(302.89,213.96) --
	(302.93,213.97) --
	(303.17,214.04) --
	(304.10,214.40) --
	(304.31,214.48) --
	(304.47,214.53) --
	(304.48,214.53) --
	(304.60,214.58) --
	(304.68,214.61) --
	(305.11,214.78) --
	(305.28,214.85) --
	(305.76,215.07) --
	(305.86,215.12) --
	(306.15,215.24) --
	(307.03,215.53) --
	(307.16,215.60) --
	(307.37,215.70) --
	(308.42,216.07) --
	(308.63,216.18) --
	(308.79,216.26) --
	(309.08,216.38) --
	(309.55,216.59) --
	(309.72,216.66) --
	(310.36,216.92) --
	(310.45,216.96) --
	(310.80,217.15) --
	(310.98,217.23) --
	(311.09,217.27) --
	(311.49,217.42) --
	(311.71,217.50) --
	(311.87,217.56) --
	(312.39,217.84) --
	(312.68,217.99) --
	(312.75,218.02) --
	(314.09,218.56) --
	(314.12,218.57) --
	(314.55,218.71) --
	(314.84,218.84) --
	(315.05,218.90) --
	(315.28,218.98) --
	(315.54,219.09) --
	(315.61,219.12) --
	(315.65,219.13) --
	(316.29,219.37) --
	(316.37,219.39) --
	(316.58,219.47) --
	(316.61,219.48) --
	(316.72,219.52) --
	(317.13,219.68) --
	(318.60,220.33) --
	(319.14,220.56) --
	(320.71,221.16) --
	(321.34,221.42) --
	(321.47,221.48) --
	(322.56,221.95) --
	(322.64,221.98) --
	(322.93,222.10) --
	(323.98,222.52) --
	(324.06,222.56) --
	(324.53,222.78) --
	(324.82,222.91) --
	(325.46,223.17) --
	(325.85,223.33) --
	(325.87,223.34) --
	(325.91,223.34) --
	(326.16,223.46) --
	(326.46,223.57) --
	(327.01,223.82) --
	(327.50,224.06) --
	(327.78,224.16) --
	(328.37,224.39) --
	(328.92,224.61) --
	(330.17,225.15) --
	(330.38,225.24) --
	(330.88,225.42) --
	(333.53,226.62) --
	(333.56,226.63) --
	(333.69,226.70) --
	(333.82,226.78) --
	(333.83,226.78) --
	(335.38,227.53) --
	(335.47,227.58) --
	(337.65,228.52) --
	(338.50,228.87) --
	(338.83,228.96) --
	(338.87,228.99) --
	(339.04,229.08) --
	(339.21,229.15) --
	(339.40,229.23) --
	(339.55,229.29) --
	(339.76,229.37) --
	(339.84,229.40) --
	(340.15,229.53) --
	(340.28,229.58) --
	(340.65,229.71) --
	(341.52,230.07) --
	(342.53,230.56) --
	(343.79,231.08) --
	(343.95,231.15) --
	(345.11,231.69) --
	(345.20,231.73) --
	(345.78,231.99) --
	(346.49,232.31) --
	(346.80,232.47) --
	(346.94,232.52) --
	(347.30,232.67) --
	(348.63,233.25) --
	(348.84,233.35) --
	(349.59,233.76) --
	(350.01,233.97) --
	(350.27,234.09) --
	(350.57,234.23) --
	(351.61,234.73) --
	(352.50,235.16) --
	(354.60,236.10) --
	(354.77,236.18) --
	(355.03,236.30) --
	(355.06,236.31) --
	(355.38,236.46) --
	(355.44,236.49) --
	(355.66,236.61) --
	(355.69,236.61) --
	(356.06,236.78) --
	(356.08,236.79) --
	(356.57,237.02) --
	(356.86,237.14) --
	(356.99,237.20) --
	(357.04,237.22) --
	(357.10,237.25) --
	(357.28,237.33) --
	(357.68,237.52) --
	(357.71,237.53) --
	(358.36,237.85) --
	(358.45,237.89) --
	(358.80,238.04) --
	(359.58,238.38) --
	(360.14,238.61) --
	(360.17,238.62) --
	(360.28,238.68) --
	(360.68,238.85) --
	(362.54,239.69) --
	(362.92,239.87) --
	(362.99,239.91) --
	(362.99,239.91) --
	(363.68,240.21) --
	(363.71,240.22) --
	(364.25,240.52) --
	(364.57,240.68) --
	(364.77,240.77) --
	(365.00,240.87) --
	(365.26,241.04) --
	(365.84,241.33) --
	(365.84,241.33) --
	(365.92,241.37) --
	(366.04,241.42) --
	(366.47,241.62) --
	(367.05,241.89) --
	(368.20,242.45) --
	(368.39,242.54) --
	(368.50,242.59) --
	(370.21,243.39) --
	(371.17,243.81) --
	(371.58,244.00) --
	(372.02,244.21) --
	(373.58,244.87) --
	(376.96,246.54) --
	(378.52,247.26) --
	(383.85,249.83) --
	(384.44,250.15) --
	(384.92,250.40) --
	(385.30,250.56) --
	(385.73,250.74) --
	(385.96,250.84) --
	(386.70,251.22) --
	(386.81,251.27) --
	(387.17,251.44) --
	(387.19,251.44) --
	(387.24,251.46) --
	(387.39,251.54) --
	(387.67,251.67) --
	(388.26,251.97) --
	(389.16,252.36) --
	(389.82,252.64) --
	(390.40,252.95) --
	(390.44,252.98) --
	(390.79,253.21) --
	(391.07,253.39) --
	(395.16,255.52) --
	(396.00,255.92) --
	(396.15,255.99) --
	(403.52,259.54) --
	(404.02,259.83) --
	(404.02,193.21) --
	(403.52,193.10) --
	(396.15,191.59) --
	(396.00,191.58) --
	(395.16,191.49) --
	(391.07,190.75) --
	(390.79,190.69) --
	(390.44,190.64) --
	(390.40,190.63) --
	(389.82,190.53) --
	(389.16,190.43) --
	(388.26,190.20) --
	(387.67,190.10) --
	(387.39,190.05) --
	(387.24,190.01) --
	(387.19,190.01) --
	(387.17,190.00) --
	(386.81,189.94) --
	(386.70,189.92) --
	(385.96,189.70) --
	(385.73,189.67) --
	(385.30,189.58) --
	(384.92,189.52) --
	(384.44,189.44) --
	(383.85,189.36) --
	(378.52,187.97) --
	(376.96,187.68) --
	(373.58,186.94) --
	(372.02,186.66) --
	(371.58,186.58) --
	(371.17,186.48) --
	(370.21,186.18) --
	(368.50,185.79) --
	(368.39,185.78) --
	(368.20,185.73) --
	(367.05,185.54) --
	(366.47,185.42) --
	(366.04,185.30) --
	(365.92,185.27) --
	(365.84,185.25) --
	(365.84,185.25) --
	(365.26,185.12) --
	(365.00,185.06) --
	(364.77,185.00) --
	(364.57,184.94) --
	(364.25,184.87) --
	(363.71,184.72) --
	(363.68,184.71) --
	(362.99,184.56) --
	(362.99,184.56) --
	(362.92,184.54) --
	(362.54,184.46) --
	(360.68,184.05) --
	(360.28,183.92) --
	(360.17,183.88) --
	(360.14,183.88) --
	(359.58,183.76) --
	(358.80,183.59) --
	(358.45,183.50) --
	(358.36,183.48) --
	(357.71,183.31) --
	(357.68,183.31) --
	(357.28,183.23) --
	(357.10,183.18) --
	(357.04,183.16) --
	(356.99,183.15) --
	(356.86,183.10) --
	(356.57,183.03) --
	(356.08,182.90) --
	(356.06,182.90) --
	(355.69,182.80) --
	(355.66,182.80) --
	(355.44,182.75) --
	(355.38,182.73) --
	(355.06,182.65) --
	(355.03,182.64) --
	(354.77,182.59) --
	(354.60,182.56) --
	(352.50,182.02) --
	(351.61,181.81) --
	(350.57,181.53) --
	(350.27,181.44) --
	(350.01,181.39) --
	(349.59,181.29) --
	(348.84,181.08) --
	(348.63,181.04) --
	(347.30,180.74) --
	(346.94,180.63) --
	(346.80,180.60) --
	(346.49,180.52) --
	(345.78,180.31) --
	(345.20,180.18) --
	(345.11,180.16) --
	(343.95,179.88) --
	(343.79,179.84) --
	(342.53,179.55) --
	(341.52,179.33) --
	(340.65,179.17) --
	(340.28,179.02) --
	(340.15,178.96) --
	(339.84,178.86) --
	(339.76,178.82) --
	(339.55,178.78) --
	(339.40,178.76) --
	(339.21,178.74) --
	(339.04,178.70) --
	(338.87,178.68) --
	(338.83,178.66) --
	(338.50,178.58) --
	(337.65,178.29) --
	(335.47,177.71) --
	(335.38,177.68) --
	(333.83,177.24) --
	(333.82,177.24) --
	(333.69,177.20) --
	(333.56,177.17) --
	(333.53,177.16) --
	(330.88,176.48) --
	(330.38,176.35) --
	(330.17,176.29) --
	(328.92,176.00) --
	(328.37,175.84) --
	(327.78,175.64) --
	(327.50,175.55) --
	(327.01,175.42) --
	(326.46,175.32) --
	(326.16,175.25) --
	(325.91,175.20) --
	(325.87,175.20) --
	(325.85,175.19) --
	(325.46,175.13) --
	(324.82,174.98) --
	(324.53,174.93) --
	(324.06,174.82) --
	(323.98,174.80) --
	(322.93,174.54) --
	(322.64,174.46) --
	(322.56,174.44) --
	(321.47,174.10) --
	(321.34,174.07) --
	(320.71,173.87) --
	(319.14,173.48) --
	(318.60,173.34) --
	(317.13,172.94) --
	(316.72,172.82) --
	(316.61,172.81) --
	(316.58,172.80) --
	(316.37,172.74) --
	(316.29,172.72) --
	(315.65,172.55) --
	(315.61,172.53) --
	(315.54,172.52) --
	(315.28,172.44) --
	(315.05,172.37) --
	(314.84,172.32) --
	(314.55,172.24) --
	(314.12,172.13) --
	(314.09,172.13) --
	(312.75,171.77) --
	(312.68,171.75) --
	(312.39,171.66) --
	(311.87,171.50) --
	(311.71,171.46) --
	(311.49,171.39) --
	(311.09,171.28) --
	(310.98,171.25) --
	(310.80,171.20) --
	(310.45,171.10) --
	(310.36,171.07) --
	(309.72,170.92) --
	(309.55,170.87) --
	(309.08,170.76) --
	(308.79,170.69) --
	(308.63,170.64) --
	(308.42,170.58) --
	(307.37,170.23) --
	(307.16,170.16) --
	(307.03,170.09) --
	(306.15,169.83) --
	(305.86,169.74) --
	(305.76,169.71) --
	(305.28,169.58) --
	(305.11,169.54) --
	(304.68,169.40) --
	(304.60,169.37) --
	(304.48,169.34) --
	(304.47,169.33) --
	(304.31,169.29) --
	(304.10,169.23) --
	(303.17,168.95) --
	(302.93,168.88) --
	(302.89,168.88) --
	(302.61,168.80) --
	(302.52,168.79) --
	(302.45,168.78) --
	(302.30,168.73) --
	(302.20,168.71) --
	(302.08,168.67) --
	(302.06,168.66) --
	(301.86,168.62) --
	(301.81,168.60) --
	(301.34,168.41) --
	(300.69,168.22) --
	(300.07,168.01) --
	(299.22,167.81) --
	(298.89,167.74) --
	(298.71,167.68) --
	(298.48,167.60) --
	(298.45,167.59) --
	(298.21,167.50) --
	(298.13,167.47) --
	(298.09,167.46) --
	(297.99,167.43) --
	(297.11,167.14) --
	(296.61,167.00) --
	(295.85,166.72) --
	(295.12,166.43) --
	(293.94,166.08) --
	(293.90,166.07) --
	(293.61,165.97) --
	(293.49,165.93) --
	(292.94,165.77) --
	(292.85,165.76) --
	(292.78,165.72) --
	(292.77,165.72) --
	(292.36,165.59) --
	(292.20,165.54) --
	(291.58,165.36) --
	(291.45,165.31) --
	(291.29,165.25) --
	(290.44,164.93) --
	(289.99,164.78) --
	(289.71,164.71) --
	(289.61,164.68) --
	(289.30,164.60) --
	(289.13,164.54) --
	(289.03,164.51) --
	(288.86,164.44) --
	(288.45,164.32) --
	(288.16,164.21) --
	(287.76,164.11) --
	(286.95,163.86) --
	(286.70,163.79) --
	(286.04,163.59) --
	(285.67,163.45) --
	(285.44,163.36) --
	(284.88,163.17) --
	(284.84,163.16) --
	(284.81,163.14) --
	(284.81,163.14) --
	(284.80,163.14) --
	(284.75,163.11) --
	(284.43,163.01) --
	(284.26,162.95) --
	(283.80,162.82) --
	(283.74,162.80) --
	(283.34,162.65) --
	(283.28,162.64) --
	(283.10,162.60) --
	(281.60,162.07) --
	(281.52,162.04) --
	(281.16,161.89) --
	(280.63,161.70) --
	(280.60,161.69) --
	(280.50,161.65) --
	(280.47,161.64) --
	(280.41,161.61) --
	(280.36,161.59) --
	(280.21,161.53) --
	(279.90,161.42) --
	(279.83,161.40) --
	(279.36,161.22) --
	(279.03,161.12) --
	(278.99,161.10) --
	(278.93,161.08) --
	(278.83,161.05) --
	(277.37,160.51) --
	(277.28,160.48) --
	(276.77,160.29) --
	(276.73,160.28) --
	(276.14,160.03) --
	(276.03,159.99) --
	(275.50,159.78) --
	(274.88,159.56) --
	(274.85,159.55) --
	(274.31,159.34) --
	(273.48,159.10) --
	(273.29,159.00) --
	(273.01,158.91) --
	(272.91,158.87) --
	(272.40,158.63) --
	(272.31,158.61) --
	(271.51,158.38) --
	(269.91,157.84) --
	(269.75,157.78) --
	(269.30,157.63) --
	(269.12,157.57) --
	(268.75,157.44) --
	(268.51,157.36) --
	(268.32,157.28) --
	(268.15,157.20) --
	(267.95,157.14) --
	(267.83,157.09) --
	(267.74,157.07) --
	(266.70,156.71) --
	(266.69,156.71) --
	(265.54,156.24) --
	(265.48,156.21) --
	(265.11,156.07) --
	(264.94,156.00) --
	(264.43,155.81) --
	(264.14,155.71) --
	(263.47,155.47) --
	(262.57,155.16) --
	(261.39,154.71) --
	(260.16,154.24) --
	(260.08,154.22) --
	(258.99,153.78) --
	(256.94,152.92) --
	(255.74,152.43) --
	(254.34,151.90) --
	(253.89,151.74) --
	(253.69,151.68) --
	(252.46,151.14) --
	(251.97,150.97) --
	(250.69,150.46) --
	(250.38,150.35) --
	(250.22,150.29) --
	(249.74,150.09) --
	(248.30,149.50) --
	(244.78,148.02) --
	(242.95,147.32) --
	(241.65,146.80) --
	(241.27,146.62) --
	(240.80,146.42) --
	(240.03,146.15) --
	(239.31,145.84) --
	(239.29,145.83) --
	(237.41,145.07) --
	(237.40,145.07) --
	(236.22,144.55) --
	(235.24,144.17) --
	(234.25,143.77) --
	(233.34,143.45) --
	(232.39,143.03) --
	(231.34,142.62) --
	(231.15,142.56) --
	(229.14,141.64) --
	(224.00,139.28) --
	(223.59,139.09) --
	(223.54,139.07) --
	(223.29,138.97) --
	(222.58,138.65) --
	(221.63,138.25) --
	(220.93,137.97) --
	(220.33,137.73) --
	(220.17,137.66) --
	(218.17,136.81) --
	(217.68,136.58) --
	(217.43,136.47) --
	(217.33,136.42) --
	(215.87,135.77) --
	(211.50,133.87) --
	(209.65,133.08) --
	(205.19,130.98) --
	(204.99,130.88) --
	(199.87,128.62) --
	(198.04,127.77) --
	(196.35,126.93) --
	(195.71,126.63) --
	(195.62,126.59) --
	(192.71,125.33) --
	(187.92,123.15) --
	(178.49,118.78) --
	(171.43,115.43) --
	(158.01,108.85) --
	(154.64,107.17) --
	(110.88, 85.07) --
	(101.92, 80.39) --
	( 95.59, 77.11) --
	( 92.34, 75.32) --
	( 86.71, 72.24) --
	( 78.30, 67.94) --
	( 64.98, 61.14) --
	( 64.85, 61.07) --
	( 60.61, 58.89) --
	( 52.97, 54.98) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 52.97, 34.03) -- ( 52.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 60.61, 34.03) -- ( 60.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.85, 34.03) -- ( 64.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.98, 34.03) -- ( 64.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 78.30, 34.03) -- ( 78.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 86.71, 34.03) -- ( 86.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 92.34, 34.03) -- ( 92.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 95.59, 34.03) -- ( 95.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (101.92, 34.03) -- (101.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (110.88, 34.03) -- (110.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (154.64, 34.03) -- (154.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (158.01, 34.03) -- (158.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (171.43, 34.03) -- (171.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (178.49, 34.03) -- (178.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (187.92, 34.03) -- (187.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (192.71, 34.03) -- (192.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.62, 34.03) -- (195.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.71, 34.03) -- (195.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (196.35, 34.03) -- (196.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (198.04, 34.03) -- (198.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (199.87, 34.03) -- (199.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (204.99, 34.03) -- (204.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.19, 34.03) -- (205.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (209.65, 34.03) -- (209.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (211.50, 34.03) -- (211.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (215.87, 34.03) -- (215.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (217.33, 34.03) -- (217.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (217.43, 34.03) -- (217.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (217.68, 34.03) -- (217.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (218.17, 34.03) -- (218.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.17, 34.03) -- (220.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.33, 34.03) -- (220.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.93, 34.03) -- (220.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (221.63, 34.03) -- (221.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (222.58, 34.03) -- (222.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.29, 34.03) -- (223.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.54, 34.03) -- (223.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.59, 34.03) -- (223.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (224.00, 34.03) -- (224.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.14, 34.03) -- (229.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.15, 34.03) -- (231.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.34, 34.03) -- (231.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.39, 34.03) -- (232.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (233.34, 34.03) -- (233.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (234.25, 34.03) -- (234.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (235.24, 34.03) -- (235.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.22, 34.03) -- (236.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (237.40, 34.03) -- (237.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (237.41, 34.03) -- (237.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.29, 34.03) -- (239.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.31, 34.03) -- (239.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.03, 34.03) -- (240.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.80, 34.03) -- (240.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (241.27, 34.03) -- (241.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (241.65, 34.03) -- (241.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.95, 34.03) -- (242.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (244.78, 34.03) -- (244.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.30, 34.03) -- (248.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.74, 34.03) -- (249.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.22, 34.03) -- (250.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.38, 34.03) -- (250.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.69, 34.03) -- (250.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.97, 34.03) -- (251.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.46, 34.03) -- (252.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.69, 34.03) -- (253.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.89, 34.03) -- (253.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (254.34, 34.03) -- (254.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (255.74, 34.03) -- (255.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.94, 34.03) -- (256.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.99, 34.03) -- (258.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.08, 34.03) -- (260.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.16, 34.03) -- (260.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.39, 34.03) -- (261.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.57, 34.03) -- (262.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (263.47, 34.03) -- (263.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.14, 34.03) -- (264.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.43, 34.03) -- (264.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.94, 34.03) -- (264.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (265.11, 34.03) -- (265.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (265.48, 34.03) -- (265.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (265.54, 34.03) -- (265.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.69, 34.03) -- (266.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.70, 34.03) -- (266.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.75, 34.03) -- (267.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.83, 34.03) -- (267.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.95, 34.03) -- (267.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.15, 34.03) -- (268.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.31, 34.03) -- (268.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.51, 34.03) -- (268.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.75, 34.03) -- (268.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.12, 34.03) -- (269.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.30, 34.03) -- (269.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.75, 34.03) -- (269.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.91, 34.03) -- (269.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.51, 34.03) -- (271.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.31, 34.03) -- (272.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.40, 34.03) -- (272.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.91, 34.03) -- (272.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.01, 34.03) -- (273.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.29, 34.03) -- (273.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.48, 34.03) -- (273.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.31, 34.03) -- (274.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.85, 34.03) -- (274.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.88, 34.03) -- (274.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.50, 34.03) -- (275.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.03, 34.03) -- (276.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.14, 34.03) -- (276.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.73, 34.03) -- (276.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.77, 34.03) -- (276.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.28, 34.03) -- (277.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.37, 34.03) -- (277.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.83, 34.03) -- (278.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.93, 34.03) -- (278.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.99, 34.03) -- (278.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.03, 34.03) -- (279.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.36, 34.03) -- (279.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.83, 34.03) -- (279.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.90, 34.03) -- (279.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.21, 34.03) -- (280.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.36, 34.03) -- (280.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.41, 34.03) -- (280.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.47, 34.03) -- (280.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.50, 34.03) -- (280.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.60, 34.03) -- (280.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.63, 34.03) -- (280.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.16, 34.03) -- (281.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.52, 34.03) -- (281.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.60, 34.03) -- (281.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.10, 34.03) -- (283.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.28, 34.03) -- (283.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.34, 34.03) -- (283.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.74, 34.03) -- (283.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.80, 34.03) -- (283.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.26, 34.03) -- (284.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.43, 34.03) -- (284.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.75, 34.03) -- (284.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.80, 34.03) -- (284.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.81, 34.03) -- (284.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.81, 34.03) -- (284.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.84, 34.03) -- (284.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.88, 34.03) -- (284.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.44, 34.03) -- (285.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.67, 34.03) -- (285.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.04, 34.03) -- (286.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.70, 34.03) -- (286.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.95, 34.03) -- (286.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.76, 34.03) -- (287.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.16, 34.03) -- (288.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.45, 34.03) -- (288.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.86, 34.03) -- (288.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.03, 34.03) -- (289.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.13, 34.03) -- (289.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.30, 34.03) -- (289.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.61, 34.03) -- (289.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.71, 34.03) -- (289.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.99, 34.03) -- (289.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.44, 34.03) -- (290.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.29, 34.03) -- (291.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.45, 34.03) -- (291.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.58, 34.03) -- (291.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.20, 34.03) -- (292.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.36, 34.03) -- (292.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.77, 34.03) -- (292.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.78, 34.03) -- (292.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.85, 34.03) -- (292.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.94, 34.03) -- (292.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.49, 34.03) -- (293.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.61, 34.03) -- (293.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.90, 34.03) -- (293.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.94, 34.03) -- (293.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.12, 34.03) -- (295.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.85, 34.03) -- (295.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.61, 34.03) -- (296.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.11, 34.03) -- (297.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.99, 34.03) -- (297.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.09, 34.03) -- (298.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.14, 34.03) -- (298.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.21, 34.03) -- (298.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.45, 34.03) -- (298.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.48, 34.03) -- (298.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.71, 34.03) -- (298.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.89, 34.03) -- (298.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.22, 34.03) -- (299.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.07, 34.03) -- (300.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.69, 34.03) -- (300.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.34, 34.03) -- (301.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.81, 34.03) -- (301.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.86, 34.03) -- (301.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.06, 34.03) -- (302.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.08, 34.03) -- (302.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.20, 34.03) -- (302.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.30, 34.03) -- (302.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.45, 34.03) -- (302.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.52, 34.03) -- (302.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.61, 34.03) -- (302.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.89, 34.03) -- (302.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.93, 34.03) -- (302.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.17, 34.03) -- (303.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.10, 34.03) -- (304.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.31, 34.03) -- (304.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.47, 34.03) -- (304.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.48, 34.03) -- (304.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.60, 34.03) -- (304.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.68, 34.03) -- (304.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.11, 34.03) -- (305.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.28, 34.03) -- (305.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.76, 34.03) -- (305.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.86, 34.03) -- (305.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.15, 34.03) -- (306.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.03, 34.03) -- (307.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.16, 34.03) -- (307.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.37, 34.03) -- (307.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.42, 34.03) -- (308.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.63, 34.03) -- (308.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.79, 34.03) -- (308.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.08, 34.03) -- (309.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.55, 34.03) -- (309.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.72, 34.03) -- (309.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.36, 34.03) -- (310.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.45, 34.03) -- (310.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.80, 34.03) -- (310.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.98, 34.03) -- (310.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.09, 34.03) -- (311.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.49, 34.03) -- (311.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.71, 34.03) -- (311.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.87, 34.03) -- (311.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.39, 34.03) -- (312.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.68, 34.03) -- (312.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.75, 34.03) -- (312.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.09, 34.03) -- (314.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.12, 34.03) -- (314.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.55, 34.03) -- (314.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.84, 34.03) -- (314.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.05, 34.03) -- (315.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.28, 34.03) -- (315.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.54, 34.03) -- (315.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.61, 34.03) -- (315.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.65, 34.03) -- (315.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.29, 34.03) -- (316.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.37, 34.03) -- (316.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.58, 34.03) -- (316.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.61, 34.03) -- (316.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.72, 34.03) -- (316.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.13, 34.03) -- (317.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.60, 34.03) -- (318.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.14, 34.03) -- (319.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.71, 34.03) -- (320.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.34, 34.03) -- (321.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.47, 34.03) -- (321.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.56, 34.03) -- (322.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.64, 34.03) -- (322.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.93, 34.03) -- (322.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.97, 34.03) -- (323.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.06, 34.03) -- (324.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.53, 34.03) -- (324.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.82, 34.03) -- (324.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.46, 34.03) -- (325.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.85, 34.03) -- (325.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.87, 34.03) -- (325.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.91, 34.03) -- (325.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.16, 34.03) -- (326.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.46, 34.03) -- (326.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.01, 34.03) -- (327.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.50, 34.03) -- (327.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.78, 34.03) -- (327.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.37, 34.03) -- (328.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.92, 34.03) -- (328.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.17, 34.03) -- (330.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.38, 34.03) -- (330.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.88, 34.03) -- (330.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.53, 34.03) -- (333.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.56, 34.03) -- (333.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.69, 34.03) -- (333.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.82, 34.03) -- (333.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.83, 34.03) -- (333.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.38, 34.03) -- (335.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.47, 34.03) -- (335.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.65, 34.03) -- (337.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.50, 34.03) -- (338.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.83, 34.03) -- (338.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.87, 34.03) -- (338.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.04, 34.03) -- (339.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.21, 34.03) -- (339.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.40, 34.03) -- (339.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.55, 34.03) -- (339.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.76, 34.03) -- (339.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.84, 34.03) -- (339.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.15, 34.03) -- (340.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.28, 34.03) -- (340.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.65, 34.03) -- (340.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.52, 34.03) -- (341.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.53, 34.03) -- (342.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.79, 34.03) -- (343.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.95, 34.03) -- (343.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.11, 34.03) -- (345.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.20, 34.03) -- (345.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.78, 34.03) -- (345.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.49, 34.03) -- (346.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.80, 34.03) -- (346.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.94, 34.03) -- (346.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (347.30, 34.03) -- (347.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.63, 34.03) -- (348.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.84, 34.03) -- (348.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.59, 34.03) -- (349.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.01, 34.03) -- (350.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.27, 34.03) -- (350.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.57, 34.03) -- (350.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.61, 34.03) -- (351.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (352.51, 34.03) -- (352.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.60, 34.03) -- (354.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.77, 34.03) -- (354.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.03, 34.03) -- (355.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.06, 34.03) -- (355.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.38, 34.03) -- (355.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.44, 34.03) -- (355.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.66, 34.03) -- (355.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.69, 34.03) -- (355.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.06, 34.03) -- (356.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.08, 34.03) -- (356.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.57, 34.03) -- (356.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.86, 34.03) -- (356.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.99, 34.03) -- (356.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.04, 34.03) -- (357.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.10, 34.03) -- (357.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.28, 34.03) -- (357.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.68, 34.03) -- (357.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.71, 34.03) -- (357.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.36, 34.03) -- (358.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.45, 34.03) -- (358.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.80, 34.03) -- (358.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.58, 34.03) -- (359.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.14, 34.03) -- (360.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.17, 34.03) -- (360.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.28, 34.03) -- (360.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.68, 34.03) -- (360.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.54, 34.03) -- (362.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.92, 34.03) -- (362.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.99, 34.03) -- (362.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.99, 34.03) -- (362.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.68, 34.03) -- (363.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.71, 34.03) -- (363.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.25, 34.03) -- (364.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.57, 34.03) -- (364.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.77, 34.03) -- (364.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.00, 34.03) -- (365.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.26, 34.03) -- (365.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.84, 34.03) -- (365.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.84, 34.03) -- (365.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.92, 34.03) -- (365.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.04, 34.03) -- (366.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.47, 34.03) -- (366.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (367.05, 34.03) -- (367.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.20, 34.03) -- (368.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.39, 34.03) -- (368.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.50, 34.03) -- (368.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (370.21, 34.03) -- (370.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.17, 34.03) -- (371.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.58, 34.03) -- (371.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.02, 34.03) -- (372.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.58, 34.03) -- (373.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.96, 34.03) -- (376.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.52, 34.03) -- (378.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.85, 34.03) -- (383.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.44, 34.03) -- (384.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.92, 34.03) -- (384.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.30, 34.03) -- (385.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.73, 34.03) -- (385.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.96, 34.03) -- (385.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.70, 34.03) -- (386.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.81, 34.03) -- (386.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.17, 34.03) -- (387.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.19, 34.03) -- (387.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.24, 34.03) -- (387.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.39, 34.03) -- (387.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.67, 34.03) -- (387.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.26, 34.03) -- (388.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.16, 34.03) -- (389.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.82, 34.03) -- (389.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.40, 34.03) -- (390.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.45, 34.03) -- (390.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.79, 34.03) -- (390.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (391.07, 34.03) -- (391.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (395.16, 34.03) -- (395.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (396.00, 34.03) -- (396.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (396.15, 34.03) -- (396.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (403.52, 34.03) -- (403.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.02, 34.03) -- (404.02, 41.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	( 35.42,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31, 54.28) {-5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,133.28) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,212.28) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 36.01, 20.31) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 83.96, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (131.90, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (179.85, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (227.80, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (275.75, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (323.69, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (371.64, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (228.50,  4.82) {Ln(Min. Cap Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}