### 2014-10-06 LaTeX/mCitySimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-10-06 19:44:41
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.97,106.37) --
	( 61.01,109.04) --
	( 65.48,110.52) --
	( 65.62,110.57) --
	( 79.64,115.23) --
	( 88.50,118.17) --
	( 94.41,120.14) --
	( 97.84,121.27) --
	(104.50,123.49) --
	(110.01,125.32) --
	(113.94,126.62) --
	(116.56,127.49) --
	(124.94,130.28) --
	(160.01,141.93) --
	(160.15,141.97) --
	(163.56,143.10) --
	(164.50,143.42) --
	(175.08,146.93) --
	(177.68,147.80) --
	(185.12,150.27) --
	(195.04,153.56) --
	(195.05,153.57) --
	(199.79,155.14) --
	(200.09,155.24) --
	(201.70,155.77) --
	(202.42,156.01) --
	(203.15,156.26) --
	(203.24,156.29) --
	(203.92,156.51) --
	(204.08,156.56) --
	(205.70,157.10) --
	(211.30,158.97) --
	(213.02,159.54) --
	(213.22,159.60) --
	(213.28,159.62) --
	(214.88,160.15) --
	(219.39,161.65) --
	(219.87,161.81) --
	(224.51,163.35) --
	(226.38,163.97) --
	(226.89,164.14) --
	(228.22,164.58) --
	(229.80,165.11) --
	(230.53,165.35) --
	(231.54,165.69) --
	(232.29,165.94) --
	(232.55,166.02) --
	(232.60,166.04) --
	(233.03,166.18) --
	(236.50,167.34) --
	(238.44,167.98) --
	(239.50,168.33) --
	(240.76,168.75) --
	(242.87,169.45) --
	(243.82,169.77) --
	(244.87,170.11) --
	(245.90,170.46) --
	(247.14,170.87) --
	(247.15,170.87) --
	(248.00,171.16) --
	(248.59,171.35) --
	(249.15,171.54) --
	(249.91,171.79) --
	(250.14,171.87) --
	(250.71,172.06) --
	(251.21,172.22) --
	(251.61,172.35) --
	(251.88,172.44) --
	(252.52,172.66) --
	(252.98,172.81) --
	(253.12,172.85) --
	(253.36,172.94) --
	(254.91,173.45) --
	(255.02,173.49) --
	(256.32,173.92) --
	(258.61,174.68) --
	(260.03,175.15) --
	(260.14,175.19) --
	(260.80,175.41) --
	(261.97,175.79) --
	(262.19,175.87) --
	(262.48,175.96) --
	(262.99,176.14) --
	(264.29,176.57) --
	(264.50,176.64) --
	(264.53,176.65) --
	(264.97,176.79) --
	(266.10,177.17) --
	(266.13,177.18) --
	(266.44,177.28) --
	(267.28,177.56) --
	(267.71,177.70) --
	(268.38,177.92) --
	(269.11,178.17) --
	(269.74,178.38) --
	(269.87,178.42) --
	(271.02,178.80) --
	(271.10,178.83) --
	(271.68,179.02) --
	(272.40,179.26) --
	(272.79,179.39) --
	(273.63,179.67) --
	(274.52,179.96) --
	(275.28,180.22) --
	(275.59,180.32) --
	(276.13,180.50) --
	(276.31,180.56) --
	(276.77,180.71) --
	(277.30,180.89) --
	(277.98,181.12) --
	(278.04,181.13) --
	(278.45,181.27) --
	(279.17,181.51) --
	(279.30,181.55) --
	(279.52,181.62) --
	(279.66,181.67) --
	(279.69,181.68) --
	(280.15,181.83) --
	(280.53,181.96) --
	(280.72,182.02) --
	(281.20,182.18) --
	(281.36,182.24) --
	(281.45,182.27) --
	(282.80,182.71) --
	(283.05,182.80) --
	(283.89,183.08) --
	(283.98,183.11) --
	(284.07,183.14) --
	(284.62,183.32) --
	(284.86,183.40) --
	(285.03,183.46) --
	(285.13,183.49) --
	(286.00,183.78) --
	(286.57,183.97) --
	(286.59,183.98) --
	(287.25,184.19) --
	(287.81,184.38) --
	(288.54,184.62) --
	(288.59,184.64) --
	(289.22,184.85) --
	(289.76,185.03) --
	(290.76,185.36) --
	(290.86,185.39) --
	(291.20,185.51) --
	(291.32,185.54) --
	(291.81,185.71) --
	(291.88,185.73) --
	(292.21,185.84) --
	(292.36,185.89) --
	(292.42,185.91) --
	(292.49,185.93) --
	(292.51,185.94) --
	(292.62,185.98) --
	(293.21,186.17) --
	(293.91,186.41) --
	(295.25,186.85) --
	(295.45,186.92) --
	(295.51,186.94) --
	(295.93,187.08) --
	(295.99,187.10) --
	(296.10,187.13) --
	(296.47,187.26) --
	(296.65,187.32) --
	(296.99,187.43) --
	(297.01,187.43) --
	(297.04,187.45) --
	(297.05,187.45) --
	(297.09,187.46) --
	(297.13,187.47) --
	(297.22,187.51) --
	(297.31,187.53) --
	(297.31,187.53) --
	(297.71,187.67) --
	(297.96,187.75) --
	(298.34,187.88) --
	(298.57,187.95) --
	(299.04,188.11) --
	(300.58,188.62) --
	(300.69,188.66) --
	(301.32,188.87) --
	(302.10,189.13) --
	(302.23,189.17) --
	(302.50,189.26) --
	(302.82,189.36) --
	(302.98,189.42) --
	(303.49,189.59) --
	(303.87,189.71) --
	(304.04,189.77) --
	(304.05,189.77) --
	(304.17,189.82) --
	(304.21,189.83) --
	(304.88,190.05) --
	(305.44,190.24) --
	(305.52,190.26) --
	(305.61,190.29) --
	(305.86,190.37) --
	(306.19,190.49) --
	(306.32,190.53) --
	(307.91,191.06) --
	(307.96,191.07) --
	(308.47,191.24) --
	(308.57,191.27) --
	(309.22,191.49) --
	(309.44,191.56) --
	(309.51,191.59) --
	(309.67,191.64) --
	(310.00,191.75) --
	(310.25,191.83) --
	(310.76,192.00) --
	(310.92,192.06) --
	(311.04,192.09) --
	(311.16,192.14) --
	(311.41,192.22) --
	(311.41,192.22) --
	(311.44,192.23) --
	(311.54,192.26) --
	(311.68,192.31) --
	(311.68,192.31) --
	(311.88,192.38) --
	(312.22,192.49) --
	(312.92,192.72) --
	(313.12,192.79) --
	(313.27,192.84) --
	(313.53,192.93) --
	(313.83,193.02) --
	(314.24,193.16) --
	(314.31,193.18) --
	(314.45,193.23) --
	(314.63,193.29) --
	(315.00,193.41) --
	(315.21,193.48) --
	(315.23,193.49) --
	(315.37,193.53) --
	(315.48,193.57) --
	(315.51,193.58) --
	(315.62,193.62) --
	(315.70,193.64) --
	(315.79,193.67) --
	(316.13,193.79) --
	(316.22,193.82) --
	(316.38,193.87) --
	(316.61,193.95) --
	(316.93,194.05) --
	(317.07,194.10) --
	(317.20,194.14) --
	(317.36,194.20) --
	(317.75,194.33) --
	(317.76,194.33) --
	(317.79,194.34) --
	(317.86,194.36) --
	(317.89,194.37) --
	(317.98,194.40) --
	(318.43,194.55) --
	(318.60,194.61) --
	(318.85,194.69) --
	(319.11,194.78) --
	(319.21,194.81) --
	(319.52,194.91) --
	(319.81,195.01) --
	(320.44,195.22) --
	(321.85,195.69) --
	(321.91,195.71) --
	(322.04,195.75) --
	(322.30,195.84) --
	(322.60,195.94) --
	(322.75,195.99) --
	(323.02,196.08) --
	(323.10,196.10) --
	(323.28,196.16) --
	(324.05,196.42) --
	(324.42,196.54) --
	(324.61,196.60) --
	(324.72,196.64) --
	(325.08,196.76) --
	(325.99,197.06) --
	(326.58,197.26) --
	(326.72,197.31) --
	(327.61,197.60) --
	(327.91,197.70) --
	(328.67,197.95) --
	(328.94,198.04) --
	(329.07,198.09) --
	(329.52,198.24) --
	(331.17,198.78) --
	(332.00,199.06) --
	(332.27,199.15) --
	(333.07,199.42) --
	(333.19,199.45) --
	(333.76,199.64) --
	(335.51,200.23) --
	(336.07,200.41) --
	(336.75,200.64) --
	(336.88,200.68) --
	(337.54,200.90) --
	(337.92,201.03) --
	(338.29,201.15) --
	(338.32,201.16) --
	(338.37,201.17) --
	(338.79,201.32) --
	(338.87,201.34) --
	(339.41,201.52) --
	(339.54,201.56) --
	(340.26,201.80) --
	(340.28,201.81) --
	(340.36,201.84) --
	(340.58,201.91) --
	(340.90,202.02) --
	(341.07,202.07) --
	(341.38,202.18) --
	(341.48,202.21) --
	(341.53,202.23) --
	(341.57,202.24) --
	(342.00,202.38) --
	(342.06,202.40) --
	(342.60,202.58) --
	(342.92,202.69) --
	(344.39,203.18) --
	(344.81,203.31) --
	(345.03,203.39) --
	(345.51,203.55) --
	(345.56,203.56) --
	(346.27,203.80) --
	(348.09,204.40) --
	(348.19,204.44) --
	(348.32,204.48) --
	(348.52,204.55) --
	(348.65,204.59) --
	(350.04,205.05) --
	(350.29,205.13) --
	(350.39,205.17) --
	(351.49,205.53) --
	(354.19,206.43) --
	(354.41,206.50) --
	(354.96,206.69) --
	(354.99,206.69) --
	(355.28,206.79) --
	(355.84,206.98) --
	(356.71,207.27) --
	(356.75,207.28) --
	(357.50,207.53) --
	(357.53,207.54) --
	(358.04,207.71) --
	(358.62,207.90) --
	(359.31,208.13) --
	(359.62,208.23) --
	(359.75,208.28) --
	(359.85,208.31) --
	(360.61,208.56) --
	(360.64,208.57) --
	(360.71,208.60) --
	(361.91,208.99) --
	(362.60,209.22) --
	(364.24,209.77) --
	(364.46,209.84) --
	(365.70,210.25) --
	(366.29,210.45) --
	(370.53,211.86) --
	(371.02,212.02) --
	(372.06,212.37) --
	(372.07,212.37) --
	(372.38,212.47) --
	(372.60,212.54) --
	(372.90,212.65) --
	(373.16,212.73) --
	(373.53,212.86) --
	(373.80,212.94) --
	(374.46,213.16) --
	(374.81,213.28) --
	(375.77,213.60) --
	(376.36,213.80) --
	(376.51,213.85) --
	(376.93,213.98) --
	(378.03,214.35) --
	(379.36,214.79) --
	(379.36,214.79) --
	(380.09,215.03) --
	(380.12,215.04) --
	(381.03,215.35) --
	(381.47,215.49) --
	(381.75,215.58) --
	(382.36,215.79) --
	(382.44,215.82) --
	(382.57,215.86) --
	(383.02,216.01) --
	(385.04,216.68) --
	(385.16,216.72) --
	(388.87,217.95) --
	(393.56,219.51) --
	(395.71,220.22) --
	(403.26,222.73) --
	(404.02,222.98);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 52.97,167.36) --
	( 61.01,168.41) --
	( 65.48,168.80) --
	( 65.62,168.81) --
	( 79.64,170.79) --
	( 88.50,171.79) --
	( 94.41,172.49) --
	( 97.84,173.06) --
	(104.50,173.98) --
	(110.01,174.56) --
	(113.94,175.03) --
	(116.56,175.36) --
	(124.94,176.30) --
	(160.01,181.27) --
	(160.15,181.30) --
	(163.56,181.70) --
	(164.50,181.78) --
	(175.08,183.35) --
	(177.68,183.86) --
	(185.12,185.40) --
	(195.04,187.30) --
	(195.05,187.30) --
	(199.79,188.52) --
	(200.09,188.60) --
	(201.70,188.94) --
	(202.42,189.05) --
	(203.15,189.18) --
	(203.24,189.19) --
	(203.92,189.29) --
	(204.08,189.32) --
	(205.70,189.71) --
	(211.30,190.89) --
	(213.02,191.21) --
	(213.22,191.26) --
	(213.28,191.27) --
	(214.88,191.68) --
	(219.39,192.70) --
	(219.87,192.77) --
	(224.51,193.76) --
	(226.38,194.26) --
	(226.89,194.38) --
	(228.22,194.75) --
	(229.80,195.15) --
	(230.53,195.25) --
	(231.54,195.47) --
	(232.29,195.59) --
	(232.55,195.66) --
	(232.60,195.67) --
	(233.03,195.77) --
	(236.50,196.61) --
	(238.44,197.03) --
	(239.50,197.27) --
	(240.76,197.54) --
	(242.87,198.03) --
	(243.82,198.26) --
	(244.87,198.51) --
	(245.90,198.81) --
	(247.14,199.15) --
	(247.15,199.15) --
	(248.00,199.39) --
	(248.59,199.55) --
	(249.15,199.72) --
	(249.91,199.94) --
	(250.14,200.00) --
	(250.71,200.16) --
	(251.21,200.29) --
	(251.61,200.40) --
	(251.88,200.46) --
	(252.52,200.69) --
	(252.98,200.84) --
	(253.12,200.86) --
	(253.36,200.89) --
	(254.91,201.29) --
	(255.02,201.32) --
	(256.32,201.72) --
	(258.61,202.38) --
	(260.03,202.80) --
	(260.14,202.84) --
	(260.80,203.04) --
	(261.97,203.27) --
	(262.19,203.35) --
	(262.48,203.44) --
	(262.99,203.62) --
	(264.29,204.03) --
	(264.50,204.10) --
	(264.53,204.11) --
	(264.97,204.25) --
	(266.10,204.54) --
	(266.13,204.55) --
	(266.44,204.61) --
	(267.28,204.93) --
	(267.71,205.05) --
	(268.38,205.26) --
	(269.11,205.45) --
	(269.74,205.69) --
	(269.87,205.74) --
	(271.02,206.18) --
	(271.10,206.22) --
	(271.68,206.47) --
	(272.40,206.66) --
	(272.79,206.78) --
	(273.63,207.07) --
	(274.52,207.40) --
	(275.28,207.68) --
	(275.59,207.78) --
	(276.13,207.95) --
	(276.31,207.99) --
	(276.77,208.14) --
	(277.30,208.37) --
	(277.98,208.66) --
	(278.04,208.68) --
	(278.45,208.79) --
	(279.17,208.98) --
	(279.30,209.02) --
	(279.52,209.08) --
	(279.66,209.14) --
	(279.69,209.14) --
	(280.15,209.31) --
	(280.53,209.48) --
	(280.72,209.56) --
	(281.20,209.75) --
	(281.36,209.82) --
	(281.45,209.85) --
	(282.80,210.34) --
	(283.05,210.37) --
	(283.89,210.70) --
	(283.98,210.74) --
	(284.07,210.78) --
	(284.62,210.98) --
	(284.86,211.08) --
	(285.03,211.13) --
	(285.13,211.16) --
	(286.00,211.42) --
	(286.57,211.61) --
	(286.59,211.61) --
	(287.25,211.79) --
	(287.81,211.98) --
	(288.54,212.28) --
	(288.59,212.29) --
	(289.22,212.44) --
	(289.76,212.64) --
	(290.76,213.08) --
	(290.86,213.13) --
	(291.20,213.21) --
	(291.32,213.21) --
	(291.81,213.42) --
	(291.88,213.46) --
	(292.21,213.58) --
	(292.36,213.65) --
	(292.42,213.69) --
	(292.49,213.72) --
	(292.51,213.74) --
	(292.62,213.79) --
	(293.21,213.98) --
	(293.91,214.28) --
	(295.25,214.70) --
	(295.45,214.78) --
	(295.51,214.81) --
	(295.93,215.00) --
	(295.99,215.02) --
	(296.10,215.06) --
	(296.47,215.21) --
	(296.65,215.26) --
	(296.99,215.30) --
	(297.01,215.30) --
	(297.04,215.31) --
	(297.05,215.31) --
	(297.09,215.32) --
	(297.13,215.34) --
	(297.22,215.39) --
	(297.31,215.44) --
	(297.31,215.44) --
	(297.71,215.55) --
	(297.96,215.64) --
	(298.34,215.77) --
	(298.57,215.86) --
	(299.04,216.04) --
	(300.58,216.50) --
	(300.69,216.54) --
	(301.32,216.72) --
	(302.10,216.99) --
	(302.23,217.04) --
	(302.50,217.16) --
	(302.82,217.32) --
	(302.98,217.40) --
	(303.49,217.63) --
	(303.87,217.81) --
	(304.04,217.89) --
	(304.05,217.89) --
	(304.17,217.95) --
	(304.21,217.97) --
	(304.88,218.27) --
	(305.44,218.51) --
	(305.52,218.55) --
	(305.61,218.58) --
	(305.86,218.68) --
	(306.19,218.81) --
	(306.32,218.86) --
	(307.91,219.57) --
	(307.96,219.60) --
	(308.47,219.81) --
	(308.57,219.85) --
	(309.22,220.10) --
	(309.44,220.18) --
	(309.51,220.20) --
	(309.67,220.25) --
	(310.00,220.39) --
	(310.25,220.47) --
	(310.76,220.68) --
	(310.92,220.74) --
	(311.04,220.77) --
	(311.16,220.82) --
	(311.41,220.90) --
	(311.41,220.90) --
	(311.44,220.91) --
	(311.54,220.94) --
	(311.68,220.98) --
	(311.68,220.99) --
	(311.88,221.05) --
	(312.22,221.19) --
	(312.92,221.49) --
	(313.12,221.57) --
	(313.27,221.65) --
	(313.53,221.73) --
	(313.83,221.83) --
	(314.24,222.00) --
	(314.31,222.04) --
	(314.45,222.11) --
	(314.63,222.16) --
	(315.00,222.37) --
	(315.21,222.46) --
	(315.23,222.47) --
	(315.37,222.52) --
	(315.48,222.56) --
	(315.51,222.58) --
	(315.62,222.62) --
	(315.70,222.64) --
	(315.79,222.66) --
	(316.13,222.81) --
	(316.22,222.83) --
	(316.38,222.87) --
	(316.61,222.95) --
	(316.93,223.12) --
	(317.07,223.18) --
	(317.20,223.26) --
	(317.36,223.33) --
	(317.75,223.47) --
	(317.76,223.47) --
	(317.79,223.49) --
	(317.86,223.53) --
	(317.89,223.54) --
	(317.98,223.58) --
	(318.43,223.78) --
	(318.60,223.85) --
	(318.85,224.00) --
	(319.11,224.09) --
	(319.21,224.11) --
	(319.52,224.25) --
	(319.81,224.40) --
	(320.44,224.64) --
	(321.85,225.23) --
	(321.91,225.27) --
	(322.04,225.35) --
	(322.30,225.47) --
	(322.60,225.60) --
	(322.75,225.66) --
	(323.02,225.77) --
	(323.10,225.80) --
	(323.28,225.87) --
	(324.05,226.17) --
	(324.42,226.34) --
	(324.61,226.43) --
	(324.72,226.50) --
	(325.08,226.65) --
	(325.99,227.12) --
	(326.58,227.38) --
	(326.72,227.44) --
	(327.61,227.83) --
	(327.91,227.96) --
	(328.67,228.25) --
	(328.94,228.38) --
	(329.07,228.46) --
	(329.52,228.70) --
	(331.17,229.50) --
	(332.00,229.87) --
	(332.27,230.00) --
	(333.07,230.25) --
	(333.19,230.29) --
	(333.76,230.52) --
	(335.51,231.36) --
	(336.07,231.68) --
	(336.75,232.03) --
	(336.88,232.07) --
	(337.54,232.34) --
	(337.92,232.47) --
	(338.29,232.66) --
	(338.32,232.68) --
	(338.37,232.70) --
	(338.79,232.94) --
	(338.87,232.97) --
	(339.41,233.19) --
	(339.54,233.26) --
	(340.26,233.63) --
	(340.28,233.64) --
	(340.36,233.69) --
	(340.58,233.82) --
	(340.90,234.00) --
	(341.07,234.09) --
	(341.38,234.25) --
	(341.48,234.29) --
	(341.53,234.32) --
	(341.57,234.33) --
	(342.00,234.53) --
	(342.06,234.56) --
	(342.60,234.82) --
	(342.92,234.94) --
	(344.39,235.59) --
	(344.81,235.76) --
	(345.03,235.85) --
	(345.51,236.00) --
	(345.56,236.01) --
	(346.27,236.21) --
	(348.09,237.11) --
	(348.19,237.16) --
	(348.32,237.23) --
	(348.52,237.32) --
	(348.65,237.36) --
	(350.04,238.11) --
	(350.29,238.22) --
	(350.39,238.27) --
	(351.49,238.86) --
	(354.19,240.24) --
	(354.41,240.37) --
	(354.96,240.65) --
	(354.99,240.67) --
	(355.28,240.79) --
	(355.84,241.02) --
	(356.71,241.48) --
	(356.75,241.51) --
	(357.50,241.86) --
	(357.53,241.87) --
	(358.04,242.13) --
	(358.62,242.42) --
	(359.31,242.81) --
	(359.62,242.94) --
	(359.75,243.03) --
	(359.85,243.09) --
	(360.61,243.54) --
	(360.64,243.55) --
	(360.71,243.59) --
	(361.91,244.20) --
	(362.60,244.50) --
	(364.24,245.24) --
	(364.46,245.35) --
	(365.70,245.99) --
	(366.29,246.26) --
	(370.53,248.44) --
	(371.02,248.71) --
	(372.06,249.22) --
	(372.07,249.22) --
	(372.38,249.33) --
	(372.60,249.45) --
	(372.90,249.61) --
	(373.16,249.73) --
	(373.53,249.89) --
	(373.80,250.04) --
	(374.46,250.35) --
	(374.81,250.52) --
	(375.77,251.03) --
	(376.36,251.34) --
	(376.51,251.43) --
	(376.93,251.64) --
	(378.03,252.22) --
	(379.36,252.85) --
	(379.36,252.86) --
	(380.09,253.22) --
	(380.12,253.24) --
	(381.03,253.71) --
	(381.47,253.91) --
	(381.75,254.05) --
	(382.36,254.37) --
	(382.44,254.41) --
	(382.57,254.46) --
	(383.02,254.69) --
	(385.04,255.76) --
	(385.16,255.82) --
	(388.87,257.82) --
	(393.56,260.29) --
	(395.71,261.42) --
	(403.26,265.59) --
	(404.02,265.99) --
	(404.02,182.27) --
	(403.26,182.16) --
	(395.71,180.94) --
	(393.56,180.61) --
	(388.87,179.73) --
	(385.16,179.04) --
	(385.04,179.03) --
	(383.02,178.81) --
	(382.57,178.70) --
	(382.44,178.65) --
	(382.36,178.61) --
	(381.75,178.55) --
	(381.47,178.49) --
	(381.03,178.47) --
	(380.12,178.30) --
	(380.09,178.30) --
	(379.36,178.16) --
	(379.36,178.16) --
	(378.03,177.94) --
	(376.93,177.72) --
	(376.51,177.66) --
	(376.36,177.65) --
	(375.77,177.59) --
	(374.81,177.41) --
	(374.46,177.37) --
	(373.80,177.19) --
	(373.53,177.12) --
	(373.16,177.03) --
	(372.90,176.96) --
	(372.60,176.88) --
	(372.38,176.82) --
	(372.07,176.74) --
	(372.06,176.74) --
	(371.02,176.61) --
	(370.53,176.50) --
	(366.29,175.56) --
	(365.70,175.48) --
	(364.46,175.26) --
	(364.24,175.21) --
	(362.60,174.88) --
	(361.91,174.73) --
	(360.71,174.46) --
	(360.64,174.43) --
	(360.61,174.42) --
	(359.85,174.28) --
	(359.75,174.26) --
	(359.62,174.24) --
	(359.31,174.18) --
	(358.62,174.10) --
	(358.04,174.00) --
	(357.53,173.89) --
	(357.50,173.88) --
	(356.75,173.74) --
	(356.71,173.73) --
	(355.84,173.55) --
	(355.28,173.44) --
	(354.99,173.38) --
	(354.96,173.38) --
	(354.41,173.27) --
	(354.19,173.22) --
	(351.49,172.56) --
	(350.39,172.25) --
	(350.29,172.22) --
	(350.04,172.22) --
	(348.65,172.01) --
	(348.52,171.96) --
	(348.32,171.89) --
	(348.19,171.87) --
	(348.09,171.85) --
	(346.27,171.41) --
	(345.56,171.26) --
	(345.51,171.26) --
	(345.03,171.14) --
	(344.81,171.13) --
	(344.39,171.04) --
	(342.92,170.73) --
	(342.60,170.68) --
	(342.06,170.59) --
	(342.00,170.58) --
	(341.57,170.51) --
	(341.53,170.50) --
	(341.48,170.48) --
	(341.38,170.46) --
	(341.07,170.41) --
	(340.90,170.35) --
	(340.58,170.27) --
	(340.36,170.25) --
	(340.28,170.24) --
	(340.26,170.24) --
	(339.54,170.08) --
	(339.41,170.03) --
	(338.87,169.90) --
	(338.79,169.88) --
	(338.37,169.75) --
	(338.32,169.74) --
	(338.29,169.73) --
	(337.92,169.66) --
	(337.54,169.63) --
	(336.88,169.52) --
	(336.75,169.49) --
	(336.07,169.33) --
	(335.51,169.18) --
	(333.76,168.91) --
	(333.19,168.80) --
	(333.07,168.78) --
	(332.27,168.47) --
	(332.00,168.39) --
	(331.17,168.19) --
	(329.52,167.74) --
	(329.07,167.64) --
	(328.94,167.62) --
	(328.67,167.52) --
	(327.91,167.32) --
	(327.61,167.25) --
	(326.72,167.12) --
	(326.58,167.07) --
	(325.99,166.89) --
	(325.08,166.68) --
	(324.72,166.60) --
	(324.61,166.57) --
	(324.42,166.53) --
	(324.05,166.45) --
	(323.28,166.24) --
	(323.10,166.20) --
	(323.02,166.19) --
	(322.75,166.13) --
	(322.60,166.08) --
	(322.30,166.02) --
	(322.04,165.95) --
	(321.91,165.91) --
	(321.85,165.90) --
	(320.44,165.54) --
	(319.81,165.48) --
	(319.52,165.41) --
	(319.21,165.35) --
	(319.11,165.31) --
	(318.85,165.22) --
	(318.60,165.15) --
	(318.43,165.12) --
	(317.98,165.00) --
	(317.89,164.98) --
	(317.86,164.98) --
	(317.79,164.97) --
	(317.76,164.96) --
	(317.75,164.96) --
	(317.36,164.87) --
	(317.20,164.84) --
	(317.07,164.81) --
	(316.93,164.77) --
	(316.61,164.69) --
	(316.38,164.64) --
	(316.22,164.63) --
	(316.13,164.63) --
	(315.79,164.58) --
	(315.70,164.56) --
	(315.62,164.55) --
	(315.51,164.53) --
	(315.48,164.52) --
	(315.37,164.49) --
	(315.23,164.45) --
	(315.21,164.44) --
	(315.00,164.38) --
	(314.63,164.28) --
	(314.45,164.23) --
	(314.31,164.17) --
	(314.24,164.15) --
	(313.83,164.06) --
	(313.53,163.97) --
	(313.27,163.89) --
	(313.12,163.80) --
	(312.92,163.72) --
	(312.22,163.58) --
	(311.88,163.47) --
	(311.68,163.41) --
	(311.68,163.41) --
	(311.54,163.36) --
	(311.44,163.32) --
	(311.41,163.32) --
	(311.41,163.32) --
	(311.16,163.25) --
	(311.04,163.22) --
	(310.92,163.20) --
	(310.76,163.15) --
	(310.25,163.05) --
	(310.00,162.99) --
	(309.67,162.92) --
	(309.51,162.86) --
	(309.44,162.83) --
	(309.22,162.77) --
	(308.57,162.62) --
	(308.47,162.60) --
	(307.96,162.44) --
	(307.91,162.43) --
	(306.32,161.94) --
	(306.19,161.90) --
	(305.86,161.80) --
	(305.61,161.73) --
	(305.52,161.70) --
	(305.44,161.68) --
	(304.88,161.50) --
	(304.21,161.34) --
	(304.17,161.33) --
	(304.05,161.30) --
	(304.04,161.30) --
	(303.87,161.27) --
	(303.49,161.16) --
	(302.98,161.00) --
	(302.82,160.95) --
	(302.50,160.89) --
	(302.23,160.84) --
	(302.10,160.82) --
	(301.32,160.64) --
	(300.69,160.55) --
	(300.58,160.54) --
	(299.04,160.13) --
	(298.57,160.04) --
	(298.34,159.99) --
	(297.96,159.86) --
	(297.71,159.79) --
	(297.31,159.71) --
	(297.31,159.71) --
	(297.22,159.67) --
	(297.13,159.62) --
	(297.09,159.60) --
	(297.05,159.58) --
	(297.04,159.58) --
	(297.01,159.57) --
	(296.99,159.56) --
	(296.65,159.45) --
	(296.47,159.35) --
	(296.10,159.26) --
	(295.99,159.24) --
	(295.93,159.23) --
	(295.51,159.10) --
	(295.45,159.08) --
	(295.25,159.01) --
	(293.91,158.66) --
	(293.21,158.46) --
	(292.62,158.25) --
	(292.51,158.22) --
	(292.49,158.21) --
	(292.42,158.20) --
	(292.36,158.18) --
	(292.21,158.15) --
	(291.88,158.09) --
	(291.81,158.08) --
	(291.32,157.93) --
	(291.20,157.93) --
	(290.86,157.83) --
	(290.76,157.78) --
	(289.76,157.50) --
	(289.22,157.31) --
	(288.59,157.10) --
	(288.54,157.09) --
	(287.81,156.93) --
	(287.25,156.69) --
	(286.59,156.44) --
	(286.57,156.44) --
	(286.00,156.29) --
	(285.13,156.04) --
	(285.03,156.02) --
	(284.86,155.95) --
	(284.62,155.87) --
	(284.07,155.70) --
	(283.98,155.67) --
	(283.89,155.63) --
	(283.05,155.39) --
	(282.80,155.29) --
	(281.45,154.91) --
	(281.36,154.89) --
	(281.20,154.85) --
	(280.72,154.72) --
	(280.53,154.68) --
	(280.15,154.56) --
	(279.69,154.48) --
	(279.66,154.48) --
	(279.52,154.44) --
	(279.30,154.38) --
	(279.17,154.34) --
	(278.45,154.06) --
	(278.04,153.95) --
	(277.98,153.93) --
	(277.30,153.70) --
	(276.77,153.46) --
	(276.31,153.28) --
	(276.13,153.22) --
	(275.59,153.06) --
	(275.28,152.95) --
	(274.52,152.77) --
	(273.63,152.39) --
	(272.79,152.07) --
	(272.40,151.94) --
	(271.68,151.70) --
	(271.10,151.43) --
	(271.02,151.41) --
	(269.87,151.00) --
	(269.74,150.96) --
	(269.11,150.78) --
	(268.38,150.50) --
	(267.71,150.29) --
	(267.28,150.15) --
	(266.44,149.85) --
	(266.13,149.74) --
	(266.10,149.73) --
	(264.97,149.27) --
	(264.53,149.09) --
	(264.50,149.08) --
	(264.29,149.03) --
	(262.99,148.54) --
	(262.48,148.27) --
	(262.19,148.19) --
	(261.97,148.12) --
	(260.80,147.75) --
	(260.14,147.55) --
	(260.03,147.49) --
	(258.61,146.90) --
	(256.32,145.98) --
	(255.02,145.42) --
	(254.91,145.38) --
	(253.36,144.78) --
	(253.12,144.70) --
	(252.98,144.64) --
	(252.52,144.47) --
	(251.88,144.26) --
	(251.61,144.16) --
	(251.21,144.00) --
	(250.71,143.81) --
	(250.14,143.57) --
	(249.91,143.49) --
	(249.15,143.20) --
	(248.59,142.98) --
	(248.00,142.78) --
	(247.15,142.45) --
	(247.14,142.44) --
	(245.90,142.02) --
	(244.87,141.53) --
	(243.82,141.15) --
	(242.87,140.79) --
	(240.76,139.96) --
	(239.50,139.44) --
	(238.44,139.02) --
	(236.50,138.27) --
	(233.03,136.65) --
	(232.60,136.47) --
	(232.55,136.45) --
	(232.29,136.34) --
	(231.54,136.09) --
	(230.53,135.66) --
	(229.80,135.35) --
	(228.22,134.60) --
	(226.89,134.03) --
	(226.38,133.85) --
	(224.51,133.07) --
	(219.87,130.96) --
	(219.39,130.73) --
	(214.88,128.51) --
	(213.28,127.81) --
	(213.22,127.78) --
	(213.02,127.70) --
	(211.30,126.92) --
	(205.70,124.51) --
	(204.08,123.77) --
	(203.92,123.68) --
	(203.24,123.35) --
	(203.15,123.31) --
	(202.42,123.02) --
	(201.70,122.67) --
	(200.09,121.90) --
	(199.79,121.73) --
	(195.05,119.42) --
	(195.04,119.41) --
	(185.12,114.89) --
	(177.68,111.26) --
	(175.08,110.08) --
	(164.50,104.62) --
	(163.56,104.15) --
	(160.15,102.47) --
	(160.01,102.40) --
	(124.94, 84.29) --
	(116.56, 79.63) --
	(113.94, 78.09) --
	(110.01, 75.93) --
	(104.50, 73.31) --
	( 97.84, 69.48) --
	( 94.41, 67.75) --
	( 88.50, 64.75) --
	( 79.64, 59.82) --
	( 65.62, 51.87) --
	( 65.48, 51.79) --
	( 61.01, 49.34) --
	( 52.97, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 52.97,157.52) --
	( 61.01,158.71) --
	( 65.48,159.33) --
	( 65.62,159.35) --
	( 79.64,161.53) --
	( 88.50,163.05) --
	( 94.41,164.05) --
	( 97.84,164.63) --
	(104.50,165.72) --
	(110.01,166.51) --
	(113.94,167.24) --
	(116.56,167.79) --
	(124.94,169.05) --
	(160.01,175.04) --
	(160.15,175.05) --
	(163.56,175.67) --
	(164.50,175.82) --
	(175.08,178.09) --
	(177.68,178.61) --
	(185.12,180.03) --
	(195.04,182.20) --
	(195.05,182.20) --
	(199.79,183.10) --
	(200.09,183.15) --
	(201.70,183.53) --
	(202.42,183.68) --
	(203.15,183.82) --
	(203.24,183.84) --
	(203.92,183.98) --
	(204.08,184.00) --
	(205.70,184.39) --
	(211.30,185.68) --
	(213.02,186.07) --
	(213.22,186.12) --
	(213.28,186.14) --
	(214.88,186.52) --
	(219.39,187.64) --
	(219.87,187.75) --
	(224.51,188.92) --
	(226.38,189.39) --
	(226.89,189.48) --
	(228.22,189.88) --
	(229.80,190.26) --
	(230.53,190.49) --
	(231.54,190.73) --
	(232.29,190.94) --
	(232.55,191.02) --
	(232.60,191.04) --
	(233.03,191.16) --
	(236.50,192.03) --
	(238.44,192.60) --
	(239.50,192.87) --
	(240.76,193.23) --
	(242.87,193.80) --
	(243.82,193.97) --
	(244.87,194.22) --
	(245.90,194.45) --
	(247.14,194.78) --
	(247.15,194.78) --
	(248.00,195.06) --
	(248.59,195.19) --
	(249.15,195.38) --
	(249.91,195.68) --
	(250.14,195.72) --
	(250.71,195.90) --
	(251.21,196.07) --
	(251.61,196.16) --
	(251.88,196.25) --
	(252.52,196.37) --
	(252.98,196.48) --
	(253.12,196.53) --
	(253.36,196.58) --
	(254.91,197.00) --
	(255.02,197.02) --
	(256.32,197.42) --
	(258.61,198.09) --
	(260.03,198.53) --
	(260.14,198.54) --
	(260.80,198.76) --
	(261.97,199.13) --
	(262.19,199.18) --
	(262.48,199.25) --
	(262.99,199.41) --
	(264.29,199.74) --
	(264.50,199.84) --
	(264.53,199.86) --
	(264.97,199.98) --
	(266.10,200.32) --
	(266.13,200.33) --
	(266.44,200.43) --
	(267.28,200.69) --
	(267.71,200.79) --
	(268.38,200.99) --
	(269.11,201.16) --
	(269.74,201.38) --
	(269.87,201.40) --
	(271.02,201.80) --
	(271.10,201.83) --
	(271.68,202.00) --
	(272.40,202.28) --
	(272.79,202.41) --
	(273.63,202.69) --
	(274.52,202.98) --
	(275.28,203.20) --
	(275.59,203.27) --
	(276.13,203.43) --
	(276.31,203.47) --
	(276.77,203.63) --
	(277.30,203.79) --
	(277.98,204.00) --
	(278.04,204.00) --
	(278.45,204.12) --
	(279.17,204.40) --
	(279.30,204.44) --
	(279.52,204.50) --
	(279.66,204.55) --
	(279.69,204.56) --
	(280.15,204.72) --
	(280.53,204.89) --
	(280.72,204.98) --
	(281.20,205.18) --
	(281.36,205.23) --
	(281.45,205.24) --
	(282.80,205.79) --
	(283.05,205.89) --
	(283.89,206.20) --
	(283.98,206.21) --
	(284.07,206.25) --
	(284.62,206.45) --
	(284.86,206.51) --
	(285.03,206.59) --
	(285.13,206.64) --
	(286.00,206.89) --
	(286.57,207.08) --
	(286.59,207.10) --
	(287.25,207.30) --
	(287.81,207.46) --
	(288.54,207.69) --
	(288.59,207.71) --
	(289.22,207.96) --
	(289.76,208.14) --
	(290.76,208.49) --
	(290.86,208.52) --
	(291.20,208.64) --
	(291.32,208.69) --
	(291.81,208.87) --
	(291.88,208.91) --
	(292.21,209.05) --
	(292.36,209.11) --
	(292.42,209.13) --
	(292.49,209.15) --
	(292.51,209.16) --
	(292.62,209.22) --
	(293.21,209.38) --
	(293.91,209.64) --
	(295.25,210.12) --
	(295.45,210.17) --
	(295.51,210.19) --
	(295.93,210.34) --
	(295.99,210.37) --
	(296.10,210.42) --
	(296.47,210.53) --
	(296.65,210.60) --
	(296.99,210.70) --
	(297.01,210.71) --
	(297.04,210.72) --
	(297.05,210.72) --
	(297.09,210.73) --
	(297.13,210.74) --
	(297.22,210.77) --
	(297.31,210.80) --
	(297.31,210.80) --
	(297.71,210.96) --
	(297.96,211.07) --
	(298.34,211.23) --
	(298.57,211.33) --
	(299.04,211.48) --
	(300.58,212.07) --
	(300.69,212.11) --
	(301.32,212.32) --
	(302.10,212.65) --
	(302.23,212.71) --
	(302.50,212.86) --
	(302.82,212.96) --
	(302.98,213.02) --
	(303.49,213.24) --
	(303.87,213.38) --
	(304.04,213.46) --
	(304.05,213.46) --
	(304.17,213.52) --
	(304.21,213.54) --
	(304.88,213.80) --
	(305.44,214.06) --
	(305.52,214.08) --
	(305.61,214.12) --
	(305.86,214.20) --
	(306.19,214.32) --
	(306.32,214.36) --
	(307.91,214.96) --
	(307.96,214.99) --
	(308.47,215.19) --
	(308.57,215.23) --
	(309.22,215.49) --
	(309.44,215.58) --
	(309.51,215.60) --
	(309.67,215.66) --
	(310.00,215.78) --
	(310.25,215.88) --
	(310.76,216.10) --
	(310.92,216.15) --
	(311.04,216.19) --
	(311.16,216.24) --
	(311.41,216.35) --
	(311.41,216.36) --
	(311.44,216.37) --
	(311.54,216.40) --
	(311.68,216.46) --
	(311.68,216.46) --
	(311.88,216.53) --
	(312.22,216.64) --
	(312.92,216.90) --
	(313.12,216.98) --
	(313.27,217.05) --
	(313.53,217.12) --
	(313.83,217.24) --
	(314.24,217.40) --
	(314.31,217.42) --
	(314.45,217.49) --
	(314.63,217.59) --
	(315.00,217.69) --
	(315.21,217.81) --
	(315.23,217.82) --
	(315.37,217.86) --
	(315.48,217.91) --
	(315.51,217.92) --
	(315.62,217.95) --
	(315.70,217.98) --
	(315.79,218.02) --
	(316.13,218.15) --
	(316.22,218.18) --
	(316.38,218.25) --
	(316.61,218.33) --
	(316.93,218.48) --
	(317.07,218.54) --
	(317.20,218.60) --
	(317.36,218.68) --
	(317.75,218.86) --
	(317.76,218.87) --
	(317.79,218.88) --
	(317.86,218.90) --
	(317.89,218.91) --
	(317.98,218.95) --
	(318.43,219.10) --
	(318.60,219.15) --
	(318.85,219.23) --
	(319.11,219.34) --
	(319.21,219.36) --
	(319.52,219.46) --
	(319.81,219.61) --
	(320.44,219.87) --
	(321.85,220.46) --
	(321.91,220.48) --
	(322.04,220.54) --
	(322.30,220.65) --
	(322.60,220.80) --
	(322.75,220.85) --
	(323.02,220.98) --
	(323.10,221.01) --
	(323.28,221.10) --
	(324.05,221.40) --
	(324.42,221.57) --
	(324.61,221.63) --
	(324.72,221.66) --
	(325.08,221.77) --
	(325.99,222.17) --
	(326.58,222.47) --
	(326.72,222.50) --
	(327.61,222.87) --
	(327.91,223.02) --
	(328.67,223.32) --
	(328.94,223.47) --
	(329.07,223.53) --
	(329.52,223.68) --
	(331.17,224.41) --
	(332.00,224.77) --
	(332.27,224.86) --
	(333.07,225.24) --
	(333.19,225.28) --
	(333.76,225.52) --
	(335.51,226.18) --
	(336.07,226.42) --
	(336.75,226.75) --
	(336.88,226.82) --
	(337.54,227.09) --
	(337.92,227.27) --
	(338.29,227.44) --
	(338.32,227.46) --
	(338.37,227.49) --
	(338.79,227.66) --
	(338.87,227.69) --
	(339.41,227.95) --
	(339.54,228.00) --
	(340.26,228.35) --
	(340.28,228.35) --
	(340.36,228.38) --
	(340.58,228.46) --
	(340.90,228.61) --
	(341.07,228.67) --
	(341.38,228.80) --
	(341.48,228.85) --
	(341.53,228.88) --
	(341.57,228.90) --
	(342.00,229.08) --
	(342.06,229.11) --
	(342.60,229.32) --
	(342.92,229.41) --
	(344.39,230.06) --
	(344.81,230.26) --
	(345.03,230.35) --
	(345.51,230.53) --
	(345.56,230.55) --
	(346.27,230.84) --
	(348.09,231.63) --
	(348.19,231.66) --
	(348.32,231.71) --
	(348.52,231.79) --
	(348.65,231.84) --
	(350.04,232.48) --
	(350.29,232.58) --
	(350.39,232.61) --
	(351.49,233.03) --
	(354.19,234.20) --
	(354.41,234.28) --
	(354.96,234.53) --
	(354.99,234.54) --
	(355.28,234.70) --
	(355.84,234.97) --
	(356.71,235.37) --
	(356.75,235.39) --
	(357.50,235.70) --
	(357.53,235.71) --
	(358.04,235.90) --
	(358.62,236.15) --
	(359.31,236.46) --
	(359.62,236.61) --
	(359.75,236.69) --
	(359.85,236.75) --
	(360.61,237.12) --
	(360.64,237.13) --
	(360.71,237.18) --
	(361.91,237.72) --
	(362.60,238.04) --
	(364.24,238.88) --
	(364.46,239.00) --
	(365.70,239.62) --
	(366.29,239.93) --
	(370.53,241.73) --
	(371.02,241.94) --
	(372.06,242.43) --
	(372.07,242.43) --
	(372.38,242.60) --
	(372.60,242.71) --
	(372.90,242.85) --
	(373.16,243.00) --
	(373.53,243.19) --
	(373.80,243.33) --
	(374.46,243.64) --
	(374.81,243.78) --
	(375.77,244.22) --
	(376.36,244.50) --
	(376.51,244.59) --
	(376.93,244.79) --
	(378.03,245.25) --
	(379.36,245.80) --
	(379.36,245.80) --
	(380.09,246.17) --
	(380.12,246.18) --
	(381.03,246.59) --
	(381.47,246.80) --
	(381.75,246.94) --
	(382.36,247.23) --
	(382.44,247.28) --
	(382.57,247.34) --
	(383.02,247.55) --
	(385.04,248.44) --
	(385.16,248.51) --
	(388.87,250.32) --
	(393.56,252.71) --
	(395.71,253.79) --
	(403.26,257.56) --
	(404.02,257.95) --
	(404.02,188.48) --
	(403.26,188.33) --
	(395.71,186.92) --
	(393.56,186.56) --
	(388.87,185.69) --
	(385.16,184.82) --
	(385.04,184.80) --
	(383.02,184.39) --
	(382.57,184.31) --
	(382.44,184.30) --
	(382.36,184.29) --
	(381.75,184.16) --
	(381.47,184.11) --
	(381.03,184.04) --
	(380.12,183.81) --
	(380.09,183.80) --
	(379.36,183.64) --
	(379.36,183.64) --
	(378.03,183.40) --
	(376.93,183.15) --
	(376.51,183.06) --
	(376.36,183.05) --
	(375.77,182.97) --
	(374.81,182.77) --
	(374.46,182.70) --
	(373.80,182.52) --
	(373.53,182.49) --
	(373.16,182.43) --
	(372.90,182.36) --
	(372.60,182.28) --
	(372.38,182.23) --
	(372.07,182.19) --
	(372.06,182.19) --
	(371.02,182.00) --
	(370.53,181.89) --
	(366.29,181.14) --
	(365.70,181.03) --
	(364.46,180.67) --
	(364.24,180.65) --
	(362.60,180.37) --
	(361.91,180.21) --
	(360.71,179.92) --
	(360.64,179.91) --
	(360.61,179.91) --
	(359.85,179.80) --
	(359.75,179.78) --
	(359.62,179.75) --
	(359.31,179.68) --
	(358.62,179.58) --
	(358.04,179.52) --
	(357.53,179.41) --
	(357.50,179.40) --
	(356.75,179.23) --
	(356.71,179.22) --
	(355.84,179.05) --
	(355.28,178.90) --
	(354.99,178.79) --
	(354.96,178.78) --
	(354.41,178.64) --
	(354.19,178.60) --
	(351.49,177.92) --
	(350.39,177.61) --
	(350.29,177.59) --
	(350.04,177.51) --
	(348.65,177.21) --
	(348.52,177.18) --
	(348.32,177.15) --
	(348.19,177.12) --
	(348.09,177.09) --
	(346.27,176.71) --
	(345.56,176.53) --
	(345.51,176.51) --
	(345.03,176.39) --
	(344.81,176.34) --
	(344.39,176.25) --
	(342.92,175.94) --
	(342.60,175.88) --
	(342.06,175.75) --
	(342.00,175.73) --
	(341.57,175.65) --
	(341.53,175.64) --
	(341.48,175.64) --
	(341.38,175.62) --
	(341.07,175.54) --
	(340.90,175.51) --
	(340.58,175.45) --
	(340.36,175.41) --
	(340.28,175.40) --
	(340.26,175.39) --
	(339.54,175.25) --
	(339.41,175.22) --
	(338.87,175.11) --
	(338.79,175.10) --
	(338.37,175.01) --
	(338.32,174.99) --
	(338.29,174.98) --
	(337.92,174.87) --
	(337.54,174.80) --
	(336.88,174.67) --
	(336.75,174.64) --
	(336.07,174.44) --
	(335.51,174.32) --
	(333.76,173.87) --
	(333.19,173.78) --
	(333.07,173.75) --
	(332.27,173.58) --
	(332.00,173.51) --
	(331.17,173.30) --
	(329.52,172.87) --
	(329.07,172.77) --
	(328.94,172.73) --
	(328.67,172.67) --
	(327.91,172.46) --
	(327.61,172.39) --
	(326.72,172.17) --
	(326.58,172.14) --
	(325.99,172.00) --
	(325.08,171.78) --
	(324.72,171.70) --
	(324.61,171.67) --
	(324.42,171.61) --
	(324.05,171.49) --
	(323.28,171.35) --
	(323.10,171.31) --
	(323.02,171.29) --
	(322.75,171.18) --
	(322.60,171.15) --
	(322.30,171.10) --
	(322.04,171.03) --
	(321.91,171.01) --
	(321.85,171.00) --
	(320.44,170.69) --
	(319.81,170.53) --
	(319.52,170.45) --
	(319.21,170.37) --
	(319.11,170.34) --
	(318.85,170.26) --
	(318.60,170.22) --
	(318.43,170.18) --
	(317.98,170.05) --
	(317.89,170.02) --
	(317.86,170.01) --
	(317.79,169.99) --
	(317.76,169.99) --
	(317.75,169.99) --
	(317.36,169.90) --
	(317.20,169.85) --
	(317.07,169.82) --
	(316.93,169.79) --
	(316.61,169.68) --
	(316.38,169.60) --
	(316.22,169.57) --
	(316.13,169.55) --
	(315.79,169.46) --
	(315.70,169.44) --
	(315.62,169.41) --
	(315.51,169.37) --
	(315.48,169.36) --
	(315.37,169.32) --
	(315.23,169.28) --
	(315.21,169.27) --
	(315.00,169.20) --
	(314.63,169.05) --
	(314.45,168.98) --
	(314.31,168.94) --
	(314.24,168.90) --
	(313.83,168.77) --
	(313.53,168.69) --
	(313.27,168.63) --
	(313.12,168.61) --
	(312.92,168.58) --
	(312.22,168.40) --
	(311.88,168.29) --
	(311.68,168.25) --
	(311.68,168.25) --
	(311.54,168.22) --
	(311.44,168.18) --
	(311.41,168.17) --
	(311.41,168.17) --
	(311.16,168.09) --
	(311.04,168.07) --
	(310.92,168.05) --
	(310.76,168.02) --
	(310.25,167.88) --
	(310.00,167.83) --
	(309.67,167.75) --
	(309.51,167.72) --
	(309.44,167.71) --
	(309.22,167.66) --
	(308.57,167.46) --
	(308.47,167.44) --
	(307.96,167.34) --
	(307.91,167.34) --
	(306.32,166.93) --
	(306.19,166.90) --
	(305.86,166.81) --
	(305.61,166.74) --
	(305.52,166.71) --
	(305.44,166.69) --
	(304.88,166.51) --
	(304.21,166.30) --
	(304.17,166.30) --
	(304.05,166.26) --
	(304.04,166.26) --
	(303.87,166.21) --
	(303.49,166.10) --
	(302.98,165.93) --
	(302.82,165.89) --
	(302.50,165.78) --
	(302.23,165.67) --
	(302.10,165.62) --
	(301.32,165.40) --
	(300.69,165.24) --
	(300.58,165.20) --
	(299.04,164.69) --
	(298.57,164.54) --
	(298.34,164.48) --
	(297.96,164.38) --
	(297.71,164.30) --
	(297.31,164.13) --
	(297.31,164.13) --
	(297.22,164.10) --
	(297.13,164.08) --
	(297.09,164.07) --
	(297.05,164.05) --
	(297.04,164.05) --
	(297.01,164.04) --
	(296.99,164.03) --
	(296.65,163.91) --
	(296.47,163.85) --
	(296.10,163.73) --
	(295.99,163.69) --
	(295.93,163.67) --
	(295.51,163.54) --
	(295.45,163.52) --
	(295.25,163.45) --
	(293.91,163.09) --
	(293.21,162.89) --
	(292.62,162.68) --
	(292.51,162.66) --
	(292.49,162.65) --
	(292.42,162.63) --
	(292.36,162.61) --
	(292.21,162.56) --
	(291.88,162.48) --
	(291.81,162.46) --
	(291.32,162.32) --
	(291.20,162.27) --
	(290.86,162.16) --
	(290.76,162.12) --
	(289.76,161.73) --
	(289.22,161.57) --
	(288.59,161.34) --
	(288.54,161.32) --
	(287.81,161.08) --
	(287.25,160.92) --
	(286.59,160.70) --
	(286.57,160.70) --
	(286.00,160.52) --
	(285.13,160.28) --
	(285.03,160.25) --
	(284.86,160.19) --
	(284.62,160.11) --
	(284.07,159.94) --
	(283.98,159.91) --
	(283.89,159.89) --
	(283.05,159.64) --
	(282.80,159.57) --
	(281.45,159.12) --
	(281.36,159.08) --
	(281.20,159.03) --
	(280.72,158.84) --
	(280.53,158.80) --
	(280.15,158.71) --
	(279.69,158.55) --
	(279.66,158.54) --
	(279.52,158.48) --
	(279.30,158.40) --
	(279.17,158.36) --
	(278.45,158.13) --
	(278.04,157.99) --
	(277.98,157.96) --
	(277.30,157.71) --
	(276.77,157.53) --
	(276.31,157.40) --
	(276.13,157.34) --
	(275.59,157.16) --
	(275.28,157.07) --
	(274.52,156.81) --
	(273.63,156.52) --
	(272.79,156.25) --
	(272.40,156.11) --
	(271.68,155.88) --
	(271.10,155.68) --
	(271.02,155.65) --
	(269.87,155.20) --
	(269.74,155.17) --
	(269.11,154.96) --
	(268.38,154.71) --
	(267.71,154.49) --
	(267.28,154.34) --
	(266.44,154.07) --
	(266.13,153.97) --
	(266.10,153.96) --
	(264.97,153.64) --
	(264.53,153.52) --
	(264.50,153.51) --
	(264.29,153.45) --
	(262.99,153.00) --
	(262.48,152.85) --
	(262.19,152.74) --
	(261.97,152.68) --
	(260.80,152.32) --
	(260.14,152.06) --
	(260.03,152.03) --
	(258.61,151.54) --
	(256.32,150.72) --
	(255.02,150.28) --
	(254.91,150.24) --
	(253.36,149.63) --
	(253.12,149.55) --
	(252.98,149.50) --
	(252.52,149.32) --
	(251.88,149.06) --
	(251.61,148.94) --
	(251.21,148.84) --
	(250.71,148.65) --
	(250.14,148.41) --
	(249.91,148.31) --
	(249.15,148.07) --
	(248.59,147.93) --
	(248.00,147.70) --
	(247.15,147.40) --
	(247.14,147.39) --
	(245.90,146.93) --
	(244.87,146.49) --
	(243.82,146.09) --
	(242.87,145.70) --
	(240.76,144.79) --
	(239.50,144.28) --
	(238.44,143.86) --
	(236.50,143.12) --
	(233.03,141.67) --
	(232.60,141.48) --
	(232.55,141.46) --
	(232.29,141.37) --
	(231.54,141.09) --
	(230.53,140.70) --
	(229.80,140.39) --
	(228.22,139.75) --
	(226.89,139.17) --
	(226.38,138.96) --
	(224.51,138.27) --
	(219.87,136.46) --
	(219.39,136.25) --
	(214.88,134.34) --
	(213.28,133.63) --
	(213.22,133.61) --
	(213.02,133.50) --
	(211.30,132.78) --
	(205.70,130.38) --
	(204.08,129.58) --
	(203.92,129.52) --
	(203.24,129.24) --
	(203.15,129.19) --
	(202.42,128.83) --
	(201.70,128.52) --
	(200.09,127.78) --
	(199.79,127.65) --
	(195.05,125.74) --
	(195.04,125.74) --
	(185.12,121.11) --
	(177.68,117.77) --
	(175.08,116.58) --
	(164.50,111.49) --
	(163.56,111.05) --
	(160.15,109.37) --
	(160.01,109.30) --
	(124.94, 91.80) --
	(116.56, 87.52) --
	(113.94, 86.20) --
	(110.01, 84.11) --
	(104.50, 81.25) --
	( 97.84, 77.97) --
	( 94.41, 76.21) --
	( 88.50, 73.15) --
	( 79.64, 68.59) --
	( 65.62, 61.16) --
	( 65.48, 61.09) --
	( 61.01, 58.86) --
	( 52.97, 54.68) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 52.97, 34.03) -- ( 52.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 61.01, 34.03) -- ( 61.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 65.48, 34.03) -- ( 65.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 65.62, 34.03) -- ( 65.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.64, 34.03) -- ( 79.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 88.50, 34.03) -- ( 88.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 94.41, 34.03) -- ( 94.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 97.84, 34.03) -- ( 97.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (104.50, 34.03) -- (104.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (110.01, 34.03) -- (110.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (113.94, 34.03) -- (113.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (116.56, 34.03) -- (116.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (124.94, 34.03) -- (124.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (160.01, 34.03) -- (160.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (160.15, 34.03) -- (160.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (163.56, 34.03) -- (163.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (164.50, 34.03) -- (164.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (175.08, 34.03) -- (175.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (177.68, 34.03) -- (177.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (185.12, 34.03) -- (185.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.04, 34.03) -- (195.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.05, 34.03) -- (195.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (199.79, 34.03) -- (199.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (200.09, 34.03) -- (200.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (201.70, 34.03) -- (201.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (202.42, 34.03) -- (202.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.15, 34.03) -- (203.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.24, 34.03) -- (203.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.92, 34.03) -- (203.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (204.08, 34.03) -- (204.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.70, 34.03) -- (205.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (211.30, 34.03) -- (211.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.02, 34.03) -- (213.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.22, 34.03) -- (213.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.28, 34.03) -- (213.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.88, 34.03) -- (214.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.39, 34.03) -- (219.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.87, 34.03) -- (219.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (224.51, 34.03) -- (224.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.38, 34.03) -- (226.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.89, 34.03) -- (226.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (228.22, 34.03) -- (228.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.80, 34.03) -- (229.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (230.53, 34.03) -- (230.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.54, 34.03) -- (231.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.29, 34.03) -- (232.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.55, 34.03) -- (232.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.60, 34.03) -- (232.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (233.03, 34.03) -- (233.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.50, 34.03) -- (236.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (238.44, 34.03) -- (238.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.49, 34.03) -- (239.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.76, 34.03) -- (240.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.87, 34.03) -- (242.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (243.82, 34.03) -- (243.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (244.87, 34.03) -- (244.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (245.90, 34.03) -- (245.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.14, 34.03) -- (247.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.15, 34.03) -- (247.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.00, 34.03) -- (248.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.59, 34.03) -- (248.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.15, 34.03) -- (249.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.91, 34.03) -- (249.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.14, 34.03) -- (250.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.71, 34.03) -- (250.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.21, 34.03) -- (251.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.61, 34.03) -- (251.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.88, 34.03) -- (251.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.52, 34.03) -- (252.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.98, 34.03) -- (252.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.12, 34.03) -- (253.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.36, 34.03) -- (253.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (254.91, 34.03) -- (254.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (255.02, 34.03) -- (255.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.32, 34.03) -- (256.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.61, 34.03) -- (258.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.03, 34.03) -- (260.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.14, 34.03) -- (260.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.80, 34.03) -- (260.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.97, 34.03) -- (261.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.19, 34.03) -- (262.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.47, 34.03) -- (262.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.99, 34.03) -- (262.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.29, 34.03) -- (264.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.50, 34.03) -- (264.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.53, 34.03) -- (264.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.97, 34.03) -- (264.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.10, 34.03) -- (266.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.13, 34.03) -- (266.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.44, 34.03) -- (266.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.28, 34.03) -- (267.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.71, 34.03) -- (267.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.38, 34.03) -- (268.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.11, 34.03) -- (269.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.74, 34.03) -- (269.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.87, 34.03) -- (269.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.02, 34.03) -- (271.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.10, 34.03) -- (271.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.68, 34.03) -- (271.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.40, 34.03) -- (272.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.79, 34.03) -- (272.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.63, 34.03) -- (273.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.52, 34.03) -- (274.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.28, 34.03) -- (275.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.59, 34.03) -- (275.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.13, 34.03) -- (276.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.31, 34.03) -- (276.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.77, 34.03) -- (276.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.30, 34.03) -- (277.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.98, 34.03) -- (277.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.04, 34.03) -- (278.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.45, 34.03) -- (278.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.17, 34.03) -- (279.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.30, 34.03) -- (279.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.52, 34.03) -- (279.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.66, 34.03) -- (279.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.69, 34.03) -- (279.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.15, 34.03) -- (280.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.53, 34.03) -- (280.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.72, 34.03) -- (280.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.20, 34.03) -- (281.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.36, 34.03) -- (281.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.45, 34.03) -- (281.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.79, 34.03) -- (282.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.05, 34.03) -- (283.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.89, 34.03) -- (283.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.98, 34.03) -- (283.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.07, 34.03) -- (284.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.63, 34.03) -- (284.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.86, 34.03) -- (284.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.03, 34.03) -- (285.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.13, 34.03) -- (285.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.00, 34.03) -- (286.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.57, 34.03) -- (286.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.59, 34.03) -- (286.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.25, 34.03) -- (287.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.81, 34.03) -- (287.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.54, 34.03) -- (288.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.59, 34.03) -- (288.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.22, 34.03) -- (289.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.76, 34.03) -- (289.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.76, 34.03) -- (290.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.86, 34.03) -- (290.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.20, 34.03) -- (291.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.32, 34.03) -- (291.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.81, 34.03) -- (291.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.88, 34.03) -- (291.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.21, 34.03) -- (292.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.36, 34.03) -- (292.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.42, 34.03) -- (292.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.49, 34.03) -- (292.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.51, 34.03) -- (292.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.62, 34.03) -- (292.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.21, 34.03) -- (293.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.91, 34.03) -- (293.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.25, 34.03) -- (295.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.45, 34.03) -- (295.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.51, 34.03) -- (295.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.93, 34.03) -- (295.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.99, 34.03) -- (295.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.10, 34.03) -- (296.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.47, 34.03) -- (296.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.65, 34.03) -- (296.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.99, 34.03) -- (296.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.01, 34.03) -- (297.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.04, 34.03) -- (297.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.05, 34.03) -- (297.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.08, 34.03) -- (297.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.12, 34.03) -- (297.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.22, 34.03) -- (297.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.31, 34.03) -- (297.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.31, 34.03) -- (297.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.71, 34.03) -- (297.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.96, 34.03) -- (297.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.34, 34.03) -- (298.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.57, 34.03) -- (298.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.04, 34.03) -- (299.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.58, 34.03) -- (300.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.69, 34.03) -- (300.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.32, 34.03) -- (301.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.10, 34.03) -- (302.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.23, 34.03) -- (302.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.50, 34.03) -- (302.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.82, 34.03) -- (302.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.98, 34.03) -- (302.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.49, 34.03) -- (303.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.87, 34.03) -- (303.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.04, 34.03) -- (304.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.05, 34.03) -- (304.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.17, 34.03) -- (304.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.21, 34.03) -- (304.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.88, 34.03) -- (304.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.44, 34.03) -- (305.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.52, 34.03) -- (305.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.61, 34.03) -- (305.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.86, 34.03) -- (305.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.19, 34.03) -- (306.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.32, 34.03) -- (306.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.91, 34.03) -- (307.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.96, 34.03) -- (307.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.47, 34.03) -- (308.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.57, 34.03) -- (308.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.22, 34.03) -- (309.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.44, 34.03) -- (309.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.51, 34.03) -- (309.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.67, 34.03) -- (309.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.00, 34.03) -- (310.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.25, 34.03) -- (310.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.76, 34.03) -- (310.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.92, 34.03) -- (310.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.04, 34.03) -- (311.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.16, 34.03) -- (311.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.41, 34.03) -- (311.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.41, 34.03) -- (311.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.44, 34.03) -- (311.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.54, 34.03) -- (311.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.68, 34.03) -- (311.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.68, 34.03) -- (311.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.88, 34.03) -- (311.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.22, 34.03) -- (312.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.92, 34.03) -- (312.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.12, 34.03) -- (313.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.27, 34.03) -- (313.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.53, 34.03) -- (313.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.83, 34.03) -- (313.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.24, 34.03) -- (314.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.31, 34.03) -- (314.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.45, 34.03) -- (314.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.63, 34.03) -- (314.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.00, 34.03) -- (315.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.21, 34.03) -- (315.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.23, 34.03) -- (315.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.37, 34.03) -- (315.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.48, 34.03) -- (315.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.51, 34.03) -- (315.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.62, 34.03) -- (315.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.70, 34.03) -- (315.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.79, 34.03) -- (315.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.13, 34.03) -- (316.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.22, 34.03) -- (316.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.38, 34.03) -- (316.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.61, 34.03) -- (316.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.93, 34.03) -- (316.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.07, 34.03) -- (317.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.20, 34.03) -- (317.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.36, 34.03) -- (317.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.75, 34.03) -- (317.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.76, 34.03) -- (317.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.79, 34.03) -- (317.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.86, 34.03) -- (317.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.89, 34.03) -- (317.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.98, 34.03) -- (317.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.43, 34.03) -- (318.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.60, 34.03) -- (318.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.85, 34.03) -- (318.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.11, 34.03) -- (319.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.21, 34.03) -- (319.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.52, 34.03) -- (319.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.81, 34.03) -- (319.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.44, 34.03) -- (320.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.85, 34.03) -- (321.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.91, 34.03) -- (321.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.04, 34.03) -- (322.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.30, 34.03) -- (322.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.60, 34.03) -- (322.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.75, 34.03) -- (322.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.02, 34.03) -- (323.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.10, 34.03) -- (323.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.28, 34.03) -- (323.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.05, 34.03) -- (324.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.42, 34.03) -- (324.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.61, 34.03) -- (324.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.72, 34.03) -- (324.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.08, 34.03) -- (325.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.99, 34.03) -- (325.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.58, 34.03) -- (326.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.72, 34.03) -- (326.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.61, 34.03) -- (327.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.91, 34.03) -- (327.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.67, 34.03) -- (328.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.94, 34.03) -- (328.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.07, 34.03) -- (329.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.52, 34.03) -- (329.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.17, 34.03) -- (331.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (332.00, 34.03) -- (332.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (332.27, 34.03) -- (332.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.07, 34.03) -- (333.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.19, 34.03) -- (333.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.76, 34.03) -- (333.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.51, 34.03) -- (335.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.07, 34.03) -- (336.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.75, 34.03) -- (336.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.88, 34.03) -- (336.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.54, 34.03) -- (337.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.92, 34.03) -- (337.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.29, 34.03) -- (338.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.32, 34.03) -- (338.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.37, 34.03) -- (338.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.79, 34.03) -- (338.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.87, 34.03) -- (338.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.41, 34.03) -- (339.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.54, 34.03) -- (339.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.26, 34.03) -- (340.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.28, 34.03) -- (340.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.36, 34.03) -- (340.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.58, 34.03) -- (340.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.90, 34.03) -- (340.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.07, 34.03) -- (341.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.38, 34.03) -- (341.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.48, 34.03) -- (341.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.54, 34.03) -- (341.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.57, 34.03) -- (341.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.00, 34.03) -- (342.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.06, 34.03) -- (342.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.60, 34.03) -- (342.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.92, 34.03) -- (342.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.39, 34.03) -- (344.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.81, 34.03) -- (344.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.03, 34.03) -- (345.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.51, 34.03) -- (345.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.56, 34.03) -- (345.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.27, 34.03) -- (346.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.09, 34.03) -- (348.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.19, 34.03) -- (348.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.32, 34.03) -- (348.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.52, 34.03) -- (348.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.65, 34.03) -- (348.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.04, 34.03) -- (350.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.29, 34.03) -- (350.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.39, 34.03) -- (350.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.49, 34.03) -- (351.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.19, 34.03) -- (354.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.41, 34.03) -- (354.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.96, 34.03) -- (354.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.99, 34.03) -- (354.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.28, 34.03) -- (355.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.84, 34.03) -- (355.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.71, 34.03) -- (356.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.75, 34.03) -- (356.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.50, 34.03) -- (357.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.53, 34.03) -- (357.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.04, 34.03) -- (358.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.62, 34.03) -- (358.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.31, 34.03) -- (359.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.62, 34.03) -- (359.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.75, 34.03) -- (359.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.85, 34.03) -- (359.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.61, 34.03) -- (360.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.64, 34.03) -- (360.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.71, 34.03) -- (360.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (361.91, 34.03) -- (361.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.60, 34.03) -- (362.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.24, 34.03) -- (364.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.46, 34.03) -- (364.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.70, 34.03) -- (365.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.29, 34.03) -- (366.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (370.53, 34.03) -- (370.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.02, 34.03) -- (371.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.06, 34.03) -- (372.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.07, 34.03) -- (372.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.38, 34.03) -- (372.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.60, 34.03) -- (372.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.90, 34.03) -- (372.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.16, 34.03) -- (373.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.53, 34.03) -- (373.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.80, 34.03) -- (373.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.46, 34.03) -- (374.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.81, 34.03) -- (374.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (375.77, 34.03) -- (375.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.36, 34.03) -- (376.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.51, 34.03) -- (376.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.93, 34.03) -- (376.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.03, 34.03) -- (378.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (379.36, 34.03) -- (379.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (379.36, 34.03) -- (379.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (380.09, 34.03) -- (380.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (380.12, 34.03) -- (380.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.03, 34.03) -- (381.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.47, 34.03) -- (381.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.75, 34.03) -- (381.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.36, 34.03) -- (382.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.44, 34.03) -- (382.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.57, 34.03) -- (382.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.02, 34.03) -- (383.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.04, 34.03) -- (385.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.16, 34.03) -- (385.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.87, 34.03) -- (388.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.56, 34.03) -- (393.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (395.71, 34.03) -- (395.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (403.26, 34.03) -- (403.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.02, 34.03) -- (404.02, 41.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	( 35.42,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31, 44.90) {-5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,128.11) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,211.32) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 85.59, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (136.07, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (186.55, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (237.03, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (287.51, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (337.99, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (388.47, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (228.50,  4.82) {Ln(Min. City Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}