### 2014-10-09 LaTeX/mCapSimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-10-06 20:10:12
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.97,105.17) --
	( 60.61,107.79) --
	( 64.85,109.25) --
	( 64.98,109.30) --
	( 78.30,113.88) --
	( 86.71,116.78) --
	( 92.34,118.71) --
	( 95.59,119.83) --
	(101.92,122.01) --
	(110.88,125.09) --
	(154.64,140.15) --
	(158.01,141.31) --
	(171.43,145.93) --
	(178.49,148.35) --
	(187.92,151.60) --
	(192.71,153.25) --
	(195.62,154.25) --
	(195.71,154.28) --
	(196.35,154.50) --
	(198.04,155.08) --
	(199.87,155.71) --
	(204.99,157.47) --
	(205.19,157.54) --
	(209.65,159.08) --
	(211.50,159.71) --
	(215.87,161.22) --
	(217.33,161.72) --
	(217.43,161.75) --
	(217.68,161.84) --
	(218.17,162.01) --
	(220.17,162.70) --
	(220.33,162.75) --
	(220.93,162.96) --
	(221.63,163.20) --
	(222.58,163.53) --
	(223.29,163.77) --
	(223.54,163.86) --
	(223.59,163.87) --
	(224.00,164.02) --
	(229.14,165.78) --
	(231.15,166.48) --
	(231.34,166.54) --
	(232.39,166.90) --
	(233.34,167.23) --
	(234.24,167.54) --
	(235.24,167.88) --
	(236.22,168.22) --
	(237.40,168.63) --
	(237.41,168.63) --
	(239.29,169.28) --
	(239.31,169.28) --
	(240.03,169.53) --
	(240.80,169.79) --
	(241.27,169.96) --
	(241.65,170.09) --
	(242.95,170.54) --
	(244.78,171.17) --
	(248.30,172.38) --
	(249.74,172.87) --
	(250.22,173.04) --
	(250.38,173.09) --
	(250.69,173.20) --
	(251.97,173.64) --
	(252.46,173.81) --
	(253.69,174.23) --
	(253.89,174.30) --
	(254.34,174.45) --
	(255.73,174.94) --
	(256.94,175.35) --
	(258.99,176.06) --
	(260.08,176.43) --
	(260.16,176.46) --
	(261.39,176.88) --
	(262.57,177.29) --
	(263.47,177.60) --
	(264.14,177.83) --
	(264.43,177.93) --
	(264.94,178.10) --
	(265.11,178.16) --
	(265.48,178.29) --
	(265.54,178.31) --
	(266.69,178.71) --
	(266.70,178.71) --
	(267.74,179.07) --
	(267.83,179.10) --
	(267.95,179.14) --
	(268.15,179.21) --
	(268.32,179.26) --
	(268.51,179.33) --
	(268.75,179.41) --
	(269.12,179.54) --
	(269.30,179.60) --
	(269.75,179.76) --
	(269.91,179.81) --
	(271.51,180.36) --
	(272.31,180.64) --
	(272.40,180.67) --
	(272.91,180.84) --
	(273.01,180.88) --
	(273.29,180.98) --
	(273.48,181.04) --
	(274.31,181.33) --
	(274.85,181.51) --
	(274.88,181.52) --
	(275.50,181.74) --
	(276.03,181.92) --
	(276.14,181.95) --
	(276.73,182.16) --
	(276.77,182.17) --
	(277.28,182.35) --
	(277.37,182.38) --
	(278.83,182.88) --
	(278.93,182.92) --
	(278.99,182.94) --
	(279.03,182.95) --
	(279.36,183.06) --
	(279.83,183.23) --
	(279.90,183.25) --
	(280.21,183.36) --
	(280.36,183.41) --
	(280.41,183.43) --
	(280.47,183.45) --
	(280.50,183.46) --
	(280.60,183.49) --
	(280.63,183.50) --
	(281.16,183.68) --
	(281.52,183.81) --
	(281.60,183.84) --
	(283.10,184.35) --
	(283.28,184.41) --
	(283.34,184.44) --
	(283.74,184.57) --
	(283.80,184.59) --
	(284.26,184.75) --
	(284.43,184.81) --
	(284.75,184.92) --
	(284.80,184.94) --
	(284.81,184.94) --
	(284.81,184.94) --
	(284.84,184.95) --
	(284.88,184.96) --
	(285.44,185.16) --
	(285.67,185.24) --
	(286.04,185.36) --
	(286.70,185.59) --
	(286.95,185.68) --
	(287.76,185.96) --
	(288.16,186.09) --
	(288.45,186.19) --
	(288.86,186.33) --
	(289.03,186.39) --
	(289.13,186.43) --
	(289.30,186.48) --
	(289.61,186.59) --
	(289.71,186.63) --
	(289.99,186.72) --
	(290.44,186.88) --
	(291.29,187.17) --
	(291.45,187.22) --
	(291.58,187.27) --
	(292.20,187.48) --
	(292.36,187.54) --
	(292.77,187.68) --
	(292.78,187.68) --
	(292.85,187.71) --
	(292.94,187.74) --
	(293.49,187.93) --
	(293.61,187.97) --
	(293.90,188.07) --
	(293.94,188.08) --
	(295.12,188.49) --
	(295.85,188.74) --
	(296.61,189.00) --
	(297.11,189.17) --
	(297.99,189.47) --
	(298.09,189.51) --
	(298.13,189.52) --
	(298.21,189.55) --
	(298.45,189.63) --
	(298.48,189.64) --
	(298.71,189.72) --
	(298.89,189.79) --
	(299.22,189.90) --
	(300.07,190.19) --
	(300.69,190.40) --
	(301.34,190.63) --
	(301.81,190.79) --
	(301.86,190.81) --
	(302.06,190.87) --
	(302.08,190.88) --
	(302.20,190.93) --
	(302.30,190.96) --
	(302.45,191.01) --
	(302.52,191.03) --
	(302.61,191.06) --
	(302.89,191.16) --
	(302.93,191.17) --
	(303.17,191.26) --
	(304.10,191.58) --
	(304.31,191.65) --
	(304.47,191.71) --
	(304.48,191.71) --
	(304.60,191.75) --
	(304.68,191.78) --
	(305.11,191.93) --
	(305.28,191.98) --
	(305.76,192.15) --
	(305.86,192.18) --
	(306.15,192.28) --
	(307.03,192.59) --
	(307.16,192.63) --
	(307.37,192.70) --
	(308.42,193.06) --
	(308.63,193.14) --
	(308.79,193.19) --
	(309.08,193.29) --
	(309.55,193.45) --
	(309.72,193.51) --
	(310.36,193.73) --
	(310.45,193.76) --
	(310.80,193.88) --
	(310.98,193.95) --
	(311.09,193.98) --
	(311.49,194.12) --
	(311.71,194.20) --
	(311.87,194.25) --
	(312.39,194.43) --
	(312.68,194.53) --
	(312.75,194.55) --
	(314.09,195.02) --
	(314.12,195.02) --
	(314.55,195.17) --
	(314.84,195.27) --
	(315.05,195.34) --
	(315.28,195.42) --
	(315.54,195.51) --
	(315.61,195.54) --
	(315.65,195.55) --
	(316.29,195.77) --
	(316.37,195.80) --
	(316.58,195.87) --
	(316.61,195.88) --
	(316.72,195.92) --
	(317.13,196.06) --
	(318.60,196.57) --
	(319.14,196.75) --
	(320.71,197.29) --
	(321.34,197.51) --
	(321.47,197.55) --
	(322.56,197.93) --
	(322.64,197.96) --
	(322.93,198.06) --
	(323.98,198.42) --
	(324.06,198.44) --
	(324.53,198.61) --
	(324.82,198.71) --
	(325.46,198.93) --
	(325.85,199.06) --
	(325.87,199.07) --
	(325.91,199.08) --
	(326.16,199.17) --
	(326.46,199.27) --
	(327.01,199.46) --
	(327.50,199.63) --
	(327.78,199.72) --
	(328.37,199.93) --
	(328.92,200.12) --
	(330.17,200.55) --
	(330.38,200.62) --
	(330.88,200.79) --
	(333.53,201.71) --
	(333.56,201.71) --
	(333.69,201.76) --
	(333.82,201.80) --
	(333.83,201.81) --
	(335.38,202.34) --
	(335.47,202.37) --
	(337.65,203.12) --
	(338.50,203.41) --
	(338.83,203.53) --
	(338.87,203.54) --
	(339.04,203.60) --
	(339.21,203.66) --
	(339.40,203.72) --
	(339.55,203.77) --
	(339.76,203.85) --
	(339.84,203.88) --
	(340.15,203.98) --
	(340.28,204.03) --
	(340.65,204.15) --
	(341.52,204.45) --
	(342.53,204.80) --
	(343.79,205.23) --
	(343.95,205.29) --
	(345.11,205.69) --
	(345.20,205.72) --
	(345.78,205.92) --
	(346.49,206.16) --
	(346.80,206.27) --
	(346.94,206.32) --
	(347.30,206.44) --
	(348.63,206.90) --
	(348.84,206.97) --
	(349.59,207.23) --
	(350.01,207.38) --
	(350.27,207.46) --
	(350.57,207.57) --
	(351.61,207.93) --
	(352.50,208.23) --
	(354.60,208.95) --
	(354.77,209.01) --
	(355.03,209.10) --
	(355.06,209.11) --
	(355.38,209.22) --
	(355.44,209.24) --
	(355.66,209.32) --
	(355.69,209.33) --
	(356.06,209.46) --
	(356.08,209.46) --
	(356.57,209.63) --
	(356.86,209.73) --
	(356.99,209.78) --
	(357.04,209.79) --
	(357.10,209.81) --
	(357.28,209.88) --
	(357.68,210.01) --
	(357.71,210.02) --
	(358.36,210.25) --
	(358.45,210.28) --
	(358.80,210.40) --
	(359.58,210.67) --
	(360.14,210.86) --
	(360.17,210.87) --
	(360.28,210.91) --
	(360.68,211.05) --
	(362.54,211.69) --
	(362.92,211.82) --
	(362.99,211.84) --
	(362.99,211.84) --
	(363.68,212.08) --
	(363.71,212.09) --
	(364.25,212.28) --
	(364.57,212.39) --
	(364.77,212.45) --
	(365.00,212.53) --
	(365.26,212.62) --
	(365.84,212.82) --
	(365.84,212.82) --
	(365.92,212.85) --
	(366.04,212.89) --
	(366.47,213.04) --
	(367.05,213.24) --
	(368.20,213.63) --
	(368.39,213.70) --
	(368.50,213.74) --
	(370.21,214.32) --
	(371.17,214.65) --
	(371.58,214.80) --
	(372.02,214.95) --
	(373.58,215.48) --
	(376.96,216.65) --
	(378.52,217.19) --
	(383.85,219.02) --
	(384.44,219.22) --
	(384.92,219.39) --
	(385.30,219.52) --
	(385.73,219.66) --
	(385.96,219.74) --
	(386.70,220.00) --
	(386.81,220.04) --
	(387.17,220.16) --
	(387.19,220.17) --
	(387.24,220.18) --
	(387.39,220.24) --
	(387.67,220.34) --
	(388.26,220.54) --
	(389.16,220.85) --
	(389.82,221.07) --
	(390.40,221.27) --
	(390.44,221.29) --
	(390.79,221.41) --
	(391.07,221.50) --
	(395.16,222.91) --
	(396.00,223.20) --
	(396.15,223.25) --
	(403.52,225.79) --
	(404.02,225.96);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 52.97,165.07) --
	( 60.61,166.03) --
	( 64.85,166.51) --
	( 64.98,166.54) --
	( 78.30,168.33) --
	( 86.71,169.61) --
	( 92.34,170.44) --
	( 95.59,171.03) --
	(101.92,171.81) --
	(110.88,172.89) --
	(154.64,180.05) --
	(158.01,180.45) --
	(171.43,182.73) --
	(178.49,184.08) --
	(187.92,185.93) --
	(192.71,186.77) --
	(195.62,187.27) --
	(195.71,187.28) --
	(196.35,187.38) --
	(198.04,187.65) --
	(199.87,188.05) --
	(204.99,189.01) --
	(205.19,189.03) --
	(209.65,190.06) --
	(211.50,190.52) --
	(215.87,191.59) --
	(217.33,191.90) --
	(217.43,191.93) --
	(217.68,191.97) --
	(218.17,192.07) --
	(220.17,192.44) --
	(220.33,192.49) --
	(220.93,192.59) --
	(221.63,192.75) --
	(222.58,192.98) --
	(223.29,193.13) --
	(223.54,193.18) --
	(223.59,193.19) --
	(224.00,193.30) --
	(229.14,194.56) --
	(231.15,194.97) --
	(231.34,194.99) --
	(232.39,195.37) --
	(233.34,195.60) --
	(234.24,195.69) --
	(235.24,195.85) --
	(236.22,196.18) --
	(237.40,196.49) --
	(237.41,196.49) --
	(239.29,197.05) --
	(239.31,197.06) --
	(240.03,197.23) --
	(240.80,197.39) --
	(241.27,197.46) --
	(241.65,197.55) --
	(242.95,197.83) --
	(244.78,198.38) --
	(248.30,199.35) --
	(249.74,199.77) --
	(250.22,199.91) --
	(250.38,199.94) --
	(250.69,200.06) --
	(251.97,200.42) --
	(252.46,200.53) --
	(253.69,200.79) --
	(253.89,200.86) --
	(254.34,200.96) --
	(255.73,201.43) --
	(256.94,201.82) --
	(258.99,202.54) --
	(260.08,202.88) --
	(260.16,202.90) --
	(261.39,203.25) --
	(262.57,203.64) --
	(263.47,203.91) --
	(264.14,204.11) --
	(264.43,204.22) --
	(264.94,204.38) --
	(265.11,204.45) --
	(265.48,204.57) --
	(265.54,204.59) --
	(266.69,205.01) --
	(266.70,205.01) --
	(267.74,205.35) --
	(267.83,205.37) --
	(267.95,205.45) --
	(268.15,205.52) --
	(268.32,205.56) --
	(268.51,205.63) --
	(268.75,205.74) --
	(269.12,205.89) --
	(269.30,205.97) --
	(269.75,206.12) --
	(269.91,206.17) --
	(271.51,206.69) --
	(272.31,207.01) --
	(272.40,207.06) --
	(272.91,207.29) --
	(273.01,207.33) --
	(273.29,207.43) --
	(273.48,207.49) --
	(274.31,207.72) --
	(274.85,207.89) --
	(274.88,207.90) --
	(275.50,208.11) --
	(276.03,208.30) --
	(276.14,208.34) --
	(276.73,208.55) --
	(276.77,208.57) --
	(277.28,208.75) --
	(277.37,208.78) --
	(278.83,209.26) --
	(278.93,209.30) --
	(278.99,209.33) --
	(279.03,209.34) --
	(279.36,209.43) --
	(279.83,209.54) --
	(279.90,209.57) --
	(280.21,209.70) --
	(280.36,209.75) --
	(280.41,209.78) --
	(280.47,209.80) --
	(280.50,209.81) --
	(280.60,209.85) --
	(280.63,209.85) --
	(281.16,210.05) --
	(281.52,210.20) --
	(281.60,210.24) --
	(283.10,210.77) --
	(283.28,210.82) --
	(283.34,210.83) --
	(283.74,210.94) --
	(283.80,210.97) --
	(284.26,211.19) --
	(284.43,211.24) --
	(284.75,211.33) --
	(284.80,211.34) --
	(284.81,211.34) --
	(284.81,211.34) --
	(284.84,211.35) --
	(284.88,211.37) --
	(285.44,211.54) --
	(285.67,211.63) --
	(286.04,211.78) --
	(286.70,212.02) --
	(286.95,212.12) --
	(287.76,212.51) --
	(288.16,212.63) --
	(288.45,212.70) --
	(288.86,212.81) --
	(289.03,212.87) --
	(289.13,212.93) --
	(289.30,212.99) --
	(289.61,213.12) --
	(289.71,213.16) --
	(289.99,213.26) --
	(290.44,213.43) --
	(291.29,213.72) --
	(291.45,213.79) --
	(291.58,213.85) --
	(292.20,214.08) --
	(292.36,214.13) --
	(292.77,214.30) --
	(292.78,214.30) --
	(292.85,214.32) --
	(292.94,214.36) --
	(293.49,214.59) --
	(293.61,214.64) --
	(293.90,214.77) --
	(293.94,214.79) --
	(295.12,215.21) --
	(295.85,215.51) --
	(296.61,215.83) --
	(297.11,216.04) --
	(297.99,216.40) --
	(298.09,216.45) --
	(298.13,216.47) --
	(298.21,216.50) --
	(298.45,216.60) --
	(298.48,216.61) --
	(298.71,216.72) --
	(298.89,216.80) --
	(299.22,216.90) --
	(300.07,217.25) --
	(300.69,217.51) --
	(301.34,217.77) --
	(301.81,217.95) --
	(301.86,217.97) --
	(302.06,218.06) --
	(302.08,218.06) --
	(302.20,218.12) --
	(302.30,218.16) --
	(302.45,218.22) --
	(302.52,218.25) --
	(302.61,218.29) --
	(302.89,218.40) --
	(302.93,218.42) --
	(303.17,218.52) --
	(304.10,218.84) --
	(304.31,218.92) --
	(304.47,218.97) --
	(304.48,218.97) --
	(304.60,219.01) --
	(304.68,219.04) --
	(305.11,219.23) --
	(305.28,219.30) --
	(305.76,219.46) --
	(305.86,219.49) --
	(306.15,219.58) --
	(307.03,219.96) --
	(307.16,220.02) --
	(307.37,220.07) --
	(308.42,220.47) --
	(308.63,220.55) --
	(308.79,220.61) --
	(309.08,220.73) --
	(309.55,220.92) --
	(309.72,220.98) --
	(310.36,221.22) --
	(310.45,221.25) --
	(310.80,221.37) --
	(310.98,221.46) --
	(311.09,221.48) --
	(311.49,221.66) --
	(311.71,221.78) --
	(311.87,221.84) --
	(312.39,222.06) --
	(312.68,222.19) --
	(312.75,222.21) --
	(314.09,222.81) --
	(314.12,222.82) --
	(314.55,223.00) --
	(314.84,223.11) --
	(315.05,223.18) --
	(315.28,223.25) --
	(315.54,223.34) --
	(315.61,223.36) --
	(315.65,223.37) --
	(316.29,223.65) --
	(316.37,223.69) --
	(316.58,223.78) --
	(316.61,223.79) --
	(316.72,223.82) --
	(317.13,223.99) --
	(318.60,224.64) --
	(319.14,224.92) --
	(320.71,225.62) --
	(321.34,225.89) --
	(321.47,225.94) --
	(322.56,226.40) --
	(322.64,226.44) --
	(322.93,226.57) --
	(323.98,227.03) --
	(324.06,227.07) --
	(324.53,227.28) --
	(324.82,227.40) --
	(325.46,227.67) --
	(325.85,227.84) --
	(325.87,227.86) --
	(325.91,227.88) --
	(326.16,227.97) --
	(326.46,228.10) --
	(327.01,228.35) --
	(327.50,228.52) --
	(327.78,228.67) --
	(328.37,228.96) --
	(328.92,229.20) --
	(330.17,229.74) --
	(330.38,229.85) --
	(330.88,230.08) --
	(333.53,231.22) --
	(333.56,231.23) --
	(333.69,231.29) --
	(333.82,231.35) --
	(333.83,231.35) --
	(335.38,232.05) --
	(335.47,232.09) --
	(337.65,233.08) --
	(338.50,233.45) --
	(338.83,233.59) --
	(338.87,233.61) --
	(339.04,233.68) --
	(339.21,233.75) --
	(339.40,233.82) --
	(339.55,233.88) --
	(339.76,233.97) --
	(339.84,234.00) --
	(340.15,234.13) --
	(340.28,234.18) --
	(340.65,234.38) --
	(341.52,234.84) --
	(342.53,235.30) --
	(343.79,235.93) --
	(343.95,236.02) --
	(345.11,236.58) --
	(345.20,236.63) --
	(345.78,236.90) --
	(346.49,237.20) --
	(346.80,237.34) --
	(346.94,237.42) --
	(347.30,237.59) --
	(348.63,238.22) --
	(348.84,238.32) --
	(349.59,238.65) --
	(350.01,238.84) --
	(350.27,238.96) --
	(350.57,239.10) --
	(351.61,239.64) --
	(352.50,240.05) --
	(354.60,241.08) --
	(354.77,241.16) --
	(355.03,241.32) --
	(355.06,241.34) --
	(355.38,241.48) --
	(355.44,241.51) --
	(355.66,241.60) --
	(355.69,241.61) --
	(356.06,241.77) --
	(356.08,241.77) --
	(356.57,242.01) --
	(356.86,242.17) --
	(356.99,242.24) --
	(357.04,242.27) --
	(357.10,242.29) --
	(357.28,242.37) --
	(357.68,242.57) --
	(357.71,242.59) --
	(358.36,242.94) --
	(358.45,242.99) --
	(358.80,243.19) --
	(359.58,243.57) --
	(360.14,243.86) --
	(360.17,243.87) --
	(360.28,243.93) --
	(360.68,244.11) --
	(362.54,244.96) --
	(362.92,245.17) --
	(362.99,245.21) --
	(362.99,245.21) --
	(363.68,245.56) --
	(363.71,245.57) --
	(364.25,245.86) --
	(364.57,246.01) --
	(364.77,246.11) --
	(365.00,246.26) --
	(365.26,246.42) --
	(365.84,246.74) --
	(365.84,246.74) --
	(365.92,246.77) --
	(366.04,246.82) --
	(366.47,247.02) --
	(367.05,247.32) --
	(368.20,247.89) --
	(368.39,247.96) --
	(368.50,248.02) --
	(370.21,248.86) --
	(371.17,249.37) --
	(371.58,249.56) --
	(372.02,249.77) --
	(373.58,250.63) --
	(376.96,252.21) --
	(378.52,252.92) --
	(383.85,255.70) --
	(384.44,256.07) --
	(384.92,256.30) --
	(385.30,256.48) --
	(385.73,256.69) --
	(385.96,256.80) --
	(386.70,257.17) --
	(386.81,257.23) --
	(387.17,257.43) --
	(387.19,257.44) --
	(387.24,257.46) --
	(387.39,257.54) --
	(387.67,257.70) --
	(388.26,258.02) --
	(389.16,258.43) --
	(389.82,258.73) --
	(390.40,259.04) --
	(390.44,259.06) --
	(390.79,259.24) --
	(391.07,259.36) --
	(395.16,261.33) --
	(396.00,261.76) --
	(396.15,261.83) --
	(403.52,265.76) --
	(404.02,265.99) --
	(404.02,186.34) --
	(403.52,186.23) --
	(396.15,184.68) --
	(396.00,184.66) --
	(395.16,184.56) --
	(391.07,183.90) --
	(390.79,183.85) --
	(390.44,183.80) --
	(390.40,183.79) --
	(389.82,183.70) --
	(389.16,183.57) --
	(388.26,183.43) --
	(387.67,183.36) --
	(387.39,183.28) --
	(387.24,183.25) --
	(387.19,183.24) --
	(387.17,183.24) --
	(386.81,183.18) --
	(386.70,183.16) --
	(385.96,183.06) --
	(385.73,183.03) --
	(385.30,182.96) --
	(384.92,182.88) --
	(384.44,182.77) --
	(383.85,182.65) --
	(378.52,181.64) --
	(376.96,181.44) --
	(373.58,180.91) --
	(372.02,180.68) --
	(371.58,180.65) --
	(371.17,180.59) --
	(370.21,180.45) --
	(368.50,180.08) --
	(368.39,180.06) --
	(368.20,180.02) --
	(367.05,179.77) --
	(366.47,179.65) --
	(366.04,179.54) --
	(365.92,179.53) --
	(365.84,179.52) --
	(365.84,179.52) --
	(365.26,179.44) --
	(365.00,179.39) --
	(364.77,179.33) --
	(364.57,179.26) --
	(364.25,179.16) --
	(363.71,179.01) --
	(363.68,179.01) --
	(362.99,178.88) --
	(362.99,178.88) --
	(362.92,178.86) --
	(362.54,178.77) --
	(360.68,178.40) --
	(360.28,178.34) --
	(360.17,178.33) --
	(360.14,178.33) --
	(359.58,178.20) --
	(358.80,177.96) --
	(358.45,177.91) --
	(358.36,177.90) --
	(357.71,177.77) --
	(357.68,177.77) --
	(357.28,177.69) --
	(357.10,177.64) --
	(357.04,177.62) --
	(356.99,177.60) --
	(356.86,177.57) --
	(356.57,177.50) --
	(356.08,177.44) --
	(356.06,177.44) --
	(355.69,177.39) --
	(355.66,177.39) --
	(355.44,177.35) --
	(355.38,177.33) --
	(355.06,177.25) --
	(355.03,177.24) --
	(354.77,177.18) --
	(354.60,177.14) --
	(352.50,176.83) --
	(351.61,176.61) --
	(350.57,176.40) --
	(350.27,176.31) --
	(350.01,176.25) --
	(349.59,176.10) --
	(348.84,175.91) --
	(348.63,175.90) --
	(347.30,175.60) --
	(346.94,175.50) --
	(346.80,175.46) --
	(346.49,175.36) --
	(345.78,175.18) --
	(345.20,175.07) --
	(345.11,175.06) --
	(343.95,174.79) --
	(343.79,174.78) --
	(342.53,174.52) --
	(341.52,174.36) --
	(340.65,174.13) --
	(340.28,174.04) --
	(340.15,174.01) --
	(339.84,173.92) --
	(339.76,173.90) --
	(339.55,173.83) --
	(339.40,173.79) --
	(339.21,173.73) --
	(339.04,173.70) --
	(338.87,173.64) --
	(338.83,173.63) --
	(338.50,173.54) --
	(337.65,173.34) --
	(335.47,172.82) --
	(335.38,172.79) --
	(333.83,172.41) --
	(333.82,172.41) --
	(333.69,172.37) --
	(333.56,172.35) --
	(333.53,172.35) --
	(330.88,171.78) --
	(330.38,171.66) --
	(330.17,171.61) --
	(328.92,171.27) --
	(328.37,171.13) --
	(327.78,170.94) --
	(327.50,170.85) --
	(327.01,170.66) --
	(326.46,170.48) --
	(326.16,170.42) --
	(325.91,170.35) --
	(325.87,170.35) --
	(325.85,170.34) --
	(325.46,170.28) --
	(324.82,170.17) --
	(324.53,170.12) --
	(324.06,170.03) --
	(323.98,170.02) --
	(322.93,169.78) --
	(322.64,169.70) --
	(322.56,169.67) --
	(321.47,169.34) --
	(321.34,169.32) --
	(320.71,169.14) --
	(319.14,168.58) --
	(318.60,168.45) --
	(317.13,168.04) --
	(316.72,167.88) --
	(316.61,167.84) --
	(316.58,167.83) --
	(316.37,167.76) --
	(316.29,167.72) --
	(315.65,167.51) --
	(315.61,167.51) --
	(315.54,167.50) --
	(315.28,167.39) --
	(315.05,167.32) --
	(314.84,167.28) --
	(314.55,167.23) --
	(314.12,167.07) --
	(314.09,167.07) --
	(312.75,166.67) --
	(312.68,166.66) --
	(312.39,166.57) --
	(311.87,166.38) --
	(311.71,166.32) --
	(311.49,166.25) --
	(311.09,166.15) --
	(310.98,166.13) --
	(310.80,166.08) --
	(310.45,166.02) --
	(310.36,166.01) --
	(309.72,165.79) --
	(309.55,165.75) --
	(309.08,165.61) --
	(308.79,165.54) --
	(308.63,165.48) --
	(308.42,165.44) --
	(307.37,165.12) --
	(307.16,165.08) --
	(307.03,165.02) --
	(306.15,164.82) --
	(305.86,164.72) --
	(305.76,164.70) --
	(305.28,164.59) --
	(305.11,164.57) --
	(304.68,164.44) --
	(304.60,164.42) --
	(304.48,164.39) --
	(304.47,164.39) --
	(304.31,164.34) --
	(304.10,164.27) --
	(303.17,163.99) --
	(302.93,163.92) --
	(302.89,163.90) --
	(302.61,163.82) --
	(302.52,163.80) --
	(302.45,163.78) --
	(302.30,163.73) --
	(302.20,163.70) --
	(302.08,163.66) --
	(302.06,163.66) --
	(301.86,163.61) --
	(301.81,163.60) --
	(301.34,163.43) --
	(300.69,163.25) --
	(300.07,163.04) --
	(299.22,162.75) --
	(298.89,162.65) --
	(298.71,162.61) --
	(298.48,162.53) --
	(298.45,162.52) --
	(298.21,162.43) --
	(298.13,162.41) --
	(298.09,162.40) --
	(297.99,162.37) --
	(297.11,162.13) --
	(296.61,161.93) --
	(295.85,161.71) --
	(295.12,161.51) --
	(293.94,161.16) --
	(293.90,161.15) --
	(293.61,161.05) --
	(293.49,161.00) --
	(292.94,160.81) --
	(292.85,160.79) --
	(292.78,160.77) --
	(292.77,160.76) --
	(292.36,160.65) --
	(292.20,160.60) --
	(291.58,160.39) --
	(291.45,160.35) --
	(291.29,160.30) --
	(290.44,160.04) --
	(289.99,159.91) --
	(289.71,159.82) --
	(289.61,159.78) --
	(289.30,159.65) --
	(289.13,159.58) --
	(289.03,159.53) --
	(288.86,159.47) --
	(288.45,159.34) --
	(288.16,159.21) --
	(287.76,159.03) --
	(286.95,158.71) --
	(286.70,158.61) --
	(286.04,158.39) --
	(285.67,158.25) --
	(285.44,158.18) --
	(284.88,157.97) --
	(284.84,157.96) --
	(284.81,157.95) --
	(284.81,157.95) --
	(284.80,157.95) --
	(284.75,157.94) --
	(284.43,157.86) --
	(284.26,157.81) --
	(283.80,157.65) --
	(283.74,157.63) --
	(283.34,157.47) --
	(283.28,157.45) --
	(283.10,157.38) --
	(281.60,156.86) --
	(281.52,156.84) --
	(281.16,156.72) --
	(280.63,156.53) --
	(280.60,156.52) --
	(280.50,156.49) --
	(280.47,156.48) --
	(280.41,156.46) --
	(280.36,156.44) --
	(280.21,156.38) --
	(279.90,156.27) --
	(279.83,156.25) --
	(279.36,156.10) --
	(279.03,155.99) --
	(278.99,155.97) --
	(278.93,155.95) --
	(278.83,155.92) --
	(277.37,155.41) --
	(277.28,155.37) --
	(276.77,155.18) --
	(276.73,155.17) --
	(276.14,154.91) --
	(276.03,154.89) --
	(275.50,154.75) --
	(274.88,154.53) --
	(274.85,154.52) --
	(274.31,154.33) --
	(273.48,153.97) --
	(273.29,153.90) --
	(273.01,153.83) --
	(272.91,153.80) --
	(272.40,153.63) --
	(272.31,153.60) --
	(271.51,153.32) --
	(269.91,152.88) --
	(269.75,152.84) --
	(269.30,152.67) --
	(269.12,152.60) --
	(268.75,152.46) --
	(268.51,152.38) --
	(268.32,152.31) --
	(268.15,152.25) --
	(267.95,152.15) --
	(267.83,152.10) --
	(267.74,152.07) --
	(266.70,151.65) --
	(266.69,151.65) --
	(265.54,151.21) --
	(265.48,151.19) --
	(265.11,151.03) --
	(264.94,150.97) --
	(264.43,150.80) --
	(264.14,150.73) --
	(263.47,150.51) --
	(262.57,150.18) --
	(261.39,149.72) --
	(260.16,149.25) --
	(260.08,149.22) --
	(258.99,148.77) --
	(256.94,147.97) --
	(255.73,147.55) --
	(254.34,147.05) --
	(253.89,146.87) --
	(253.69,146.77) --
	(252.46,146.32) --
	(251.97,146.14) --
	(250.69,145.55) --
	(250.38,145.47) --
	(250.22,145.41) --
	(249.74,145.23) --
	(248.30,144.72) --
	(244.78,143.23) --
	(242.95,142.41) --
	(241.65,141.90) --
	(241.27,141.77) --
	(240.80,141.53) --
	(240.03,141.24) --
	(239.31,140.92) --
	(239.29,140.91) --
	(237.41,140.17) --
	(237.40,140.16) --
	(236.22,139.67) --
	(235.24,139.19) --
	(234.24,138.83) --
	(233.34,138.46) --
	(232.39,137.97) --
	(231.34,137.53) --
	(231.15,137.48) --
	(229.14,136.70) --
	(224.00,134.47) --
	(223.59,134.28) --
	(223.54,134.27) --
	(223.29,134.18) --
	(222.58,133.90) --
	(221.63,133.38) --
	(220.93,133.02) --
	(220.33,132.73) --
	(220.17,132.66) --
	(218.17,131.78) --
	(217.68,131.60) --
	(217.43,131.51) --
	(217.33,131.46) --
	(215.87,130.73) --
	(211.50,128.68) --
	(209.65,127.80) --
	(205.19,125.59) --
	(204.99,125.51) --
	(199.87,122.96) --
	(198.04,122.14) --
	(196.35,121.30) --
	(195.71,120.97) --
	(195.62,120.92) --
	(192.71,119.45) --
	(187.92,116.99) --
	(178.49,112.56) --
	(171.43,108.91) --
	(158.01,101.79) --
	(154.64,100.08) --
	(110.88, 76.82) --
	(101.92, 71.87) --
	( 95.59, 68.45) --
	( 92.34, 66.61) --
	( 86.71, 63.49) --
	( 78.30, 58.97) --
	( 64.98, 51.78) --
	( 64.85, 51.70) --
	( 60.61, 49.34) --
	( 52.97, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 52.97,155.19) --
	( 60.61,156.53) --
	( 64.85,157.27) --
	( 64.98,157.28) --
	( 78.30,159.47) --
	( 86.71,160.71) --
	( 92.34,161.58) --
	( 95.59,162.05) --
	(101.92,163.22) --
	(110.88,164.79) --
	(154.64,172.89) --
	(158.01,173.52) --
	(171.43,176.11) --
	(178.49,177.51) --
	(187.92,179.51) --
	(192.71,180.54) --
	(195.62,181.28) --
	(195.71,181.30) --
	(196.35,181.45) --
	(198.04,181.86) --
	(199.87,182.28) --
	(204.99,183.41) --
	(205.19,183.47) --
	(209.65,184.61) --
	(211.50,185.09) --
	(215.87,186.10) --
	(217.33,186.45) --
	(217.43,186.49) --
	(217.68,186.54) --
	(218.17,186.65) --
	(220.17,187.17) --
	(220.33,187.22) --
	(220.93,187.38) --
	(221.63,187.54) --
	(222.58,187.83) --
	(223.29,188.00) --
	(223.54,188.05) --
	(223.59,188.07) --
	(224.00,188.19) --
	(229.14,189.51) --
	(231.15,190.06) --
	(231.34,190.12) --
	(232.39,190.40) --
	(233.34,190.68) --
	(234.24,191.01) --
	(235.24,191.26) --
	(236.22,191.57) --
	(237.40,191.97) --
	(237.41,191.97) --
	(239.29,192.53) --
	(239.31,192.53) --
	(240.03,192.72) --
	(240.80,192.93) --
	(241.27,193.06) --
	(241.65,193.15) --
	(242.95,193.53) --
	(244.78,194.03) --
	(248.30,195.14) --
	(249.74,195.57) --
	(250.22,195.74) --
	(250.38,195.79) --
	(250.69,195.89) --
	(251.97,196.35) --
	(252.46,196.50) --
	(253.69,196.89) --
	(253.89,196.94) --
	(254.34,197.06) --
	(255.73,197.56) --
	(256.94,197.95) --
	(258.99,198.48) --
	(260.08,198.81) --
	(260.16,198.84) --
	(261.39,199.27) --
	(262.57,199.62) --
	(263.47,199.90) --
	(264.14,200.12) --
	(264.43,200.22) --
	(264.94,200.38) --
	(265.11,200.45) --
	(265.48,200.56) --
	(265.54,200.57) --
	(266.69,200.88) --
	(266.70,200.88) --
	(267.74,201.28) --
	(267.83,201.30) --
	(267.95,201.35) --
	(268.15,201.41) --
	(268.32,201.47) --
	(268.51,201.54) --
	(268.75,201.61) --
	(269.12,201.74) --
	(269.30,201.81) --
	(269.75,201.95) --
	(269.91,202.03) --
	(271.51,202.58) --
	(272.31,202.84) --
	(272.40,202.88) --
	(272.91,203.02) --
	(273.01,203.06) --
	(273.29,203.17) --
	(273.48,203.20) --
	(274.31,203.44) --
	(274.85,203.62) --
	(274.88,203.63) --
	(275.50,203.81) --
	(276.03,203.99) --
	(276.14,204.04) --
	(276.73,204.19) --
	(276.77,204.20) --
	(277.28,204.39) --
	(277.37,204.42) --
	(278.83,204.90) --
	(278.93,204.94) --
	(278.99,204.95) --
	(279.03,204.97) --
	(279.36,205.12) --
	(279.83,205.26) --
	(279.90,205.28) --
	(280.21,205.37) --
	(280.36,205.42) --
	(280.41,205.43) --
	(280.47,205.44) --
	(280.50,205.45) --
	(280.60,205.49) --
	(280.63,205.50) --
	(281.16,205.65) --
	(281.52,205.76) --
	(281.60,205.78) --
	(283.10,206.25) --
	(283.28,206.33) --
	(283.34,206.35) --
	(283.74,206.50) --
	(283.80,206.52) --
	(284.26,206.69) --
	(284.43,206.76) --
	(284.75,206.88) --
	(284.80,206.90) --
	(284.81,206.91) --
	(284.81,206.91) --
	(284.84,206.92) --
	(284.88,206.93) --
	(285.44,207.14) --
	(285.67,207.22) --
	(286.04,207.36) --
	(286.70,207.63) --
	(286.95,207.73) --
	(287.76,208.02) --
	(288.16,208.16) --
	(288.45,208.28) --
	(288.86,208.45) --
	(289.03,208.52) --
	(289.13,208.56) --
	(289.30,208.64) --
	(289.61,208.76) --
	(289.71,208.80) --
	(289.99,208.88) --
	(290.44,209.04) --
	(291.29,209.36) --
	(291.45,209.42) --
	(291.58,209.46) --
	(292.20,209.69) --
	(292.36,209.73) --
	(292.77,209.90) --
	(292.78,209.91) --
	(292.85,209.93) --
	(292.94,209.95) --
	(293.49,210.13) --
	(293.61,210.18) --
	(293.90,210.31) --
	(293.94,210.32) --
	(295.12,210.69) --
	(295.85,210.96) --
	(296.61,211.26) --
	(297.11,211.43) --
	(297.99,211.77) --
	(298.09,211.82) --
	(298.13,211.84) --
	(298.21,211.88) --
	(298.45,211.97) --
	(298.48,211.98) --
	(298.71,212.08) --
	(298.89,212.18) --
	(299.22,212.26) --
	(300.07,212.54) --
	(300.69,212.82) --
	(301.34,213.11) --
	(301.81,213.30) --
	(301.86,213.32) --
	(302.06,213.44) --
	(302.08,213.45) --
	(302.20,213.50) --
	(302.30,213.52) --
	(302.45,213.59) --
	(302.52,213.63) --
	(302.61,213.67) --
	(302.89,213.80) --
	(302.93,213.81) --
	(303.17,213.89) --
	(304.10,214.24) --
	(304.31,214.32) --
	(304.47,214.39) --
	(304.48,214.39) --
	(304.60,214.44) --
	(304.68,214.47) --
	(305.11,214.62) --
	(305.28,214.68) --
	(305.76,214.84) --
	(305.86,214.87) --
	(306.15,214.99) --
	(307.03,215.33) --
	(307.16,215.39) --
	(307.37,215.47) --
	(308.42,215.91) --
	(308.63,216.00) --
	(308.79,216.07) --
	(309.08,216.15) --
	(309.55,216.36) --
	(309.72,216.43) --
	(310.36,216.68) --
	(310.45,216.72) --
	(310.80,216.85) --
	(310.98,216.92) --
	(311.09,216.97) --
	(311.49,217.14) --
	(311.71,217.26) --
	(311.87,217.33) --
	(312.39,217.54) --
	(312.68,217.63) --
	(312.75,217.66) --
	(314.09,218.31) --
	(314.12,218.32) --
	(314.55,218.48) --
	(314.84,218.60) --
	(315.05,218.68) --
	(315.28,218.77) --
	(315.54,218.83) --
	(315.61,218.85) --
	(315.65,218.87) --
	(316.29,219.12) --
	(316.37,219.15) --
	(316.58,219.26) --
	(316.61,219.28) --
	(316.72,219.32) --
	(317.13,219.49) --
	(318.60,220.09) --
	(319.14,220.29) --
	(320.71,220.94) --
	(321.34,221.16) --
	(321.47,221.23) --
	(322.56,221.66) --
	(322.64,221.71) --
	(322.93,221.87) --
	(323.98,222.34) --
	(324.06,222.38) --
	(324.53,222.58) --
	(324.82,222.69) --
	(325.46,222.96) --
	(325.85,223.17) --
	(325.87,223.18) --
	(325.91,223.20) --
	(326.16,223.30) --
	(326.46,223.44) --
	(327.01,223.67) --
	(327.50,223.87) --
	(327.78,223.98) --
	(328.37,224.21) --
	(328.92,224.41) --
	(330.17,224.96) --
	(330.38,225.06) --
	(330.88,225.26) --
	(333.53,226.46) --
	(333.56,226.47) --
	(333.69,226.53) --
	(333.82,226.59) --
	(333.83,226.59) --
	(335.38,227.23) --
	(335.47,227.26) --
	(337.65,228.26) --
	(338.50,228.58) --
	(338.83,228.72) --
	(338.87,228.74) --
	(339.04,228.81) --
	(339.21,228.91) --
	(339.40,228.99) --
	(339.55,229.06) --
	(339.76,229.11) --
	(339.84,229.13) --
	(340.15,229.30) --
	(340.28,229.35) --
	(340.65,229.49) --
	(341.52,229.90) --
	(342.53,230.36) --
	(343.79,230.93) --
	(343.95,231.01) --
	(345.11,231.48) --
	(345.20,231.51) --
	(345.78,231.75) --
	(346.49,232.11) --
	(346.80,232.22) --
	(346.94,232.30) --
	(347.30,232.45) --
	(348.63,233.08) --
	(348.84,233.17) --
	(349.59,233.49) --
	(350.01,233.69) --
	(350.27,233.84) --
	(350.57,233.99) --
	(351.61,234.45) --
	(352.50,234.88) --
	(354.60,235.89) --
	(354.77,235.97) --
	(355.03,236.10) --
	(355.06,236.11) --
	(355.38,236.26) --
	(355.44,236.29) --
	(355.66,236.39) --
	(355.69,236.41) --
	(356.06,236.58) --
	(356.08,236.59) --
	(356.57,236.82) --
	(356.86,236.97) --
	(356.99,237.03) --
	(357.04,237.05) --
	(357.10,237.08) --
	(357.28,237.18) --
	(357.68,237.39) --
	(357.71,237.41) --
	(358.36,237.73) --
	(358.45,237.77) --
	(358.80,237.91) --
	(359.58,238.25) --
	(360.14,238.49) --
	(360.17,238.49) --
	(360.28,238.54) --
	(360.68,238.74) --
	(362.54,239.57) --
	(362.92,239.73) --
	(362.99,239.76) --
	(362.99,239.76) --
	(363.68,240.09) --
	(363.71,240.11) --
	(364.25,240.39) --
	(364.57,240.53) --
	(364.77,240.61) --
	(365.00,240.72) --
	(365.26,240.83) --
	(365.84,241.13) --
	(365.84,241.13) --
	(365.92,241.18) --
	(366.04,241.25) --
	(366.47,241.48) --
	(367.05,241.72) --
	(368.20,242.22) --
	(368.39,242.31) --
	(368.50,242.36) --
	(370.21,243.12) --
	(371.17,243.56) --
	(371.58,243.77) --
	(372.02,244.00) --
	(373.58,244.77) --
	(376.96,246.43) --
	(378.52,247.22) --
	(383.85,249.72) --
	(384.44,250.00) --
	(384.92,250.25) --
	(385.30,250.43) --
	(385.73,250.63) --
	(385.96,250.75) --
	(386.70,251.19) --
	(386.81,251.26) --
	(387.17,251.47) --
	(387.19,251.47) --
	(387.24,251.49) --
	(387.39,251.56) --
	(387.67,251.69) --
	(388.26,251.93) --
	(389.16,252.31) --
	(389.82,252.63) --
	(390.40,252.99) --
	(390.44,253.01) --
	(390.79,253.23) --
	(391.07,253.39) --
	(395.16,255.39) --
	(396.00,255.83) --
	(396.15,255.90) --
	(403.52,259.58) --
	(404.02,259.80) --
	(404.02,193.19) --
	(403.52,193.11) --
	(396.15,191.61) --
	(396.00,191.61) --
	(395.16,191.48) --
	(391.07,190.82) --
	(390.79,190.75) --
	(390.44,190.67) --
	(390.40,190.66) --
	(389.82,190.53) --
	(389.16,190.35) --
	(388.26,190.22) --
	(387.67,190.10) --
	(387.39,190.05) --
	(387.24,190.02) --
	(387.19,190.00) --
	(387.17,190.00) --
	(386.81,189.88) --
	(386.70,189.84) --
	(385.96,189.65) --
	(385.73,189.61) --
	(385.30,189.54) --
	(384.92,189.44) --
	(384.44,189.37) --
	(383.85,189.24) --
	(378.52,187.98) --
	(376.96,187.66) --
	(373.58,186.93) --
	(372.02,186.65) --
	(371.58,186.49) --
	(371.17,186.38) --
	(370.21,186.10) --
	(368.50,185.76) --
	(368.39,185.75) --
	(368.20,185.70) --
	(367.05,185.49) --
	(366.47,185.33) --
	(366.04,185.27) --
	(365.92,185.25) --
	(365.84,185.23) --
	(365.84,185.23) --
	(365.26,185.08) --
	(365.00,185.01) --
	(364.77,184.98) --
	(364.57,184.96) --
	(364.25,184.88) --
	(363.71,184.76) --
	(363.68,184.76) --
	(362.99,184.59) --
	(362.99,184.59) --
	(362.92,184.58) --
	(362.54,184.47) --
	(360.68,184.03) --
	(360.28,183.93) --
	(360.17,183.89) --
	(360.14,183.89) --
	(359.58,183.75) --
	(358.80,183.55) --
	(358.45,183.44) --
	(358.36,183.41) --
	(357.71,183.25) --
	(357.68,183.25) --
	(357.28,183.16) --
	(357.10,183.11) --
	(357.04,183.09) --
	(356.99,183.08) --
	(356.86,183.03) --
	(356.57,182.97) --
	(356.08,182.88) --
	(356.06,182.87) --
	(355.69,182.77) --
	(355.66,182.76) --
	(355.44,182.71) --
	(355.38,182.69) --
	(355.06,182.60) --
	(355.03,182.58) --
	(354.77,182.52) --
	(354.60,182.49) --
	(352.50,182.01) --
	(351.61,181.76) --
	(350.57,181.44) --
	(350.27,181.37) --
	(350.01,181.30) --
	(349.59,181.21) --
	(348.84,181.06) --
	(348.63,181.03) --
	(347.30,180.67) --
	(346.94,180.58) --
	(346.80,180.53) --
	(346.49,180.45) --
	(345.78,180.29) --
	(345.20,180.15) --
	(345.11,180.12) --
	(343.95,179.83) --
	(343.79,179.81) --
	(342.53,179.51) --
	(341.52,179.27) --
	(340.65,179.03) --
	(340.28,178.93) --
	(340.15,178.90) --
	(339.84,178.84) --
	(339.76,178.82) --
	(339.55,178.77) --
	(339.40,178.72) --
	(339.21,178.67) --
	(339.04,178.66) --
	(338.87,178.60) --
	(338.83,178.60) --
	(338.50,178.51) --
	(337.65,178.32) --
	(335.47,177.66) --
	(335.38,177.66) --
	(333.83,177.22) --
	(333.82,177.21) --
	(333.69,177.18) --
	(333.56,177.16) --
	(333.53,177.15) --
	(330.88,176.40) --
	(330.38,176.27) --
	(330.17,176.19) --
	(328.92,175.83) --
	(328.37,175.67) --
	(327.78,175.48) --
	(327.50,175.42) --
	(327.01,175.27) --
	(326.46,175.15) --
	(326.16,175.09) --
	(325.91,175.03) --
	(325.87,175.02) --
	(325.85,175.02) --
	(325.46,174.98) --
	(324.82,174.87) --
	(324.53,174.78) --
	(324.06,174.68) --
	(323.98,174.67) --
	(322.93,174.38) --
	(322.64,174.29) --
	(322.56,174.27) --
	(321.47,173.96) --
	(321.34,173.91) --
	(320.71,173.74) --
	(319.14,173.37) --
	(318.60,173.23) --
	(317.13,172.83) --
	(316.72,172.72) --
	(316.61,172.69) --
	(316.58,172.68) --
	(316.37,172.62) --
	(316.29,172.61) --
	(315.65,172.41) --
	(315.61,172.39) --
	(315.54,172.36) --
	(315.28,172.26) --
	(315.05,172.19) --
	(314.84,172.14) --
	(314.55,172.06) --
	(314.12,171.98) --
	(314.09,171.97) --
	(312.75,171.61) --
	(312.68,171.61) --
	(312.39,171.50) --
	(311.87,171.35) --
	(311.71,171.29) --
	(311.49,171.25) --
	(311.09,171.16) --
	(310.98,171.13) --
	(310.80,171.04) --
	(310.45,170.94) --
	(310.36,170.92) --
	(309.72,170.76) --
	(309.55,170.73) --
	(309.08,170.61) --
	(308.79,170.53) --
	(308.63,170.50) --
	(308.42,170.44) --
	(307.37,170.08) --
	(307.16,169.99) --
	(307.03,169.96) --
	(306.15,169.70) --
	(305.86,169.60) --
	(305.76,169.57) --
	(305.28,169.43) --
	(305.11,169.37) --
	(304.68,169.29) --
	(304.60,169.26) --
	(304.48,169.24) --
	(304.47,169.24) --
	(304.31,169.20) --
	(304.10,169.14) --
	(303.17,168.86) --
	(302.93,168.78) --
	(302.89,168.77) --
	(302.61,168.67) --
	(302.52,168.64) --
	(302.45,168.62) --
	(302.30,168.58) --
	(302.20,168.55) --
	(302.08,168.52) --
	(302.06,168.52) --
	(301.86,168.48) --
	(301.81,168.46) --
	(301.34,168.30) --
	(300.69,168.09) --
	(300.07,167.93) --
	(299.22,167.66) --
	(298.89,167.58) --
	(298.71,167.52) --
	(298.48,167.45) --
	(298.45,167.44) --
	(298.21,167.36) --
	(298.13,167.33) --
	(298.09,167.31) --
	(297.99,167.28) --
	(297.11,166.97) --
	(296.61,166.83) --
	(295.85,166.56) --
	(295.12,166.33) --
	(293.94,165.91) --
	(293.90,165.90) --
	(293.61,165.83) --
	(293.49,165.80) --
	(292.94,165.62) --
	(292.85,165.60) --
	(292.78,165.59) --
	(292.77,165.59) --
	(292.36,165.45) --
	(292.20,165.42) --
	(291.58,165.18) --
	(291.45,165.14) --
	(291.29,165.09) --
	(290.44,164.80) --
	(289.99,164.65) --
	(289.71,164.55) --
	(289.61,164.51) --
	(289.30,164.42) --
	(289.13,164.37) --
	(289.03,164.34) --
	(288.86,164.30) --
	(288.45,164.13) --
	(288.16,164.03) --
	(287.76,163.91) --
	(286.95,163.73) --
	(286.70,163.63) --
	(286.04,163.40) --
	(285.67,163.30) --
	(285.44,163.21) --
	(284.88,162.98) --
	(284.84,162.95) --
	(284.81,162.93) --
	(284.81,162.93) --
	(284.80,162.93) --
	(284.75,162.91) --
	(284.43,162.81) --
	(284.26,162.77) --
	(283.80,162.62) --
	(283.74,162.59) --
	(283.34,162.46) --
	(283.28,162.45) --
	(283.10,162.38) --
	(281.60,161.86) --
	(281.52,161.82) --
	(281.16,161.69) --
	(280.63,161.49) --
	(280.60,161.48) --
	(280.50,161.44) --
	(280.47,161.43) --
	(280.41,161.41) --
	(280.36,161.39) --
	(280.21,161.34) --
	(279.90,161.21) --
	(279.83,161.18) --
	(279.36,161.02) --
	(279.03,160.90) --
	(278.99,160.89) --
	(278.93,160.88) --
	(278.83,160.85) --
	(277.37,160.31) --
	(277.28,160.26) --
	(276.77,160.09) --
	(276.73,160.07) --
	(276.14,159.87) --
	(276.03,159.84) --
	(275.50,159.62) --
	(274.88,159.38) --
	(274.85,159.37) --
	(274.31,159.22) --
	(273.48,158.96) --
	(273.29,158.86) --
	(273.01,158.72) --
	(272.91,158.70) --
	(272.40,158.57) --
	(272.31,158.54) --
	(271.51,158.23) --
	(269.91,157.66) --
	(269.75,157.60) --
	(269.30,157.43) --
	(269.12,157.37) --
	(268.75,157.24) --
	(268.51,157.16) --
	(268.32,157.08) --
	(268.15,157.04) --
	(267.95,156.96) --
	(267.83,156.90) --
	(267.74,156.87) --
	(266.70,156.51) --
	(266.69,156.51) --
	(265.54,156.06) --
	(265.48,156.04) --
	(265.11,155.87) --
	(264.94,155.82) --
	(264.43,155.64) --
	(264.14,155.54) --
	(263.47,155.27) --
	(262.57,154.94) --
	(261.39,154.51) --
	(260.16,154.01) --
	(260.08,153.99) --
	(258.99,153.54) --
	(256.94,152.72) --
	(255.73,152.25) --
	(254.34,151.68) --
	(253.89,151.52) --
	(253.69,151.45) --
	(252.46,150.93) --
	(251.97,150.72) --
	(250.69,150.23) --
	(250.38,150.10) --
	(250.22,150.06) --
	(249.74,149.87) --
	(248.30,149.31) --
	(244.78,147.88) --
	(242.95,147.21) --
	(241.65,146.64) --
	(241.27,146.46) --
	(240.80,146.27) --
	(240.03,145.97) --
	(239.31,145.69) --
	(239.29,145.68) --
	(237.41,144.89) --
	(237.40,144.88) --
	(236.22,144.44) --
	(235.24,144.03) --
	(234.24,143.66) --
	(233.34,143.28) --
	(232.39,142.91) --
	(231.34,142.45) --
	(231.15,142.37) --
	(229.14,141.49) --
	(224.00,139.16) --
	(223.59,138.95) --
	(223.54,138.94) --
	(223.29,138.82) --
	(222.58,138.54) --
	(221.63,138.11) --
	(220.93,137.81) --
	(220.33,137.57) --
	(220.17,137.50) --
	(218.17,136.61) --
	(217.68,136.38) --
	(217.43,136.24) --
	(217.33,136.19) --
	(215.87,135.57) --
	(211.50,133.70) --
	(209.65,132.94) --
	(205.19,130.92) --
	(204.99,130.83) --
	(199.87,128.59) --
	(198.04,127.74) --
	(196.35,126.88) --
	(195.71,126.55) --
	(195.62,126.51) --
	(192.71,125.09) --
	(187.92,122.99) --
	(178.49,118.59) --
	(171.43,115.31) --
	(158.01,108.67) --
	(154.64,106.95) --
	(110.88, 84.87) --
	(101.92, 80.24) --
	( 95.59, 76.90) --
	( 92.34, 75.13) --
	( 86.71, 72.15) --
	( 78.30, 67.89) --
	( 64.98, 61.01) --
	( 64.85, 60.94) --
	( 60.61, 58.72) --
	( 52.97, 54.81) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 52.97, 34.03) -- ( 52.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 60.61, 34.03) -- ( 60.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.85, 34.03) -- ( 64.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.98, 34.03) -- ( 64.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 78.30, 34.03) -- ( 78.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 86.71, 34.03) -- ( 86.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 92.34, 34.03) -- ( 92.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 95.59, 34.03) -- ( 95.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (101.92, 34.03) -- (101.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (110.88, 34.03) -- (110.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (154.64, 34.03) -- (154.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (158.01, 34.03) -- (158.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (171.43, 34.03) -- (171.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (178.49, 34.03) -- (178.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (187.92, 34.03) -- (187.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (192.71, 34.03) -- (192.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.62, 34.03) -- (195.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.71, 34.03) -- (195.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (196.35, 34.03) -- (196.35, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (198.04, 34.03) -- (198.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (199.87, 34.03) -- (199.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (204.99, 34.03) -- (204.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.19, 34.03) -- (205.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (209.65, 34.03) -- (209.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (211.50, 34.03) -- (211.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (215.87, 34.03) -- (215.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (217.33, 34.03) -- (217.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (217.43, 34.03) -- (217.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (217.68, 34.03) -- (217.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (218.17, 34.03) -- (218.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.17, 34.03) -- (220.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.33, 34.03) -- (220.33, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.93, 34.03) -- (220.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (221.63, 34.03) -- (221.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (222.58, 34.03) -- (222.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.29, 34.03) -- (223.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.54, 34.03) -- (223.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.59, 34.03) -- (223.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (224.00, 34.03) -- (224.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.14, 34.03) -- (229.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.15, 34.03) -- (231.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.34, 34.03) -- (231.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.39, 34.03) -- (232.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (233.34, 34.03) -- (233.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (234.25, 34.03) -- (234.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (235.24, 34.03) -- (235.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.22, 34.03) -- (236.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (237.40, 34.03) -- (237.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (237.41, 34.03) -- (237.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.29, 34.03) -- (239.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.31, 34.03) -- (239.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.03, 34.03) -- (240.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.80, 34.03) -- (240.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (241.27, 34.03) -- (241.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (241.65, 34.03) -- (241.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.95, 34.03) -- (242.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (244.78, 34.03) -- (244.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.30, 34.03) -- (248.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.74, 34.03) -- (249.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.22, 34.03) -- (250.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.38, 34.03) -- (250.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.69, 34.03) -- (250.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.97, 34.03) -- (251.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.46, 34.03) -- (252.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.69, 34.03) -- (253.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.89, 34.03) -- (253.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (254.34, 34.03) -- (254.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (255.74, 34.03) -- (255.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.94, 34.03) -- (256.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.99, 34.03) -- (258.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.08, 34.03) -- (260.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.16, 34.03) -- (260.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.39, 34.03) -- (261.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.57, 34.03) -- (262.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (263.47, 34.03) -- (263.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.14, 34.03) -- (264.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.43, 34.03) -- (264.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.94, 34.03) -- (264.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (265.11, 34.03) -- (265.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (265.48, 34.03) -- (265.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (265.54, 34.03) -- (265.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.69, 34.03) -- (266.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.70, 34.03) -- (266.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.75, 34.03) -- (267.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.83, 34.03) -- (267.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.95, 34.03) -- (267.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.15, 34.03) -- (268.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.31, 34.03) -- (268.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.51, 34.03) -- (268.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.75, 34.03) -- (268.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.12, 34.03) -- (269.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.30, 34.03) -- (269.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.75, 34.03) -- (269.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.91, 34.03) -- (269.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.51, 34.03) -- (271.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.31, 34.03) -- (272.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.40, 34.03) -- (272.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.91, 34.03) -- (272.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.01, 34.03) -- (273.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.29, 34.03) -- (273.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.48, 34.03) -- (273.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.32, 34.03) -- (274.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.85, 34.03) -- (274.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.88, 34.03) -- (274.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.50, 34.03) -- (275.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.03, 34.03) -- (276.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.14, 34.03) -- (276.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.73, 34.03) -- (276.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.77, 34.03) -- (276.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.28, 34.03) -- (277.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.37, 34.03) -- (277.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.83, 34.03) -- (278.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.93, 34.03) -- (278.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.99, 34.03) -- (278.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.03, 34.03) -- (279.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.36, 34.03) -- (279.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.83, 34.03) -- (279.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.90, 34.03) -- (279.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.21, 34.03) -- (280.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.36, 34.03) -- (280.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.41, 34.03) -- (280.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.47, 34.03) -- (280.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.50, 34.03) -- (280.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.60, 34.03) -- (280.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.63, 34.03) -- (280.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.16, 34.03) -- (281.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.52, 34.03) -- (281.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.60, 34.03) -- (281.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.10, 34.03) -- (283.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.28, 34.03) -- (283.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.34, 34.03) -- (283.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.74, 34.03) -- (283.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.80, 34.03) -- (283.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.26, 34.03) -- (284.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.43, 34.03) -- (284.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.75, 34.03) -- (284.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.80, 34.03) -- (284.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.81, 34.03) -- (284.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.81, 34.03) -- (284.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.84, 34.03) -- (284.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.88, 34.03) -- (284.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.44, 34.03) -- (285.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.67, 34.03) -- (285.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.04, 34.03) -- (286.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.70, 34.03) -- (286.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.95, 34.03) -- (286.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.76, 34.03) -- (287.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.16, 34.03) -- (288.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.45, 34.03) -- (288.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.86, 34.03) -- (288.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.03, 34.03) -- (289.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.13, 34.03) -- (289.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.30, 34.03) -- (289.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.61, 34.03) -- (289.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.71, 34.03) -- (289.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.99, 34.03) -- (289.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.44, 34.03) -- (290.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.29, 34.03) -- (291.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.45, 34.03) -- (291.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.57, 34.03) -- (291.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.20, 34.03) -- (292.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.36, 34.03) -- (292.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.77, 34.03) -- (292.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.78, 34.03) -- (292.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.85, 34.03) -- (292.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.94, 34.03) -- (292.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.49, 34.03) -- (293.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.61, 34.03) -- (293.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.90, 34.03) -- (293.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.94, 34.03) -- (293.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.12, 34.03) -- (295.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.85, 34.03) -- (295.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.61, 34.03) -- (296.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.11, 34.03) -- (297.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.99, 34.03) -- (297.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.09, 34.03) -- (298.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.13, 34.03) -- (298.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.21, 34.03) -- (298.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.45, 34.03) -- (298.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.48, 34.03) -- (298.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.71, 34.03) -- (298.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.89, 34.03) -- (298.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.22, 34.03) -- (299.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.07, 34.03) -- (300.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.69, 34.03) -- (300.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.34, 34.03) -- (301.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.81, 34.03) -- (301.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.86, 34.03) -- (301.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.06, 34.03) -- (302.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.08, 34.03) -- (302.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.20, 34.03) -- (302.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.30, 34.03) -- (302.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.45, 34.03) -- (302.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.52, 34.03) -- (302.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.61, 34.03) -- (302.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.89, 34.03) -- (302.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.93, 34.03) -- (302.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.17, 34.03) -- (303.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.10, 34.03) -- (304.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.31, 34.03) -- (304.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.47, 34.03) -- (304.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.48, 34.03) -- (304.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.60, 34.03) -- (304.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.68, 34.03) -- (304.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.11, 34.03) -- (305.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.28, 34.03) -- (305.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.76, 34.03) -- (305.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.86, 34.03) -- (305.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.15, 34.03) -- (306.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.03, 34.03) -- (307.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.16, 34.03) -- (307.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.37, 34.03) -- (307.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.42, 34.03) -- (308.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.63, 34.03) -- (308.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.79, 34.03) -- (308.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.08, 34.03) -- (309.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.55, 34.03) -- (309.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.72, 34.03) -- (309.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.36, 34.03) -- (310.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.45, 34.03) -- (310.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.80, 34.03) -- (310.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.98, 34.03) -- (310.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.09, 34.03) -- (311.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.49, 34.03) -- (311.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.71, 34.03) -- (311.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.87, 34.03) -- (311.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.39, 34.03) -- (312.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.68, 34.03) -- (312.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.75, 34.03) -- (312.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.09, 34.03) -- (314.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.12, 34.03) -- (314.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.55, 34.03) -- (314.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.84, 34.03) -- (314.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.05, 34.03) -- (315.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.28, 34.03) -- (315.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.54, 34.03) -- (315.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.61, 34.03) -- (315.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.65, 34.03) -- (315.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.29, 34.03) -- (316.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.37, 34.03) -- (316.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.58, 34.03) -- (316.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.61, 34.03) -- (316.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.72, 34.03) -- (316.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.13, 34.03) -- (317.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.60, 34.03) -- (318.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.14, 34.03) -- (319.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.71, 34.03) -- (320.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.34, 34.03) -- (321.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.47, 34.03) -- (321.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.56, 34.03) -- (322.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.64, 34.03) -- (322.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.93, 34.03) -- (322.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.97, 34.03) -- (323.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.06, 34.03) -- (324.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.53, 34.03) -- (324.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.82, 34.03) -- (324.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.46, 34.03) -- (325.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.85, 34.03) -- (325.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.87, 34.03) -- (325.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.91, 34.03) -- (325.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.16, 34.03) -- (326.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.46, 34.03) -- (326.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.01, 34.03) -- (327.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.50, 34.03) -- (327.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.78, 34.03) -- (327.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.37, 34.03) -- (328.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.92, 34.03) -- (328.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.17, 34.03) -- (330.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.38, 34.03) -- (330.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.88, 34.03) -- (330.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.53, 34.03) -- (333.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.56, 34.03) -- (333.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.69, 34.03) -- (333.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.82, 34.03) -- (333.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.83, 34.03) -- (333.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.38, 34.03) -- (335.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.47, 34.03) -- (335.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.65, 34.03) -- (337.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.49, 34.03) -- (338.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.83, 34.03) -- (338.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.87, 34.03) -- (338.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.04, 34.03) -- (339.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.21, 34.03) -- (339.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.40, 34.03) -- (339.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.55, 34.03) -- (339.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.76, 34.03) -- (339.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.84, 34.03) -- (339.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.15, 34.03) -- (340.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.28, 34.03) -- (340.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.65, 34.03) -- (340.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.52, 34.03) -- (341.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.53, 34.03) -- (342.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.79, 34.03) -- (343.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (343.95, 34.03) -- (343.95, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.11, 34.03) -- (345.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.20, 34.03) -- (345.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.78, 34.03) -- (345.78, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.49, 34.03) -- (346.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.80, 34.03) -- (346.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.94, 34.03) -- (346.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (347.30, 34.03) -- (347.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.63, 34.03) -- (348.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.84, 34.03) -- (348.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.59, 34.03) -- (349.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.01, 34.03) -- (350.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.27, 34.03) -- (350.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.57, 34.03) -- (350.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.61, 34.03) -- (351.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (352.51, 34.03) -- (352.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.60, 34.03) -- (354.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.77, 34.03) -- (354.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.03, 34.03) -- (355.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.06, 34.03) -- (355.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.38, 34.03) -- (355.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.44, 34.03) -- (355.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.66, 34.03) -- (355.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.69, 34.03) -- (355.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.06, 34.03) -- (356.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.08, 34.03) -- (356.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.57, 34.03) -- (356.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.86, 34.03) -- (356.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.99, 34.03) -- (356.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.04, 34.03) -- (357.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.10, 34.03) -- (357.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.28, 34.03) -- (357.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.68, 34.03) -- (357.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.71, 34.03) -- (357.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.36, 34.03) -- (358.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.45, 34.03) -- (358.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.80, 34.03) -- (358.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.58, 34.03) -- (359.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.14, 34.03) -- (360.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.17, 34.03) -- (360.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.28, 34.03) -- (360.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.68, 34.03) -- (360.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.54, 34.03) -- (362.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.92, 34.03) -- (362.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.99, 34.03) -- (362.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.99, 34.03) -- (362.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.68, 34.03) -- (363.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (363.71, 34.03) -- (363.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.25, 34.03) -- (364.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.57, 34.03) -- (364.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.77, 34.03) -- (364.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.99, 34.03) -- (364.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.26, 34.03) -- (365.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.84, 34.03) -- (365.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.84, 34.03) -- (365.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.92, 34.03) -- (365.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.04, 34.03) -- (366.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.47, 34.03) -- (366.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (367.05, 34.03) -- (367.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.20, 34.03) -- (368.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.39, 34.03) -- (368.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (368.50, 34.03) -- (368.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (370.20, 34.03) -- (370.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.17, 34.03) -- (371.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.58, 34.03) -- (371.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.02, 34.03) -- (372.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.58, 34.03) -- (373.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.96, 34.03) -- (376.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.52, 34.03) -- (378.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.85, 34.03) -- (383.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.44, 34.03) -- (384.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (384.92, 34.03) -- (384.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.30, 34.03) -- (385.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.73, 34.03) -- (385.73, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.96, 34.03) -- (385.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.70, 34.03) -- (386.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (386.81, 34.03) -- (386.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.17, 34.03) -- (387.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.19, 34.03) -- (387.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.24, 34.03) -- (387.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.39, 34.03) -- (387.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.67, 34.03) -- (387.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.26, 34.03) -- (388.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.16, 34.03) -- (389.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.82, 34.03) -- (389.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.40, 34.03) -- (390.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.44, 34.03) -- (390.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (390.79, 34.03) -- (390.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (391.07, 34.03) -- (391.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (395.16, 34.03) -- (395.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (396.00, 34.03) -- (396.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (396.15, 34.03) -- (396.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (403.52, 34.03) -- (403.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.02, 34.03) -- (404.02, 41.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	( 35.42,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31, 53.75) {-5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,132.99) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,212.23) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 36.01, 20.31) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 83.96, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (131.90, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (179.85, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (227.80, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (275.75, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (323.69, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (371.64, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (228.50,  4.82) {Ln(Min. Capital Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}