### 2014-10-09 LaTeX/mCityCoefPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-10-06 20:01:15
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] (124.66, 27.42) rectangle (277.03,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] (124.66, 27.42) rectangle (277.04,277.03);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (200.93, 40.79) -- (213.04, 40.79);

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (200.64, 63.08) -- (202.41, 63.08);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (198.74, 85.37) -- (203.35, 85.37);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (148.07,107.66) -- (170.40,107.66);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (131.59,129.94) -- (270.11,129.94);

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (196.74,152.23) -- (231.20,152.23);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (143.41,174.52) -- (189.28,174.52);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (200.75,196.80) -- (202.64,196.80);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (190.64,219.09) -- (251.11,219.09);

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (160.16,241.38) -- (204.59,241.38);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,draw opacity=0.30,line width= 0.3pt,line join=round,fill=fillColor,fill opacity=0.30] (203.83,263.66) -- (223.65,263.66);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (201.90, 40.79) -- (212.07, 40.79);

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (200.78, 63.08) -- (202.27, 63.08);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (199.11, 85.37) -- (202.98, 85.37);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (149.87,107.66) -- (168.61,107.66);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (142.72,129.94) -- (258.97,129.94);

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (199.51,152.23) -- (228.43,152.23);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (147.09,174.52) -- (185.60,174.52);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (200.90,196.80) -- (202.49,196.80);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (195.50,219.09) -- (246.25,219.09);

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (163.73,241.38) -- (201.02,241.38);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 1.1pt,line join=round,fill=fillColor] (205.42,263.66) -- (222.05,263.66);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}
\definecolor[named]{fillColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round,fill=fillColor] (200.31, 27.42) -- (200.31,277.03);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (213.74,263.66) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (182.37,241.38) circle (  2.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (220.87,219.09) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (201.69,196.80) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (166.35,174.52) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (213.97,152.23) circle (  2.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (200.85,129.94) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}
\definecolor[named]{fillColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (159.24,107.66) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}
\definecolor[named]{fillColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (201.04, 85.37) circle (  2.85);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}
\definecolor[named]{fillColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (201.52, 63.08) circle (  2.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (206.98, 40.79) circle (  2.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.04, 39.68) --
	(213.04, 41.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.04, 40.79) --
	(200.93, 40.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.93, 39.68) --
	(200.93, 41.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.41, 61.97) --
	(202.41, 64.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.41, 63.08) --
	(200.64, 63.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.64, 61.97) --
	(200.64, 64.20);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.35, 84.25) --
	(203.35, 86.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.35, 85.37) --
	(198.74, 85.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.74, 84.25) --
	(198.74, 86.48);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.6pt,line join=round] (170.40,106.54) --
	(170.40,108.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (170.40,107.66) --
	(148.07,107.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.07,106.54) --
	(148.07,108.77);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (270.11,128.83) --
	(270.11,131.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (270.11,129.94) --
	(131.59,129.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (131.59,128.83) --
	(131.59,131.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.20,151.11) --
	(231.20,153.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.20,152.23) --
	(196.74,152.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (196.74,151.11) --
	(196.74,153.34);
\definecolor[named]{drawColor}{rgb}{0.87,0.18,0.15}

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.28,173.40) --
	(189.28,175.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.28,174.52) --
	(143.41,174.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.41,173.40) --
	(143.41,175.63);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.64,195.69) --
	(202.64,197.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.64,196.80) --
	(200.75,196.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.75,195.69) --
	(200.75,197.92);
\definecolor[named]{drawColor}{rgb}{0.59,0.59,0.59}

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.11,217.97) --
	(251.11,220.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.11,219.09) --
	(190.64,219.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.64,217.97) --
	(190.64,220.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (204.59,240.26) --
	(204.59,242.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (204.59,241.38) --
	(160.16,241.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (160.16,240.26) --
	(160.16,242.49);
\definecolor[named]{drawColor}{rgb}{0.21,0.56,0.75}

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.65,262.55) --
	(223.65,264.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.65,263.66) --
	(203.83,263.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.83,262.55) --
	(203.83,264.78);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] (124.66, 27.42) --
	(124.66,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55, 37.49) {World GDP Growth$_{t}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55, 59.78) {Resource Rents/GDP$_{t}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55, 82.06) {Democracy$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,104.35) {Ln(Inflation)$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,126.64) {Upper Income};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,148.92) {Number of conflicts$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,171.21) {Area$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,193.50) {Duration$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,215.78) {Type$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,238.07) {Intensity$_{t-1}$};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (117.55,260.36) {Ln(Min. City Dist.)$_{t-1}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] (124.66, 27.42) --
	(277.03, 27.42);
\end{scope}
\end{tikzpicture}