### 2014-10-09 LaTeX/mCitySimPlot.tex

% Created by tikzDevice version 0.7.0 on 2014-10-06 20:01:25
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (433.62,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}

\path[fill=fillColor] ( 35.42, 34.03) rectangle (421.57,277.03);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.97,107.25) --
	( 61.01,109.92) --
	( 65.48,111.40) --
	( 65.62,111.44) --
	( 79.64,116.10) --
	( 88.50,119.04) --
	( 94.41,121.00) --
	( 97.84,122.14) --
	(104.50,124.35) --
	(110.01,126.18) --
	(113.94,127.48) --
	(116.56,128.35) --
	(124.94,131.13) --
	(160.01,142.77) --
	(160.15,142.82) --
	(163.56,143.95) --
	(164.50,144.26) --
	(175.08,147.77) --
	(177.68,148.64) --
	(185.12,151.10) --
	(195.04,154.40) --
	(195.05,154.40) --
	(199.79,155.97) --
	(200.09,156.07) --
	(201.70,156.61) --
	(202.42,156.85) --
	(203.15,157.09) --
	(203.24,157.12) --
	(203.92,157.34) --
	(204.08,157.39) --
	(205.70,157.93) --
	(211.30,159.79) --
	(213.02,160.36) --
	(213.22,160.43) --
	(213.28,160.45) --
	(214.88,160.98) --
	(219.39,162.48) --
	(219.87,162.64) --
	(224.51,164.18) --
	(226.38,164.80) --
	(226.89,164.97) --
	(228.22,165.41) --
	(229.80,165.93) --
	(230.53,166.18) --
	(231.54,166.51) --
	(232.29,166.76) --
	(232.55,166.84) --
	(232.60,166.86) --
	(233.03,167.01) --
	(236.50,168.16) --
	(238.44,168.80) --
	(239.50,169.15) --
	(240.76,169.57) --
	(242.87,170.27) --
	(243.82,170.58) --
	(244.87,170.93) --
	(245.90,171.27) --
	(247.14,171.69) --
	(247.15,171.69) --
	(248.00,171.97) --
	(248.59,172.17) --
	(249.15,172.35) --
	(249.91,172.60) --
	(250.14,172.68) --
	(250.71,172.87) --
	(251.21,173.04) --
	(251.61,173.17) --
	(251.88,173.26) --
	(252.52,173.47) --
	(252.98,173.63) --
	(253.12,173.67) --
	(253.36,173.75) --
	(254.91,174.27) --
	(255.02,174.30) --
	(256.32,174.73) --
	(258.61,175.49) --
	(260.03,175.96) --
	(260.14,176.00) --
	(260.80,176.22) --
	(261.97,176.61) --
	(262.19,176.68) --
	(262.48,176.78) --
	(262.99,176.95) --
	(264.29,177.38) --
	(264.50,177.45) --
	(264.53,177.46) --
	(264.97,177.60) --
	(266.10,177.98) --
	(266.13,177.99) --
	(266.44,178.09) --
	(267.28,178.37) --
	(267.71,178.51) --
	(268.38,178.73) --
	(269.11,178.98) --
	(269.74,179.19) --
	(269.87,179.23) --
	(271.02,179.61) --
	(271.10,179.64) --
	(271.68,179.83) --
	(272.40,180.07) --
	(272.79,180.20) --
	(273.63,180.48) --
	(274.52,180.77) --
	(275.28,181.03) --
	(275.59,181.13) --
	(276.13,181.31) --
	(276.31,181.37) --
	(276.77,181.52) --
	(277.30,181.69) --
	(277.98,181.92) --
	(278.04,181.94) --
	(278.45,182.08) --
	(279.17,182.32) --
	(279.30,182.36) --
	(279.52,182.43) --
	(279.66,182.48) --
	(279.69,182.49) --
	(280.15,182.64) --
	(280.53,182.77) --
	(280.72,182.83) --
	(281.20,182.99) --
	(281.36,183.04) --
	(281.45,183.07) --
	(282.80,183.52) --
	(283.05,183.60) --
	(283.89,183.88) --
	(283.98,183.91) --
	(284.07,183.94) --
	(284.62,184.13) --
	(284.86,184.21) --
	(285.04,184.26) --
	(285.13,184.29) --
	(286.00,184.58) --
	(286.57,184.77) --
	(286.59,184.78) --
	(287.25,185.00) --
	(287.81,185.18) --
	(288.54,185.43) --
	(288.59,185.44) --
	(289.22,185.65) --
	(289.76,185.83) --
	(290.76,186.16) --
	(290.86,186.20) --
	(291.20,186.31) --
	(291.32,186.35) --
	(291.81,186.51) --
	(291.88,186.53) --
	(292.21,186.64) --
	(292.36,186.70) --
	(292.42,186.71) --
	(292.49,186.74) --
	(292.51,186.74) --
	(292.62,186.78) --
	(293.21,186.98) --
	(293.91,187.21) --
	(295.25,187.65) --
	(295.45,187.72) --
	(295.51,187.74) --
	(295.93,187.88) --
	(295.99,187.90) --
	(296.10,187.93) --
	(296.47,188.06) --
	(296.65,188.12) --
	(296.99,188.23) --
	(297.01,188.24) --
	(297.04,188.25) --
	(297.05,188.25) --
	(297.09,188.26) --
	(297.13,188.28) --
	(297.22,188.31) --
	(297.31,188.34) --
	(297.31,188.34) --
	(297.71,188.47) --
	(297.96,188.55) --
	(298.34,188.68) --
	(298.57,188.75) --
	(299.05,188.91) --
	(300.58,189.42) --
	(300.69,189.46) --
	(301.32,189.67) --
	(302.10,189.93) --
	(302.23,189.97) --
	(302.50,190.06) --
	(302.82,190.16) --
	(302.98,190.22) --
	(303.49,190.39) --
	(303.87,190.51) --
	(304.04,190.57) --
	(304.05,190.57) --
	(304.17,190.61) --
	(304.21,190.63) --
	(304.88,190.85) --
	(305.44,191.04) --
	(305.52,191.06) --
	(305.61,191.09) --
	(305.86,191.17) --
	(306.19,191.28) --
	(306.32,191.33) --
	(307.91,191.85) --
	(307.96,191.87) --
	(308.47,192.04) --
	(308.57,192.07) --
	(309.22,192.29) --
	(309.44,192.36) --
	(309.51,192.38) --
	(309.67,192.44) --
	(310.00,192.55) --
	(310.25,192.63) --
	(310.76,192.80) --
	(310.92,192.85) --
	(311.04,192.89) --
	(311.16,192.93) --
	(311.41,193.01) --
	(311.41,193.02) --
	(311.44,193.03) --
	(311.54,193.06) --
	(311.68,193.10) --
	(311.68,193.11) --
	(311.88,193.17) --
	(312.22,193.29) --
	(312.92,193.52) --
	(313.12,193.58) --
	(313.27,193.63) --
	(313.53,193.72) --
	(313.83,193.82) --
	(314.24,193.95) --
	(314.31,193.98) --
	(314.45,194.03) --
	(314.63,194.09) --
	(315.00,194.21) --
	(315.21,194.28) --
	(315.23,194.28) --
	(315.37,194.33) --
	(315.48,194.37) --
	(315.51,194.38) --
	(315.62,194.41) --
	(315.70,194.44) --
	(315.79,194.47) --
	(316.13,194.58) --
	(316.22,194.61) --
	(316.38,194.67) --
	(316.61,194.74) --
	(316.93,194.85) --
	(317.07,194.89) --
	(317.20,194.94) --
	(317.36,194.99) --
	(317.75,195.12) --
	(317.76,195.12) --
	(317.79,195.13) --
	(317.86,195.16) --
	(317.89,195.17) --
	(317.98,195.20) --
	(318.43,195.34) --
	(318.60,195.40) --
	(318.85,195.48) --
	(319.11,195.57) --
	(319.21,195.61) --
	(319.52,195.71) --
	(319.81,195.80) --
	(320.44,196.01) --
	(321.85,196.48) --
	(321.91,196.50) --
	(322.04,196.54) --
	(322.30,196.63) --
	(322.60,196.73) --
	(322.75,196.78) --
	(323.02,196.87) --
	(323.10,196.90) --
	(323.28,196.96) --
	(324.05,197.21) --
	(324.42,197.33) --
	(324.61,197.40) --
	(324.72,197.43) --
	(325.08,197.55) --
	(325.99,197.85) --
	(326.58,198.05) --
	(326.72,198.10) --
	(327.61,198.39) --
	(327.91,198.49) --
	(328.67,198.74) --
	(328.94,198.83) --
	(329.07,198.88) --
	(329.52,199.03) --
	(331.17,199.57) --
	(332.00,199.85) --
	(332.27,199.94) --
	(333.07,200.21) --
	(333.19,200.24) --
	(333.76,200.43) --
	(335.51,201.02) --
	(336.07,201.20) --
	(336.75,201.43) --
	(336.88,201.47) --
	(337.54,201.69) --
	(337.92,201.81) --
	(338.29,201.94) --
	(338.32,201.95) --
	(338.37,201.96) --
	(338.79,202.10) --
	(338.87,202.13) --
	(339.41,202.31) --
	(339.54,202.35) --
	(340.26,202.59) --
	(340.28,202.60) --
	(340.36,202.62) --
	(340.58,202.70) --
	(340.90,202.80) --
	(341.07,202.86) --
	(341.38,202.96) --
	(341.48,203.00) --
	(341.53,203.01) --
	(341.57,203.03) --
	(342.00,203.17) --
	(342.06,203.19) --
	(342.60,203.37) --
	(342.92,203.47) --
	(344.39,203.96) --
	(344.81,204.10) --
	(345.03,204.17) --
	(345.51,204.33) --
	(345.56,204.35) --
	(346.27,204.59) --
	(348.09,205.19) --
	(348.19,205.22) --
	(348.32,205.27) --
	(348.52,205.33) --
	(348.65,205.38) --
	(350.04,205.84) --
	(350.29,205.92) --
	(350.39,205.95) --
	(351.49,206.32) --
	(354.19,207.21) --
	(354.41,207.29) --
	(354.96,207.47) --
	(354.99,207.48) --
	(355.28,207.58) --
	(355.84,207.76) --
	(356.71,208.05) --
	(356.75,208.06) --
	(357.50,208.31) --
	(357.53,208.32) --
	(358.04,208.49) --
	(358.62,208.68) --
	(359.31,208.91) --
	(359.62,209.01) --
	(359.75,209.06) --
	(359.85,209.09) --
	(360.61,209.34) --
	(360.64,209.35) --
	(360.71,209.38) --
	(361.91,209.77) --
	(362.60,210.01) --
	(364.24,210.55) --
	(364.46,210.62) --
	(365.70,211.03) --
	(366.29,211.23) --
	(370.53,212.64) --
	(371.02,212.80) --
	(372.06,213.14) --
	(372.07,213.15) --
	(372.38,213.25) --
	(372.60,213.32) --
	(372.90,213.42) --
	(373.16,213.51) --
	(373.53,213.63) --
	(373.80,213.72) --
	(374.46,213.94) --
	(374.81,214.06) --
	(375.77,214.38) --
	(376.36,214.57) --
	(376.51,214.62) --
	(376.93,214.76) --
	(378.03,215.12) --
	(379.36,215.57) --
	(379.36,215.57) --
	(380.09,215.81) --
	(380.12,215.82) --
	(381.03,216.12) --
	(381.47,216.27) --
	(381.75,216.36) --
	(382.36,216.56) --
	(382.44,216.59) --
	(382.57,216.63) --
	(383.02,216.78) --
	(385.04,217.45) --
	(385.16,217.49) --
	(388.87,218.72) --
	(393.56,220.28) --
	(395.71,220.99) --
	(403.26,223.50) --
	(404.02,223.75);
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.30] ( 52.97,170.47) --
	( 61.01,171.42) --
	( 65.48,171.92) --
	( 65.62,171.94) --
	( 79.64,173.62) --
	( 88.50,174.72) --
	( 94.41,175.36) --
	( 97.84,175.84) --
	(104.50,176.60) --
	(110.01,177.24) --
	(113.94,177.93) --
	(116.56,178.20) --
	(124.94,179.24) --
	(160.01,184.17) --
	(160.15,184.18) --
	(163.56,184.51) --
	(164.50,184.59) --
	(175.08,186.23) --
	(177.68,186.60) --
	(185.12,187.82) --
	(195.04,189.45) --
	(195.05,189.45) --
	(199.79,190.58) --
	(200.09,190.62) --
	(201.70,190.93) --
	(202.42,191.10) --
	(203.15,191.25) --
	(203.24,191.25) --
	(203.92,191.30) --
	(204.08,191.31) --
	(205.70,191.75) --
	(211.30,192.92) --
	(213.02,193.28) --
	(213.22,193.33) --
	(213.28,193.35) --
	(214.88,193.63) --
	(219.39,194.53) --
	(219.87,194.68) --
	(224.51,195.65) --
	(226.38,196.05) --
	(226.89,196.15) --
	(228.22,196.41) --
	(229.80,196.85) --
	(230.53,197.01) --
	(231.54,197.24) --
	(232.29,197.36) --
	(232.55,197.40) --
	(232.60,197.41) --
	(233.03,197.50) --
	(236.50,198.30) --
	(238.44,198.81) --
	(239.50,199.08) --
	(240.76,199.42) --
	(242.87,200.00) --
	(243.82,200.19) --
	(244.87,200.39) --
	(245.90,200.62) --
	(247.14,200.93) --
	(247.15,200.93) --
	(248.00,201.17) --
	(248.59,201.32) --
	(249.15,201.48) --
	(249.91,201.67) --
	(250.14,201.73) --
	(250.71,201.90) --
	(251.21,202.02) --
	(251.61,202.16) --
	(251.88,202.27) --
	(252.52,202.50) --
	(252.98,202.60) --
	(253.12,202.63) --
	(253.36,202.70) --
	(254.91,203.17) --
	(255.02,203.20) --
	(256.32,203.41) --
	(258.61,204.14) --
	(260.03,204.46) --
	(260.14,204.50) --
	(260.80,204.66) --
	(261.97,205.06) --
	(262.19,205.14) --
	(262.48,205.20) --
	(262.99,205.39) --
	(264.29,205.73) --
	(264.50,205.78) --
	(264.53,205.79) --
	(264.97,205.90) --
	(266.10,206.25) --
	(266.13,206.27) --
	(266.44,206.39) --
	(267.28,206.57) --
	(267.71,206.73) --
	(268.38,206.99) --
	(269.11,207.18) --
	(269.74,207.29) --
	(269.87,207.31) --
	(271.02,207.71) --
	(271.10,207.73) --
	(271.68,207.94) --
	(272.40,208.09) --
	(272.79,208.15) --
	(273.63,208.42) --
	(274.52,208.72) --
	(275.28,208.98) --
	(275.59,209.10) --
	(276.13,209.32) --
	(276.31,209.37) --
	(276.77,209.50) --
	(277.30,209.67) --
	(277.98,209.90) --
	(278.04,209.92) --
	(278.45,210.08) --
	(279.17,210.31) --
	(279.30,210.34) --
	(279.52,210.40) --
	(279.66,210.45) --
	(279.69,210.46) --
	(280.15,210.58) --
	(280.53,210.74) --
	(280.72,210.81) --
	(281.20,210.99) --
	(281.36,211.04) --
	(281.45,211.05) --
	(282.80,211.51) --
	(283.05,211.59) --
	(283.89,211.98) --
	(283.98,212.01) --
	(284.07,212.04) --
	(284.62,212.16) --
	(284.86,212.24) --
	(285.04,212.27) --
	(285.13,212.29) --
	(286.00,212.55) --
	(286.57,212.69) --
	(286.59,212.70) --
	(287.25,212.92) --
	(287.81,213.15) --
	(288.54,213.43) --
	(288.59,213.45) --
	(289.22,213.66) --
	(289.76,213.90) --
	(290.76,214.26) --
	(290.86,214.30) --
	(291.20,214.44) --
	(291.32,214.50) --
	(291.81,214.76) --
	(291.88,214.79) --
	(292.21,214.88) --
	(292.36,214.93) --
	(292.42,214.94) --
	(292.49,214.96) --
	(292.51,214.97) --
	(292.62,215.01) --
	(293.21,215.25) --
	(293.91,215.52) --
	(295.25,215.94) --
	(295.45,215.99) --
	(295.51,216.01) --
	(295.93,216.15) --
	(295.99,216.18) --
	(296.10,216.23) --
	(296.47,216.40) --
	(296.65,216.46) --
	(296.99,216.55) --
	(297.01,216.56) --
	(297.04,216.58) --
	(297.05,216.58) --
	(297.09,216.61) --
	(297.13,216.63) --
	(297.22,216.69) --
	(297.31,216.73) --
	(297.31,216.73) --
	(297.71,216.87) --
	(297.96,216.95) --
	(298.34,217.07) --
	(298.57,217.15) --
	(299.05,217.32) --
	(300.58,217.91) --
	(300.69,217.95) --
	(301.32,218.15) --
	(302.10,218.48) --
	(302.23,218.52) --
	(302.50,218.58) --
	(302.82,218.68) --
	(302.98,218.74) --
	(303.49,218.99) --
	(303.87,219.11) --
	(304.04,219.19) --
	(304.05,219.19) --
	(304.17,219.25) --
	(304.21,219.25) --
	(304.88,219.55) --
	(305.44,219.84) --
	(305.52,219.87) --
	(305.61,219.90) --
	(305.86,219.96) --
	(306.19,220.13) --
	(306.32,220.15) --
	(307.91,220.71) --
	(307.96,220.73) --
	(308.47,220.90) --
	(308.57,220.93) --
	(309.22,221.23) --
	(309.44,221.32) --
	(309.51,221.37) --
	(309.67,221.40) --
	(310.00,221.57) --
	(310.25,221.71) --
	(310.76,221.86) --
	(310.92,221.92) --
	(311.04,221.97) --
	(311.16,222.02) --
	(311.41,222.13) --
	(311.41,222.13) --
	(311.44,222.14) --
	(311.54,222.18) --
	(311.68,222.24) --
	(311.68,222.24) --
	(311.88,222.33) --
	(312.22,222.47) --
	(312.92,222.75) --
	(313.12,222.83) --
	(313.27,222.87) --
	(313.53,222.95) --
	(313.83,223.10) --
	(314.24,223.23) --
	(314.31,223.26) --
	(314.45,223.33) --
	(314.63,223.41) --
	(315.00,223.57) --
	(315.21,223.65) --
	(315.23,223.66) --
	(315.37,223.70) --
	(315.48,223.72) --
	(315.51,223.72) --
	(315.62,223.77) --
	(315.70,223.81) --
	(315.79,223.85) --
	(316.13,224.06) --
	(316.22,224.13) --
	(316.38,224.22) --
	(316.61,224.27) --
	(316.93,224.41) --
	(317.07,224.44) --
	(317.20,224.49) --
	(317.36,224.54) --
	(317.75,224.75) --
	(317.76,224.76) --
	(317.79,224.77) --
	(317.86,224.78) --
	(317.89,224.79) --
	(317.98,224.81) --
	(318.43,224.96) --
	(318.60,225.02) --
	(318.85,225.15) --
	(319.11,225.26) --
	(319.21,225.29) --
	(319.52,225.41) --
	(319.81,225.55) --
	(320.44,225.84) --
	(321.85,226.35) --
	(321.91,226.39) --
	(322.04,226.46) --
	(322.30,226.62) --
	(322.60,226.79) --
	(322.75,226.88) --
	(323.02,227.04) --
	(323.10,227.08) --
	(323.28,227.19) --
	(324.05,227.52) --
	(324.42,227.66) --
	(324.61,227.72) --
	(324.72,227.76) --
	(325.08,227.86) --
	(325.99,228.30) --
	(326.58,228.58) --
	(326.72,228.65) --
	(327.61,229.10) --
	(327.91,229.23) --
	(328.67,229.53) --
	(328.94,229.64) --
	(329.07,229.70) --
	(329.52,229.91) --
	(331.17,230.54) --
	(332.00,230.87) --
	(332.27,230.97) --
	(333.07,231.28) --
	(333.19,231.32) --
	(333.76,231.57) --
	(335.51,232.30) --
	(336.07,232.58) --
	(336.75,232.91) --
	(336.88,232.96) --
	(337.54,233.20) --
	(337.92,233.41) --
	(338.29,233.66) --
	(338.32,233.68) --
	(338.37,233.71) --
	(338.79,233.92) --
	(338.87,233.95) --
	(339.41,234.15) --
	(339.54,234.20) --
	(340.26,234.46) --
	(340.28,234.46) --
	(340.36,234.50) --
	(340.58,234.59) --
	(340.90,234.71) --
	(341.07,234.78) --
	(341.38,234.92) --
	(341.48,234.98) --
	(341.53,235.00) --
	(341.57,235.03) --
	(342.00,235.27) --
	(342.06,235.30) --
	(342.60,235.57) --
	(342.92,235.69) --
	(344.39,236.26) --
	(344.81,236.45) --
	(345.03,236.56) --
	(345.51,236.76) --
	(345.56,236.78) --
	(346.27,237.20) --
	(348.09,238.13) --
	(348.19,238.20) --
	(348.32,238.28) --
	(348.52,238.35) --
	(348.65,238.39) --
	(350.04,239.05) --
	(350.29,239.17) --
	(350.39,239.24) --
	(351.49,239.80) --
	(354.19,240.95) --
	(354.41,241.02) --
	(354.96,241.26) --
	(354.99,241.28) --
	(355.28,241.43) --
	(355.84,241.70) --
	(356.71,242.13) --
	(356.75,242.15) --
	(357.50,242.49) --
	(357.53,242.50) --
	(358.04,242.80) --
	(358.62,243.06) --
	(359.31,243.37) --
	(359.62,243.47) --
	(359.75,243.52) --
	(359.85,243.55) --
	(360.61,243.94) --
	(360.64,243.96) --
	(360.71,244.00) --
	(361.91,244.47) --
	(362.60,244.81) --
	(364.24,245.62) --
	(364.46,245.74) --
	(365.70,246.32) --
	(366.29,246.62) --
	(370.53,248.75) --
	(371.02,248.94) --
	(372.06,249.52) --
	(372.07,249.52) --
	(372.38,249.70) --
	(372.60,249.79) --
	(372.90,249.94) --
	(373.16,250.08) --
	(373.53,250.23) --
	(373.80,250.37) --
	(374.46,250.74) --
	(374.81,250.92) --
	(375.77,251.40) --
	(376.36,251.70) --
	(376.51,251.79) --
	(376.93,252.03) --
	(378.03,252.55) --
	(379.36,253.32) --
	(379.36,253.33) --
	(380.09,253.71) --
	(380.12,253.73) --
	(381.03,254.16) --
	(381.47,254.34) --
	(381.75,254.50) --
	(382.36,254.82) --
	(382.44,254.87) --
	(382.57,254.96) --
	(383.02,255.16) --
	(385.04,256.33) --
	(385.16,256.40) --
	(388.87,257.98) --
	(393.56,260.47) --
	(395.71,261.52) --
	(403.26,265.59) --
	(404.02,265.99) --
	(404.02,182.81) --
	(403.26,182.73) --
	(395.71,181.79) --
	(393.56,181.46) --
	(388.87,180.68) --
	(385.16,180.00) --
	(385.04,179.96) --
	(383.02,179.75) --
	(382.57,179.67) --
	(382.44,179.66) --
	(382.36,179.64) --
	(381.75,179.59) --
	(381.47,179.52) --
	(381.03,179.41) --
	(380.12,179.26) --
	(380.09,179.26) --
	(379.36,179.10) --
	(379.36,179.10) --
	(378.03,178.84) --
	(376.93,178.70) --
	(376.51,178.67) --
	(376.36,178.66) --
	(375.77,178.52) --
	(374.81,178.34) --
	(374.46,178.30) --
	(373.80,178.15) --
	(373.53,178.09) --
	(373.16,178.02) --
	(372.90,177.97) --
	(372.60,177.93) --
	(372.38,177.89) --
	(372.07,177.86) --
	(372.06,177.86) --
	(371.02,177.63) --
	(370.53,177.50) --
	(366.29,176.69) --
	(365.70,176.59) --
	(364.46,176.36) --
	(364.24,176.31) --
	(362.60,175.96) --
	(361.91,175.81) --
	(360.71,175.55) --
	(360.64,175.55) --
	(360.61,175.54) --
	(359.85,175.40) --
	(359.75,175.38) --
	(359.62,175.36) --
	(359.31,175.29) --
	(358.62,175.15) --
	(358.04,175.03) --
	(357.53,174.93) --
	(357.50,174.92) --
	(356.75,174.77) --
	(356.71,174.76) --
	(355.84,174.61) --
	(355.28,174.53) --
	(354.99,174.45) --
	(354.96,174.44) --
	(354.41,174.32) --
	(354.19,174.27) --
	(351.49,173.63) --
	(350.39,173.47) --
	(350.29,173.45) --
	(350.04,173.40) --
	(348.65,173.12) --
	(348.52,173.09) --
	(348.32,173.07) --
	(348.19,173.04) --
	(348.09,173.02) --
	(346.27,172.69) --
	(345.56,172.63) --
	(345.51,172.62) --
	(345.03,172.50) --
	(344.81,172.46) --
	(344.39,172.37) --
	(342.92,172.08) --
	(342.60,171.98) --
	(342.06,171.86) --
	(342.00,171.85) --
	(341.57,171.77) --
	(341.53,171.76) --
	(341.48,171.75) --
	(341.38,171.73) --
	(341.07,171.67) --
	(340.90,171.64) --
	(340.58,171.55) --
	(340.36,171.51) --
	(340.28,171.49) --
	(340.26,171.48) --
	(339.54,171.28) --
	(339.41,171.24) --
	(338.87,171.08) --
	(338.79,171.06) --
	(338.37,170.94) --
	(338.32,170.93) --
	(338.29,170.92) --
	(337.92,170.81) --
	(337.54,170.77) --
	(336.88,170.59) --
	(336.75,170.55) --
	(336.07,170.33) --
	(335.51,170.22) --
	(333.76,169.79) --
	(333.19,169.68) --
	(333.07,169.68) --
	(332.27,169.47) --
	(332.00,169.41) --
	(331.17,169.20) --
	(329.52,168.69) --
	(329.07,168.59) --
	(328.94,168.56) --
	(328.67,168.48) --
	(327.91,168.38) --
	(327.61,168.29) --
	(326.72,168.13) --
	(326.58,168.12) --
	(325.99,168.06) --
	(325.08,167.81) --
	(324.72,167.72) --
	(324.61,167.68) --
	(324.42,167.61) --
	(324.05,167.47) --
	(323.28,167.37) --
	(323.10,167.31) --
	(323.02,167.30) --
	(322.75,167.20) --
	(322.60,167.17) --
	(322.30,167.12) --
	(322.04,167.03) --
	(321.91,167.02) --
	(321.85,167.01) --
	(320.44,166.60) --
	(319.81,166.49) --
	(319.52,166.44) --
	(319.21,166.38) --
	(319.11,166.36) --
	(318.85,166.30) --
	(318.60,166.26) --
	(318.43,166.24) --
	(317.98,166.17) --
	(317.89,166.14) --
	(317.86,166.13) --
	(317.79,166.12) --
	(317.76,166.11) --
	(317.75,166.11) --
	(317.36,166.04) --
	(317.20,165.97) --
	(317.07,165.93) --
	(316.93,165.90) --
	(316.61,165.79) --
	(316.38,165.72) --
	(316.22,165.65) --
	(316.13,165.64) --
	(315.79,165.57) --
	(315.70,165.57) --
	(315.62,165.54) --
	(315.51,165.53) --
	(315.48,165.53) --
	(315.37,165.51) --
	(315.23,165.50) --
	(315.21,165.50) --
	(315.00,165.43) --
	(314.63,165.31) --
	(314.45,165.25) --
	(314.31,165.20) --
	(314.24,165.17) --
	(313.83,165.06) --
	(313.53,164.99) --
	(313.27,164.87) --
	(313.12,164.81) --
	(312.92,164.74) --
	(312.22,164.53) --
	(311.88,164.43) --
	(311.68,164.35) --
	(311.68,164.35) --
	(311.54,164.30) --
	(311.44,164.29) --
	(311.41,164.29) --
	(311.41,164.29) --
	(311.16,164.26) --
	(311.04,164.22) --
	(310.92,164.18) --
	(310.76,164.12) --
	(310.25,163.95) --
	(310.00,163.87) --
	(309.67,163.76) --
	(309.51,163.70) --
	(309.44,163.67) --
	(309.22,163.60) --
	(308.57,163.38) --
	(308.47,163.35) --
	(307.96,163.16) --
	(307.91,163.14) --
	(306.32,162.74) --
	(306.19,162.69) --
	(305.86,162.59) --
	(305.61,162.52) --
	(305.52,162.49) --
	(305.44,162.46) --
	(304.88,162.30) --
	(304.21,162.10) --
	(304.17,162.08) --
	(304.05,162.05) --
	(304.04,162.04) --
	(303.87,162.00) --
	(303.49,161.95) --
	(302.98,161.79) --
	(302.82,161.75) --
	(302.50,161.60) --
	(302.23,161.57) --
	(302.10,161.55) --
	(301.32,161.31) --
	(300.69,161.08) --
	(300.58,161.05) --
	(299.05,160.72) --
	(298.57,160.58) --
	(298.34,160.51) --
	(297.96,160.38) --
	(297.71,160.30) --
	(297.31,160.19) --
	(297.31,160.19) --
	(297.22,160.17) --
	(297.13,160.15) --
	(297.09,160.14) --
	(297.05,160.13) --
	(297.04,160.13) --
	(297.01,160.12) --
	(296.99,160.11) --
	(296.65,160.04) --
	(296.47,160.00) --
	(296.10,159.87) --
	(295.99,159.83) --
	(295.93,159.82) --
	(295.51,159.67) --
	(295.45,159.65) --
	(295.25,159.59) --
	(293.91,159.27) --
	(293.21,158.98) --
	(292.62,158.85) --
	(292.51,158.82) --
	(292.49,158.81) --
	(292.42,158.79) --
	(292.36,158.76) --
	(292.21,158.73) --
	(291.88,158.60) --
	(291.81,158.58) --
	(291.32,158.39) --
	(291.20,158.36) --
	(290.86,158.25) --
	(290.76,158.21) --
	(289.76,157.85) --
	(289.22,157.64) --
	(288.59,157.49) --
	(288.54,157.48) --
	(287.81,157.20) --
	(287.25,157.08) --
	(286.59,156.84) --
	(286.57,156.83) --
	(286.00,156.63) --
	(285.13,156.33) --
	(285.04,156.28) --
	(284.86,156.20) --
	(284.62,156.11) --
	(284.07,155.96) --
	(283.98,155.93) --
	(283.89,155.89) --
	(283.05,155.58) --
	(282.80,155.54) --
	(281.45,155.14) --
	(281.36,155.11) --
	(281.20,155.07) --
	(280.72,154.94) --
	(280.53,154.88) --
	(280.15,154.75) --
	(279.69,154.60) --
	(279.66,154.59) --
	(279.52,154.55) --
	(279.30,154.46) --
	(279.17,154.42) --
	(278.45,154.23) --
	(278.04,154.06) --
	(277.98,154.03) --
	(277.30,153.77) --
	(276.77,153.59) --
	(276.31,153.43) --
	(276.13,153.37) --
	(275.59,153.18) --
	(275.28,153.08) --
	(274.52,152.92) --
	(273.63,152.60) --
	(272.79,152.35) --
	(272.40,152.29) --
	(271.68,152.02) --
	(271.10,151.79) --
	(271.02,151.75) --
	(269.87,151.37) --
	(269.74,151.31) --
	(269.11,151.10) --
	(268.38,150.82) --
	(267.71,150.63) --
	(267.28,150.49) --
	(266.44,150.19) --
	(266.13,150.09) --
	(266.10,150.08) --
	(264.97,149.73) --
	(264.53,149.58) --
	(264.50,149.57) --
	(264.29,149.50) --
	(262.99,148.98) --
	(262.48,148.79) --
	(262.19,148.70) --
	(261.97,148.61) --
	(260.80,148.23) --
	(260.14,147.97) --
	(260.03,147.91) --
	(258.61,147.39) --
	(256.32,146.54) --
	(255.02,146.07) --
	(254.91,146.03) --
	(253.36,145.49) --
	(253.12,145.41) --
	(252.98,145.36) --
	(252.52,145.15) --
	(251.88,144.89) --
	(251.61,144.78) --
	(251.21,144.61) --
	(250.71,144.40) --
	(250.14,144.19) --
	(249.91,144.08) --
	(249.15,143.82) --
	(248.59,143.55) --
	(248.00,143.31) --
	(247.15,143.00) --
	(247.14,142.99) --
	(245.90,142.54) --
	(244.87,142.05) --
	(243.82,141.66) --
	(242.87,141.28) --
	(240.76,140.36) --
	(239.50,139.84) --
	(238.44,139.43) --
	(236.50,138.65) --
	(233.03,137.25) --
	(232.60,137.08) --
	(232.55,137.06) --
	(232.29,136.94) --
	(231.54,136.62) --
	(230.53,136.14) --
	(229.80,135.84) --
	(228.22,135.16) --
	(226.89,134.52) --
	(226.38,134.30) --
	(224.51,133.47) --
	(219.87,131.27) --
	(219.39,131.04) --
	(214.88,129.12) --
	(213.28,128.37) --
	(213.22,128.34) --
	(213.02,128.24) --
	(211.30,127.42) --
	(205.70,125.04) --
	(204.08,124.36) --
	(203.92,124.31) --
	(203.24,124.02) --
	(203.15,123.98) --
	(202.42,123.65) --
	(201.70,123.29) --
	(200.09,122.49) --
	(199.79,122.31) --
	(195.05,119.88) --
	(195.04,119.88) --
	(185.12,115.12) --
	(177.68,111.43) --
	(175.08,110.16) --
	(164.50,105.01) --
	(163.56,104.57) --
	(160.15,102.91) --
	(160.01,102.85) --
	(124.94, 84.46) --
	(116.56, 79.86) --
	(113.94, 78.49) --
	(110.01, 76.18) --
	(104.50, 73.14) --
	( 97.84, 69.47) --
	( 94.41, 67.70) --
	( 88.50, 64.37) --
	( 79.64, 59.56) --
	( 65.62, 51.98) --
	( 65.48, 51.89) --
	( 61.01, 49.61) --
	( 52.97, 45.08) --
	cycle;
\definecolor[named]{fillColor}{rgb}{0.20,0.20,0.20}

\path[fill=fillColor,fill opacity=0.50] ( 52.97,159.47) --
	( 61.01,160.59) --
	( 65.48,161.23) --
	( 65.62,161.26) --
	( 79.64,163.39) --
	( 88.50,164.57) --
	( 94.41,165.69) --
	( 97.84,166.20) --
	(104.50,167.06) --
	(110.01,168.00) --
	(113.94,168.69) --
	(116.56,169.17) --
	(124.94,170.52) --
	(160.01,176.92) --
	(160.15,176.94) --
	(163.56,177.51) --
	(164.50,177.71) --
	(175.08,179.80) --
	(177.68,180.16) --
	(185.12,181.57) --
	(195.04,183.80) --
	(195.05,183.80) --
	(199.79,184.69) --
	(200.09,184.75) --
	(201.70,185.11) --
	(202.42,185.27) --
	(203.15,185.46) --
	(203.24,185.48) --
	(203.92,185.62) --
	(204.08,185.66) --
	(205.70,185.99) --
	(211.30,187.30) --
	(213.02,187.66) --
	(213.22,187.71) --
	(213.28,187.73) --
	(214.88,188.04) --
	(219.39,189.00) --
	(219.87,189.14) --
	(224.51,190.27) --
	(226.38,190.71) --
	(226.89,190.77) --
	(228.22,191.06) --
	(229.80,191.48) --
	(230.53,191.63) --
	(231.54,191.83) --
	(232.29,192.09) --
	(232.55,192.17) --
	(232.60,192.19) --
	(233.03,192.29) --
	(236.50,193.20) --
	(238.44,193.74) --
	(239.50,193.98) --
	(240.76,194.31) --
	(242.87,194.91) --
	(243.82,195.16) --
	(244.87,195.48) --
	(245.90,195.82) --
	(247.14,196.17) --
	(247.15,196.18) --
	(248.00,196.50) --
	(248.59,196.63) --
	(249.15,196.79) --
	(249.91,197.01) --
	(250.14,197.09) --
	(250.71,197.23) --
	(251.21,197.34) --
	(251.61,197.49) --
	(251.88,197.55) --
	(252.52,197.67) --
	(252.98,197.76) --
	(253.12,197.80) --
	(253.36,197.88) --
	(254.91,198.29) --
	(255.02,198.30) --
	(256.32,198.66) --
	(258.61,199.32) --
	(260.03,199.67) --
	(260.14,199.69) --
	(260.80,199.89) --
	(261.97,200.20) --
	(262.19,200.28) --
	(262.48,200.35) --
	(262.99,200.50) --
	(264.29,200.85) --
	(264.50,200.95) --
	(264.53,200.97) --
	(264.97,201.14) --
	(266.10,201.49) --
	(266.13,201.51) --
	(266.44,201.63) --
	(267.28,201.85) --
	(267.71,202.01) --
	(268.38,202.15) --
	(269.11,202.33) --
	(269.74,202.52) --
	(269.87,202.56) --
	(271.02,202.98) --
	(271.10,203.00) --
	(271.68,203.15) --
	(272.40,203.42) --
	(272.79,203.55) --
	(273.63,203.81) --
	(274.52,204.12) --
	(275.28,204.39) --
	(275.59,204.51) --
	(276.13,204.65) --
	(276.31,204.70) --
	(276.77,204.86) --
	(277.30,204.96) --
	(277.98,205.21) --
	(278.04,205.24) --
	(278.45,205.38) --
	(279.17,205.57) --
	(279.30,205.61) --
	(279.52,205.68) --
	(279.66,205.73) --
	(279.69,205.73) --
	(280.15,205.88) --
	(280.53,206.00) --
	(280.72,206.07) --
	(281.20,206.23) --
	(281.36,206.28) --
	(281.45,206.31) --
	(282.80,206.75) --
	(283.05,206.86) --
	(283.89,207.18) --
	(283.98,207.21) --
	(284.07,207.24) --
	(284.62,207.46) --
	(284.86,207.55) --
	(285.04,207.61) --
	(285.13,207.65) --
	(286.00,207.95) --
	(286.57,208.14) --
	(286.59,208.15) --
	(287.25,208.36) --
	(287.81,208.58) --
	(288.54,208.78) --
	(288.59,208.79) --
	(289.22,208.99) --
	(289.76,209.13) --
	(290.76,209.50) --
	(290.86,209.53) --
	(291.20,209.68) --
	(291.32,209.73) --
	(291.81,209.89) --
	(291.88,209.92) --
	(292.21,210.06) --
	(292.36,210.13) --
	(292.42,210.16) --
	(292.49,210.17) --
	(292.51,210.18) --
	(292.62,210.20) --
	(293.21,210.39) --
	(293.91,210.61) --
	(295.25,211.13) --
	(295.45,211.21) --
	(295.51,211.24) --
	(295.93,211.39) --
	(295.99,211.41) --
	(296.10,211.45) --
	(296.47,211.59) --
	(296.65,211.64) --
	(296.99,211.74) --
	(297.01,211.75) --
	(297.04,211.76) --
	(297.05,211.76) --
	(297.09,211.78) --
	(297.13,211.81) --
	(297.22,211.85) --
	(297.31,211.88) --
	(297.31,211.88) --
	(297.71,212.06) --
	(297.96,212.13) --
	(298.34,212.29) --
	(298.57,212.42) --
	(299.05,212.60) --
	(300.58,213.21) --
	(300.69,213.25) --
	(301.32,213.49) --
	(302.10,213.78) --
	(302.23,213.83) --
	(302.50,213.93) --
	(302.82,214.09) --
	(302.98,214.15) --
	(303.49,214.33) --
	(303.87,214.46) --
	(304.04,214.52) --
	(304.05,214.53) --
	(304.17,214.56) --
	(304.21,214.57) --
	(304.88,214.80) --
	(305.44,215.00) --
	(305.52,215.03) --
	(305.61,215.07) --
	(305.86,215.17) --
	(306.19,215.29) --
	(306.32,215.33) --
	(307.91,215.92) --
	(307.96,215.93) --
	(308.47,216.11) --
	(308.57,216.15) --
	(309.22,216.41) --
	(309.44,216.48) --
	(309.51,216.51) --
	(309.67,216.58) --
	(310.00,216.73) --
	(310.25,216.84) --
	(310.76,217.04) --
	(310.92,217.08) --
	(311.04,217.15) --
	(311.16,217.18) --
	(311.41,217.26) --
	(311.41,217.26) --
	(311.44,217.27) --
	(311.54,217.31) --
	(311.68,217.37) --
	(311.68,217.37) --
	(311.88,217.45) --
	(312.22,217.58) --
	(312.92,217.88) --
	(313.12,217.96) --
	(313.27,218.01) --
	(313.53,218.11) --
	(313.83,218.21) --
	(314.24,218.39) --
	(314.31,218.42) --
	(314.45,218.47) --
	(314.63,218.54) --
	(315.00,218.69) --
	(315.21,218.76) --
	(315.23,218.77) --
	(315.37,218.83) --
	(315.48,218.87) --
	(315.51,218.88) --
	(315.62,218.94) --
	(315.70,218.97) --
	(315.79,219.01) --
	(316.13,219.13) --
	(316.22,219.18) --
	(316.38,219.25) --
	(316.61,219.34) --
	(316.93,219.46) --
	(317.07,219.50) --
	(317.20,219.56) --
	(317.36,219.63) --
	(317.75,219.81) --
	(317.76,219.81) --
	(317.79,219.82) --
	(317.86,219.85) --
	(317.89,219.87) --
	(317.98,219.89) --
	(318.43,220.06) --
	(318.60,220.13) --
	(318.85,220.21) --
	(319.11,220.33) --
	(319.21,220.36) --
	(319.52,220.49) --
	(319.81,220.61) --
	(320.44,220.89) --
	(321.85,221.43) --
	(321.91,221.45) --
	(322.04,221.50) --
	(322.30,221.59) --
	(322.60,221.73) --
	(322.75,221.78) --
	(323.02,221.89) --
	(323.10,221.92) --
	(323.28,221.97) --
	(324.05,222.30) --
	(324.42,222.44) --
	(324.61,222.53) --
	(324.72,222.58) --
	(325.08,222.73) --
	(325.99,223.18) --
	(326.58,223.43) --
	(326.72,223.49) --
	(327.61,223.81) --
	(327.91,223.95) --
	(328.67,224.23) --
	(328.94,224.34) --
	(329.07,224.39) --
	(329.52,224.58) --
	(331.17,225.26) --
	(332.00,225.59) --
	(332.27,225.71) --
	(333.07,226.08) --
	(333.19,226.13) --
	(333.76,226.44) --
	(335.51,227.21) --
	(336.07,227.44) --
	(336.75,227.74) --
	(336.88,227.79) --
	(337.54,228.09) --
	(337.92,228.27) --
	(338.29,228.43) --
	(338.32,228.44) --
	(338.37,228.47) --
	(338.79,228.65) --
	(338.87,228.69) --
	(339.41,228.95) --
	(339.54,229.03) --
	(340.26,229.31) --
	(340.28,229.33) --
	(340.36,229.36) --
	(340.58,229.46) --
	(340.90,229.60) --
	(341.07,229.67) --
	(341.38,229.82) --
	(341.48,229.87) --
	(341.53,229.89) --
	(341.57,229.91) --
	(342.00,230.08) --
	(342.06,230.11) --
	(342.60,230.32) --
	(342.92,230.44) --
	(344.39,231.06) --
	(344.81,231.24) --
	(345.03,231.34) --
	(345.51,231.53) --
	(345.56,231.55) --
	(346.27,231.84) --
	(348.09,232.60) --
	(348.19,232.65) --
	(348.32,232.69) --
	(348.52,232.77) --
	(348.65,232.83) --
	(350.04,233.41) --
	(350.29,233.53) --
	(350.39,233.57) --
	(351.49,234.02) --
	(354.19,235.25) --
	(354.41,235.35) --
	(354.96,235.62) --
	(354.99,235.64) --
	(355.28,235.76) --
	(355.84,236.01) --
	(356.71,236.45) --
	(356.75,236.48) --
	(357.50,236.93) --
	(357.53,236.94) --
	(358.04,237.18) --
	(358.62,237.46) --
	(359.31,237.76) --
	(359.62,237.90) --
	(359.75,237.96) --
	(359.85,238.02) --
	(360.61,238.35) --
	(360.64,238.36) --
	(360.71,238.39) --
	(361.91,238.92) --
	(362.60,239.27) --
	(364.24,240.07) --
	(364.46,240.16) --
	(365.70,240.71) --
	(366.29,241.00) --
	(370.53,242.95) --
	(371.02,243.20) --
	(372.06,243.72) --
	(372.07,243.72) --
	(372.38,243.88) --
	(372.60,244.00) --
	(372.90,244.15) --
	(373.16,244.28) --
	(373.53,244.45) --
	(373.80,244.56) --
	(374.46,244.87) --
	(374.81,245.06) --
	(375.77,245.55) --
	(376.36,245.83) --
	(376.51,245.90) --
	(376.93,246.11) --
	(378.03,246.60) --
	(379.36,247.20) --
	(379.36,247.20) --
	(380.09,247.56) --
	(380.12,247.58) --
	(381.03,247.95) --
	(381.47,248.18) --
	(381.75,248.28) --
	(382.36,248.58) --
	(382.44,248.63) --
	(382.57,248.68) --
	(383.02,248.88) --
	(385.04,249.84) --
	(385.16,249.90) --
	(388.87,251.64) --
	(393.56,253.73) --
	(395.71,254.78) --
	(403.26,258.41) --
	(404.02,258.78) --
	(404.02,189.12) --
	(403.26,189.00) --
	(395.71,187.56) --
	(393.56,187.20) --
	(388.87,186.37) --
	(385.16,185.62) --
	(385.04,185.60) --
	(383.02,185.20) --
	(382.57,185.13) --
	(382.44,185.10) --
	(382.36,185.07) --
	(381.75,184.98) --
	(381.47,184.90) --
	(381.03,184.85) --
	(380.12,184.68) --
	(380.09,184.68) --
	(379.36,184.57) --
	(379.36,184.57) --
	(378.03,184.35) --
	(376.93,184.17) --
	(376.51,184.09) --
	(376.36,184.06) --
	(375.77,183.96) --
	(374.81,183.76) --
	(374.46,183.67) --
	(373.80,183.49) --
	(373.53,183.44) --
	(373.16,183.39) --
	(372.90,183.34) --
	(372.60,183.30) --
	(372.38,183.28) --
	(372.07,183.23) --
	(372.06,183.23) --
	(371.02,183.00) --
	(370.53,182.90) --
	(366.29,182.01) --
	(365.70,181.91) --
	(364.46,181.68) --
	(364.24,181.61) --
	(362.60,181.33) --
	(361.91,181.20) --
	(360.71,180.94) --
	(360.64,180.93) --
	(360.61,180.92) --
	(359.85,180.78) --
	(359.75,180.74) --
	(359.62,180.70) --
	(359.31,180.61) --
	(358.62,180.42) --
	(358.04,180.29) --
	(357.53,180.20) --
	(357.50,180.20) --
	(356.75,180.11) --
	(356.71,180.09) --
	(355.84,179.88) --
	(355.28,179.78) --
	(354.99,179.71) --
	(354.96,179.70) --
	(354.41,179.60) --
	(354.19,179.57) --
	(351.49,178.95) --
	(350.39,178.69) --
	(350.29,178.67) --
	(350.04,178.64) --
	(348.65,178.28) --
	(348.52,178.24) --
	(348.32,178.21) --
	(348.19,178.18) --
	(348.09,178.15) --
	(346.27,177.73) --
	(345.56,177.59) --
	(345.51,177.59) --
	(345.03,177.50) --
	(344.81,177.45) --
	(344.39,177.37) --
	(342.92,177.09) --
	(342.60,176.99) --
	(342.06,176.84) --
	(342.00,176.81) --
	(341.57,176.73) --
	(341.53,176.72) --
	(341.48,176.71) --
	(341.38,176.70) --
	(341.07,176.63) --
	(340.90,176.58) --
	(340.58,176.50) --
	(340.36,176.44) --
	(340.28,176.41) --
	(340.26,176.40) --
	(339.54,176.24) --
	(339.41,176.20) --
	(338.87,176.07) --
	(338.79,176.05) --
	(338.37,175.97) --
	(338.32,175.96) --
	(338.29,175.95) --
	(337.92,175.84) --
	(337.54,175.76) --
	(336.88,175.64) --
	(336.75,175.61) --
	(336.07,175.46) --
	(335.51,175.33) --
	(333.76,174.95) --
	(333.19,174.78) --
	(333.07,174.75) --
	(332.27,174.53) --
	(332.00,174.48) --
	(331.17,174.26) --
	(329.52,173.83) --
	(329.07,173.73) --
	(328.94,173.70) --
	(328.67,173.61) --
	(327.91,173.43) --
	(327.61,173.34) --
	(326.72,173.08) --
	(326.58,173.04) --
	(325.99,172.91) --
	(325.08,172.70) --
	(324.72,172.63) --
	(324.61,172.62) --
	(324.42,172.57) --
	(324.05,172.46) --
	(323.28,172.27) --
	(323.10,172.22) --
	(323.02,172.20) --
	(322.75,172.14) --
	(322.60,172.09) --
	(322.30,172.02) --
	(322.04,171.94) --
	(321.91,171.92) --
	(321.85,171.91) --
	(320.44,171.52) --
	(319.81,171.28) --
	(319.52,171.22) --
	(319.21,171.14) --
	(319.11,171.10) --
	(318.85,170.99) --
	(318.60,170.95) --
	(318.43,170.89) --
	(317.98,170.75) --
	(317.89,170.72) --
	(317.86,170.71) --
	(317.79,170.71) --
	(317.76,170.70) --
	(317.75,170.70) --
	(317.36,170.64) --
	(317.20,170.61) --
	(317.07,170.58) --
	(316.93,170.55) --
	(316.61,170.45) --
	(316.38,170.41) --
	(316.22,170.39) --
	(316.13,170.36) --
	(315.79,170.31) --
	(315.70,170.29) --
	(315.62,170.27) --
	(315.51,170.25) --
	(315.48,170.24) --
	(315.37,170.22) --
	(315.23,170.19) --
	(315.21,170.18) --
	(315.00,170.14) --
	(314.63,170.04) --
	(314.45,169.99) --
	(314.31,169.95) --
	(314.24,169.93) --
	(313.83,169.83) --
	(313.53,169.75) --
	(313.27,169.68) --
	(313.12,169.64) --
	(312.92,169.59) --
	(312.22,169.43) --
	(311.88,169.36) --
	(311.68,169.32) --
	(311.68,169.32) --
	(311.54,169.27) --
	(311.44,169.24) --
	(311.41,169.23) --
	(311.41,169.23) --
	(311.16,169.18) --
	(311.04,169.16) --
	(310.92,169.13) --
	(310.76,169.07) --
	(310.25,168.92) --
	(310.00,168.83) --
	(309.67,168.76) --
	(309.51,168.72) --
	(309.44,168.70) --
	(309.22,168.66) --
	(308.57,168.53) --
	(308.47,168.52) --
	(307.96,168.35) --
	(307.91,168.33) --
	(306.32,167.83) --
	(306.19,167.78) --
	(305.86,167.65) --
	(305.61,167.56) --
	(305.52,167.54) --
	(305.44,167.51) --
	(304.88,167.36) --
	(304.21,167.16) --
	(304.17,167.15) --
	(304.05,167.13) --
	(304.04,167.12) --
	(303.87,167.09) --
	(303.49,166.97) --
	(302.98,166.82) --
	(302.82,166.77) --
	(302.50,166.68) --
	(302.23,166.59) --
	(302.10,166.55) --
	(301.32,166.30) --
	(300.69,166.08) --
	(300.58,166.04) --
	(299.05,165.49) --
	(298.57,165.30) --
	(298.34,165.21) --
	(297.96,165.07) --
	(297.71,165.02) --
	(297.31,164.88) --
	(297.31,164.88) --
	(297.22,164.85) --
	(297.13,164.81) --
	(297.09,164.80) --
	(297.05,164.79) --
	(297.04,164.79) --
	(297.01,164.78) --
	(296.99,164.78) --
	(296.65,164.70) --
	(296.47,164.63) --
	(296.10,164.50) --
	(295.99,164.48) --
	(295.93,164.47) --
	(295.51,164.37) --
	(295.45,164.35) --
	(295.25,164.28) --
	(293.91,163.93) --
	(293.21,163.71) --
	(292.62,163.48) --
	(292.51,163.45) --
	(292.49,163.44) --
	(292.42,163.42) --
	(292.36,163.39) --
	(292.21,163.34) --
	(291.88,163.28) --
	(291.81,163.25) --
	(291.32,163.09) --
	(291.20,163.06) --
	(290.86,162.93) --
	(290.76,162.90) --
	(289.76,162.58) --
	(289.22,162.41) --
	(288.59,162.24) --
	(288.54,162.22) --
	(287.81,161.98) --
	(287.25,161.77) --
	(286.59,161.56) --
	(286.57,161.55) --
	(286.00,161.34) --
	(285.13,161.02) --
	(285.04,160.98) --
	(284.86,160.92) --
	(284.62,160.82) --
	(284.07,160.64) --
	(283.98,160.62) --
	(283.89,160.59) --
	(283.05,160.27) --
	(282.80,160.20) --
	(281.45,159.74) --
	(281.36,159.72) --
	(281.20,159.66) --
	(280.72,159.52) --
	(280.53,159.44) --
	(280.15,159.32) --
	(279.69,159.16) --
	(279.66,159.15) --
	(279.52,159.10) --
	(279.30,159.03) --
	(279.17,159.01) --
	(278.45,158.83) --
	(278.04,158.67) --
	(277.98,158.65) --
	(277.30,158.39) --
	(276.77,158.21) --
	(276.31,158.02) --
	(276.13,157.96) --
	(275.59,157.78) --
	(275.28,157.68) --
	(274.52,157.44) --
	(273.63,157.14) --
	(272.79,156.90) --
	(272.40,156.80) --
	(271.68,156.52) --
	(271.10,156.29) --
	(271.02,156.28) --
	(269.87,155.91) --
	(269.74,155.86) --
	(269.11,155.66) --
	(268.38,155.37) --
	(267.71,155.15) --
	(267.28,155.00) --
	(266.44,154.66) --
	(266.13,154.57) --
	(266.10,154.55) --
	(264.97,154.19) --
	(264.53,154.04) --
	(264.50,154.03) --
	(264.29,153.96) --
	(262.99,153.57) --
	(262.48,153.41) --
	(262.19,153.29) --
	(261.97,153.18) --
	(260.80,152.79) --
	(260.14,152.50) --
	(260.03,152.48) --
	(258.61,151.94) --
	(256.32,151.04) --
	(255.02,150.67) --
	(254.91,150.65) --
	(253.36,150.08) --
	(253.12,149.99) --
	(252.98,149.94) --
	(252.52,149.83) --
	(251.88,149.57) --
	(251.61,149.49) --
	(251.21,149.34) --
	(250.71,149.16) --
	(250.14,148.92) --
	(249.91,148.82) --
	(249.15,148.56) --
	(248.59,148.38) --
	(248.00,148.19) --
	(247.15,147.89) --
	(247.14,147.88) --
	(245.90,147.34) --
	(244.87,146.93) --
	(243.82,146.50) --
	(242.87,146.14) --
	(240.76,145.44) --
	(239.50,145.01) --
	(238.44,144.54) --
	(236.50,143.73) --
	(233.03,142.41) --
	(232.60,142.21) --
	(232.55,142.19) --
	(232.29,142.09) --
	(231.54,141.79) --
	(230.53,141.39) --
	(229.80,141.08) --
	(228.22,140.35) --
	(226.89,139.79) --
	(226.38,139.59) --
	(224.51,138.88) --
	(219.87,136.94) --
	(219.39,136.76) --
	(214.88,134.81) --
	(213.28,134.08) --
	(213.22,134.04) --
	(213.02,133.96) --
	(211.30,133.22) --
	(205.70,130.66) --
	(204.08,129.95) --
	(203.92,129.87) --
	(203.24,129.55) --
	(203.15,129.52) --
	(202.42,129.19) --
	(201.70,128.88) --
	(200.09,128.20) --
	(199.79,128.09) --
	(195.05,125.93) --
	(195.04,125.93) --
	(185.12,121.41) --
	(177.68,117.89) --
	(175.08,116.76) --
	(164.50,111.80) --
	(163.56,111.33) --
	(160.15,109.66) --
	(160.01,109.60) --
	(124.94, 92.19) --
	(116.56, 87.81) --
	(113.94, 86.49) --
	(110.01, 84.50) --
	(104.50, 81.76) --
	( 97.84, 78.33) --
	( 94.41, 76.61) --
	( 88.50, 73.60) --
	( 79.64, 69.03) --
	( 65.62, 61.78) --
	( 65.48, 61.68) --
	( 61.01, 59.42) --
	( 52.97, 55.16) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 52.97, 34.03) -- ( 52.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 61.01, 34.03) -- ( 61.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 65.48, 34.03) -- ( 65.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 65.62, 34.03) -- ( 65.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.64, 34.03) -- ( 79.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 88.50, 34.03) -- ( 88.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 94.41, 34.03) -- ( 94.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 97.84, 34.03) -- ( 97.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (104.50, 34.03) -- (104.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (110.01, 34.03) -- (110.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (113.94, 34.03) -- (113.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (116.56, 34.03) -- (116.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (124.94, 34.03) -- (124.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (160.01, 34.03) -- (160.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (160.15, 34.03) -- (160.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (163.56, 34.03) -- (163.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (164.50, 34.03) -- (164.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (175.08, 34.03) -- (175.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (177.68, 34.03) -- (177.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (185.12, 34.03) -- (185.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.04, 34.03) -- (195.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.05, 34.03) -- (195.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (199.79, 34.03) -- (199.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (200.09, 34.03) -- (200.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (201.70, 34.03) -- (201.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (202.42, 34.03) -- (202.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.15, 34.03) -- (203.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.24, 34.03) -- (203.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.92, 34.03) -- (203.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (204.08, 34.03) -- (204.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (205.70, 34.03) -- (205.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (211.30, 34.03) -- (211.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.02, 34.03) -- (213.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.22, 34.03) -- (213.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.28, 34.03) -- (213.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.88, 34.03) -- (214.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.39, 34.03) -- (219.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.87, 34.03) -- (219.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (224.51, 34.03) -- (224.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.38, 34.03) -- (226.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.89, 34.03) -- (226.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (228.22, 34.03) -- (228.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.80, 34.03) -- (229.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (230.53, 34.03) -- (230.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.54, 34.03) -- (231.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.29, 34.03) -- (232.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.55, 34.03) -- (232.55, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.60, 34.03) -- (232.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (233.03, 34.03) -- (233.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.50, 34.03) -- (236.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (238.44, 34.03) -- (238.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (239.50, 34.03) -- (239.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.76, 34.03) -- (240.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.87, 34.03) -- (242.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (243.82, 34.03) -- (243.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (244.87, 34.03) -- (244.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (245.90, 34.03) -- (245.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.14, 34.03) -- (247.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.15, 34.03) -- (247.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.00, 34.03) -- (248.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (248.59, 34.03) -- (248.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.15, 34.03) -- (249.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.91, 34.03) -- (249.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.14, 34.03) -- (250.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.71, 34.03) -- (250.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.21, 34.03) -- (251.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.61, 34.03) -- (251.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.88, 34.03) -- (251.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.52, 34.03) -- (252.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.98, 34.03) -- (252.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.12, 34.03) -- (253.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.36, 34.03) -- (253.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (254.91, 34.03) -- (254.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (255.02, 34.03) -- (255.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.32, 34.03) -- (256.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.61, 34.03) -- (258.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.03, 34.03) -- (260.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.14, 34.03) -- (260.14, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.80, 34.03) -- (260.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.97, 34.03) -- (261.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.19, 34.03) -- (262.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.48, 34.03) -- (262.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (262.99, 34.03) -- (262.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.29, 34.03) -- (264.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.50, 34.03) -- (264.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.53, 34.03) -- (264.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.97, 34.03) -- (264.97, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.10, 34.03) -- (266.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.13, 34.03) -- (266.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (266.44, 34.03) -- (266.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.28, 34.03) -- (267.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.71, 34.03) -- (267.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.38, 34.03) -- (268.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.11, 34.03) -- (269.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.74, 34.03) -- (269.74, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.87, 34.03) -- (269.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.02, 34.03) -- (271.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.10, 34.03) -- (271.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.68, 34.03) -- (271.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.40, 34.03) -- (272.40, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.79, 34.03) -- (272.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (273.64, 34.03) -- (273.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (274.52, 34.03) -- (274.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.28, 34.03) -- (275.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.59, 34.03) -- (275.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.13, 34.03) -- (276.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.31, 34.03) -- (276.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (276.77, 34.03) -- (276.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.30, 34.03) -- (277.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.98, 34.03) -- (277.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.04, 34.03) -- (278.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (278.45, 34.03) -- (278.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.17, 34.03) -- (279.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.30, 34.03) -- (279.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.52, 34.03) -- (279.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.66, 34.03) -- (279.66, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.69, 34.03) -- (279.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.15, 34.03) -- (280.15, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.53, 34.03) -- (280.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.72, 34.03) -- (280.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.20, 34.03) -- (281.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.36, 34.03) -- (281.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.45, 34.03) -- (281.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (282.80, 34.03) -- (282.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.05, 34.03) -- (283.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.89, 34.03) -- (283.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.98, 34.03) -- (283.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.07, 34.03) -- (284.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.63, 34.03) -- (284.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (284.86, 34.03) -- (284.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.04, 34.03) -- (285.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (285.13, 34.03) -- (285.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.00, 34.03) -- (286.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.57, 34.03) -- (286.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.59, 34.03) -- (286.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.25, 34.03) -- (287.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.81, 34.03) -- (287.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.54, 34.03) -- (288.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.59, 34.03) -- (288.59, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.22, 34.03) -- (289.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (289.76, 34.03) -- (289.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.76, 34.03) -- (290.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.86, 34.03) -- (290.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.20, 34.03) -- (291.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.32, 34.03) -- (291.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.81, 34.03) -- (291.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.88, 34.03) -- (291.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.21, 34.03) -- (292.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.36, 34.03) -- (292.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.42, 34.03) -- (292.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.49, 34.03) -- (292.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.51, 34.03) -- (292.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.62, 34.03) -- (292.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.21, 34.03) -- (293.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (293.91, 34.03) -- (293.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.25, 34.03) -- (295.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.45, 34.03) -- (295.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.51, 34.03) -- (295.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.93, 34.03) -- (295.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (295.99, 34.03) -- (295.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.10, 34.03) -- (296.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.47, 34.03) -- (296.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.65, 34.03) -- (296.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (296.99, 34.03) -- (296.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.01, 34.03) -- (297.01, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.04, 34.03) -- (297.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.05, 34.03) -- (297.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.09, 34.03) -- (297.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.13, 34.03) -- (297.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.22, 34.03) -- (297.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.31, 34.03) -- (297.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.31, 34.03) -- (297.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.71, 34.03) -- (297.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.96, 34.03) -- (297.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.34, 34.03) -- (298.34, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.57, 34.03) -- (298.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.04, 34.03) -- (299.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.58, 34.03) -- (300.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.69, 34.03) -- (300.69, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (301.32, 34.03) -- (301.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.10, 34.03) -- (302.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.23, 34.03) -- (302.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.50, 34.03) -- (302.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.82, 34.03) -- (302.82, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.98, 34.03) -- (302.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.49, 34.03) -- (303.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.87, 34.03) -- (303.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.04, 34.03) -- (304.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.05, 34.03) -- (304.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.17, 34.03) -- (304.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.21, 34.03) -- (304.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.88, 34.03) -- (304.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.44, 34.03) -- (305.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.52, 34.03) -- (305.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.61, 34.03) -- (305.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (305.86, 34.03) -- (305.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.19, 34.03) -- (306.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.32, 34.03) -- (306.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.91, 34.03) -- (307.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (307.96, 34.03) -- (307.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.47, 34.03) -- (308.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (308.57, 34.03) -- (308.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.22, 34.03) -- (309.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.44, 34.03) -- (309.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.51, 34.03) -- (309.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.67, 34.03) -- (309.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.00, 34.03) -- (310.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.25, 34.03) -- (310.25, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.76, 34.03) -- (310.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.92, 34.03) -- (310.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.04, 34.03) -- (311.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.16, 34.03) -- (311.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.41, 34.03) -- (311.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.41, 34.03) -- (311.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.44, 34.03) -- (311.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.54, 34.03) -- (311.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.68, 34.03) -- (311.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.68, 34.03) -- (311.68, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (311.88, 34.03) -- (311.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.22, 34.03) -- (312.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.92, 34.03) -- (312.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.12, 34.03) -- (313.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.27, 34.03) -- (313.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.53, 34.03) -- (313.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.83, 34.03) -- (313.83, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.24, 34.03) -- (314.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.31, 34.03) -- (314.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.45, 34.03) -- (314.45, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.63, 34.03) -- (314.63, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.00, 34.03) -- (315.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.21, 34.03) -- (315.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.23, 34.03) -- (315.23, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.37, 34.03) -- (315.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.48, 34.03) -- (315.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.51, 34.03) -- (315.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.62, 34.03) -- (315.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.70, 34.03) -- (315.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.79, 34.03) -- (315.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.13, 34.03) -- (316.13, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.22, 34.03) -- (316.22, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.38, 34.03) -- (316.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.61, 34.03) -- (316.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.93, 34.03) -- (316.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.07, 34.03) -- (317.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.20, 34.03) -- (317.20, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.36, 34.03) -- (317.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.75, 34.03) -- (317.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.76, 34.03) -- (317.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.79, 34.03) -- (317.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.86, 34.03) -- (317.86, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.89, 34.03) -- (317.89, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.98, 34.03) -- (317.98, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.43, 34.03) -- (318.43, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.60, 34.03) -- (318.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (318.85, 34.03) -- (318.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.11, 34.03) -- (319.11, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.21, 34.03) -- (319.21, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.52, 34.03) -- (319.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.81, 34.03) -- (319.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.44, 34.03) -- (320.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.85, 34.03) -- (321.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.91, 34.03) -- (321.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.04, 34.03) -- (322.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.30, 34.03) -- (322.30, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.60, 34.03) -- (322.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (322.75, 34.03) -- (322.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.02, 34.03) -- (323.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.10, 34.03) -- (323.10, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.28, 34.03) -- (323.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.05, 34.03) -- (324.05, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.42, 34.03) -- (324.42, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.61, 34.03) -- (324.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (324.72, 34.03) -- (324.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.08, 34.03) -- (325.08, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.99, 34.03) -- (325.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.58, 34.03) -- (326.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.72, 34.03) -- (326.72, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.61, 34.03) -- (327.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.91, 34.03) -- (327.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.67, 34.03) -- (328.67, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (328.94, 34.03) -- (328.94, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.07, 34.03) -- (329.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.52, 34.03) -- (329.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.17, 34.03) -- (331.17, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (332.00, 34.03) -- (332.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (332.27, 34.03) -- (332.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.07, 34.03) -- (333.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.19, 34.03) -- (333.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.76, 34.03) -- (333.76, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (335.51, 34.03) -- (335.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.07, 34.03) -- (336.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.75, 34.03) -- (336.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.88, 34.03) -- (336.88, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.54, 34.03) -- (337.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (337.92, 34.03) -- (337.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.29, 34.03) -- (338.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.32, 34.03) -- (338.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.37, 34.03) -- (338.37, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.79, 34.03) -- (338.79, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (338.87, 34.03) -- (338.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.41, 34.03) -- (339.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (339.54, 34.03) -- (339.54, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.26, 34.03) -- (340.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.28, 34.03) -- (340.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.36, 34.03) -- (340.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.58, 34.03) -- (340.58, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.90, 34.03) -- (340.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.07, 34.03) -- (341.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.38, 34.03) -- (341.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.48, 34.03) -- (341.48, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.53, 34.03) -- (341.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (341.57, 34.03) -- (341.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.00, 34.03) -- (342.00, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.06, 34.03) -- (342.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.60, 34.03) -- (342.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.92, 34.03) -- (342.92, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.39, 34.03) -- (344.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.81, 34.03) -- (344.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.03, 34.03) -- (345.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.51, 34.03) -- (345.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (345.56, 34.03) -- (345.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (346.27, 34.03) -- (346.27, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.09, 34.03) -- (348.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.19, 34.03) -- (348.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.32, 34.03) -- (348.32, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.52, 34.03) -- (348.52, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.65, 34.03) -- (348.65, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.04, 34.03) -- (350.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.29, 34.03) -- (350.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (350.39, 34.03) -- (350.39, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (351.49, 34.03) -- (351.49, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.19, 34.03) -- (354.19, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.41, 34.03) -- (354.41, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.96, 34.03) -- (354.96, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (354.99, 34.03) -- (354.99, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.28, 34.03) -- (355.28, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.84, 34.03) -- (355.84, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.71, 34.03) -- (356.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (356.75, 34.03) -- (356.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.50, 34.03) -- (357.50, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (357.53, 34.03) -- (357.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.04, 34.03) -- (358.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (358.62, 34.03) -- (358.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.31, 34.03) -- (359.31, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.62, 34.03) -- (359.62, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.75, 34.03) -- (359.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (359.85, 34.03) -- (359.85, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.61, 34.03) -- (360.61, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.64, 34.03) -- (360.64, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.71, 34.03) -- (360.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (361.91, 34.03) -- (361.91, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (362.60, 34.03) -- (362.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.24, 34.03) -- (364.24, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (364.46, 34.03) -- (364.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.70, 34.03) -- (365.70, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.29, 34.03) -- (366.29, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (370.53, 34.03) -- (370.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.02, 34.03) -- (371.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.06, 34.03) -- (372.06, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.07, 34.03) -- (372.07, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.38, 34.03) -- (372.38, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.60, 34.03) -- (372.60, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (372.90, 34.03) -- (372.90, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.16, 34.03) -- (373.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.53, 34.03) -- (373.53, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (373.80, 34.03) -- (373.80, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.46, 34.03) -- (374.46, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (374.81, 34.03) -- (374.81, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (375.77, 34.03) -- (375.77, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.36, 34.03) -- (376.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.51, 34.03) -- (376.51, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.93, 34.03) -- (376.93, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (378.03, 34.03) -- (378.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (379.36, 34.03) -- (379.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (379.36, 34.03) -- (379.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (380.09, 34.03) -- (380.09, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (380.12, 34.03) -- (380.12, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.03, 34.03) -- (381.03, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.47, 34.03) -- (381.47, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (381.75, 34.03) -- (381.75, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.36, 34.03) -- (382.36, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.44, 34.03) -- (382.44, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.57, 34.03) -- (382.57, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.02, 34.03) -- (383.02, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.04, 34.03) -- (385.04, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (385.16, 34.03) -- (385.16, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.87, 34.03) -- (388.87, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.56, 34.03) -- (393.56, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (395.71, 34.03) -- (395.71, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (403.26, 34.03) -- (403.26, 41.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (404.02, 34.03) -- (404.02, 41.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	( 35.42,277.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31, 45.41) {-5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,128.66) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 28.31,211.91) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.42, 34.03) --
	(421.57, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 85.59, 20.31) {1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (136.07, 20.31) {2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (186.55, 20.31) {3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (237.03, 20.31) {4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (287.51, 20.31) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (337.99, 20.31) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (388.47, 20.31) {7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (228.50,  4.82) {Ln(Min. City Dist.)$_{t}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,289.08);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 19.11,155.53) {\% $\Delta$ Ln(GDP)$_{t}$};
\end{scope}
\end{tikzpicture}