### 2014-10-28 LaTeX/coverLetter/coverLetter.tex

\documentclass[letterpaper]{article}
\usepackage{graphicx,fullpage}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage[T1]{fontenc}
\usepackage[sc,osf]{mathpazo}

\geometry{
  body={6.5in, 8.5in},
  left=1.0in,
  top=1.25in
}
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}
\setlength\parindent{0em}

% Make lists without bullets
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
  }
}{
  \end{list}
}

\begin{document}
\thispagestyle{empty}

  \begin{minipage}{0.64\linewidth}
Shahryar Minhas \\
Department of Political Science \\
Duke University \\
140 Science Drive, 208 Gross Hall \\
Durham, NC, 27708\\
United States
\end{minipage}
\begin{minipage}{0.45\linewidth}
  \begin{tabular}{lr}
    Phone: & (248) 675-7345 \\
    Fax: & (919) 660-4330\\
    Email: & \href{mailto:shahryar.minhas@duke.edu}{\tt shahryar.minhas@duke.edu} 
    % Website:& \href{http://sites.google.com/site/nilsmetternich/}{\tt ucl.ac.uk/spp/people}
  \end{tabular}
\end{minipage}

\vspace{1.5in}

{Editorial Team of the Journal of Conflict Resolution via submission portal}

\vspace{0.5in}

Dear Colleagues:\\[1ex]

This letter accompanies our submission of a manuscript for your consideration. The manuscript ``Enemy at the Gates: Variation in Economic Growth from Civil Conflict'' applies a subnational spatial analysis to explain variation in macroeconomic performance amongst countries experiencing intrastate conflict. Our theoretical argument, bridges the economic development literature on the importance of cities with extant literature on the effect of armed conflict to provide a novel explanation for the paradox of high macroeconomic growth in conflict ridden countries. We think this is a very timely topic and an important contribution to an ongoing debate about the economic effects of instrastate conflict.\\[1ex]

We look forward to your evaluation of this paper.\\[1ex]

Respectfully submitted,

\vspace{.1in}

\includegraphics [scale=.8]{/Users/janus829/Dropbox/Finances/signature.png}

\vskip 0.5in
\hrule

\end{document}\bye