### 2015-04-17 LaTeX/Presentation/mpsa15.tex

%!TEX TS-program = xelatex
\documentclass[10pt, compress]{beamer}

\usetheme[usetitleprogressbar]{m}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{coordsys} % for number lines
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{dcolumn}
\usepackage{caption}
\usepackage{subfig}
\usepackage{tikz}

% easy commands for number propers
\makeatletter
\def\input@path{{/Users/janus829/Dropbox/Research/Rhodium/Graphics/}}
\makeatother
\graphicspath{{/Users/janus829/Dropbox/Research/Rhodium/Graphics/}}

\title[Enemy at the Gates]{Variation in Economic Growth From Civil Conflict}
\author[Minhas \& Radford]{Shahryar Minhas \& Benjamin Radford}
\institute[Duke University]
{
{\emph{shahryar.minhas@duke.edu \& benjamin.radford@duke.edu}} \\
\medskip
Duke University 
}
\date{\today}

\begin{document}

\frame{\titlepage}

\input{PreezeText}

\plain{Thanks}

\end{document}