### 2015-08-17 LaTeX/Revision/editor.tex
rev letter

\section{Editor}

In revising your paper you should consider the full range of questions raised by each of the reviewers as they offer thoughtful comments and suggestions for improving the paper. I want to highlight some of those points to emphasize the need for your careful attention to them as you work on revisions.

Regarding R1, there are three issues to highlight. 
\begin{enumerate}
\item  First, this reviewer would like to see your main model be run on a fixed effects setup with standard errors clustered on country. \\

\textcolor{blue}{\emph{stuff}}

\item Second, this reviewer would like to see country year observations without conflict included in the analyses instead of being excluded. I should note that R2 raises these two concerns as well which suggests that it will be critical to address these issues in your revisions. \\

\textcolor{blue}{\emph{stuff}}

\item Third, this reviewer would like to see you estimate a baseline model for the impact of conflict on growth in which GDP growth is regressed on a dummy for armed conflict in a country. After that, you can then move onto your central argument and compare your findings to this baseline set of findings. \\

\textcolor{blue}{\emph{stuff}}

\end{enumerate}

Regarding R2, there is one additional issue to highlight (points one and two overlap with R1).

\begin{enumerate}
\item The third point raised is to run a robustness check on the more disaggregated ACLED data set to see if your findings hold up. \\

\textcolor{blue}{\emph{stuff}}

\end{enumerate}