### 2015-08-17 LaTeX/Revision/reviseLetter.tex

\documentclass[12pt,onesided,fullpage]{amsart}

%%% PAGE DIMENSIONS
\usepackage{multirow}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry} 
%\pdfpagewidth=8.5in % for pdflatex
%\pdfpageheight=11in % for pdflatex

\usepackage{graphicx} % support the \includegraphics command and options

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array,amsmath,graphicx,psfrag,amssymb,subfigure,tabularx,booktabs}
\usepackage{mparhack}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{multicol}
\usepackage{color}
\usepackage{dcolumn}
\usepackage{hyperref}
\usepackage{stmaryrd}
\setcitestyle{authordate,round,semicolon,aysep={,},yysep={,}}
\bibpunct[:]{(}{)}{;}{a}{,}{,}

\doublespacing

\begin{document}
\singlespacing

\title[JCR-14-0358​]{JCR-14-0358​ entitled ``Enemy at the Gates: Variation in Economic Growth from Civil Conflict''}

\date{\today~~Version 0.01}
\maketitle

$\hspace{-5mm}$Dear Professor Huth, \\ [1ex]

We first would like to thank you for the opportunity to revise and resubmit our manuscript. We believe the manuscript has greatly benefitted from the Reviewers' helpful and thoughtful comments. We have thoroughly revised the manuscript, taking seriously each individual point raised by the Reviewers. The revision memo is organized by first responding to your comments and then addressing the reviewers' points. Our comments and responses are shown in \textcolor{blue}{\emph{BLUE}} below each point.

We hope you agree that the manuscript has greatly improved through this helpful process and we are looking forward to your response.\\ [1ex]

Sincerely, \\ [1ex]

The Authors.

\input{editor.tex}
\input{rev1.tex}
\input{rev2.tex}

\newpage\tiny
\end{document}\bye