### 2015-08-26 LaTeX/coverLetter/rr_coverLetter.tex
added cover letter for resubmission

\documentclass[letterpaper]{article}
\usepackage{graphicx,fullpage}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage[T1]{fontenc}
\usepackage[sc,osf]{mathpazo}

\geometry{
  body={6.5in, 8.5in},
  left=1.0in,
  top=1.25in
}
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}
\setlength\parindent{0em}

% Make lists without bullets
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
  }
}{
  \end{list}
}

\begin{document}
\thispagestyle{empty}

  \begin{minipage}{0.64\linewidth}
Shahryar Minhas \\
Department of Political Science \\
Duke University \\
140 Science Drive, 208 Gross Hall \\
Durham, NC, 27708\\
United States
\end{minipage}
\begin{minipage}{0.45\linewidth}
  \begin{tabular}{lr}
    Phone: & (248) 675-7345 \\
    Fax: & (919) 660-4330\\
    Email: & \href{mailto:shahryar.minhas@duke.edu}{\tt shahryar.minhas@duke.edu} 
    % Website:& \href{http://sites.google.com/site/nilsmetternich/}{\tt ucl.ac.uk/spp/people}
  \end{tabular}
\end{minipage}

\vspace{1.5in}

{Editorial Team of the Journal of Conflict Resolution via submission portal}

\vspace{0.5in}

Dear Colleagues:\\[1ex]

This letter accompanies our resubmission of the manuscript ``Enemy at the Gates: Variation in Economic Growth from Civil Conflict'' (Manuscript ID: JCR- 14-0358), which applies a subnational spatial analysis to explain variation in macroeconomic performance amongst countries experiencing intrastate conflict. We feel confident that the revisions have made this a stronger manuscript and continue to think that the topic addressed by this manuscript is timely and serves an important contribution to the ongoing debate about the economic effects of instrastate conflict.\\[1ex]

We look forward to your evaluation of this paper.\\[1ex]

Respectfully submitted,

\vspace{.1in}

\includegraphics [scale=.8]{/Users/s7m/Dropbox/Finances/signature.png}

\vskip 0.5in
\hrule

\end{document}\bye