### 2015-10-25 LaTeX/Revision2/reviseLetter2.tex

\documentclass[12pt,onesided,fullpage]{amsart}

%%% PAGE DIMENSIONS
\usepackage{multirow}
\usepackage{placeins}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry} 
%\pdfpagewidth=8.5in % for pdflatex
%\pdfpageheight=11in % for pdflatex

\usepackage{graphicx} % support the \includegraphics command and options

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array,amsmath,graphicx,psfrag,amssymb,subfigure,tabularx,booktabs}
\usepackage{mparhack}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{multicol}
\usepackage{color}
\usepackage{dcolumn}
\usepackage{hyperref}
\usepackage{stmaryrd}
\setcitestyle{authordate,round,semicolon,aysep={,},yysep={,}}
\bibpunct[:]{(}{)}{;}{a}{,}{,}

\doublespacing

\begin{document}
\singlespacing

\title[JCR-14-0358.R1]{JCR-14-0358.R1 entitled ``Enemy at the Gates: Variation in Economic Growth from Civil Conflict''}

\date{\today}
\maketitle

$\hspace{-5mm}$Dear Professor Huth, \\ [1ex]

We first would like to thank you for the opportunity to revise and resubmit our manuscript once again. We have incorporated each of the comments made by R2 into the revised manuscript. The revision memo is organized in response to the bulleted set of questions that were posted with the decision letter. Our comments and responses are shown in \textcolor{blue}{\emph{BLUE}} below each point.

We hope you agree that the manuscript has greatly improved through this helpful process and we are looking forward to your response.\\ [1ex]

Sincerely, \\ [1ex]

The Authors.

\input{response.tex}

\newpage\tiny
\end{document}\bye