### 2015-11-27 LaTeX/RhodiumAppendix.tex
new version of manuscript and letter, ready to send

\newpage
\section{Appendix}
\label{appendix}

\subsection{Full Panel Fixed Effects Analysis}

While we contend that random effects are preferred to fixed effects for the reasons outlined in the body of the paper, we have nonetheless repeated our analysis using fixed effects on a full country-year panel dataset. The results are presented in Tables~\ref{tab:cityFullPiecewiseFE} and \ref{tab:capFullPiecewiseFE}. These analyses are performed on the full panel of data and therefore are missing conflict-specific covariates that are unavailable for countries not experiencing civil war. When utilizing this alternative approach we still find strong support for the hypothesis that countries experiencing conflict that is more proximate to capital or major cities are more likely to see significant reductions in GDP growth than conflicts.

\input{cityFullPiecewiseFE.tex}
\FloatBarrier

\input{capFullPiecewiseFE.tex}
\FloatBarrier

\newpage
\subsection{ACLED Analysis}
\label{acled}

The Armed Conflict Location and Event Dataset provides an alternative source of information on the subnational spatial distribution of armed conflict \citep{raleigh:linke:etal:2010}. This dataset is, at the time of writing, limited to Africa and therefore was not selected for the primary analysis presented in the text. It does, however, offer us a valuable opportunity to validate our results. Here, we have replicated the primary model described in Section~\ref{empirics}.

In order to match our existing data structure, it was necessary to aggregate the ACLED data to the country-year level. We did this by first subsetting ACLED to the years 1989-2008 and then selecting only conflict sites with at least 25 fatalities in each given year. The 25 fatalities threshold is intended to mirror the PRIO coding criteria and to prevent very low-fatality events from biasing our estimates of conflict location toward high-population areas. Covariates created with PRIO but unavailable in ACLED are omitted. The analysis procedure then continues as described in Section~\ref{empirics}: the minimum distance measure is calculated as the natural logarithm of the average distance in kilometers from any conflict site to the nearest major city (or capital). The results are presented in Table~\ref{tab:acledRegResults} below.

\input{acledModels.tex}
\FloatBarrier
\newpage

\subsection{Estimating Models Separately for High and Low Intensity Conflicts}

Here instead of treating conflict intensity as a control, we re-do our primary random effect regression models estimating the effect of distance on growth, but restricting to the civil conflicts coded as wars and then a separate model for civil conflicts coded as low intensity events. In both low intensity and high intensity cases we find that the conflict distance variables remain significant and in the expected direction, and the $\beta$ estimate of our distance variables is noticeably higher when using high intensity versus low intensity civil conflict cases. The results are presented in Table~\ref{tab:modHiLoIntensity} below.

\input{modelHiLoIntensity.tex}
\FloatBarrier
\newpage

\subsection{Baseline Effect of Civil Conflict on Growth}

In Table \ref{tab:baseline} below, we show the fixed effects regression results for the model in which we utilize a full country year panel in order to define a baseline effect of civil war on economic growth.

\input{baseline.tex}
\FloatBarrier
\newpage

\subsection{Conflict Distance Models in Tabular Format}

Here we present the results of our models measuring the effect of conflict distance in a tabular format, results for distance from the capital city are shown in the first column of Table \ref{tab:capminDistMod} and for any major city in the second column.

\input{capminDistRegMain.tex}

\FloatBarrier
\newpage