### 2015-11-28 LaTeX/Revision2/reviseLetter2.tex
minor edits to letter, still need to recompile

\documentclass[12pt,onesided,fullpage]{amsart}

%%% PAGE DIMENSIONS
\usepackage{multirow}
\usepackage{placeins}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry} 
%\pdfpagewidth=8.5in % for pdflatex
%\pdfpageheight=11in % for pdflatex

\usepackage{graphicx} % support the \includegraphics command and options

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array,amsmath,graphicx,psfrag,amssymb,subfigure,tabularx,booktabs}
\usepackage{mparhack}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{multicol}
\usepackage{color}
\usepackage{dcolumn}
\usepackage{hyperref}
\usepackage{stmaryrd}
\setcitestyle{authordate,round,semicolon,aysep={,},yysep={,}}
\bibpunct[:]{(}{)}{;}{a}{,}{,}

% -------------------- customizations -------------------- %

% references to graphics
\makeatletter
\def\input@path{{/Users/janus829/Dropbox/Research/Rhodium/Graphics/}, {/Users/Ben/Dropbox/Rhodium/Graphics/}, {/Users/s7m/Dropbox/Research/Rhodium/Graphics/}}
\makeatother
\graphicspath{{/Users/janus829/Dropbox/Research/Rhodium/Graphics/}, {/Users/Ben/Dropbox/Rhodium/Graphics/}, {/Users/s7m/Dropbox/Research/Rhodium/Graphics/}}

% -------------------------------------------------------- %

\doublespacing

\begin{document}
\singlespacing

\title[JCR-14-0358.R1]{JCR-14-0358.R1 entitled ``Enemy at the Gates: Variation in Economic Growth from Civil Conflict''}

\date{\today}
\maketitle

$\hspace{-5mm}$Dear Professor Huth, \\ [1ex]

Thank you for the opportunity to revise and resubmit our manuscript. We have addressed the remaining comments made by R2 and have made the relevant modifications to the revised manuscript. The revision memo is organized in response to the bulleted set of questions that were posted with the decision letter. Our comments and responses are shown in \textcolor{blue}{\emph{BLUE}} below each point.

We are looking forward to your response.\\ [1ex]

Sincerely, \\ [1ex]

The Authors.

\input{response.tex}

\newpage\tiny
\end{document}\bye