### 2016-01-11 LaTeX/EnemyAtTheGates_MinhasRadford_OnlineAppendix.tex
Added online appendix.

\documentclass[12pt,onesided]{amsart}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% PREAMBLE %%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% -------------------- defaults -------------------- %
% load lots o' packages

% layout control
\usepackage{geometry}
\geometry{verbose,tmargin=1.25in,bmargin=1.25in,lmargin=1.1in,rmargin=1.1in}
\usepackage[figuresright]{rotating}
\newenvironment{amssidewaysfigure}
  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
  {\end{minipage}\end{sidewaysfigure}}
\usepackage{parallel}
\usepackage{parcolumns}

% math typesetting
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

\usepackage[%
decimalsymbol=.,
digitsep=fullstop
]{siunitx}

% to adapt caption style
\usepackage[font={small},labelfont=bf]{caption}

% references
\usepackage{natbib}
% \usepackage[natbib=true, style=authoryear]{biblatex}
% \bibliography{master.bib}

% footnotes at bottom
\usepackage[bottom]{footmisc}

% to change enumeration symbols begin{enumerate}[(a)]
\usepackage{enumerate}

% to make enumerations and itemizations within paragraphs or
% lines. f.i. begin{inparaenum} for (a) is (b) and (c)
\usepackage{paralist}

% to colorize links in document. See color specification below
\usepackage[x11names]{xcolor}

% load the hyper-references package and set document info
\usepackage[pdftex]{hyperref}

% graphics stuff
\usepackage{subfig}
\usepackage{graphicx}
\usepackage[space]{grffile} % allows us to specify directories that have spaces
\usepackage[section]{placeins} % prevents floats from moving past a \FloatBarrier or section
\usepackage{tikz}

% Spacing
\usepackage[doublespacing]{setspace}

% define clickable links and their colors
\hypersetup{
	unicode=false,          % non-Latin characters in Acrobat's bookmarks
	pdftoolbar=true,        % show Acrobat's toolbar?
	pdfmenubar=true,        % show Acrobat's menu?
	pdffitwindow=false,     % window fit to page when opened
	pdfstartview={FitH},    % fits the width of the page to the window
	pdfnewwindow=true,%
	pdfauthor={Minhas and Radford},%
	pdftitle={Enemy at the Gates},%
	colorlinks,%
	citecolor=black,%
	filecolor=black,%
	linkcolor=black,%
	urlcolor=RoyalBlue4%
	}
% -------------------------------------------------- %

% -------------------- title -------------------- %

\title[Enemy at the Gates]{Enemy at the Gates: Variation in Economic Growth from Civil Conflict}
\date{\today}

\author[Minhas]{Shahryar Minhas}
\address{Shahryar Minhas: Department of Political Science}
\curraddr{Duke University, Durham, NC, 27708, USA}
\email{shahryar.minhas@duke.edu}

\author[Radford]{Benjamin J. Radford}
\address{Benjamin J. Radford: Department of Political Science}
\curraddr{Duke University, Durham, NC, 27708, USA}
\email{benjamin.radford@duke.edu}

% \thanks{We are grateful for comments on earlier versions of this paper received at the 72$^{nd}$ annual Midwest Political Science Association Conference in Chicago, April 2-6 2014. }

% ----------------------------------------------- %

% -------------------- customizations -------------------- %

% easy commands for number propers
\makeatletter
\def\input@path{{/Users/janus829/Dropbox/Research/Rhodium/Graphics/}, {/Users/Ben/Dropbox/Rhodium/Graphics/}, {/Users/s7m/Dropbox/Research/Rhodium/Graphics/}, {C:/Users/Ben/Dropbox/Rhodium/Graphics}}
\makeatother
\graphicspath{{/Users/janus829/Dropbox/Research/Rhodium/Graphics/}, {/Users/Ben/Dropbox/Rhodium/Graphics/}, {/Users/s7m/Dropbox/Research/Rhodium/Graphics/}, {C:/Users/Ben/Dropbox/Rhodium/Graphics/}}

% -------------------------------------------------------- %

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% DOCUMENT %%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\doublespacing

\begin{document}
% Wordcount: 9814
\maketitle

\begin{abstract}
\singlespacing
\input{Abstract}
\end{abstract}

\newpage
\newpage\setcounter{page}{1}

%%%%% Appendix %%%%%
\input{RhodiumOnlineAppendix}
%%%%%%%%%%%%%%%%%%%%%%

\newpage

% \bibliographystyle{elsarticle-harv} 
\bibliographystyle{jpr}
\bibliography{master.bib}

\end{document}